### 2017-01-20 analysis/index.Rmd
Files commited by wflow_commit.

---
title: "Home"
output:
  html_document:
    toc: false
---

Click on this [link](first-analysis.html) to see my results.

The goal of this new template is to simplify the setup and maintenance of a research website.
Specifically,

*  Easier to build and extend the website using the new tools in the [rmarkdown][] package and [latest RStudio release][rstudio]
*  Easier to deploy the website with Git and GitHub by only using one branch

[rmarkdown]: http://rmarkdown.rstudio.com/rmarkdown_websites.htm
[rstudio]: https://www.rstudio.com/products/rstudio/download/preview/

### 2017-02-01 analysis/index.Rmd
Files commited by wflow_commit.

@@
 
-Click on this [link](first-analysis.html) to see my results.
-
-The goal of this new template is to simplify the setup and maintenance of a research website.
-Specifically,
-
-*  Easier to build and extend the website using the new tools in the [rmarkdown][] package and [latest RStudio release][rstudio]
-*  Easier to deploy the website with Git and GitHub by only using one branch
-
-[rmarkdown]: http://rmarkdown.rstudio.com/rmarkdown_websites.htm
-[rstudio]: https://www.rstudio.com/products/rstudio/download/preview/
+Click on this [orthogonal CV](miss_position.html) to see my example.

### 2017-02-01 analysis/miss_position.Rmd
Files commited by wflow_commit.

---
title: "corss validation of missing value"
author: "wei wang"
date: 2017-01-30
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`

## 5-fold CV

start from five fold and can easily extend to 10 fold or K -fold.

### code

To generate the cross validation data for missing value, I randomly pick the missing possition for each fold.

the code is like the following:

```{r}

cross_plot = function(Y,k_fold = 5){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = matrix(sample(P,P),ncol = k_fold)
  rowindex = matrix(sample(N,N),ncol = k_fold)

  missing= array(0,dim = c(k_fold,N,P))
  foldindex = array(0,dim = c(k_fold,k_fold,2))
  for(i in 1:k_fold){
    for(j in 1:k_fold){
      foldindex[i,j,1] = i
      foldindex[i,j,2] = (i+j) %% k_fold
    }
  }
  foldindex[which(foldindex == 0)] = k_fold
  for(i in 1:k_fold){
    missing[i, , ] = Y
    for(j in 1:k_fold){
      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
    }
  }
  par(mfrow = c(2,3), mar = (c(5,4,4,2)-1.9))
  image(Y,main = "data Y")
  for(i in 1:k_fold){
    image(missing[i,,],main = "missing data")
  }
  return(missing)
}

```

### block design

Starting from simplest case, where Y is $5 \times 5$ matrix.

```{r}
N = 5
P = 5
Y = matrix(rep(1,N*P),ncol =  P)
g = cross_plot(Y)
```

### permute rows and columns

for larger matrix

```{r}
N = 50
P = 100
Y = matrix(rep(1,N*P),ncol =  P)
g = cross_plot(Y)
```
## Session Information

```{r session-info}
```

### 2017-02-02 analysis/miss_position.Rmd
Files commited by wflow_commit.

@@
 **Code version:** `r workflowr::extract_commit(".", 1)$sha1`
+
+## 3-fold CV
+
+take 3-fold CV as example.
+We want to divide the data $Y$ in to 3 parts to do 3-fold CV.
+We randomly divide the the columns into 3 sets and the rows into 3 sets as well.
+The data matrix $Y$ is divided into 9 partition (by row and column permutation):
+
+$$Y = \begin{pmatrix} Y_{11} & Y_{12} & Y_{13} \\ Y_{21} & Y_{22} & Y_{23} \\ Y_{31} & Y_{32} & Y_{33} \end{pmatrix}$$
+
+Then $Y_{(1)} = \{  Y_{11},  Y_{22} , Y_{33}\}$, $Y_{(2)} = \{  Y_{12},  Y_{23} , Y_{31}\}$ and $Y_{(3)} = \{  Y_{13},  Y_{21} , Y_{32}\}$ are orthogonal to each other.
+Then the data matrix $Y$ is marked as:
+$$Y = \begin{pmatrix} Y_{(1)} & Y_{(2)} & Y_{(3)} \\ Y_{(3)} & Y_{(1)} & Y_{(2)} \\ Y_{(2)} & Y_{(3)} & Y_{(1)} \end{pmatrix}$$
+
+For complete data set, we would use scheme 1.
+The imputation methods we consider are based on the model with low rank structure: $LF^T$, so each $Y_{(i)}$ in scheme 1 contains equally one set of indices of $L_{.k}$ and $F_{k.}$.
+This also helps to make sure that all $i$'s and $j$'s are included into each $Y_{-(i)}$.
+Otherwise, we can't make prediction when some times whole columns or rows are missing.
+
+$$Y=  \begin{pmatrix} Y_{11} & Y_{12} \\ Y_{21} & Y_{22}  \end{pmatrix}= \begin{bmatrix} L^{(1)}F^{(1)} & L^{(1)}F^{(2)} \\ L^{(2)}F^{(1)} & L^{(2)}F^{(2)} \end{bmatrix} + E = \left[ \begin{array}{c} L^{(1)} \\ L^{(2)} \end{array} \right] \times \left[ \begin{array}{cc} F^{(1)}& F^{(2)} \end{array} \right] + E$$
+
+In this case, $Y_{(1)} = \{  Y_{11},  Y_{22} \}$, $Y_{(2)} = \{  Y_{12},  Y_{21} \}$, and each of $L^{(1)},F^{(1)},L^{(2)},F^{(2)}$ shows up in both $Y_{(1)}$ and $Y_{(2)}$ so the information contains in each $Y_{(i)}$ are in balance.
 

### 2017-02-07 analysis/first-analysis.Rmd
Files commited by wflow_commit.

---
title: "outline of questions"
author: "Wei Wang"
date: 2017-02-07
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`

## list of questions

let's write the $X$ into $2 \times2$ partitions.

$$X=  \begin{pmatrix} X_{11} & X_{12} \\ X_{21} & X_{22}  \end{pmatrix}= \left[ \begin{array}{c} U_{1} \\ U_{2} \end{array} \right] \Gamma \left[ \begin{array}{cc} V_{1}^{T}& V_{2}^{T} \end{array} \right] $$

### 1 algorithm to find original svd

we want to find a solution for

\begin{eqnarray}
&\min_{U,V,\Gamma}& ||X_{11} - U_1 \Gamma V_1^T||_2 + ||X_{22} - U_2 \Gamma V_2^T||_2 \\
&st& U_1 U_1^T +  U_2 U_2^T = I \\
& & V_1 V_1^T +  V_2 V_2^T = I
\end{eqnarray}

#### 1.1 convexity of this problem

- We are not sure if this optimazation problem is convex or not.

#### 1.2 uniqueness of the solution

- The solution of the above problem is unique or not?

#### 1.3 computational complexity

- To solve this problem might be more computationaly intensive comparing with BCV which just need to calculate svd and general inverse.

### 2 separate svd vs original svd

we know that

\begin{eqnarray}
X_{11} = U_1 \Gamma V_1^T \\
X_{22} = U_2 \Gamma V_2^T
\end{eqnarray}

but $U_1 \Gamma V_1^T$ is not SVD for $X_{11}$. On the other hand, if we combine the svd for $X_{11} = \tilde{U}_1 \Sigma_{11}\tilde{V}_1$ and that for $X_{22} = \tilde{U}_2 \Sigma_{22}\tilde{V}_2$, the combining columns $\left[ \begin{array}{c} \tilde{U}_{1} \\ \tilde{U}_{2} \end{array}  \right]$ is not the original svd $\left[ \begin{array}{c} U_{1} \\ U_{2} \end{array} \right]$. It's because $\tilde{U}_{1}$ and $\tilde{U}_{2}$ just contains local information for $X_{11}$ and $X_{22}$.

### 2017-02-13 analysis/sd_chr1_asinh.Rmd
Files commited by wflow_commit.

---
title: "FLASH on chrommarks_chipcounts_example_regions"
author: "wei wang"
date: 2017-02-09
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`

```{r,cache=TRUE,eval=TRUE,message=FALSE,warning=FALSE}

#GM12878 = readRDS("~/HG/flash_workflow/data/chrommarks_chipcounts_example_regions/GM12878_10marks_meanChIP_chr1_region_asinh.rds")
#Y = GM12878[,1:9]
#g_flash = flashr::greedy(Y, K = 8, flash_para = list(partype = "var_col"))
#b_flash = flashr::backfitting(Y,initial_list = g_flash, flash_para = list(partype = "var_col"), maxiter_bf = 10)
#FL_GM = b_flash
#saveRDS(FL_GM, "~/HG/flash_workflow/data/chrommarks_chipcounts_example_regions/FL_GM_var_col.rds")
FL_GM = readRDS("~/HG/flash_workflow/data/chrommarks_chipcounts_example_regions/FL_GM_var_col.rds")
```

```{r,cache=TRUE,eval=TRUE,message=FALSE,warning=FALSE}

HUVEC = readRDS("~/HG/flash_workflow/data/chrommarks_chipcounts_example_regions/HUVEC_10marks_meanChIP_chr1_region_asinh.rds")
#Y = HUVEC[,1:9]
#g_flash = flashr::greedy(Y, K = 8, flash_para = list(partype = "var_col"))
#b_flash = flashr::backfitting(Y,initial_list = g_flash, flash_para = list(partype = "var_col"), maxiter_bf = 10)
#FL_HUVEC = b_flash
#saveRDS(FL_HUVEC, "~/HG/flash_workflow/data/chrommarks_chipcounts_example_regions/FL_HUVEC_var_col.rds")
```

```{r,cache=TRUE,eval=TRUE,message=FALSE,warning=FALSE}
#K562 = readRDS("~/HG/flash_workflow/data/chrommarks_chipcounts_example_regions/K562_10marks_meanChIP_chr1_region_asinh.rds")
#Y = K562[,1:9]
#g_flash = flashr::greedy(Y, K = 8, flash_para = list(partype = "var_col"))
#b_flash = flashr::backfitting(Y,initial_list = g_flash, flash_para = list(partype = "var_col"), maxiter_bf = 10)
#FL_K562 = b_flash
#saveRDS(FL_K562, "~/HG/flash_workflow/data/chrommarks_chipcounts_example_regions/FL_K562_var_col.rds")
FL_K562 = readRDS("~/HG/flash_workflow/data/chrommarks_chipcounts_example_regions/FL_K562_var_col.rds")
```

```{r,cache=TRUE,eval=TRUE,message=FALSE,warning=FALSE}
#NHLF = readRDS("~/HG/flash_workflow/data/chrommarks_chipcounts_example_regions/NHLF_10marks_meanChIP_chr1_region_asinh.rds")
#Y = NHLF[,1:9]
#g_flash = flashr::greedy(Y, K = 8, flash_para = list(partype = "var_col"))
#b_flash = flashr::backfitting(Y,initial_list = g_flash, flash_para = list(partype = "var_col"), maxiter_bf = 10)
#FL_NHLF = b_flash
#saveRDS(FL_NHLF, "~/HG/flash_workflow/data/chrommarks_chipcounts_example_regions/FL_NHLF_var_col.rds")
FL_NHLF = readRDS("~/HG/flash_workflow/data/chrommarks_chipcounts_example_regions/FL_NHLF_var_col.rds")
```

## factor plot

### GM12878

```{r,echo=FALSE}
par(mfrow = c(2,2))
for(i in 1:4) barplot(FL_GM$f[,i],names = colnames(HUVEC)[1:9],main = paste("factor",i,":GM12878"),las = 2,cex.names = 0.6)
```

### K562

```{r,echo=FALSE}
par(mfrow = c(2,3))
for(i in 1:5) barplot(FL_K562$f[,i],names = colnames(HUVEC)[1:9],main = paste("factor",i,"K562"),las = 2,cex.names = 0.6)
```

### NHLF

```{r,echo=FALSE}
par(mfrow = c(2,3))
for(i in 1:5) barplot(FL_NHLF$f[,i],names = colnames(HUVEC)[1:9],main = paste("factor",i,"NHLF"),las = 2,cex.names = 0.6)
```

## Session Information

```{r session-info}
```

### 2017-02-14 code/rcctemplate.R
ignore the data

@@ -1,0 +1,61 @@
+# Biobase is for accessing expressionSet objects
+library(Biobase)
+
+# Load Buettner data
+devtools::install_github("jhsiao999/singleCellRNASeqHumanTungiPSC")
+library(singleCellRNASeqHumanTungiPSC)
+
+# Extract expression data
+eset <- get(data("HumanTungiPSC"))
+
+# Access the count matrix
+counts <- exprs(eset)
+
+# to get the logcount
+# log_counts = log(counts)
+# log_counts[which(log_counts == -Inf)] = NA
+library(limma)
+library(ashr)
+source("~/HG/flash/Rcode/flash.R")
+source("~/HG/flash/Rcode/greedy.R")
+source("~/HG/flash/Rcode/backfitting.R")
+voom_ouptut <- voom(counts, normalization = "none")
+cpm_counts = voom_ouptut$E
+log_cpm = log(cpm_counts)
+
+g_flash = flashr::greedy(log_cpm, K = 30, flash_para = list(partype = "var_col"))
+saveRDS(g_flash, "./logcpm/gflash_var_col.rds")
+b_flash = flashr::backfitting(Y,initial_list = g_flash, flash_para = list(partype = "var_col"), maxiter_bf = 6)
+saveRDS(b_flash, "./logcpm/bflash_var_col.rds")
+
+
+
+
+
+source("~/HG/flash/Rcode/flash.R")
+source("~/HG/flash/Rcode/greedy.R")
+source("~/HG/flash/Rcode/backfitting.R")
+
+saveRDS(result, "./NBSFAout/output.rds")
+
+
+#!/bin/bash
+
+#SBATCH --job-name=arrayJob
+#SBATCH --output=/home/weidong/HG/flash/data/singlecell/HumanTungiPSC/outlog/arrayJob_%A_%a.out
+#SBATCH --error=/home/weidong/HG/flash/data/singlecell/HumanTungiPSC/outlog/arrayJob_%A_%a.err
+#SBATCH --array=1
+#SBATCH --time=32:00:00
+#SBATCH --partition=broadwl
+#SBATCH --ntasks=1
+#SBATCH --mem-per-cpu=4000
+
+
+######################
+# Begin work section #
+######################
+
+# Print this sub-job's task ID
+cd /home/weidong/HG/flash/data/singlecell/HumanTungiPSC
+Rscript --verbose /home/weidong/HG/flash/data/singlecell/HumanTungiPSC/logcpm.R
+

### 2017-03-13 analysis/simulationrank1.Rmd
Files commited by wflow_commit.

---
title: "simulation for rankone case"
author: "Wei Wang"
date: 2017-03-13
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`

## simulation rank one

### simulated data

```{r,echo=FALSE}
library("MASS")
betaO=function(P,betapi,betasigma,propt){
  idx=rmultinom(P,size=1,prob=betapi)
  K=length(betapi)
  s=array(0,dim=c(K,K))
  diag(s)=betasigma
  bnorm=mvrnorm(P,rep(0,K),s)
  betaO=apply(bnorm*t(idx),1,sum)
  betaO=betaO*rbinom(P,1,propt)
  return(betaO)
}

datamaker = function(N,P,l_pi,l_se,l_sp,f_pi,f_se,f_sp,sigmae){
  # here I would like to fix the sparsity of L and F which is simple
  # if we need to do futher experiment
  L_true = betaO(N,l_pi,l_se,l_sp)
  F_true = betaO(P,f_pi,f_se,f_sp)
  E = matrix(rnorm(N*P,0,sigmae),ncol = P)
  Y = L_true %*% t(F_true) + E
  return(list(Y = Y,L_true = L_true,F_true = F_true))
}
# I think we can try some situation
# pve is small due to the

# this can be a example of the bi sparse factor and loadings

CVPMD=function(Y,c_u,c_v){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = matrix(sample(P,P),ncol = 5)
  rowindex = matrix(sample(N,N),ncol = 5)

  missing= array(0,dim = c(5,N,P))
  foldindex = array(0,dim = c(5,5,2))
  for(i in 1:5){
    for(j in 1:5){
      foldindex[i,j,1] = i
      foldindex[i,j,2] = (i+j) %% 5
    }
  }
  foldindex[which(foldindex == 0)] = 5
  for(i in 1:5){
    missing[i, , ] = Y
    for(j in 1:5){
      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
    }
  }
  n_u = length(c_u)
  n_v = length(c_v)
  CVRMSE = array(0,dim = c(n_u,n_v))
  minrmse = Inf
  opt_u = 0
  opt_v = 0
  for(t_u in 1:n_u){
    for(t_v in 1:n_v){
      rmse = rep(0,5)
      for(i in 1:5){
        out = PMD(missing[i,,], sumabsu = c_u[t_u], sumabsv = c_v[t_v])
        misshat = out$d * out$u %*% t(out$v)
        for(j in 1:5){
          rmse[i] = rmse[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] - 
                                     misshat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2)
        }
        rmse[i] = sqrt(rmse[i] / (N * P/5))
      }
      CVRMSE[t_u,t_v] = mean(rmse)
      if(CVRMSE[t_u,t_v] < minrmse){
        minrmse = CVRMSE[t_u,t_v]
        opt_u = c_u[t_u]
        opt_v = c_v[t_v]
      }
    }
  }
  return(list(opt_u = opt_u,opt_v = opt_v))
}

PMA.wrapper = function(Y,ngrids = 10){
  library(PMA)
  N = dim(Y)[1]
  P = dim(Y)[2]
  c_u = seq(1,sqrt(N),length.out = ngrids)
  c_v = seq(1,sqrt(P),length.out = ngrids)
  cvout = CVPMD(Y,c_u,c_v)
  out = PMD(Y,sumabsu = cvout$opt_u, sumabsv = cvout$opt_v )
  residual_PMD = Y - out$d * out$u %*% t(out$v)
  return(list(d = out$d, u = out$u, v = out$v, residual = residual_PMD))
}
```

## big signal

```{r}
set.seed(99)
Data = datamaker(60,100,c(0.5,0.1,0.2,0.15,0.05),c(0.01,0.5,1,2,5),0.5,c(0.6,0.05,0.3,0.025,0.025),c(0.01,0.5,1,2,5),0.5,sqrt(1))
dim(Data$Y)
gf = flashr::flash(Data$Y)
plot(as.vector(Data$L_true %*% t(Data$F_true)),as.vector(gf$l %*% t(gf$f)),pch = 16,col = "red")
abline(0,1,col = "red")
PMDlog = capture.output({gPMD = invisible( PMA.wrapper(Data$Y)) })
```

```{r}
plot(as.vector(Data$L_true %*% t(Data$F_true)),as.vector(gf$l %*% t(gf$f)),pch = 16,col = "red")
abline(0,1,col = "red")
plot(as.vector(Data$L_true %*% t(Data$F_true)),as.vector( (gPMD$u * gPMD$d) %*% t(gPMD$v) ),pch = 16,col = "blue")
sqrt(mean((gf$l %*% t(gf$f) - Data$L_true %*% t(Data$F_true))^2))/sqrt(mean((Data$L_true %*% t(Data$F_true))^2))
sqrt(mean(((gPMD$u * gPMD$d) %*% t(gPMD$v) - Data$L_true %*% t(Data$F_true))^2))/sqrt(mean((Data$L_true %*% t(Data$F_true))^2))
plot(as.vector( (gPMD$u * gPMD$d) %*% t(gPMD$v) ),as.vector(gf$l %*% t(gf$f)),pch = 16,col = "blue")
```

## Session Information

```{r session-info}
```

### 2017-03-13 analysis/simulation2.Rmd
Files commited by wflow_commit.

---
title: "Simulation with rank 2"
author: "Wei Wang"
date: 2017-03-13
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`

## simulated data

```{r}
set.seed(99)
library("MASS")
betaO=function(P,betapi,betasigma,propt){
  idx=rmultinom(P,size=1,prob=betapi)
  K=length(betapi)
  s=array(0,dim=c(K,K))
  diag(s)=betasigma
  bnorm=mvrnorm(P,rep(0,K),s)
  betaO=apply(bnorm*t(idx),1,sum)
  betaO=betaO*rbinom(P,1,propt)
  return(betaO)
}
N = 180
P = 300
K = 2
L_true = array(0,dim = c(N,K))
F_true = array(0,dim = c(P,K))
F_true[,1] = rnorm(P,0,1)
F_true[,2] = rnorm(P,0,1)
# F_true = scale(F_true)
rho = rnorm(N,1.96,0.6)
L_true[,1] = rho * sin(pi*(seq(1,N)/(N/2)))
L_true[,2] = rho * cos(pi*(seq(1,N)/(N/2)))
group1_index = c(31:60,121:150)
group2_index = c(1:15,76:105,166:180)
group3_index = seq(1:180)[-(union(group1_index,group2_index))]
# G = L_true %*% t(F_true)
# generate Y
E = matrix(rnorm(N*P,0,1),nrow=N)
Y = L_true %*% t(F_true) + E
par(mfrow = c(1,1), mar = c(5,4,4,2) - 1.9,pty="s")
plot(L_true[,1],L_true[,2],col = "green")
points(L_true[group1_index,1],L_true[group1_index,2],col = "green")
points(L_true[group2_index,1],L_true[group2_index,2],col = "red")
points(L_true[group3_index,1],L_true[group3_index,2],col = "blue")
plot(L_true[,1])
plot(L_true[,2])
g_flash = flashr::greedy(Y,K = 5)
gPMD = PMA::PMD(Y,K = 2)
gsvd = svd(Y,nv = 2,nu = 2)
```

```{r}
par(mfrow = c(1,1), mar = c(5,4,4,2) - 1.9,pty="s")
plot(g_flash$l[,1],g_flash$l[,2],col = "green")
points(g_flash$l[group1_index,1],g_flash$l[group1_index,2],col = "green")
points(g_flash$l[group2_index,1],g_flash$l[group2_index,2],col = "red")
points(g_flash$l[group3_index,1],g_flash$l[group3_index,2],col = "blue")
plot(L_true[,1],L_true[,2],col = "green")
points(L_true[group1_index,1],L_true[group1_index,2],col = "green")
points(L_true[group2_index,1],L_true[group2_index,2],col = "red")
points(L_true[group3_index,1],L_true[group3_index,2],col = "blue")
plot(gPMD$u[,1],gPMD$u[,2],col = "green")
points(gPMD$u[group1_index,1],gPMD$u[group1_index,2],col = "green")
points(gPMD$u[group2_index,1],gPMD$u[group2_index,2],col = "red")
points(gPMD$u[group3_index,1],gPMD$u[group3_index,2],col = "blue")
plot(gsvd$u[,1],gsvd$u[,2],col = "green")
points(gsvd$u[group1_index,1],gsvd$u[group1_index,2],col = "green")
points(gsvd$u[group2_index,1],gsvd$u[group2_index,2],col = "red")
points(gsvd$u[group3_index,1],gsvd$u[group3_index,2],col = "blue")
```

## Session Information

```{r session-info}
```

### 2017-03-14 analysis/MissingTumor.Rmd
Files commited by wflow_commit.

---
title: "missing data imputation: Tumor data"
author: "wei wang"
date: 2017-03-13
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

try the new version of the wrokflowr package

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```

### 2017-03-15 analysis/simulation2.Rmd
Files commited by wflow_commit.

@@
 
+###
+Here I would like to add a cv function for PMD when K larger than 1
+
+In the PMA R package we can find that they use the parameter sumabs to control the sparsity, so we use this one as well.
+
+```{r}
+CVPMD=function(Y,c_s,K){
+  N = dim(Y)[1]
+  P = dim(Y)[2]
+  colindex = matrix(sample(P,P),ncol = 5)
+  rowindex = matrix(sample(N,N),ncol = 5)
+
+  missing= array(0,dim = c(5,N,P))
+  foldindex = array(0,dim = c(5,5,2))
+  for(i in 1:5){
+    for(j in 1:5){
+      foldindex[i,j,1] = i
+      foldindex[i,j,2] = (i+j) %% 5
+    }
+  }
+  foldindex[which(foldindex == 0)] = 5
+  for(i in 1:5){
+    missing[i, , ] = Y
+    for(j in 1:5){
+      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
+    }
+  }
+  n_s = length(c_s)
+  CVRMSE = rep(NA,n_s)
+  minrmse = Inf
+  opt_s = 0
+  # here we only need one layer
+    for(t_s in 1:n_s){
+      rmse = rep(0,5)
+      for(i in 1:5){
+        out = PMD(missing[i,,], sumabs = c_s[t_s], sumabsv = NULL, sumabsu = NULL,K = K)
+        misshat = out$u %*%  diag(out$d) %*% t(out$v)
+        for(j in 1:5){
+          rmse[i] = rmse[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] - 
+                                     misshat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2)
+        }
+        rmse[i] = sqrt(rmse[i] / (N * P/5))
+      }
+      CVRMSE[t_s] = mean(rmse)
+      if(CVRMSE[t_s] < minrmse){
+        minrmse = CVRMSE[t_s]
+        opt_s = c_s[t_s]
+      }
+    }
+
+  return(list(opt_s = opt_s ))
+}
+
+PMA.wrapper = function(Y,ngrids = 10,K){
+  library(PMA)
+  N = dim(Y)[1]
+  P = dim(Y)[2]
+  c_s = seq(0.2,0.8,len=ngrids)
+  cvout = CVPMD(Y,c_s, K = K)
+  out = PMD(Y,sumabsu = NULL, sumabsv = NULL, sumabs = cvout$opt_s ,K = K)
+  residual_PMD = Y -  out$u %*%  diag(out$d) %*% t(out$v)
+  return(list(d = out$d, u = out$u, v = out$v, residual = residual_PMD))
+}
+```
+
 ## simulated data
@@
 }
-N = 180
-P = 300
+N = 100
+P = 200
 K = 2
@@
 F_true = array(0,dim = c(P,K))
-F_true[,1] = rnorm(P,0,1)
-F_true[,2] = rnorm(P,0,1)
+F_true[,1] = rnorm(P,0,0.3)
+F_true[,2] = rnorm(P,0,0.3)
 # F_true = scale(F_true)
-rho = rnorm(N,1.96,0.6)
-L_true[,1] = rho * sin(pi*(seq(1,N)/(N/2)))
-L_true[,2] = rho * cos(pi*(seq(1,N)/(N/2)))
-group1_index = c(31:60,121:150)
-group2_index = c(1:15,76:105,166:180)
-group3_index = seq(1:180)[-(union(group1_index,group2_index))]
-# G = L_true %*% t(F_true)
+group1_index = c(1:30)
+group2_index = c(71:100)
+group3_index = seq(1:90)[-(union(group1_index,group2_index))]
+L_true[group1_index,1] = betaO(30,c(0.9,0.1),c(0.001,0.1),1)
+L_true[group2_index,1] = betaO(30,c(0.9,0.1),c(0.5,0.75),1)
+L_true[group3_index,1] = betaO(40,c(0.9,0.1),c(1,1.5),1)
+L_true[group1_index,2] = betaO(30,c(0.9,0.1),c(1,1.5),1)
+L_true[group2_index,2] = betaO(30,c(0.9,0.1),c(0.5,0.75),1)
+L_true[group3_index,2] = betaO(40,c(0.9,0.1),c(0.001,0.1),1)
 # generate Y
@@
 g_flash = flashr::greedy(Y,K = 5)
-gPMD = PMA::PMD(Y,K = 2)
+g_flash = flashr::backfitting(Y,initial_list = g_flash,maxiter_bf = 10)
+gPMD = PMA.wrapper(Y,ngrids = 10,K = 2)
 gsvd = svd(Y,nv = 2,nu = 2)
@@
 points(gsvd$u[group3_index,1],gsvd$u[group3_index,2],col = "blue")
+plot(gsvd$d)
+```
+
+```{r}
+par(mfrow = c(1,1), mar = c(5,4,4,2) - 1.9,pty="s")
+plot(as.vector(L_true %*% t(F_true)),as.vector(g_flash$l %*% t(g_flash$f)),pch = 16,col = "red")
+abline(0,1,col = "red")
+plot(as.vector(L_true %*% t(F_true)),as.vector( (gPMD$u * gPMD$d) %*% t(gPMD$v) ),pch = 16,col = "blue")
+abline(0,1,col = "red")
+sqrt(mean((g_flash$l %*% t(g_flash$f) - L_true %*% t(F_true))^2))/sqrt(mean((L_true %*% t(F_true))^2))
+sqrt(mean((gPMD$u %*% diag(gPMD$d) %*% t(gPMD$v) - L_true %*% t(F_true))^2))/sqrt(mean((L_true %*% t(F_true))^2))
+plot(as.vector( gPMD$u %*% diag(gPMD$d) %*% t(gPMD$v) ),as.vector(g_flash$l %*% t(g_flash$f)),pch = 16,col = "blue")
+abline(0,1,col = "red")
+plot(as.vector( gsvd$u %*% diag(gsvd$d[1:2]) %*% t(gsvd$v) ),as.vector(g_flash$l %*% t(g_flash$f)),pch = 16,col = "blue")
+abline(0,1,col = "red")
 ```

### 2017-03-16 analysis/draft.Rmd
Files commited by wflow_commit.

--- 2017-03-14 analysis/MissingTumor.Rmd
@@
 ---
-title: "missing data imputation: Tumor data"
-author: "wei wang"
-date: 2017-03-13
+title: "this is for the result part of the draft"
+author: "Wei Wang"
+date: 2017-03-15
 output: html_document
@@
 
-try the new version of the wrokflowr package
+### data analysis on the normalized data
+
+#### missing data imputation
+
+#### gene expression data
+
+### further issues
+
+#### NMF
+
+#### tensor flash
 

### 2017-03-17 analysis/draft.Rmd
Files commited by wflow_commit.

@@
 
-### data analysis on the normalized data
+## data analysis on the normalized data
 
-#### missing data imputation
+### missing data imputation
+
+#### Tumors data
+
+We test the performance of our method on the real data where there is no ground truth.
+The performance is assessed by the predictive Mean Square Error (MSE) for the missing value, where the true value is know and chosen to be missing.
+This data set is a microarray data from 'denoiseR' package containing 43 brain tumors of four types and 356 continous variable related with the expression data.
+we set the rank equal to 20 for FLASH, PMD and softImpute, since we find that the rank of the low rank structure is almost always smaller than 20, whihc is around 15-20.
+
+![tumor data](~/HG/ash-sfa/Rcode/postmean/flash_simulation/denoise/denoiseRdata/tumordata/boxplot.png)
+
+#### Movie data
+
+The MovieLens data set is available at the GroupLens Research group which is usualy used for the exploration of recommender systems.
+The idea of the prediction is that users with similar preference on movies might have similar ratings so that we can predict tha ratings of certain user based on other users sharing some common interests.
+the MovieLens 100K data set we used here is a data set with about 100K ratings from 1000 user and 1700 movies. The ratings are integers from 0 to 5 represent preference of users.
+
+#### presidential address
 
@@
 
-### further issues
+#### Breast cancer data
 
-#### NMF
+#### do I need add one more into it?
 
-#### tensor flash
+click here
+
+### gene expression data
+
+## further issues
+
+### NMF
+
+### tensor flash
+
+### rank decision
+
+there is somethig we don't fully understand.
 

### 2017-03-17 analysis/index.Rmd
Files commited by wflow_commit.

@@
 
+### simulation studies
+
+### Orthogonal CV
+
 Click on this [orthogonal CV](miss_position.html) to see my example.
+
+Click on this [Questions on Algorithm](first-analysis.html) to see more details.
+
+### flash on the chipseq data from Kevin
+
+Click on this [three data sets on the 4500 rows chr1 var_col](sd_chr1_asinh.html) to see more details.
+
+[ns](draft.html)
+
+data set: /project/mstephens/data/internal_supp/gtex-v6-sumstat-hdf5
+
+[5:25]  
+on midway. you can focus on `MatrixEQTLSumStats.Portable.h5`
+
+[5:26]  
+for the 16000 times 44 data-set, stored in a table called `max` (edited)
+
+[5:26]  
+To process the data, https://github.com/stephenslab/gtex-eqtls/blob/master/doc/notes/SumstatsDB.md
+
+[5:27]  
+and the relevent scripts in this MD file: https://github.com/stephenslab/lab-resource/blob/master/utilities/src/SumstatQuery.R
+
+[5:30]  
+you should use this line `GetSS("max", "/project/mstephens/gtex/analysis/april2015/query/MatrixEQTLSumStats.h5")`
+
+----- October 20th, 2016 -----
+gaow [6:06 PM] 
+https://github.com/stephenslab/lab-resource/wiki/Tissue-Color-Palette
+
+[6:06]  
+see "load to R"

### 2017-03-21 analysis/draft.Rmd
Files commited by wflow_commit.

@@
 The MovieLens data set is available at the GroupLens Research group which is usualy used for the exploration of recommender systems.
-The idea of the prediction is that users with similar preference on movies might have similar ratings so that we can predict tha ratings of certain user based on other users sharing some common interests.
-the MovieLens 100K data set we used here is a data set with about 100K ratings from 1000 user and 1700 movies. The ratings are integers from 0 to 5 represent preference of users.
+The idea of the prediction is that users with similar preference on movies might have similar ratings so that we can predict tha ratings of certain user based on other users sharing some common preference.
+There are several data sets available for stable bench mark and we choose the smallest one.
+The MovieLens 100K data set we used here is a data set with about 100K ratings from 1000 user and 1700 movies.
+The ratings are integers from 0 to 5 represent preference of users.
+In this data, all the ratings are nonnegative, so we center and scale the data for each users.
+We focus on the structure that the users share common prefernce on movies, so the preference of each user on each movies remains the same.
+
+![Movie data](~/HG/ash-sfa/Rcode/postmean/flash_simulation/missingvalue/ML100K/boxplotrowscale.png)
 
@@
 
-#### gene expression data
+This data set is a contingency table cross-tabulating with 13 rows and 836 columns, where rows stands for US presidents from 1940 to 2009 and columns are correspoding the words they used in thier inaugural addresses.
+This data set comes from 'denoiseR' R package.
+We make one more pre-processing procedure before apply all the methods.
+The pre-precessed data is center and scale on both rows ans columns.
+It is because the words and the frequency of those words vary a lot among different presidents.
+This is data is different from movie ratings data of which the ratings can only take values from 0 to 5.
+Some words show up more frequently than others, such as "a" and "the", so we should center and scale the data not only by rows but also by coumns.
+
+![president data](~/HG/ash-sfa/Rcode/postmean/flash_simulation/denoise/denoiseRdata/presidentdata/biscaleboxplot.png)
+
+#### GTEx eQTL summary statistic
+
+This data set is summary statistics (Z-score) from eQTL analysis.
+
+![GTEx data](~/HG/ash-sfa/Rcode/postmean/flash_simulation/missingvalue/GTEXboxplot.png)
+
+This data set is
 
@@
 
-#### do I need add one more into it?
+The Breast cancer data is from West et al.
+(2007).
+There are 251 samples and 226 genes in this data set.
+For those algorithm we choose, we set the rank equal to 50 which is close the the rank of the estimation from the NSFA results.
+After some try of FLASH, we find that the rank picked by FLASH is also around 40-50.
+Following the pre-processing procedure from NSFA, we also use the centered data.
 
-click here
+![Breast Cancer data](~/HG/ash-sfa/Rcode/postmean/flash_simulation/missingvalue/Breastcancer/boxplot_centered.png)
 
@@
 
-## further issues
+To show the flexibility of FLASH, we take the summary statistics from the expression Quantitative Trait Loci (eQTLs) of 44 human Tissues. The original data is from Genotype Tissue Expression
+(GTEx) project. The aim of GTEx project "s to increase our understanding of how changes in our genes contribute to common human diseases, in order to improve health care for future generations."
+We try to understand the " which SNPs are eQTLs (i.e.
+associated with expression) in which tissues.Answering this could help distinguish regulatory regions and mechanisms that are specific to a few tissues vs shared among many tissues.It could also help with analyses that aim to integrate eQTL results with GWAS results to help identify the tissues that are most relevant to any specific complex disease (e.g.
+[18, 19])."
+
+We apply FLASH on the matrix of summary statistics which are Z-scores.
+
+"We analysed the 16,069 genes for which univariate effect estimates were available for all 44 tissues we considered; the filtering criteria used ensure that these genes show at least some indication of expression in all 44 tissues."
+
+As input to mash we use a matrix of eQTL effect estimates ˆbjr, and
+corresponding standard errors ˆsij , where the rows j index different SNP-gene
+pairs and the columns r index tissues (or cell types). We used the effect
+estimates and standard errors for candidate local (“cis”) eQTLs for each gene,
+distributed by the GTEx project (v6 release). These were obtained by
+(univariate) single-SNP analyses in each tissue by applying MatrixEQTL [20] on
+expression levels that have been rank-transformed to the corresponding
+quantiles of a standard normal distribution. Thus the effect size estimates are in
+units of standard deviations on this transformed scale. Because, like most eQTL
+analyses, these estimates were obtained by single-SNP analysis, the estimated
+effects for each SNP actually reflect the effects of both the SNP itself and other
+SNPs in LD with it (see Supplementary Text for discussion). We analysed the
+16,069 genes for which univariate effect estimates were available for all 44 tissues
+we considered; the filtering criteria used ensure that these genes show at least
+some indication of expression in all 44 tissues.
+
+## other potential extension
 
@@
 
+In ashr, there is a option that we can choose the prior to be half uniform or half normal, which leads to the final estimate is non-negative.
+In this case, we apply FLASH for non-negative matrix factorization with proper intial values.
+For example, we can use NMF result as intial value and apply FLASH with backfitting algorithm to get a flexible sparse solutions of factors and loadings.
+
 ### tensor flash
 
-### rank decision
+In our frame work, it is easy to extend out method to tensor factorization. And the benfit of FLASH is we don't need to estimate the tuning parmeters in each mode.
+
+### rank decision issue
 

### 2017-03-24 analysis/draft.Rmd
Files commited by wflow_commit.

@@
 This data set is summary statistics (Z-score) from eQTL analysis.
+We apply FLASH on the matrix of summary statistics which are Z-scores with , where the rows index different SNP-gene pairs and the columns index tissues (or cell types).
+We used the Z-scores for candidate local (“cis”) eQTLs for each gene, distributed by the GTEx project (v6 release).
 
@@
 Following the pre-processing procedure from NSFA, we also use the centered data.
+Since this input data set is Z-scores data, we didn't use any pre-processing procedure.
 
@@
 We apply FLASH on the matrix of summary statistics which are Z-scores.
-
 "We analysed the 16,069 genes for which univariate effect estimates were available for all 44 tissues we considered; the filtering criteria used ensure that these genes show at least some indication of expression in all 44 tissues."
-
-As input to mash we use a matrix of eQTL effect estimates ˆbjr, and
-corresponding standard errors ˆsij , where the rows j index different SNP-gene
-pairs and the columns r index tissues (or cell types). We used the effect
-estimates and standard errors for candidate local (“cis”) eQTLs for each gene,
-distributed by the GTEx project (v6 release). These were obtained by
-(univariate) single-SNP analyses in each tissue by applying MatrixEQTL [20] on
-expression levels that have been rank-transformed to the corresponding
-quantiles of a standard normal distribution. Thus the effect size estimates are in
-units of standard deviations on this transformed scale. Because, like most eQTL
-analyses, these estimates were obtained by single-SNP analysis, the estimated
-effects for each SNP actually reflect the effects of both the SNP itself and other
-SNPs in LD with it (see Supplementary Text for discussion). We analysed the
-16,069 genes for which univariate effect estimates were available for all 44 tissues
-we considered; the filtering criteria used ensure that these genes show at least
-some indication of expression in all 44 tissues.
 

### 2017-03-29 analysis/draft.Rmd
Files commited by wflow_commit.

@@
 
+```{r}
+library(ggplot2)
+movie_res = readRDS("../data/missingvalue/box_movies.rds")
+tumor_res = readRDS("../data/missingvalue/box_tumor.rds")
+president_res = readRDS("../data/missingvalue/box_president.rds")
+breast_res = readRDS("../data/missingvalue/box_breast.rds")
+gtex_res = readRDS("../data/missingvalue/box_gtex.rds")
+plot_res = function(output,title = "data"){
+  rmse = as.vector(output)
+  N = dim(output)[1]
+  methods = rep(c("flash","NBSFA","PMD","softImpute"), each = N)
+  df = data.frame(RMSE = rmse, Method = methods )
+  p<-ggplot(df, aes(x=Method, y=RMSE, color=Method)) +
+  #geom_boxplot()
+  geom_violin()+
+  ggtitle(title)+theme_gray()
+  p
+}
+pm = plot_res(movie_res,"MovieLens")
+pp = plot_res(president_res,"Presidents")
+pg = plot_res(gtex_res,"GTEx Z-scores")
+pt = plot_res(tumor_res,"Tumors")
+pb = plot_res(breast_res,"Breast Cancer")
+library(cowplot)
+firstrow = plot_grid(pm, pp,pg, labels = c("A","B","C"), ncol = 3,rel_widths = c(1, 1))
+secondrow <- plot_grid(pt, pb, labels = c("D","E"), ncol = 2,rel_widths = c(1, 1))
+plot_grid(firstrow, secondrow, ncol = 1,rel_widths = c(1, 0.6))
+
+```
+
 ### gene expression data
@@
 
-there is somethig we don't fully understand.
+there is somethig we don't fully understand. the simplest case is comparing rank 0 and rank 1 case.
 

### 2017-03-30 analysis/GTExresult.Rmd
Files commited by wflow_commit.

---
title: "result for GTEx data and"
author: "wei wang"
date: 2017-03-30
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

The result for FLASH on GTEx data is on midway cluster

## Zsocre data

To get the zscore data:

```
scp username@midway2.rcc.uchicago.edu:/project/mstephens/flash/eQTLzscore/gtexEQTL_zscore.rds ./
```

and then open `Rstudio`

```
load("gtexEQTL_zscore.rds")
```

the matrix `zscore` is the data matrix.

## FLASH result

### constant variance

```
scp username@midway2.rcc.uchicago.edu:/project/mstephens/flash/eQTLzscore/bflashconst.rds ./
```

### column-wise variance

```
scp username@midway2.rcc.uchicago.edu:/project/mstephens/flash/eQTLzscore/bflashvarcol.rds ./
```

### constant variance + known variance structure

```
scp username@midway2.rcc.uchicago.edu:/project/mstephens/flash/eQTLzscore/bflashnoisyConst.rds ./
```

### column-wise variance + known variance structure

```
scp username@midway2.rcc.uchicago.edu:/project/mstephens/flash/eQTLzscore/bflashnoisycol.rds ./
```

### load the result

we load the column-wise variance reslut which we use in our paper as example

```
load("bflashvarcol.rds")
```

and the list `b_flash` is the reslut of backfitting FLASH with estimation of $l$ $f$ and $\hat{\Sigma}_e$

## plot the barplot

this is from Gao:

data set: /project/mstephens/data/internal_supp/gtex-v6-sumstat-hdf5

[5:25]  
on midway. you can focus on `MatrixEQTLSumStats.Portable.h5`

[5:26]  
for the 16000 times 44 data-set, stored in a table called `max` (edited)

[5:26]  
To process the data, https://github.com/stephenslab/gtex-eqtls/blob/master/doc/notes/SumstatsDB.md

[5:27]  
and the relevent scripts in this MD file: https://github.com/stephenslab/lab-resource/blob/master/utilities/src/SumstatQuery.R

[5:30]  
you should use this line `GetSS("max", "/project/mstephens/gtex/analysis/april2015/query/MatrixEQTLSumStats.h5")`

----- October 20th, 2016 -----
gaow [6:06 PM] 
https://github.com/stephenslab/lab-resource/wiki/Tissue-Color-Palette

[6:06]  
see "load to R"

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```

### 2017-03-30 analysis/index.Rmd
Files commited by wflow_commit.

--- 2017-02-01 analysis/index.Rmd
@@
 
+### simulation studies
+
+### Orthogonal CV
+
 Click on this [orthogonal CV](miss_position.html) to see my example.
+
+Click on this [Questions on Algorithm](first-analysis.html) to see more details.
+
+### flash on the chipseq data from Kevin
+
+Click on this [three data sets on the 4500 rows chr1 var_col](sd_chr1_asinh.html) to see more details.
+
+[ns](draft.html)
+
+Click on this [load reslut of GTEx data](GTExresult.html) to see more details.

### 2017-04-03 analysis/draft.Rmd
Files commited by wflow_commit.

@@
 
+```{r}
+dat = read.table('../data/GTExColors.txt', sep = '\t', comment.char = '')
+colordata = dat[c(1:6,9:18,21:23,26:30,32,33,35,36,38:53),1:2]
+# barplot(rep(1, nrow(dat)), col = as.character(dat[,2]))
+# just plot for loadings
+plot.factor = function(Y,g,k = 1,loading_label = FALSE,color.fill = NULL){
+  library(ggplot2)
+  N = dim(Y)[1]
+  P = dim(Y)[2]
+  sample_name = row.names(Y)
+  variable_name = colnames(Y)
+  # plot the expectation of PVE 
+  ssY = sum(Y^2)
+  K = dim(g$l)[2]
+  pve = (sapply(seq(1,K),function(x){ sum(g$l[,x]^2 %*% t(g$f[,x]^2)) }))/ssY
+  pve_dat = data.frame(index = seq(1,K),EPVE = pve)
+  plot_pve <- ggplot(pve_dat, aes(index, EPVE)) + geom_point(size = 4)  + geom_line(linetype = "dotdash")
+
+  # plot the loadings
+  if(loading_label == TRUE){
+    plot_l = plot_factor(g$l[,k],pve[k],N,k,f_lables = row.names(Y),y_lab = "loading values",color.fill = color.fill)
+  }else{
+    plot_l = plot_factor(g$l[,k],pve[k],N,k,f_lables =NA,y_lab = "loading values",color.fill = color.fill)
+  }
+  return(list(plot_pve = plot_pve,plot_l = plot_l))
+}
+
+#' plot_f
+#'
+#' plot the factor 
+#' @return list of factor, loading and variance of noise matrix
+#'  \itemize{
+#'   \item{\code{plot_f}} {is a ggplot object for the factors}
+#'  }
+#' @param f factor to plot
+#' @param pve pve for this factor
+#' @param P the lenght of this factor
+#' @param k the order of the factor
+#' @param f_lables the lables for the factor 
+#' @param y_lab is the name of the Y axis
+#' @details plot_f is to plot the factor as barplot.
+#'
+#' @importFrom ggplot2 ggplot
+#'
+
+plot_factor = function(f,pve,P,k,f_lables = NA,y_lab = "factor values",color.fill = NA){
+  library(ggplot2)
+  if(any(is.na(f_lables))){
+    f_dat <- data.frame(variable = 1:P, Factor = f,
+                      sign.f = factor(sign(f)),
+                      hjust = factor(sign(f)))
+
+    plot_f = ggplot(f_dat,aes(x = variable,y = Factor,fill = factor(variable))) +
+      geom_bar(stat = "identity",width = 0.5,color = "black") +
+      scale_fill_manual(values=color.fill) + theme(legend.position="none") +
+      labs(title = paste("factor",k,"with PVE=",round(pve,3)), y = y_lab ,x = "index")
+  }else{
+    f_dat <- data.frame(variable = 1:P, Factor = f,
+                      sign.f = factor(sign(f)),
+                      variablenames = f_lables,
+                      hjust = factor(sign(f)))
+
+    # 120% lim 
+    range_f = max(f) - min(f)
+    upper_f = max(f,0) + 0.15 * range_f
+    lower_f = min(f,0) - 0.15 * range_f
+
+    plot_f = ggplot(f_dat,aes(x = factor(variable),y = Factor,fill = factor(variable))) +
+      geom_bar(stat = "identity",width = 0.5,color = "black") +
+      scale_fill_manual(values=color.fill) + theme(legend.position="none") +
+      geom_text(size = 2.75,angle = 90,hjust = as.character(f_dat$hjust),
+                nudge_y = sign(f_dat$Factor) * 0.1 * mean(abs(f_dat$Factor)) ) +
+      ylim(lower_f,upper_f) +
+      labs(title = paste("factor",k,"with PVE=",round(pve,3) ), y = y_lab,x = "index")
+  }
+  return(plot_f)
+}
+load("../data/GTExdata/bflashvarcol.rds")
+load("../data/GTExdata/gtexEQTL_zscore.rds")
+plot.factor(Y = t(zscore),g = b_flash,k = 1,loading_label = FALSE,color.fill = as.character(colordata[,2]))
+
+# plot_f = ggplot(f_dat,aes(x = factor(variable),y = Factor,color = factor(variable))) +
+#       geom_bar(stat = "identity",width = 0.5,fill = "white") +
+#       scale_color_manual(values=color.fill) + theme(legend.position="none")
+
+# ggplot(f_dat,aes(x = factor(variable),y = Factor,fill = factor(variable))) +
+#       geom_bar(stat = "identity",width = 0.5,color = "black") +
+#       scale_fill_manual(values=color.fill) + theme(legend.position="none")
+
+```
+
 ## other potential extension

### 2017-04-03 analysis/draft.Rmd
Files commited by wflow_commit.

@@
 
-    plot_f = ggplot(f_dat,aes(x = variable,y = Factor,fill = factor(variable))) +
-      geom_bar(stat = "identity",width = 0.5,color = "black") +
-      scale_fill_manual(values=color.fill) + theme(legend.position="none") +
-      labs(title = paste("factor",k,"with PVE=",round(pve,3)), y = y_lab ,x = "index")
+    plot_f = ggplot(f_dat,aes(x = variable,y = Factor,fill = factor(variable),color = factor(variable))) +
+      geom_bar(stat = "identity",width = 0.6) +scale_color_manual(values=color.fill)+
+      scale_fill_manual(values=color.fill) + theme_grey()+ theme(legend.position="none",plot.background = element_rect()) + 
+      labs(title = paste("factor",k,"with PVE=",max(0.001,round(pve,3))), y = y_lab ,x = "index",color = "index",fill="index")
   }else{
@@
 
-    plot_f = ggplot(f_dat,aes(x = factor(variable),y = Factor,fill = factor(variable))) +
-      geom_bar(stat = "identity",width = 0.5,color = "black") +
-      scale_fill_manual(values=color.fill) + theme(legend.position="none") +
+    plot_f = ggplot(f_dat,aes(x = variable,y = Factor,fill = factor(variable),color = factor(variable))) +
+      geom_bar(stat = "identity",width = 0.6) +scale_color_manual(values=color.fill)+
+      scale_fill_manual(values=color.fill) + theme_grey()+ theme(legend.position="none",plot.background = element_rect())  +
       geom_text(size = 2.75,angle = 90,hjust = as.character(f_dat$hjust),
@@
       ylim(lower_f,upper_f) +
-      labs(title = paste("factor",k,"with PVE=",round(pve,3) ), y = y_lab,x = "index")
+      labs(title = paste("factor",k,"with PVE=",max(0.001,round(pve,3)) ), y = y_lab,x = "index")
   }
@@
 load("../data/GTExdata/gtexEQTL_zscore.rds")
-plot.factor(Y = t(zscore),g = b_flash,k = 1,loading_label = FALSE,color.fill = as.character(colordata[,2]))
+# plot.factor(Y = t(zscore),g = b_flash,k = 1,loading_label = FALSE,color.fill = as.character(colordata[,2]))
 
@@
 #       scale_fill_manual(values=color.fill) + theme(legend.position="none")
+pic = list()
+for(i in 1:26){
+  pic[[i]] = plot.factor(Y = t(zscore),g = b_flash,k = i,
+                         loading_label = FALSE,color.fill = as.character(colordata[,2]))$plot_l
+}
+
+library(cowplot)
+plot_grid(pic[[1]], pic[[2]],pic[[3]],
+          pic[[4]],pic[[5]],pic[[6]],
+          pic[[7]],pic[[8]],pic[[9]],
+          labels = c("","","","","","","","",""), ncol = 3)
+
+plot_grid(pic[[10]], pic[[11]],pic[[12]],
+          pic[[13]],pic[[14]],pic[[15]],
+          pic[[16]],pic[[17]],pic[[18]],
+          labels = c("","","","","","","","",""), ncol = 3)
+plot_grid(pic[[19]], pic[[20]],pic[[21]],
+          pic[[22]],pic[[23]],pic[[24]],
+          pic[[25]],pic[[26]],
+          labels = c("","","","","","","","",""), ncol = 3)
+line1 = plot_grid(pic[[19]], pic[[20]],pic[[21]],
+          labels = c("","",""), ncol = 3)
+line2 = plot_grid(pic[[22]],pic[[23]],pic[[24]],
+          labels = c("","",""), ncol = 3)
+pic_26 =  pic[[26]] + theme(legend.direction = "horizontal",legend.position = "right")+guides(col = guide_legend(ncol = 8))
+line3 = plot_grid(pic[[25]],pic_26,
+          labels = c("",""), ncol = 2,rel_widths = c(1, 2))
+
+plot_grid(line1,line2,line3,labels = c("","",""), ncol = 1)
+# need one for the lengend
 

### 2017-04-05 analysis/simulationrank1.Rmd
Files commited by wflow_commit.

@@
 
-```{r}
+```
 set.seed(99)
@@
 
-```{r}
+```
 plot(as.vector(Data$L_true %*% t(Data$F_true)),as.vector(gf$l %*% t(gf$f)),pch = 16,col = "red")
@@
 
+## different scale of sparsity
+
+### sparse
+
+```{r}
+library(cowplot)
+set.seed(99)
+L_se = c( 0.25, 0.5, 1, 2, 4)
+L_pi = c(1, 1,1,1,1)
+L_pi = L_pi / sum(L_pi)
+N = 200
+P = 300
+Data = datamaker(N,P,L_pi,L_se,0.1,c(1),c(1),1,sqrt(1))
+dim(Data$Y)
+gf = flashr::flash(Data$Y)
+#plot(as.vector(Data$L_true %*% t(Data$F_true)),as.vector(gf$l %*% t(gf$f)),pch = 16,col = "red")
+#abline(0,1,col = "red")
+PMDlog = capture.output({gPMD = invisible( PMA.wrapper(Data$Y)) })
+#plot(as.vector(Data$L_true %*% t(Data$F_true)),as.vector( (gPMD$u * gPMD$d) %*% t(gPMD$v) ),pch = 16,col = "blue")
+#plot(Data$L_true/sqrt(sum((Data$L_true)^2)),gf$l/sqrt(sum((gf$l)^2)))
+#plot(Data$L_true/sqrt(sum((Data$L_true)^2)),gPMD$u/sqrt(sum((gPMD$u )^2)))
+#plot(Data$L_true/sqrt(sum((Data$L_true)^2)),gf$l/sqrt(sum((gf$l)^2)),ylim = c(-0.1,0.1),xlim = c(-0.1,0.1))
+#plot(Data$L_true/sqrt(sum((Data$L_true)^2)),gPMD$u/sqrt(sum((gPMD$u )^2)),ylim = c(-0.1,0.1),xlim = c(-0.1,0.1))
+#plot(gPMD$u/sqrt(sum((gPMD$u )^2)),gf$l/sqrt(sum((gf$l)^2)))
+#abline(0,1)
+
+library(ggplot2)
+library(grid)
+# zoom bounding box
+xlim <- c(-0.08,0.08); ylim <- c(-0.08,0.08)
+
+mat = cbind(Data$L_true,gf$l,gPMD$u)
+mat = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x] * sign(cor(mat[,x],mat[,1])) })
+mat1 = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x]/sqrt(sum((mat[,x])^2)) } )
+colnames(mat1) = c("truth","FLASH","PMD")
+df1 = data.frame(mat1)
+
+base1 <- ggplot(df1, aes(x = truth, y = FLASH)) +
+  geom_point()  +
+  theme_bw()+ geom_abline(intercept = 0,slope = 1)
+zoomed1 = ggplot(df1, aes(x = truth, y = FLASH)) + 
+    geom_point() + coord_cartesian(xlim=xlim, ylim=ylim)+ theme_grey()+ geom_abline(intercept = 0,slope = 1)
+
+g1 <- ggplotGrob(zoomed1)
+base1 +annotation_custom(grob = g1, xmin = -0.6, xmax = -0.1, ymin = -0.05, ymax = 0.45)
+
+base2 <- ggplot(df1, aes(x = truth, y = PMD)) +
+  geom_point()  +
+  theme_bw()+ geom_abline(intercept = 0,slope = 1)
+zoomed2 = ggplot(df1, aes(x = truth, y = PMD)) + 
+    geom_point() + coord_cartesian(xlim=xlim, ylim=ylim)+ theme_grey() + geom_abline(intercept = 0,slope = 1)
+
+g2 <- ggplotGrob(zoomed2)
+base2 +annotation_custom(grob = g2, xmin = -0.6, xmax = -0.1, ymin = -0.05, ymax = 0.45)
+plot_grid(base1, base2, zoomed1,zoomed2, labels = "AUTO", ncol = 2)
+
+gsfamix_1 = readRDS("../data/simulation/rankone/highsparse/sfamix_res.rds")
+gsvd = svd(Data$Y)
+gsfa = readRDS("../data/simulation/rankone/highsparse/sfa_res.rds")
+
+# to get the KSVD result
+# try other methods
+library("R.matlab")
+# saveRDS(Data,file = "../data/simulation/rankone/mediansparse/test_1.rds")
+# write the data into matlab
+# writeMat("../data/simulation/rankone/highsparse/test_1.mat", data = Data)
+# KSVD
+res_KSVD = readMat("../data/simulation/rankone/highsparse/KSVDtest_1.mat",sparseMatrixClass="matrix")
+Dic = res_KSVD$result[[1]]
+loading_Dic = res_KSVD$result[[2]]
+MSE_KSVD = sqrt(mean(( Dic %*% loading_Dic - Data$L_true%*%t(Data$F_true))^2))/sqrt(mean((Data$L_true%*%t(Data$F_true))^2))
+
+library(ggplot2)
+library(grid)
+
+mat = cbind(Data$L_true,gf$l,gPMD$u,gsfa$L,as.vector(gsfamix_1$F),as.vector(Dic),gsvd$u[,1])
+mat = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x] * sign(cor(mat[,x],mat[,1])) })
+mat1 = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x]/sqrt(sum((mat[,x])^2)) } )
+colnames(mat1) = c("truth","FLASH","PMD","SFA","SFAmix","KSVD","SVD")
+df1 = data.frame(mat1)
+xybound = c( -0.6, -0.1, -0.05, 0.45)
+z_size = c(-0.08,0.08,-0.08,0.08)
+zoom_plot = function(df1,i,zoom_size,bound){
+  xlim <- c(zoom_size[1],zoom_size[2])
+  ylim <- c(zoom_size[3],zoom_size[4])
+
+  base1 <- ggplot(df1, aes(x = truth, y = df1[,i])) +
+    geom_point()  + labs(y = colnames(df1)[i]) +
+    theme_bw()+ geom_abline(intercept = 0,slope = 1)
+  zoomed1 = ggplot(df1, aes(x = truth, y = df1[,i]))  + labs(y = colnames(df1)[i]) + 
+    geom_point() + coord_cartesian(xlim=xlim, ylim=ylim)+ theme_grey()+ geom_abline(intercept = 0,slope = 1)
+
+  g1 <- ggplotGrob(zoomed1)
+  pic1 = base1 +annotation_custom(grob = g1,  xmin = bound[1], xmax = bound[2], ymin = bound[3], ymax = bound[4])
+  return(pic1)
+}
+
+pic_list = list() 
+for(i in 2:7){
+  pic_list[[(i - 1)]] = zoom_plot(df1,i,z_size,xybound)
+}
+plot_grid(pic_list[[1]], pic_list[[2]], 
+          pic_list[[3]], pic_list[[4]], 
+          pic_list[[5]], pic_list[[6]],  labels = "AUTO", ncol = 2)
+```
+
+### intermediate sparse
+
+```{r}
+library(cowplot)
+set.seed(99)
+L_se = c( 0.25, 0.5, 1, 2, 4)
+L_pi = c(1, 1,1,1,1)
+L_pi = L_pi / sum(L_pi)
+N = 200
+P = 300
+Data = datamaker(N,P,L_pi,L_se,0.7,c(1),c(1),1,sqrt(16))
+dim(Data$Y)
+gf = flashr::flash(Data$Y)
+#plot(as.vector(Data$L_true %*% t(Data$F_true)),as.vector(gf$l %*% t(gf$f)),pch = 16,col = "red")
+#abline(0,1,col = "red")
+PMDlog = capture.output({gPMD = invisible( PMA.wrapper(Data$Y)) })
+#plot(as.vector(Data$L_true %*% t(Data$F_true)),as.vector( (gPMD$u * gPMD$d) %*% t(gPMD$v) ),pch = 16,col = "blue")
+#plot(Data$L_true/sqrt(sum((Data$L_true)^2)),gf$l/sqrt(sum((gf$l)^2)))
+#plot(Data$L_true/sqrt(sum((Data$L_true)^2)),gPMD$u/sqrt(sum((gPMD$u )^2)))
+#plot(Data$L_true/sqrt(sum((Data$L_true)^2)),gf$l/sqrt(sum((gf$l)^2)),ylim = c(-0.1,0.1),xlim = c(-0.1,0.1))
+#plot(Data$L_true/sqrt(sum((Data$L_true)^2)),gPMD$u/sqrt(sum((gPMD$u )^2)),ylim = c(-0.1,0.1),xlim = c(-0.1,0.1))
+#plot(gPMD$u/sqrt(sum((gPMD$u )^2)),gf$l/sqrt(sum((gf$l)^2)))
+#abline(0,1)
+
+sqrt(mean((gf$l %*% t(gf$f) - Data$L_true %*% t(Data$F_true))^2))/sqrt(mean((Data$L_true %*% t(Data$F_true))^2))
+sqrt(mean(((gPMD$u * gPMD$d) %*% t(gPMD$v) - Data$L_true %*% t(Data$F_true))^2))/sqrt(mean((Data$L_true %*% t(Data$F_true))^2))
+# result from sfa
+
+gsfamix_1 = readRDS("../data/simulation/rankone/mediansparse/sfamix_2.rds")
+gsfamix_2 = readRDS("../data/simulation/rankone/mediansparse/sfamix_3.rds")
+gsfa = readRDS("../data/simulation/rankone/mediansparse/sfa_res.rds")
+
+# to get the KSVD result
+# try other methods
+library("R.matlab")
+# saveRDS(Data,file = "../data/simulation/rankone/mediansparse/test_1.rds")
+# write the data into matlab
+# writeMat("../data/simulation/rankone/mediansparse/test_1.mat", data = Data)
+# KSVD
+res_KSVD = readMat("../data/simulation/rankone/mediansparse/KSVDtest_1.mat",sparseMatrixClass="matrix")
+Dic = res_KSVD$result[[1]]
+loading_Dic = res_KSVD$result[[2]]
+MSE_KSVD = sqrt(mean(( Dic %*% loading_Dic - Data$L_true%*%t(Data$F_true))^2))/sqrt(mean((Data$L_true%*%t(Data$F_true))^2))
+
+library(ggplot2)
+library(grid)
+# zoom bounding box
+xlim <- c(-0.1,0.1); ylim <- c(-0.1,0.1)
+
+mat = cbind(Data$L_true,gf$l,gPMD$u,gsfa$L,as.vector(gsfamix_1$F),as.vector(gsfamix_2$F),as.vector(Dic))
+mat = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x] * sign(cor(mat[,x],mat[,1])) })
+mat1 = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x]/sqrt(sum((mat[,x])^2)) } )
+colnames(mat1) = c("truth","FLASH","PMD","SFA","SFAmix.1","SFAmix.2","KSVD")
+df1 = data.frame(mat1)
+
+base1 <- ggplot(df1, aes(x = truth, y = FLASH)) +
+  geom_point()  +
+  theme_bw()+ geom_abline(intercept = 0,slope = 1)
+zoomed1 = ggplot(df1, aes(x = truth, y = FLASH)) + 
+    geom_point() + coord_cartesian(xlim=xlim, ylim=ylim)+ theme_grey()+ geom_abline(intercept = 0,slope = 1)
+
+g1 <- ggplotGrob(zoomed1)
+pic1 = base1 +annotation_custom(grob = g1,  xmin = -0.35, xmax = 0, ymin = 0.05, ymax = 0.4)
+
+mat = cbind(Data$L_true,gf$l,gPMD$u,gsfa$L,as.vector(gsfamix_1$F),as.vector(gsfamix_2$F),as.vector(Dic))
+mat = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x] * sign(cor(mat[,x],mat[,1])) })
+mat1 = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x]/sqrt(sum((mat[,x])^2)) } )
+colnames(mat1) = c("truth","FLASH","PMD","SFA","SFAmix.1","SFAmix.2","KSVD")
+df1 = data.frame(mat1)
+xybound = c(-0.35,0,0.03,0.38)
+z_size = c(-0.07,0.07,-0.07,0.07)
+zoom_plot = function(df1,i,zoom_size,bound){
+  xlim <- c(zoom_size[1],zoom_size[2])
+  ylim <- c(zoom_size[3],zoom_size[4])
+
+  base1 <- ggplot(df1, aes(x = truth, y = df1[,i])) +
+    geom_point()  + labs(y = colnames(df1)[i]) +
+    theme_bw()+ geom_abline(intercept = 0,slope = 1)
+  zoomed1 = ggplot(df1, aes(x = truth, y = df1[,i]))  + labs(y = colnames(df1)[i]) + 
+    geom_point() + coord_cartesian(xlim=xlim, ylim=ylim)+ theme_grey()+ geom_abline(intercept = 0,slope = 1)
+
+  g1 <- ggplotGrob(zoomed1)
+  pic1 = base1 +annotation_custom(grob = g1,  xmin = bound[1], xmax = bound[2], ymin = bound[3], ymax = bound[4])
+  return(pic1)
+}
+
+pic_list = list() 
+for(i in 2:7){
+  pic_list[[(i - 1)]] = zoom_plot(df1,i,z_size,xybound)
+}
+plot_grid(pic_list[[1]], pic_list[[2]], 
+          pic_list[[3]], pic_list[[4]], 
+          pic_list[[5]], pic_list[[6]],  labels = "AUTO", ncol = 2)
+```
+
+### dense
+
+```{r}
+library(cowplot)
+set.seed(99)
+L_se = c( 0.25, 0.5, 1, 2, 4)
+L_pi = c(1, 1,1,1,1)
+L_pi = L_pi / sum(L_pi)
+N = 200
+P = 300
+Data = datamaker(N,P,L_pi,L_se,1,c(1),c(1),1,sqrt(100))
+dim(Data$Y)
+gf = flashr::flash(Data$Y)
+#plot(as.vector(Data$L_true %*% t(Data$F_true)),as.vector(gf$l %*% t(gf$f)),pch = 16,col = "red")
+#abline(0,1,col = "red")
+PMDlog = capture.output({gPMD = invisible( PMA.wrapper(Data$Y)) })
+#plot(as.vector(Data$L_true %*% t(Data$F_true)),as.vector( (gPMD$u * gPMD$d) %*% t(gPMD$v) ),pch = 16,col = "blue")
+#plot(Data$L_true/sqrt(sum((Data$L_true)^2)),gf$l/sqrt(sum((gf$l)^2)))
+#plot(Data$L_true/sqrt(sum((Data$L_true)^2)),gPMD$u/sqrt(sum((gPMD$u )^2)))
+#plot(Data$L_true/sqrt(sum((Data$L_true)^2)),gf$l/sqrt(sum((gf$l)^2)),ylim = c(-0.1,0.1),xlim = c(-0.1,0.1))
+#plot(Data$L_true/sqrt(sum((Data$L_true)^2)),gPMD$u/sqrt(sum((gPMD$u )^2)),ylim = c(-0.1,0.1),xlim = c(-0.1,0.1))
+#plot(gPMD$u/sqrt(sum((gPMD$u )^2)),gf$l/sqrt(sum((gf$l)^2)))
+#abline(0,1)
+
+sqrt(mean((gf$l %*% t(gf$f) - Data$L_true %*% t(Data$F_true))^2))/sqrt(mean((Data$L_true %*% t(Data$F_true))^2))
+sqrt(mean(((gPMD$u * gPMD$d) %*% t(gPMD$v) - Data$L_true %*% t(Data$F_true))^2))/sqrt(mean((Data$L_true %*% t(Data$F_true))^2))
+sqrt(mean((gsfa$L %*% gsfa$F - Data$L_true %*% t(Data$F_true))^2))/sqrt(mean((Data$L_true %*% t(Data$F_true))^2))
+# result from sfa
+
+gsfamix_1 = readRDS("../data/simulation/rankone/dense/sfamix_1.rds")
+gsfamix_2 = readRDS("../data/simulation/rankone/dense/sfamix_2.rds")
+gsfa = readRDS("../data/simulation/rankone/dense/sfa_res.rds")
+
+# to get the KSVD result
+# try other methods
+library("R.matlab")
+# saveRDS(Data,file = "../data/simulation/rankone/mediansparse/test_1.rds")
+# write the data into matlab
+# writeMat("../data/simulation/rankone/dense/test_1.mat", data = Data)
+# KSVD
+res_KSVD = readMat("../data/simulation/rankone/dense/KSVDtest_1.mat",sparseMatrixClass="matrix")
+Dic = res_KSVD$result[[1]]
+loading_Dic = res_KSVD$result[[2]]
+MSE_KSVD = sqrt(mean(( Dic %*% loading_Dic - Data$L_true%*%t(Data$F_true))^2))/sqrt(mean((Data$L_true%*%t(Data$F_true))^2))
+
+library(ggplot2)
+library(grid)
+# zoom bounding box
+xlim <- c(-0.1,0.1); ylim <- c(-0.1,0.1)
+
+mat = cbind(Data$L_true,gf$l,gPMD$u,gsfa$L,as.vector(gsfamix_1$F),rep(as.vector(gsfamix_2$F),N),as.vector(Dic))
+mat = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x] * sign(cor(mat[,x],mat[,1])) })
+mat1 = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x]/sqrt(sum((mat[,x])^2)) } )
+mat1[,6] = 0
+colnames(mat1) = c("truth","FLASH","PMD","SFA","SFAmix.1","SFAmix.2","KSVD")
+df1 = data.frame(mat1)
+xybound = c(-0.35, -0.05, 0.03, 0.38)
+z_size = c(-0.07,0.07,-0.07,0.07)
+zoom_plot = function(df1,i,zoom_size,bound){
+  xlim <- c(zoom_size[1],zoom_size[2])
+  ylim <- c(zoom_size[3],zoom_size[4])
+
+  base1 <- ggplot(df1, aes(x = truth, y = df1[,i])) +
+    geom_point()  + labs(y = colnames(df1)[i]) + coord_cartesian(xlim=c(-0.3,0.3), ylim= c(-0.36,0.36) ) + 
+    theme_bw()+ geom_abline(intercept = 0,slope = 1)
+  zoomed1 = ggplot(df1, aes(x = truth, y = df1[,i]))  + labs(y = colnames(df1)[i]) + 
+    geom_point() + coord_cartesian(xlim=xlim, ylim=ylim)+ theme_grey()+ geom_abline(intercept = 0,slope = 1)
+
+  g1 <- ggplotGrob(zoomed1)
+  pic1 = base1 +annotation_custom(grob = g1,  xmin = bound[1], xmax = bound[2], ymin = bound[3], ymax = bound[4])
+  return(pic1)
+}
+
+pic_list = list() 
+for(i in 2:7){
+  pic_list[[(i - 1)]] = zoom_plot(df1,i,z_size,xybound)
+}
+plot_grid(pic_list[[1]], pic_list[[2]], 
+          pic_list[[3]], pic_list[[4]], 
+          pic_list[[5]], pic_list[[6]],  labels = "AUTO", ncol = 2)
+```
+
 ## Session Information

### 2017-04-05 analysis/simulation3.Rmd
Files commited by wflow_commit.

---
title: "simulation for the sparse factor loadings"
author: "Wei Wang"
date: 2017-04-04
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

```{r}

CVPMD=function(Y,c_s,K){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = matrix(sample(P,P),ncol = 5)
  rowindex = matrix(sample(N,N),ncol = 5)

  missing= array(0,dim = c(5,N,P))
  foldindex = array(0,dim = c(5,5,2))
  for(i in 1:5){
    for(j in 1:5){
      foldindex[i,j,1] = i
      foldindex[i,j,2] = (i+j) %% 5
    }
  }
  foldindex[which(foldindex == 0)] = 5
  for(i in 1:5){
    missing[i, , ] = Y
    for(j in 1:5){
      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
    }
  }
  n_s = length(c_s)
  CVRMSE = rep(NA,n_s)
  minrmse = Inf
  opt_s = 0
  # here we only need one layer
  for(t_s in 1:n_s){
    rmse = rep(0,5)
    for(i in 1:5){
      out = PMD(missing[i,,], sumabs = c_s[t_s], sumabsv = NULL, sumabsu = NULL,K = K)
      misshat = out$u %*%  diag(out$d) %*% t(out$v)
      for(j in 1:5){
        rmse[i] = rmse[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] - 
                                   misshat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2)
      }
      rmse[i] = sqrt(rmse[i] / (N * P/5))
    }
    CVRMSE[t_s] = mean(rmse)
    if(CVRMSE[t_s] < minrmse){
      minrmse = CVRMSE[t_s]
      opt_s = c_s[t_s]
    }
  }

  return(list(opt_s = opt_s ))
}

PMA.wrapper = function(Y,ngrids = 10,K){
  library(PMA)
  N = dim(Y)[1]
  P = dim(Y)[2]
  c_s = seq(0.2,0.8,len=ngrids)
  cvout = CVPMD(Y,c_s, K = K)
  out = PMD(Y,sumabsu = NULL, sumabsv = NULL, sumabs = cvout$opt_s ,K = K)
  residual_PMD = Y -  out$u %*%  diag(out$d) %*% t(out$v)
  return(list(d = out$d, u = out$u, v = out$v, residual = residual_PMD))
}

```

```{r}
set.seed(99)
N = 150
P = 240
K = 3
L_true = array(0,dim = c(N,K))
F_true = array(0,dim = c(P,K))
F_true[1:80,1] = rnorm(80,0,0.5)
F_true[81:160,2] = rnorm(80,0,1)
F_true[161:240,3] = rnorm(80,0,2)
L_true[1:10,1] = rnorm(10,0,2)
L_true[11:60,2] = rnorm(50,0,1)
L_true[61:150,3] = rnorm(90,0,0.5)
G = L_true %*% t(F_true)
# generate Y
E = matrix(rnorm(N*P,0,2),nrow=N)
Y = L_true %*% t(F_true) + E
g_flash = flashr::greedy(Y,K = 5)
g_b_flash = flashr::backfitting(Y,initial_list = g_flash)

outcv = PMA.wrapper(Y,10,K = 3)
```

```{r}
par(mfrow = c(2,2))
image( t(apply(t(g_b_flash$l %*% t(g_b_flash$f)),2,rev)),col = grey(seq(0,1,length.out = 5)),main = "FLASH")
image(t(apply(t(outcv$u %*% diag(outcv$d) %*% t(outcv$v)),2,rev)),col = grey(seq(0,1,length.out = 5)),main = "PMD")
image(t(apply(t(Y),2,rev)),col = grey(seq(0,1,length.out = 5)),main = "observed data")
image(t(apply(t(G),2,rev)),col = grey(seq(0,1,length.out = 5)),main = "underlying structure")
par(mfrow = c(2,2))
image( t(apply(t(g_b_flash$l %*% t(g_b_flash$f)),2,rev)),col =blues9,main = "FLASH")
image(t(apply(t(outcv$u %*% diag(outcv$d) %*% t(outcv$v)),2,rev)),col = blues9,main = "PMD")
image(t(apply(t(Y),2,rev)),col = blues9,main = "observed data")
image(t(apply(t(G),2,rev)),col = blues9,main = "underlying structure")
```

```{r}
library(ggplot2)
library(cowplot)
mat = cbind(L_true,g_b_flash$l,outcv$u)
mat[,8] = -mat[,8]
mat[,9] = -mat[,9]
mat[,6] = -mat[,6]
mat1 = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x]/sqrt(sum((mat[,x])^2)) } )
colnames(mat1) = c("truth.l1","truth.l2","truth.l3","FLASH.l3","FLASH.l2","FLASH.l1","PMD.u3","PMD.u2","PMD.u1")
df1 = data.frame(mat1)
p1_1_l <- ggplot(df1, aes(truth.l1, FLASH.l1)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("loading 1 (FLASH)")
p1_2_l <- ggplot(df1, aes(truth.l2, FLASH.l2)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("loading 2 (FLASH)")
p1_3_l <- ggplot(df1, aes(truth.l3, FLASH.l3)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("loading 3 (FLASH)")
p2_1_l <- ggplot(df1, aes(truth.l1, PMD.u1)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("loading 1 (PMD)")
p2_2_l <- ggplot(df1, aes(truth.l2, PMD.u2)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("loading 2 (PMD)")
p2_3_l <- ggplot(df1, aes(truth.l3, PMD.u3)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("loading 3 (PMD)")

library(ggplot2)
library(cowplot)
mat = cbind(F_true,g_b_flash$f,outcv$v)
mat[,8] = -mat[,8]
mat[,9] = -mat[,9]
mat[,6] = -mat[,6]
mat1 = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x]/sqrt(sum((mat[,x])^2)) } )
colnames(mat1) = c("truth.f1","truth.f2","truth.f3","FLASH.f3","FLASH.f2","FLASH.f1","PMD.v3","PMD.v2","PMD.v1")
df1 = data.frame(mat1)
p1_1 <- ggplot(df1, aes(truth.f1, FLASH.f1)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("factor 1 (FLASH)")
p1_2 <- ggplot(df1, aes(truth.f2, FLASH.f2)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("factor 2 (FLASH)")
p1_3 <- ggplot(df1, aes(truth.f3, FLASH.f3)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("factor 3 (FLASH)")
p2_1 <- ggplot(df1, aes(truth.f1, PMD.v1)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("factor 1 (PMD)")
p2_2 <- ggplot(df1, aes(truth.f2, PMD.v2)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("factor 2 (PMD)")
p2_3 <- ggplot(df1, aes(truth.f3, PMD.v3)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("factor 3 (PMD)")
plot_grid(p1_1_l, p2_1_l, p1_1, p2_1,
          p1_2_l, p2_2_l, p1_2, p2_2,
          p1_3_l, p2_3_l, p1_3, p2_3,labels = "" , ncol = 4)
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```

### 2017-04-07 analysis/DNAdata.Rmd
Files commited by wflow_commit.

--- 2017-04-05 analysis/simulation3.Rmd
@@
 ---
-title: "simulation for the sparse factor loadings"
-author: "Wei Wang"
-date: 2017-04-04
+title: "Untitled"
+author: "First Last"
+date: YYYY-MM-DD
 output: html_document
@@
 ```{r}
-
-CVPMD=function(Y,c_s,K){
-  N = dim(Y)[1]
-  P = dim(Y)[2]
-  colindex = matrix(sample(P,P),ncol = 5)
-  rowindex = matrix(sample(N,N),ncol = 5)
-
-  missing= array(0,dim = c(5,N,P))
-  foldindex = array(0,dim = c(5,5,2))
-  for(i in 1:5){
-    for(j in 1:5){
-      foldindex[i,j,1] = i
-      foldindex[i,j,2] = (i+j) %% 5
-    }
-  }
-  foldindex[which(foldindex == 0)] = 5
-  for(i in 1:5){
-    missing[i, , ] = Y
-    for(j in 1:5){
-      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
-    }
-  }
-  n_s = length(c_s)
-  CVRMSE = rep(NA,n_s)
-  minrmse = Inf
-  opt_s = 0
-  # here we only need one layer
-  for(t_s in 1:n_s){
-    rmse = rep(0,5)
-    for(i in 1:5){
-      out = PMD(missing[i,,], sumabs = c_s[t_s], sumabsv = NULL, sumabsu = NULL,K = K)
-      misshat = out$u %*%  diag(out$d) %*% t(out$v)
-      for(j in 1:5){
-        rmse[i] = rmse[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] - 
-                                   misshat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2)
-      }
-      rmse[i] = sqrt(rmse[i] / (N * P/5))
-    }
-    CVRMSE[t_s] = mean(rmse)
-    if(CVRMSE[t_s] < minrmse){
-      minrmse = CVRMSE[t_s]
-      opt_s = c_s[t_s]
-    }
-  }
-
-  return(list(opt_s = opt_s ))
-}
-
-PMA.wrapper = function(Y,ngrids = 10,K){
-  library(PMA)
-  N = dim(Y)[1]
-  P = dim(Y)[2]
-  c_s = seq(0.2,0.8,len=ngrids)
-  cvout = CVPMD(Y,c_s, K = K)
-  out = PMD(Y,sumabsu = NULL, sumabsv = NULL, sumabs = cvout$opt_s ,K = K)
-  residual_PMD = Y -  out$u %*%  diag(out$d) %*% t(out$v)
-  return(list(d = out$d, u = out$u, v = out$v, residual = residual_PMD))
-}
-
-```
-
-```{r}
-set.seed(99)
-N = 150
-P = 240
-K = 3
-L_true = array(0,dim = c(N,K))
-F_true = array(0,dim = c(P,K))
-F_true[1:80,1] = rnorm(80,0,0.5)
-F_true[81:160,2] = rnorm(80,0,1)
-F_true[161:240,3] = rnorm(80,0,2)
-L_true[1:10,1] = rnorm(10,0,2)
-L_true[11:60,2] = rnorm(50,0,1)
-L_true[61:150,3] = rnorm(90,0,0.5)
-G = L_true %*% t(F_true)
-# generate Y
-E = matrix(rnorm(N*P,0,2),nrow=N)
-Y = L_true %*% t(F_true) + E
-g_flash = flashr::greedy(Y,K = 5)
-g_b_flash = flashr::backfitting(Y,initial_list = g_flash)
-
-outcv = PMA.wrapper(Y,10,K = 3)
-```
-
-```{r}
-par(mfrow = c(2,2))
-image( t(apply(t(g_b_flash$l %*% t(g_b_flash$f)),2,rev)),col = grey(seq(0,1,length.out = 5)),main = "FLASH")
-image(t(apply(t(outcv$u %*% diag(outcv$d) %*% t(outcv$v)),2,rev)),col = grey(seq(0,1,length.out = 5)),main = "PMD")
-image(t(apply(t(Y),2,rev)),col = grey(seq(0,1,length.out = 5)),main = "observed data")
-image(t(apply(t(G),2,rev)),col = grey(seq(0,1,length.out = 5)),main = "underlying structure")
-par(mfrow = c(2,2))
-image( t(apply(t(g_b_flash$l %*% t(g_b_flash$f)),2,rev)),col =blues9,main = "FLASH")
-image(t(apply(t(outcv$u %*% diag(outcv$d) %*% t(outcv$v)),2,rev)),col = blues9,main = "PMD")
-image(t(apply(t(Y),2,rev)),col = blues9,main = "observed data")
-image(t(apply(t(G),2,rev)),col = blues9,main = "underlying structure")
-```
-
-```{r}
-library(ggplot2)
-library(cowplot)
-mat = cbind(L_true,g_b_flash$l,outcv$u)
-mat[,8] = -mat[,8]
-mat[,9] = -mat[,9]
-mat[,6] = -mat[,6]
-mat1 = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x]/sqrt(sum((mat[,x])^2)) } )
-colnames(mat1) = c("truth.l1","truth.l2","truth.l3","FLASH.l3","FLASH.l2","FLASH.l1","PMD.u3","PMD.u2","PMD.u1")
-df1 = data.frame(mat1)
-p1_1_l <- ggplot(df1, aes(truth.l1, FLASH.l1)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("loading 1 (FLASH)")
-p1_2_l <- ggplot(df1, aes(truth.l2, FLASH.l2)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("loading 2 (FLASH)")
-p1_3_l <- ggplot(df1, aes(truth.l3, FLASH.l3)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("loading 3 (FLASH)")
-p2_1_l <- ggplot(df1, aes(truth.l1, PMD.u1)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("loading 1 (PMD)")
-p2_2_l <- ggplot(df1, aes(truth.l2, PMD.u2)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("loading 2 (PMD)")
-p2_3_l <- ggplot(df1, aes(truth.l3, PMD.u3)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("loading 3 (PMD)")
-
-library(ggplot2)
-library(cowplot)
-mat = cbind(F_true,g_b_flash$f,outcv$v)
-mat[,8] = -mat[,8]
-mat[,9] = -mat[,9]
-mat[,6] = -mat[,6]
-mat1 = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x]/sqrt(sum((mat[,x])^2)) } )
-colnames(mat1) = c("truth.f1","truth.f2","truth.f3","FLASH.f3","FLASH.f2","FLASH.f1","PMD.v3","PMD.v2","PMD.v1")
-df1 = data.frame(mat1)
-p1_1 <- ggplot(df1, aes(truth.f1, FLASH.f1)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("factor 1 (FLASH)")
-p1_2 <- ggplot(df1, aes(truth.f2, FLASH.f2)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("factor 2 (FLASH)")
-p1_3 <- ggplot(df1, aes(truth.f3, FLASH.f3)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("factor 3 (FLASH)")
-p2_1 <- ggplot(df1, aes(truth.f1, PMD.v1)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("factor 1 (PMD)")
-p2_2 <- ggplot(df1, aes(truth.f2, PMD.v2)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("factor 2 (PMD)")
-p2_3 <- ggplot(df1, aes(truth.f3, PMD.v3)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("factor 3 (PMD)")
-plot_grid(p1_1_l, p2_1_l, p1_1, p2_1,
-          p1_2_l, p2_2_l, p1_2, p2_2,
-          p1_3_l, p2_3_l, p1_3, p2_3,labels = "" , ncol = 4)
+library(readxl)
+Data = read_excel("../data/DNAdata/sd03.xls")
 ```

### 2017-04-11 analysis/simulation3.Rmd
Files commited by wflow_commit.

@@
 ```{r}
-par(mfrow = c(2,2))
-image( t(apply(t(g_b_flash$l %*% t(g_b_flash$f)),2,rev)),col = grey(seq(0,1,length.out = 5)),main = "FLASH")
-image(t(apply(t(outcv$u %*% diag(outcv$d) %*% t(outcv$v)),2,rev)),col = grey(seq(0,1,length.out = 5)),main = "PMD")
-image(t(apply(t(Y),2,rev)),col = grey(seq(0,1,length.out = 5)),main = "observed data")
-image(t(apply(t(G),2,rev)),col = grey(seq(0,1,length.out = 5)),main = "underlying structure")
-par(mfrow = c(2,2))
-image( t(apply(t(g_b_flash$l %*% t(g_b_flash$f)),2,rev)),col =blues9,main = "FLASH")
-image(t(apply(t(outcv$u %*% diag(outcv$d) %*% t(outcv$v)),2,rev)),col = blues9,main = "PMD")
-image(t(apply(t(Y),2,rev)),col = blues9,main = "observed data")
-image(t(apply(t(G),2,rev)),col = blues9,main = "underlying structure")
+library(reshape2)
+melted_cormat_1 <- melt(G, na.rm = TRUE)
+melted_cormat_2 <- melt(g_b_flash$l %*% t(g_b_flash$f), na.rm = TRUE)
+melted_cormat_3 <- melt(outcv$u %*% diag(outcv$d) %*% t(outcv$v), na.rm = TRUE)
+melted_cormat_4 <- melt(Y, na.rm = TRUE)
+# Heatmap
+library(ggplot2)
+p1 = ggplot(data = melted_cormat_1, aes(Var2, Var1, fill = value))+
+ geom_tile(color = "white")+
+ scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
+   midpoint = 0, limit = c(-2,2), space = "Lab") + labs(title = "truth", y = "sample", x = "variable") +
+  theme_minimal() + theme(legend.position="none",plot.background = element_rect())
+
+p2 = ggplot(data = melted_cormat_2, aes(Var2, Var1, fill = value))+
+ geom_tile(color = "white")+
+ scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
+   midpoint = 0, limit = c(-2,2), space = "Lab") + labs(title = "FLASH",y = "sample", x = "variable") +
+  theme_minimal() + theme(legend.position="none",plot.background = element_rect())
+
+p3 = ggplot(data = melted_cormat_3, aes(Var2, Var1, fill = value))+
+ geom_tile(color = "white")+
+ scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
+   midpoint = 0, limit = c(-2,2), space = "Lab") + labs(title = "PMD",y = "sample", x = "variable") +
+  theme_minimal() + theme(legend.position="none",plot.background = element_rect())
+
+p4 = ggplot(data = melted_cormat_4, aes(Var2, Var1, fill = value))+
+ geom_tile(color = "white")+
+ scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
+   midpoint = 0, limit = c(-2,2), space = "Lab") + labs(title = "observed data",y = "sample", x = "variable") +
+  theme_minimal() + theme(legend.position="none",plot.background = element_rect())
+
+library(cowplot)
+plot_grid(p1,p4,p3,p2, ncol = 2)
 ```

### 2017-04-13 analysis/DNAdata.Rmd
Files commited by wflow_commit.

@@
 Data = read_excel("../data/DNAdata/sd03.xls")
+data = as.matrix(Data[,-c(1,2)])
+#saveRDS(data,file = "../data/DNAdata/sd03.rds")
+```
+
+```
+source("~/HG/flash/Rcode/flash.R")
+source("~/HG/flash/Rcode/greedy.R")
+source("~/HG/flash/Rcode/backfitting.R")
+Y = readRDS("sd03.rds")
+Y = t(Y)
+dim(Y)
+g_flash = greedy(Y, K = 30, flash_para = list(partype = "var_col"))
+saveRDS(g_flash, "./logcpm/gflash_var_col_imput.rds")
+
+b_flash = backfitting(Y,initial_list = g_flash, flash_para = list(partype = "var_col"), maxiter_bf = 6)
+saveRDS(b_flash, "./logcpm/bflash_var_col_imput.rds")
+```
+
+```
+#!/bin/bash
+
+#SBATCH --job-name=arrayJob
+#SBATCH --output=/home/weidong/HG/flash/data/DNAdata/outlog/arrayJob_%A_%a.out
+#SBATCH --error=/home/weidong/HG/flash/data/DNAdata/outlog/arrayJob_%A_%a.err
+#SBATCH --array=1
+#SBATCH --time=32:00:00
+#SBATCH --partition=broadwl
+#SBATCH --ntasks=1
+#SBATCH --mem-per-cpu=32G
+
+######################
+# Begin work section #
+######################
+
+# Print this sub-job's task ID
+cd /home/weidong/HG/flash/data/DNAdata
+Rscript --verbose /home/weidong/HG/flash/data/DNAdata/flash_const_wrapper.R
+
+```
+
+```
+#!/bin/bash
+
+#SBATCH --job-name=arrayJob
+#SBATCH --output=/home/weidong/HG/flash/data/DNAdata/outlog/arrayJob_%A_%a.out
+#SBATCH --error=/home/weidong/HG/flash/data/DNAdata/outlog/arrayJob_%A_%a.err
+#SBATCH --array=1
+#SBATCH --time=32:00:00
+#SBATCH --partition=mstephens
+#SBATCH --account=pi-mstephens
+#SBATCH --nodes=1
+#SBATCH --ntasks-per-node=1
+#SBATCH --mem-per-cpu=32G
+
+######################
+# Begin work section #
+######################
+
+# Print this sub-job's task ID
+cd /home/weidong/HG/flash/data/DNAdata
+Rscript --verbose /home/weidong/HG/flash/data/DNAdata/flash_varcol_wrapper.R
+```
+
+```
+gc = readRDS("../data/DNAdata/b_flash_const.rds")
+gv = readRDS("../data/DNAdata/bflash_varcol.rds")
+gsvd = svd(t(data))
+library(flashr)
+library(reshape2)
+melted_data <- melt((gv$l %*% t(gv$f))[1:30,1:20000], na.rm = TRUE)
+# Heatmap
+library(ggplot2)
+p1 = ggplot(data = melted_data, aes(Var2, Var1, fill = value))+
+ geom_tile(color = "white")+
+ scale_fill_gradient2(low = "black", high = "red", mid = "blue", 
+   midpoint = 0, limit = c(-0.6,0.6), space = "Lab") + labs(title = "truth", y = "sample", x = "variable") +
+  theme_minimal() + theme(legend.position="none",plot.background = element_rect())
+
+image(t(gv$l %*% t(gv$f)),col = rainbow(7))
+
+i = 3
+par(mfrow = c(2,2))
+barplot(gc$l[,i])
+barplot(gv$l[,i])
+barplot(gsvd$u[,i])
 ```

### 2017-04-25 analysis/first-analysis.Rmd
Files commited by wflow_commit.

@@
 author: "Wei Wang"
-date: 2017-02-07
+date: 2017-04-24
 output: html_document
@@
 
-$$X=  \begin{pmatrix} X_{11} & X_{12} \\ X_{21} & X_{22}  \end{pmatrix}= \left[ \begin{array}{c} U_{1} \\ U_{2} \end{array} \right] \Gamma \left[ \begin{array}{cc} V_{1}^{T}& V_{2}^{T} \end{array} \right] $$
+$$X=  \begin{pmatrix} X_{00} & X_{01} \\ X_{10} & X_{11}  \end{pmatrix}= \left[ \begin{array}{c} U_{0} \\ U_{1} \end{array} \right] \Sigma \left[ \begin{array}{cc} V_{0}^{T}& V_{1}^{T} \end{array} \right] $$
 
 ### 1 algorithm to find original svd
+
+This is one of the old quesitons and I think I would like to keep this question.
 
@@
 \begin{eqnarray}
-&\min_{U,V,\Gamma}& ||X_{11} - U_1 \Gamma V_1^T||_2 + ||X_{22} - U_2 \Gamma V_2^T||_2 \\
-&st& U_1 U_1^T +  U_2 U_2^T = I \\
-& & V_1 V_1^T +  V_2 V_2^T = I
+&\min_{U,V,\Sigma}& ||X_{01} - U_0 \Sigma V_1^T||_2 + ||X_{10} - U_1 \Sigma V_0^T||_1 \\
+&st& ||X_{00} - U_0 \Sigma V_0^T||_2 + ||X_{11} - U_1 \Sigma V_1^T||_1 = 0\\
+&& U_1 U_1^T +  U_0 U_0^T = I \\
+& & V_1 V_1^T +  V_0 V_0^T = I
 \end{eqnarray}
@@
 
-### 2 separate svd vs original svd
+Maybe this is too ambitious? __Let's start from a simple toy example to check if it is possible.__
 
-we know that
+### 2 toy example with low rank assumption
+
+#### 2.1  2 by 2 case
+
+Here we take the simpliest example: 2 by 2 matrix $\begin{pmatrix} a & b \\ c & d  \end{pmatrix}$
+
+#### 2.1.1 BCV in 2 by 2 case
+
+In the original BCV frame work, we can use $||A - B D_k^+C||_F^2$ as objective function.
+
+As I understand here, in the simpliest case, $\begin{pmatrix} a & b \\ c & d  \end{pmatrix}$ is $2 \times 2$ matrix and we assume the this matrix is a rank one matrix.
+So we can use $a = bd^{-1}c$ to "impute" (or "predict") the value of $a$.
+
+For example, we know the the matrix is $\begin{pmatrix} a & 3 \\ 2 & 3  \end{pmatrix}$ with rank one, so we know $a = 2$.
+
+#### 2.1.1 OCV in 2 by 2 case
+
+What if we hold out two parts $b,c$ and once and hold out $a,b$ in the other time? This goes to two separate problems:
+
+- (2.1) For holding out $b,c$:
+
+If the matrix looks like $\begin{pmatrix} 2 & b \\ c & 3  \end{pmatrix}$ with rank one, we actually have more than one solutions.
+
+- (2.2) For holding our $a,d$:
+
+If the matrix looks like $\begin{pmatrix} a & 3 \\ 2 & d  \end{pmatrix}$ with rank one, we actually have more than one solutions.
+
+When we try to solve the two problems (2.1), (2.2) separately, it seems there is no gain from the hold out pattern.
+
+But what if we solve the together as we discussed previously on Orthogonal Cross Validation (OCV)?
+
+#### 2.2 Orthogonal Cross Validation (background)
+
+In OCV, we hold out each orthogonal part, and then take the average of MSE of the prediction and the hold out data, we take 3-fold as example:
+
+```{r,echo=F}
+cross_plot = function(Y,k_fold = 3){
+  N = dim(Y)[1]
+  P = dim(Y)[2]
+  colindex = matrix(sample(P,P),ncol = k_fold)
+  rowindex = matrix(sample(N,N),ncol = k_fold)
+
+  missing= array(0,dim = c(k_fold,N,P))
+  foldindex = array(0,dim = c(k_fold,k_fold,2))
+  for(i in 1:k_fold){
+    for(j in 1:k_fold){
+      foldindex[i,j,1] = i
+      foldindex[i,j,2] = (i+j) %% k_fold
+    }
+  }
+  foldindex[which(foldindex == 0)] = k_fold
+  for(i in 1:k_fold){
+    missing[i, , ] = Y
+    for(j in 1:k_fold){
+      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
+    }
+  }
+  par(mfrow = c(2,2), mar = (c(5,4,4,2)-1.9))
+  image(Y,main = "data Y")
+  for(i in 1:k_fold){
+    image(missing[i,,],main = "missing data")
+  }
+  return(missing)
+}
+
+N = 3
+P = 3
+Y = matrix(rep(1,N*P),ncol =  P)
+g = cross_plot(Y)
+```
+
+$$Y = \begin{pmatrix} Y_{11} & Y_{12} & Y_{13} \\ Y_{21} & Y_{22} & Y_{23} \\ Y_{31} & Y_{32} & Y_{33} \end{pmatrix}= \begin{pmatrix} Y_{(1)} & Y_{(2)} & Y_{(3)} \\ Y_{(3)} & Y_{(1)} & Y_{(2)} \\ Y_{(2)} & Y_{(3)} & Y_{(1)} \end{pmatrix}$$
+
+by letting $Y_{(1)} = \{  Y_{11},  Y_{22} , Y_{33}\}$, $Y_{(2)} = \{  Y_{12},  Y_{23} , Y_{31}\}$ and $Y_{(3)} = \{  Y_{13},  Y_{21} , Y_{32}\}$, which are orthogonal to each other.
+
+Then the MSE for each row and column permutation would be:
 
 \begin{eqnarray}
-X_{11} = U_1 \Gamma V_1^T \\
-X_{22} = U_2 \Gamma V_2^T
+MSE &=& MSE_1 + MSE_2 + MSE_3\\
+ &=& ||\hat{Y}_{(1)} - Y_{(1)}||_2 + ||\hat{Y}_{(2)} - Y_{(2)}||_2 + ||\hat{Y}_{(3)} - Y_{(3)}||_2
 \end{eqnarray}
 
-but $U_1 \Gamma V_1^T$ is not SVD for $X_{11}$.
-On the other hand, if we combine the svd for $X_{11} = \tilde{U}_1 \Sigma_{11}\tilde{V}_1$ and that for $X_{22} = \tilde{U}_2 \Sigma_{22}\tilde{V}_2$, the combining columns $\left[ \begin{array}{c} \tilde{U}_{1} \\ \tilde{U}_{2} \end{array}  \right]$ is not the original svd $\left[ \begin{array}{c} U_{1} \\ U_{2} \end{array} \right]$.
-It's because $\tilde{U}_{1}$ and $\tilde{U}_{2}$ just contains local information for $X_{11}$ and $X_{22}$.
+#### 2.3 2 by 2 matrix again
+
+Let's go back to 2 by 2 case:
+
+The question is can we get better idea by solving problem (2.1) and (2.2) together (by taking the average of MSE)?
+
+So I would guess that the problem goes to:
+
+\begin{eqnarray}
+& & \min_{u,v,\lambda} MSE_1 + MSE_2\\
+&s.t.& uu^T = I\\
+&& v v^T = I \\
+&& \lambda >0 \\
+&where& MSE_1 = ||a - u_0\lambda v_0||_F + ||d - u_1 \lambda v_1||_F\\
+&and& MSE_2 = ||b - u_0\lambda v_1||_F + ||c - u_1 \lambda v_0||_F
+\end{eqnarray}
+
+with assumption the $\begin{pmatrix} a & b \\ c & d  \end{pmatrix}$ is rank one matrix and SVD is as $$\begin{pmatrix} a & b \\ c & d  \end{pmatrix} = u\lambda v^T$$
+
+**So we can see that SVD solution is the solution of our problem. What's more, in this simple example, we might be able to find rank two solution which BCV can't.**

### 2017-04-25 analysis/first-analysis.Rmd
Files commited by wflow_commit.

@@
 
-This is one of the old quesitons and I think I would like to keep this question.
+__This is one of the old quesitons and I think I would like to keep this question.__
 
@@
 
-- We are not sure if this optimazation problem is convex or not.
+- We are not sure if this optimization problem is convex or not.
 
@@
 
-- To solve this problem might be more computationaly intensive comparing with BCV which just need to calculate svd and general inverse.
+- To solve this problem might be more computationally intensive comparing with BCV which just need to calculate svd and general inverse.
 
@@
 
-#### 2.1  2 by 2 case
+#### 2.1 Toy example: 2 by 2 case
 
-Here we take the simpliest example: 2 by 2 matrix $\begin{pmatrix} a & b \\ c & d  \end{pmatrix}$
+Here we take the simplest example: 2 by 2 matrix $\begin{pmatrix} a & b \\ c & d  \end{pmatrix}$
 
@@
 
-As I understand here, in the simpliest case, $\begin{pmatrix} a & b \\ c & d  \end{pmatrix}$ is $2 \times 2$ matrix and we assume the this matrix is a rank one matrix.
+As I understand here, in the simplest case, $\begin{pmatrix} a & b \\ c & d  \end{pmatrix}$ is $2 \times 2$ matrix and we assume the this matrix is a rank one matrix to make the determinant equal to zero.
 So we can use $a = bd^{-1}c$ to "impute" (or "predict") the value of $a$.
+__Thus we can get the solution.__
 
@@
 
-#### 2.1.1 OCV in 2 by 2 case
+#### 2.1.2 OCV in 2 by 2 case
 
@@
 
-But what if we solve the together as we discussed previously on Orthogonal Cross Validation (OCV)?
+**But what if we solve them together as we discussed previously using Orthogonal Cross Validation (OCV)?**
 
@@
 }
-
+set.seed(999)
 N = 3
@@
 
-Then the MSE for each row and column permutation would be:
+For each row and column permutation, we calculate the MSE as follows:
 
@@
 MSE &=& MSE_1 + MSE_2 + MSE_3\\
- &=& ||\hat{Y}_{(1)} - Y_{(1)}||_2 + ||\hat{Y}_{(2)} - Y_{(2)}||_2 + ||\hat{Y}_{(3)} - Y_{(3)}||_2
+ &=& ||\hat{Y}_{(1)} - Y_{(1)}||_F^2 + ||\hat{Y}_{(2)} - Y_{(2)}||_F^2 + ||\hat{Y}_{(3)} - Y_{(3)}||_F^2
 \end{eqnarray}
 
-#### 2.3 2 by 2 matrix again
+#### 2.3 back to 2 by 2 matrix again
 
@@
 
-The question is can we get better idea by solving problem (2.1) and (2.2) together (by taking the average of MSE)?
+- problem (2.1)
+
+\begin{eqnarray}
+& & \min_{u^*, v^*, \lambda^*, b, c} ||b - u_0^*\lambda^* v_1^*||_F^2 + ||c - u_1^* \lambda^* v_0^*||_F^2\\
+&s.t.& u^*u^{*T} = I\\
+&& v^* v^{*T} = I \\
+&& \lambda^* >0 \\
+& & ||a - u_0^*\lambda^* v_0^*||_F^2 + ||d - u_1^* \lambda^* v_1^*||_F^2 = 0
+\end{eqnarray}
+
+- problem (2.2)
+
+\begin{eqnarray}
+& & \min_{u^{**}, v^{**}, \lambda^{**}, a, d} ||a - u_0^{**}\lambda^{**} v_0^{**}||_F^2 + ||d - u_1^{**} \lambda^{**} v_1^{**}||_F^2\\
+&s.t.& u^**u^{**T} = I\\
+&& v^{**} v^{**T} = I \\
+&& \lambda^{**} >0 \\
+& & ||b - u_0^{**}\lambda^{**} v_1^{**}||_F^2 + ||c - u_1^{**} \lambda^{**} v_0^{**}||_F^2 = 0
+\end{eqnarray}
+
+**The question is can we get better idea by solving problem (2.1) and (2.2) together (by taking the average of MSE)?** It means that we want $u^*,v^*$ and $u^{**},v{**}$ to be same.
+Similarly as the MSE in OCV, we would like use the average (sum) of those two MSE.
 
@@
 && \lambda >0 \\
-&where& MSE_1 = ||a - u_0\lambda v_0||_F + ||d - u_1 \lambda v_1||_F\\
-&and& MSE_2 = ||b - u_0\lambda v_1||_F + ||c - u_1 \lambda v_0||_F
+&where& MSE_1 = ||a - u_0\lambda v_0||_F^2 + ||d - u_1 \lambda v_1||_F^2\\
+&and& MSE_2 = ||b - u_0\lambda v_1||_F^2 + ||c - u_1 \lambda v_0||_F^2
 \end{eqnarray}
@@
 **So we can see that SVD solution is the solution of our problem. What's more, in this simple example, we might be able to find rank two solution which BCV can't.**
+
+In practice, we might get $u_0^*, v_0^*, u_1^*, v_1^*$ in $MSE_1$ to predict $a,d$ and get $u_0^{**}, v_0^{**}, u_1^{**}, v_1^{**}$ in $MSE_2$ to predict $b,c$.

### 2017-04-27 analysis/meetnotes.Rmd
Files commited by wflow_commit.

---
title: "Meeting notes"
author: "Wei"
date: 2017-04-27
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`

## meeting summary 04-26

### on flash paper

#### OCV criteria

we will introduce OCV in FLASH paper for missing data impute. But the criteria of the OCV is RMSE of the data.
$$RMSE(\hat{Y},\Omega) = \sqrt{\frac{1}{|\Omega|}\sum_{ij \in \Omega}(Y_{(ij)} - \hat{Y}_{(ij)})^2}$$
 where the $\Omega$ is the set of missing index. However we discussed that there is another criteria we can use.

 First fit the model 
 $$Y_{ij} = \alpha_i +\beta_j + N(0,\sigma_i^2 \sigma_j^2)$$
 $$Z_{ij} = \frac{Y_{ij} - \alpha_i -\beta_j}{\sigma_i \sigma_j} $$
and use

$$\frac{\sqrt{\frac{1}{|\Omega|}\sum_{ij \in \Omega}(Z_{(ij)} - \hat{Z}_{(ij)})^2}}{\sqrt{\frac{1}{|\Omega|}\sum_{ij \in \Omega}(Z_{(ij)} )^2}}$$

To compare which one is better, we will try both method and compare the result. In the second criteria, we can impute the missing data and transform $Z_{ij}$ back to $Y_{ij}$ to make the results comparable.

#### On the lemma

In our meeting, Matthew mentioned that we can provide a general lemma to show that the VEM factor analysis can be convert to EB approach of normal means problem.

#### psuedo algorithm

I should add the step of estimating $\sigma^2$ into the psuedo algorithm

#### combine the marginalization on $f$ and maximization on $l$ together

The two steps are basically for one problem so we can combine these two subsections together.

#### barplot for flash result

In the Gtex data, the plot is too big to fit one page. try to break the plot into two pages.

## Session Information

```{r session-info}
```

### 2017-05-01 analysis/pairwiseplotofflashgtexzscore.Rmd
Files commited by wflow_commit.

--- 2017-04-07 analysis/DNAdata.Rmd
@@
 ---
-title: "Untitled"
-author: "First Last"
-date: YYYY-MM-DD
+title: "pairwise plot of flash result of flash Z-score"
+author: "Wei Wang"
+date: 2017-05-01
 output: html_document
@@
 ```{r}
-library(readxl)
-Data = read_excel("../data/DNAdata/sd03.xls")
+load("../data/GTExdata/bflashvarcol.rds")
+dat = read.table('../data/GTExColors.txt', sep = '\t', comment.char = '')
+colordata = dat[c(1:6,9:18,21:23,26:30,32,33,35,36,38:53),1:2]
+```
+
+```{r}
+plot(b_flash$l[,1],b_flash$l[,3],col = as.vector(colordata[,2]),pch = 16)
+text(b_flash$l[,1],b_flash$l[,3], labels=colordata[,1], cex= 0.7)
 ```

### 2017-05-01 analysis/pairwiseplotofflashgtexzscore.Rmd
Files commited by wflow_commit.

@@
 ```{r}
-plot(b_flash$l[,1],b_flash$l[,3],col = as.vector(colordata[,2]),pch = 16)
-text(b_flash$l[,1],b_flash$l[,3], labels=colordata[,1], cex= 0.7)
+i = 1
+j = 3
+plot(b_flash$l[,i],b_flash$l[,j],col = as.vector(colordata[,2]),pch = 16)
+text(b_flash$l[,i],b_flash$l[,j], labels=colordata[,1], cex= 0.7)
 ```

### 2017-05-03 analysis/pairwiseplotofflashgtexzscore.Rmd
Files commited by wflow_commit.

@@
 
+### plot of $l_1$ vs $l_3$
+
 ```{r}
@@
 
+### plot $f_1$ vs $f_3$
+
+```{r}
+i = 1
+j = 3
+plot(b_flash$f[,i],b_flash$f[,j],col = "grey",pch = 16)
+
+```
+
+## logcpm of count
+
+here I use
+- 2 brain tissues
+- testis + pituitary
+- 2 skin
+- Mammary+ adipose
+- 2 heart
+- uterus + Ovary
+
+for each tissue, we random pick 50 smaples.
+
+```{r}
+bfl = readRDS("../data/GTExdata/logcpm/sixpairssubsample/bflash_varcol.rds")
+```
+
+I run flash with K=60 and it provide 60 factors in greedy algorithm, which mean it didn't stop. I run backfitting procedure with 6 iteration due to the time limit, it still provide 60 factors.
+
+I am not sure the factor are useful a
+
+```{r}
+color_six = c(rep("#eeee00",100),rep("#aaaaaa",50),rep("#aaff99",50),
+              rep("#0000ff",50),rep("#7777ff",50),
+              rep("#ffaa00",50),rep("#33cccc",50),
+              rep("#9900ff",50),rep("#660099",50),
+              rep("#ff66ff",50),rep("#ffaaff",50))
+
+par(mfrow = c(3,3),mar = c(5,4,4,2)-1.8)
+for(i in 1:60){
+  barplot(bfl$l[,i],col = color_six,border = color_six)
+}
+```
+
 ## Session information

### 2017-05-03 analysis/pairwiseplotofflashgtexzscore.Rmd
Files commited by wflow_commit.

@@
 ---
-title: "pairwise plot of flash result of flash Z-score"
+title: "pairwise plot of flash result of flash Z-score and loading of flash on logcpm of count data"
 author: "Wei Wang"
@@
 
+## pairwise loading plot of Z-score
+
 ```{r}
@@
 plot(b_flash$f[,i],b_flash$f[,j],col = "grey",pch = 16)
-
+plot(as.vector(b_flash$l[,i] %*% t(b_flash$f[,i])),
+     as.vector(b_flash$l[,j] %*% t(b_flash$f[,j])),col = "grey",pch = 16)
 ```
@@
 
+```{r}
+gsvd = readRDS("../data/GTExdata/logcpm/sixpairssubsample/svd60.rds")
+par(mfrow = c(3,3),mar = c(5,4,4,2)-1.8)
+for(i in 1:60){
+  barplot(gsvd$u[,i],col = color_six,border = color_six)
+}
+```
+
 ## Session information

### 2017-05-22 analysis/imputestratergy.Rmd
Files commited by wflow_commit.

---
title: "impute stragergy"
author: "Wei Wang"
date: 2017-05-22
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

## try the breast cancer data

```{r}
library(R.matlab)
library(PMA)
datapath = "~/HG/ash-sfa/Rcode/postmean/flash_simulation/missingvalue/nsfaexample.mat"
Y_centered = readMat(datapath)
Y = Y_centered$Ycentered
library(flashr)
library(softImpute)
## to get the biscale data
Y_bs = biScale(Y)

para_list = attributes(Y_bs)

set.seed(999)
# set the missing position

row_index = sample(dim(Y)[1],20)
col_index = sample(dim(Y)[2],20)
Y_miss = Y[row_index,col_index]

Y_input = Y
Y_input[row_index,col_index] = NA
Y_bs_input = Y_bs
Y_bs_input[row_index,col_index] = NA

#g1 = flashr::greedy(Y_input, K  = 60)
#g2 = flashr::greedy(Y_bs_input, K  = 60)
#saveRDS(g1,"../data/missingvalue/imputescheme/g1_original.rds")
#saveRDS(g2,"../data/missingvalue/imputescheme/g2_biscale.rds")
g1 = readRDS("../data/missingvalue/imputescheme/g1_original.rds")
g2 = readRDS("../data/missingvalue/imputescheme/g2_biscale.rds")
Y_hat_1 = (g1$l %*% t(g1$f))[row_index,col_index]
sqrt(mean((Y_miss - Y_hat_1)^2))
Y_hat_2 = diag(para_list$`biScale:row`$scale) %*%  (g1$l %*% t(g1$f)) %*% diag(para_list$`biScale:column`$scale) +  (para_list$`biScale:row`$center) %*% t(rep(1,251)) + rep(1,226) %*% t(para_list$`biScale:column`$center)
sqrt(mean((Y_miss - Y_hat_2[row_index,col_index])^2))
```

```
#### for RCC
library(R.matlab)
library(PMA)
datapath = "/home/weidong/HG/flash/data/missingvalue/Breastcancer/nsfaexample.mat"
Y_centered = readMat(datapath)
Y = Y_centered$Ycentered
source("~/HG/flash/Rcode/flash.R")
source("~/HG/flash/Rcode/greedy.R")
source("~/HG/flash/Rcode/backfitting.R")
library(softImpute)
## to get the biscale data
Y_bs = biScale(Y)
para_list = attributes(Y_bs)

set.seed(99)
results = matrix(0,ncol = 2, nrow = 30)
for(i in 1:30){
  print(i)
  row_index = sample(dim(Y)[1],20)
  col_index = sample(dim(Y)[2],20)
  Y_miss = Y[row_index,col_index]

  Y_input = Y
  Y_input[row_index,col_index] = NA
  Y_bs_input = Y_bs
  Y_bs_input[row_index,col_index] = NA

  g1 = greedy(Y_input, K  = 60)
  g2 = greedy(Y_bs_input, K  = 60)

  Y_hat_1 = (g1$l %*% t(g1$f))[row_index,col_index]
  results[i,1] =  sqrt(mean((Y_miss - Y_hat_1)^2))
  Y_hat_2 = diag(para_list$`biScale:row`$scale) %*%  (g1$l %*% t(g1$f)) %*% diag(para_list$`biScale:column`$scale) +  (para_list$`biScale:row`$center) %*% t(rep(1,251)) + rep(1,226) %*% t(para_list$`biScale:column`$center)
  results[i,2] = sqrt(mean((Y_miss - Y_hat_2[row_index,col_index])^2))
}
saveRDS(results, "./results.rds")
```

```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```

### 2017-06-12 analysis/ObjGap.rmd
Files commited by wflow_commit.

--- 2017-05-22 analysis/imputestratergy.Rmd
@@
 ---
-title: "impute stragergy"
+title: "'gap' issue"
 author: "Wei Wang"
-date: 2017-05-22
+date: 
 output: html_document
@@
 
-## try the breast cancer data
+## introduction of the problem.
 
-```{r}
-library(R.matlab)
-library(PMA)
-datapath = "~/HG/ash-sfa/Rcode/postmean/flash_simulation/missingvalue/nsfaexample.mat"
-Y_centered = readMat(datapath)
-Y = Y_centered$Ycentered
-library(flashr)
-library(softImpute)
-## to get the biscale data
-Y_bs = biScale(Y)
+In the old code of flash. we found that the greedy algorithm
 
-para_list = attributes(Y_bs)
+### old version documnt
 
-set.seed(999)
-# set the missing position
+The old version document is in `ash-sfa` git repo [ash-sfa repo](https://github.com/stephenslab/ash-sfa).
 
-row_index = sample(dim(Y)[1],20)
-col_index = sample(dim(Y)[2],20)
-Y_miss = Y[row_index,col_index]
-
-Y_input = Y
-Y_input[row_index,col_index] = NA
-Y_bs_input = Y_bs
-Y_bs_input[row_index,col_index] = NA
-
-#g1 = flashr::greedy(Y_input, K  = 60)
-#g2 = flashr::greedy(Y_bs_input, K  = 60)
-#saveRDS(g1,"../data/missingvalue/imputescheme/g1_original.rds")
-#saveRDS(g2,"../data/missingvalue/imputescheme/g2_biscale.rds")
-g1 = readRDS("../data/missingvalue/imputescheme/g1_original.rds")
-g2 = readRDS("../data/missingvalue/imputescheme/g2_biscale.rds")
-Y_hat_1 = (g1$l %*% t(g1$f))[row_index,col_index]
-sqrt(mean((Y_miss - Y_hat_1)^2))
-Y_hat_2 = diag(para_list$`biScale:row`$scale) %*%  (g1$l %*% t(g1$f)) %*% diag(para_list$`biScale:column`$scale) +  (para_list$`biScale:row`$center) %*% t(rep(1,251)) + rep(1,226) %*% t(para_list$`biScale:column`$center)
-sqrt(mean((Y_miss - Y_hat_2[row_index,col_index])^2))
-```
-
-```
-#### for RCC
-library(R.matlab)
-library(PMA)
-datapath = "/home/weidong/HG/flash/data/missingvalue/Breastcancer/nsfaexample.mat"
-Y_centered = readMat(datapath)
-Y = Y_centered$Ycentered
-source("~/HG/flash/Rcode/flash.R")
-source("~/HG/flash/Rcode/greedy.R")
-source("~/HG/flash/Rcode/backfitting.R")
-library(softImpute)
-## to get the biscale data
-Y_bs = biScale(Y)
-para_list = attributes(Y_bs)
-
-set.seed(99)
-results = matrix(0,ncol = 2, nrow = 30)
-for(i in 1:30){
-  print(i)
-  row_index = sample(dim(Y)[1],20)
-  col_index = sample(dim(Y)[2],20)
-  Y_miss = Y[row_index,col_index]
-
-  Y_input = Y
-  Y_input[row_index,col_index] = NA
-  Y_bs_input = Y_bs
-  Y_bs_input[row_index,col_index] = NA
-
-  g1 = greedy(Y_input, K  = 60)
-  g2 = greedy(Y_bs_input, K  = 60)
-
-  Y_hat_1 = (g1$l %*% t(g1$f))[row_index,col_index]
-  results[i,1] =  sqrt(mean((Y_miss - Y_hat_1)^2))
-  Y_hat_2 = diag(para_list$`biScale:row`$scale) %*%  (g1$l %*% t(g1$f)) %*% diag(para_list$`biScale:column`$scale) +  (para_list$`biScale:row`$center) %*% t(rep(1,251)) + rep(1,226) %*% t(para_list$`biScale:column`$center)
-  results[i,2] = sqrt(mean((Y_miss - Y_hat_2[row_index,col_index])^2))
-}
-saveRDS(results, "./results.rds")
-```
-
-```
+In the document, we have several observations:
 

### 2017-06-12 analysis/ObjGap.rmd
Files commited by wflow_commit.

@@
 
-In the old code of flash. we found that the greedy algorithm
+In the old code of flash. we found that the greedy algorithm doesn't increase the lower bound of the log-likelihood. So we turn to understand the likelihood of rank-one and rank-zero model.
+
+Here I will rerun all the representative code, and just provide the original document describing this issue.
 
@@
 
+1.
+In the rank one FLASH, we can see that our algorithm monotonically increase that lower-bound of the log-likelihood.
+[show case in old code](https://github.com/stephenslab/ash-sfa/blob/master/Rcode/postmean/flash/simulation/convergereport.pdf)
+2.
+In the greedy algorithm, the objective function is not increase in each time adding one more factor.
+We thought the penalty term in `ash` might cause this problem, and we find one case that the penalty term does have effect in rank one model.
+After we change `method = "fdr"` to `method = "shrink"` in `ash`, the objective function monotonically increase in the algorithm.
+But when we change the `mehtod ="shrink"` in `greedy` and `backfitting` function, the log-likelihood still doesn't increase in each time by adding a new factor.
+[show case in old code](https://github.com/stephenslab/ash-sfa/blob/master/Rcode/postmean/flash/simulation/testKfactorVEM.pdf).
+3.
+Then we know that change `method = shrink` can guarantee that rank one and backfitting(if not add or remove factor) algorithm monotonic increase.
+The only problem is adding or removing factor, which cause the log-likelihood not monotonic.
+Then we turn to the simplest case of adding or removing factor: rank-one vs rank-zero model.
+In this document, one thing we can't understand well is that the prediction result seems not corresponding to the objective value.
+[show case in old code](https://github.com/stephenslab/ash-sfa/blob/master/Rcode/postmean/flash/simulation/objcheck.pdf)
+
+We will rerun the analysis in this document. The above observations just show the old documents.
+
+The following is the main part of this document.
+
+## examples
+
+We will start from the problem in greedy algorithm:
+
+```{r}
+sim_K = function(K, N, P, SF, SL, signal,noise){
+  E = matrix(rnorm(N*P,0,noise),nrow=N)
+  Y = E
+  L_true = array(0, dim = c(N,K))
+  F_true = array(0, dim = c(P,K))
+
+  for(k in 1:K){
+    lstart = rnorm(N, 0, signal)
+    fstart = rnorm(P, 0, signal)
+
+    index = sample(seq(1:N),(N*SL))
+    lstart[index] = 0
+    index = sample(seq(1:P),(P*SF))
+    fstart[index] = 0
+
+    L_true[,k] = lstart
+    F_true[,k] = fstart
+
+    Y = Y + lstart %*% t(fstart)
+  }
+  return(list(Y = Y, L_true = L_true, F_true = F_true, Error = E))
+}
+```
+
 ## Session information

### 2017-06-13 analysis/ObjGap.rmd
Files commited by wflow_commit.

@@
 
-## introduction of the problem.
+## 0 introduction of the problem.
 
@@
 
-### old version documnt
+### 0.1 old version documnt
 
@@
 
-In the document, we have several observations:
+In the document, we have several observations (**please ignore this section since it is just a record for the old code, I will show all the cases in the next section**):
 
@@
 
-## examples
+## 1 summary of the result
+
+- section 2.1 shows that in the greedy algorithm doesn't always increase the lower bound of the log-likelihood when we add one more factor.
+When we add one more factor the marginal (conditional) likelihood, $E_q \log\{P(Y|L,F)\} = -\frac{1}{2\sigma^2}(Y_{ij} - \sum_k l_{ik}f_{kj})^2$, would increase, but the $\log(\frac{P(L)}{q(L)}) = \sum_k \log(\frac{P(L_k)}{q(L_k)})$ would decrease since:
+\begin{eqnarray} E_q log(\frac{P}{q}) \leq log E_q (\frac{P}{q}) = 0 \end{eqnarray}
+
+> I think the KL(q || p) is not continuous as q,p -> delta_0 [point mass at 0]
+> eg q_n = N(0,1/n) and p_n = N(0,2/n)
+> I think lim_{n \rightarrow \inf} KL(q_n, p_n) may not be 0?
+> If that is right then the lower bound F(g; qhat(g)) is not necessarily continuous in g (even though likelihood is continuous in g I think)…
+> maybe this explains the gap; and may also tell us what to do about it.
+> —— Matthew's comments
+
+$$lim_{n \rightarrow \inf} KL(q_n, p_n) = -\frac{1}{2}\{\log2 - \frac{1}{2}\} \neq 0$$
+
+- section 2.2 shows that rank one FLASH with 4 different method: `method = "fdr"`, `method = "shrink"`, `method = "spike and slab"`, `method = "uniform with point mass"`.
+We can see that the lower bound of the log likelihood is increase if we apply different methods without penalty term.
+The objective value is not always increase in `method = "fdr"` case.
+
+- In section 2.2, there is an interesting observation that the RMSE results don't always agree wit the value of the objective function, which means that the method with larger objective function value doesn't has smaller RMSE value (better in prediction).
+
+- In section 2.3 we use `method = "shrink"` to the same data in section 2.1.2 where the loglikelihood is not monotonically increase, and we find that the loglikelihood still decrease when we add fifth factor.
+So this is probably because of the penalty term.
+
+- Section 2.4 shows that the `method = "fdr"` performs very well in rank zero case.
+
+- In section 2.3 and 2.12, we can see that rank 4 model have larger value of lower bound but rank 5 model has better performance in RMSE.
+
+## 2 examples
+
+### 2.1 greedy algorithm
 
@@
 ```{r}
+library(flashr)
 sim_K = function(K, N, P, SF, SL, signal,noise){
@@
 
+#### 2.1.1
+
+"fdr" method looks fine in many cases.
+
+```{r}
+Greedy_showcase = function(K=5,N = 100, P = 200, SF = 0.8, SL = 0.8, signal = 1,noise = 1){
+  set.seed(99)
+  data = sim_K(K,N, P, SF , SL , signal ,noise )
+  Y = data$Y
+  E = data$Error
+  g1 = greedy(Y,K = 10)
+  plot(g1$clik_vec + cumsum(g1$priorpost_vec),main = "fdr method")
+  # show the prediction
+  rmse_5 = sqrt(mean((Y - g1$l %*% t(g1$f) -E)^2)) / sqrt(mean((Y - E)^2))
+  rmse_4 = sqrt(mean((Y - g1$l[,1:4] %*% t(g1$f[,1:4]) -E)^2)) / sqrt(mean((Y - E)^2))
+  return(list(obj_track = g1$clik_vec + cumsum(g1$priorpost_vec),RMSE = c(rmse_4,rmse_5)))
+}
+Greedy_showcase(K=5,N = 100, P = 200, SF = 0.8, SL = 0.8, signal = 1,noise = 1)
+```
+
+#### 2.1.2
+
+But in some case, the lower bound is decrease when we add one more factor:
+
+```{r}
+Greedy_showcase(K=5,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 1,noise = 3)
+```
+
+We guess that it is because the penalty term in `ash` we use.
+
+### 2.2 rank one model with different method
+
+Here we use four different methods: fdr, shrink, uniform without penalty and spike and slab.
+
+```{r}
+rank1test = function(K,N, P, SF , SL , signal ,noise ){
+  set.seed(99)
+  par(mfrow = c(3, 2))
+  par(cex = 0.6)
+  par(mar = c(3, 3, 0.8, 0.8), oma = c(1, 1, 1, 1))
+  data = sim_K(K,N, P, SF , SL , signal ,noise )
+  Y = data$Y
+  E = data$Error
+  # "l" is for lower bound
+  g3_1 = flash(Y,objtype = "l")
+  plot(g3_1$obj_val_track,main = "fdr method",type = "l")
+  g3_2 = flash(Y,objtype = "l",ash_para = list( prior = "uniform",pointmass = TRUE,method = NULL))
+  plot(g3_2$obj_val_track,main = "uniform + pointmass method",type = "l")
+  g3_3 = flash(Y,objtype = "l",ash_para = list(method = "shrink"))
+  plot(g3_3$obj_val_track,main = "shrink method",type = "l")
+  g3_4 = flash(Y,objtype = "l",ash_para = list( prior = "uniform",pointmass = TRUE,method = NULL,mixsd = c(1)))
+  plot(g3_4$obj_val_track,main = "spike and slab method",type = "l")
+  obj = c(g3_1$obj_val,g3_2$obj_val,g3_3$obj_val,g3_4$obj_val)
+  RMSE = c(sqrt(mean((Y - g3_1$l %*% t(g3_1$f) -E)^2)) / sqrt(mean((Y - E)^2)),
+           sqrt(mean((Y - g3_2$l %*% t(g3_2$f) -E)^2)) / sqrt(mean((Y - E)^2)),
+           sqrt(mean((Y - g3_3$l %*% t(g3_3$f) -E)^2)) / sqrt(mean((Y - E)^2)),
+           sqrt(mean((Y - g3_4$l %*% t(g3_4$f) -E)^2)) / sqrt(mean((Y - E)^2)))
+  plot(RMSE, col =c("magenta","cyan","yellow","green"), pch = 16, cex = 2,main = "RMSE")
+  legend("top", cex = 1,
+       legend = c("fdr","Uniform","shrink","SpikeSlab"),
+       col = c("magenta","cyan","yellow","green"),
+       pch=16)
+  plot(obj, col =c("magenta","cyan","yellow","green"), pch = 16, cex = 2,main = "objective value")
+  legend("top",  cex = 1,
+       legend = c("fdr","Uniform","shrink","SpikeSlab"),
+       col = c("magenta","cyan","yellow","green"),
+       pch=16)
+}
+
+rank1test(1,100,200,0.5,0.5,1,1)
+```
+
+we can see that the RMSE doesn't agree with the value of objective function.
+
+```{r}
+rank1test(1,100,200,0.8,0.8,1,2)
+```
+
+We can see that the objective function is not increasing in `method = "fdr"`
+
+```{r}
+rank1test(1,20,30,0.5,0.6,1,1)
+```
+
+#### very sparse case
+
+```{r}
+rank1test(1,100,300,0.9,0.9,1,1)
+```
+
+```{r}
+rank1test(1,100,300,0.95,0.99,1,0.5)
+```
+
+### 2.3 apply greedy without penalty
+
+```{r}
+Greedy_showcase_1 = function(K=5,N = 100, P = 200, SF = 0.8, SL = 0.8, signal = 1,noise = 1,flash_para = list(ash_para = list(method = "shrink"))){
+  set.seed(99)
+  data = sim_K(K,N, P, SF , SL , signal ,noise )
+  Y = data$Y
+  E = data$Error
+  g1 = greedy(Y,K = 10,flash_para)
+  plot(g1$clik_vec + cumsum(g1$priorpost_vec),main = "shrink method")
+  # show the prediction
+  rmse_5 = sqrt(mean((Y - g1$l %*% t(g1$f) -E)^2)) / sqrt(mean((Y - E)^2))
+  rmse_4 = sqrt(mean((Y - g1$l[,1:4] %*% t(g1$f[,1:4]) -E)^2)) / sqrt(mean((Y - E)^2))
+  return(list(obj_track = g1$clik_vec + cumsum(g1$priorpost_vec),RMSE = c(rmse_4,rmse_5)))
+}
+Greedy_showcase_1(K=5,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 1,noise = 3,flash_para = list(ash_para = list(method = "shrink")))
+```
+
+by using the `method = "shrink"`, we can see that the objective function getting large but the lower bound is still not monotonically increasing. So it's not because of the penalty we use in `ash`.
+
+### 2.4 rank zero case
+
+```{r}
+rank1test_0 = function(K,N, P, SF , SL , signal ,noise ){
+  set.seed(99)
+  par(mfrow = c(3, 2))
+  par(cex = 0.6)
+  par(mar = c(3, 3, 0.8, 0.8), oma = c(1, 1, 1, 1))
+  data = sim_K(K,N, P, SF , SL , signal ,noise )
+  Y = data$Y
+  E = data$Error
+  # "l" is for lower bound
+  g3_1 = flash(Y,objtype = "l")
+  plot(g3_1$obj_val_track,main = "fdr method",type = "l")
+  g3_2 = flash(Y,objtype = "l",ash_para = list( prior = "uniform",pointmass = TRUE,method = NULL))
+  plot(g3_2$obj_val_track,main = "uniform + pointmass method",type = "l")
+  g3_3 = flash(Y,objtype = "l",ash_para = list(method = "shrink"))
+  plot(g3_3$obj_val_track,main = "shrink method",type = "l")
+  g3_4 = flash(Y,objtype = "l",ash_para = list( prior = "uniform",pointmass = TRUE,method = NULL,mixsd = c(1)))
+  plot(g3_4$obj_val_track,main = "spike and slab method",type = "l")
+  obj = c(g3_1$obj_val,g3_2$obj_val,g3_3$obj_val,g3_4$obj_val)
+  RMSE = c(sqrt(mean((Y - g3_1$l %*% t(g3_1$f) -E)^2)) ,
+           sqrt(mean((Y - g3_2$l %*% t(g3_2$f) -E)^2)) ,
+           sqrt(mean((Y - g3_3$l %*% t(g3_3$f) -E)^2)) ,
+           sqrt(mean((Y - g3_4$l %*% t(g3_4$f) -E)^2)) )
+  plot(RMSE, col =c("magenta","cyan","yellow","green"), pch = 16, cex = 2,main = "RMSE")
+  legend("top", cex = 1,
+       legend = c("fdr","Uniform","shrink","SpikeSlab"),
+       col = c("magenta","cyan","yellow","green"),
+       pch=16)
+  plot(obj, col =c("magenta","cyan","yellow","green"), pch = 16, cex = 2,main = "objective value")
+  legend("top",  cex = 1,
+       legend = c("fdr","Uniform","shrink","SpikeSlab"),
+       col = c("magenta","cyan","yellow","green"),
+       pch=16)
+}
+
+rank1test_0(1,20,60,1,0.6,1,1)
+```
+
+```{r}
+rank1test_0(1,100,200,1,0.6,1,2)
+```
+
+```{r}
+rank1test_0(1,60,100,1,0.6,1,0.5)
+```
+
 ## Session information

### 2017-06-13 analysis/ObjGap.rmd
Files commited by wflow_commit.

@@
 author: "Wei Wang"
-date: 
+date: 2017-06-12
 output: html_document
@@
 
+#### special case
+
+```{r}
+rank1test(1,10,200,0.5,0.6,1,1)
+```
+
+this case
+
+```{r}
+rank1test(1,10,200,0.5,0.5,1,3)
+```
+
+this case show that the `shink` and `uniform + point mass` are not same.
+
+```{r}
+rank1test(1,10,200,0.5,0.5,1,2.5)
+```
+
+In this case, `fdr` method tend to get rank zero estimation, which has highest value of objective function, but rank one model with `spike and slab` has better performance in prediction.
+
 #### very sparse case
@@
 ```
-
+test
 ## Session information

### 2017-06-19 analysis/Objcheck_1.Rmd
Files commited by wflow_commit.

---
title: "objective function checking"
author: "Wei Wang"
date: YYYY-MM-DD
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

## rank 1 vs rank 0

*we focus on the "shrink" method now*

As Matthew suggested, we focus on the rank 0 vs rank 1 case.

```{r}
library(flashr)
sim_K = function(K, N, P, SF, SL, signal,noise){
  E = matrix(rnorm(N*P,0,noise),nrow=N)
  Y = E
  L_true = array(0, dim = c(N,K))
  F_true = array(0, dim = c(P,K))

  for(k in 1:K){
    lstart = rnorm(N, 0, signal)
    fstart = rnorm(P, 0, signal)

    index = sample(seq(1:N),(N*SL))
    lstart[index] = 0
    index = sample(seq(1:P),(P*SF))
    fstart[index] = 0

    L_true[,k] = lstart
    F_true[,k] = fstart

    Y = Y + lstart %*% t(fstart)
  }
  return(list(Y = Y, L_true = L_true, F_true = F_true, Error = E))
}
```

### simulation: rank 0

In this section, we will use the flash since it is a rank one model.

There are code from `flashr` package to get the objective funtion, and we will use them to calculate the objective function for rank one case.

```{r,echo=FALSE}
#' title prior and posterior part in objective function
#'
#' description prior and posterior part in objective function
#'
#' @return PrioPost the value for the proir and posterior in objectice function
#' @param mat matrix of flash.data in ash output which is for posterior
#' @param fit_g in fitted.g in ash output which is for prior
#' @keywords internal
#'
# two parts for the likelihood
Fval = function(mat,fit_g){
  # make sure the first row included in the index
  nonzeroindex = unique(c(1,which(fit_g$pi != 0)))
  #prepare for the posterior part
  mat_postmean = mat$comp_postmean[nonzeroindex,]
  mat_postmean2 = mat$comp_postmean2[nonzeroindex,]
  mat_postvar = mat_postmean2 - mat_postmean^2
  mat_postprob = mat$comp_postprob[nonzeroindex,]
  # figure our the dimension
  K = dim(mat_postprob)[1]
  N = dim(mat_postprob)[2]
  if(is.vector(mat_postprob)){
    K = 1
    N = length(mat_postprob)
  }
  # prepare for the prior
  prior_pi = fit_g$pi[nonzeroindex]
  prior_var = (fit_g$sd[nonzeroindex])^2
  mat_priorvar = matrix(rep(prior_var,N),ncol = N)
  mat_priorprob = matrix(rep(prior_pi,N),ncol = N)
  # to get the value
  varodds = mat_priorvar / mat_postvar
  varodds[1,] = 1 # deal with 0/0
  probodds = mat_priorprob / mat_postprob
  ssrodds = mat_postmean2 / mat_priorvar
  ssrodds[1,] = 1  # deal with 0/0
  priorpost = mat_postprob * (log(probodds) - (1/2)*log(varodds) - (1/2)* (ssrodds -1))
  # in case of mat_postprob = 0
  priorpost[which(mat_postprob< 1e-100)] = 0
  PrioPost = sum(priorpost)
  return(list(PrioPost = PrioPost))
}

#' title conditional likelihood
#'
#' description conditional likelihood in objective function
#'
#' @return c_lik for the onditional likelihood in objectice function
#' @param N  dimension of residual matrix
#' @param P  dimension of residual matrix
#' @param sigmae2_v  residual matrix
#' @param sigmae2  variance structure of error
#' @keywords internal
#'
# this version we need to know the truth of sigmae2, we can use sigmae2_v as the truth
C_likelihood = function(N,P,sigmae2_v,sigmae2){
  if(is.matrix(sigmae2)){
    # print("clik as matrix of sigmae2")
    c_lik = -(1/2) * sum( log(2*pi*sigmae2) + (sigmae2_v)/(sigmae2) )
  }else if(is.vector(sigmae2) & length(sigmae2)==P){
    # print("clik as vector sigmae2")
    # change the format to fit the conditional likelihood
    sigmae2_v = colMeans(sigmae2_v)
    c_lik = -(N/2) * sum( log(2*pi*sigmae2) + (sigmae2_v)/(sigmae2) )
  } else {
    # print("clik as sigmae2 is constant")
    # change the format to fit the conditional likelihood and accelerate the computation.
    sigmae2_v = mean(sigmae2_v)
    c_lik = -(N*P)/2 * ( log(2*pi*sigmae2) + (sigmae2_v)/(sigmae2) )
    # here I want to use fully variantional inference Elogsigmae2 rather logEsigmae2
    # c_lik = -(N*P)/2 * ( log(2*pi*sigmae2_true) + (sigmae2_v)/(sigmae2_true) + log(N*P/2) - digamma(N*P/2) )
  }
  return(list(c_lik = c_lik))
}

#' title objective function in VEM
#'
#' description  objective function
#'
#' @return obj_val value of the objectice function
#' @param N  dimension of residual matrix
#' @param P  dimension of residual matrix
#' @param sigmae2_v  residual matrix
#' @param sigmae2_true  true variance structure (we use the estimated one to replace that if the truth is unknown)
#' @param par_l ash output for l
#' @param par_f ash output for f
#' @param objtype  objective function type,
#' "margin_lik" for conditional likelihood,
#' "lowerbound_lik" for full objective function
#' @keywords internal
#'

# objective function
obj = function(N,P,sigmae2_v,sigmae2,par_f,par_l,objtype = "margin_lik"){
  if(is.list(sigmae2)){
    # print("obj using kronecker product")
    sigmae2 = sigmae2$sig2_l %*% t(sigmae2$sig2_f)
  }
  if(objtype=="lowerbound_lik"){
    priopost_f = Fval(par_f$mat, par_f$g)$PrioPost
    priopost_l = Fval(par_l$mat, par_l$g)$PrioPost
    c_lik = C_likelihood(N,P,sigmae2_v,sigmae2)$c_lik
    obj_val = c_lik + priopost_l + priopost_f
  } else {
    obj_val = C_likelihood(N,P,sigmae2_v,sigmae2)$c_lik
  }
  return(obj_val)
}

```

```{r}

rank0check = function(K=1,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 0,noise = 1, mtype = "shrink"){
  data = sim_K(K,N, P, SF , SL , signal ,noise )
  Y = data$Y
  E = data$Error
  if(mtype == "fdr"){
    g1 = flash(Y,objtype = "l")
  }else{
    g1 = flash(Y,objtype = "l",ash_para = list(method = "shrink"))
  }

  RMSE = sqrt(mean((Y - g1$l %*% t(g1$f) -E)^2)) 
  obj_val = g1$obj_val
  # to get the rank zero objective value
  sigmae2_v0 = Y^2
  fit_g0 = list(pi=c(1,0,0),sd = c(0,0.1,1))
  mat0 = list(comp_postmean = matrix(0,ncol= N,nrow = 3),
              comp_postmean2 = matrix(0,ncol= N,nrow = 3),
              comp_postprob = matrix(0,ncol= N,nrow = 3))
  mat0$comp_postprob[1,] = 1
  par_l0 = list(g = fit_g0, mat = mat0)
  fit_g0 = list(pi=c(1,0,0),sd = c(0,0.1,1))
  mat0 = list(comp_postmean = matrix(0,ncol= P,nrow = 3),
              comp_postmean2 = matrix(0,ncol= P,nrow = 3),
              comp_postprob = matrix(0,ncol= P,nrow = 3))
  mat0$comp_postprob[1,] = 1
  par_f0 = list(g = fit_g0, mat = mat0)
  obj0 = obj(N,P,sigmae2_v0,sigmae2 = mean(sigmae2_v0),par_f0,par_l0,objtype = "lowerbound_lik")

  return(list(rmse = RMSE,obj_val = obj_val,obj_0 = obj0))
}

sim_rank0 = function(K=1,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 0,noise = 1,mtype = "shrink"){
  T = 200
  RMSE = rep(NA,T)
  sign_obj = rep(NA,T)
  for(t in 1:T){
    g0 = rank0check(K = 1,N, P, SF, SL , signal=0,noise,mtype = mtype)
    RMSE[t] = g0$rmse
    sign_obj[t] = sign(g0$obj_val - g0$obj_0)
  }
  par(mfrow = c(2, 2))
  par(cex = 0.6)
  par(mar = c(3, 3, 0.8, 0.8), oma = c(1, 1, 1, 1))
  boxplot(RMSE,main = "RMSE")
  hist(sign_obj, main = "sign of difference of obj value")
  hist(RMSE,breaks = 50,main = "RMSE")
  plot(sign_obj,RMSE, main = "sign of obj diff vs RMSE")
}

```

we set the 200 simulations for each case.

#### case 1:

```{r}
set.seed(99)
sim_rank0(K=1,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 0,noise = 1)
```

- top-left: boxplot of RMSE
- top-right: histogram of $\text{sign} \left\{obj(model| method = ``shrink") - obj(model|rank= 0) \right\}$
- bottom-left: histogram of RMSE
- bottom-right: scatter plot of sign-difference-objective-value vs RMSE.

we can see that in almost half of the cases, we get rank zero estimation, but there are another half getting smaller objective function and larger RMSE value. There are two extreme cases out of 200, we get larger objective value but larger RMSE as well.

In this case, we know that rank zero is the truth, so the $RMSE = 0$ is the best prediction.

This is just a simple example, and we can try more situations:

```{r}
set.seed(99)
sim_rank0(K=1,N = 20, P=200, SF = 0.5, SL = 0.5, signal = 0,noise = 1)
```

```{r}
set.seed(99)
sim_rank0(K=1,N = 20, P=30, SF = 0.5, SL = 0.5, signal = 0,noise = 1)
```

#### if the method = "fdr"

the setting is similar as case 1

```{r}
set.seed(99)
sim_rank0(K=1,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 0,noise = 1,mtype = "fdr")
```

we can see that we would get rank zero estimation in most cases.

### rank 1

we simulate rank 1 and compare it with the rank zero solution (the setting is similar with case 1):

```{r}
set.seed(99)
sim_rank0(K=1,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 1,noise = 1)
```

use the "fdr" method

```{r}
set.seed(99)
sim_rank0(K=1,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 1,noise = 1, mtype = "fdr")
```

#### very sparse

the default is "shrink" method

```{r}
set.seed(99)
sim_rank0(K=1,N = 100, P=200, SF = 0.9, SL = 0.8, signal = 1,noise = 1 )
```

#### very sparse with big noise

```{r}
set.seed(99)
sim_rank0(K=1,N = 100, P=200, SF = 0.8, SL = 0.8, signal = 1,noise = 2)
```

```{r}
set.seed(99)
sim_rank0(K=1,N = 100, P=200, SF = 0.8, SL = 0.8, signal = 1,noise = 2,mtype = "fdr")
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```

### 2017-06-19 analysis/Objcheck_1.Rmd
Files commited by wflow_commit.

@@
   for(t in 1:T){
-    g0 = rank0check(K = 1,N, P, SF, SL , signal=0,noise,mtype = mtype)
+    g0 = rank0check(K = 1,N, P, SF, SL , signal,noise,mtype = mtype)
     RMSE[t] = g0$rmse
@@
   boxplot(RMSE,main = "RMSE")
-  hist(sign_obj, main = "sign of difference of obj value")
+  hist(sign_obj, breaks = 3,main = "sign of difference of obj value")
   hist(RMSE,breaks = 50,main = "RMSE")

### 2017-06-20 analysis/Objcheck_2.Rmd
Files commited by wflow_commit.

---
title: "objective function with the penalty term"
author: "wei"
date: YYYY-MM-DD
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

In this case, we would like to try the objective function with penalty term.

I would like to modify the flash code in this document.

```{r,echo=FALSE}
# ATM function
#' title ash type model for l
#'
#' description use ash type model to maxmization
#'
#' @return list of factor, loading and variance of noise matrix
#'  \itemize{
#'   \item{\code{El}} {is a N vector for mean of loadings}
#'   \item{\code{El2}} {is a N vector for second moment of loadings}
#'  }
#' @param Y is data matrix
#' @param Ef is mean for the factor
#' @param Ef2 is second moment for the factor
#' @param sigmae2 is variance structure for noise matrix
#' @param col_var is just for column variance, column is for column, row is for row
#' @param nonnegative is flag for whether output nonnegative value
#' @param output is format of output, "mean" is for mean value, "matrix" is for flash data matrix in ash
#'
#' @keywords internal
#'

ATM_r1 = function(Y, Ef, Ef2,
                  sigmae2, col_var = "row",
                  nonnegative = FALSE, output = "mean",
                  partype = "constant",ash_para = list()){
  # this part is preparing betahat and sebeta for ash
  if(is.matrix(sigmae2)){
    # print("sigmae2 is a matrix")
    # this is for all variance are different
    sum_Ef2 = (1/sigmae2) %*% Ef2
    sum_Ef2 = as.vector(sum_Ef2)
    sebeta = sqrt(1/(sum_Ef2))
    betahat = as.vector( (Y/sigmae2) %*% Ef ) / (sum_Ef2)
    betahat=as.vector(betahat)
  } else if(is.vector(sigmae2) & length(sigmae2) == length(Ef)){
    # this is for the non-constant variance in column
    # print("sigmae2 is vector")
    if(col_var == "row"){
      sum_Ef2 = (1/sigmae2) * Ef2
      sum_Ef2 = sum(sum_Ef2)
      sebeta = sqrt(1/(sum_Ef2))
      betahat = as.vector( Y %*% (Ef/sigmae2) ) / (sum_Ef2)
      betahat=as.vector(betahat)
    }else {
      # for column
      # here I have alread change the dimension by taking transpose to Y
      sum_Ef2 = sum(Ef2)
      sebeta = sqrt( sigmae2/(sum_Ef2) )
      betahat = (t(Ef) %*% t(Y)) / (sum_Ef2)
      betahat=as.vector(betahat)
    }
  }else{
    # for constant case
    # print("sigmae2 is a constant")
    sum_Ef2 = sum(Ef2)
    sebeta = sqrt(sigmae2/(sum_Ef2))
    betahat = (t(Ef) %*% t(Y)) / (sum_Ef2)
    betahat=as.vector(betahat)
  }
  # set ash default setting up
  ash_default = list(betahat = betahat, sebeta = sebeta,
                     method = "fdr", mixcompdist = "normal")
  # ATM update
  # decide the sign for output
  if(nonnegative){
    ash_default$mixcompdist = "+uniform"
  }
  # decide the output to decide the convergence criterion
  if(output == "matrix"){
    ash_para$outputlevel = 4
    ash_para = modifyList(ash_default,ash_para)
    ATM = do.call(ashr::ash,ash_para)
    # ATM = ash(betahat, sebeta, method="fdr", mixcompdist=mixdist,outputlevel=4)
    Ef = ATM$result$PosteriorMean
    SDf = ATM$result$PosteriorSD
    Ef2 = SDf^2 + Ef^2
    mat_post = ATM$flash_data
    fit_g = ATM$fitted_g
    return(list(Ef = Ef,
                Ef2 = Ef2,
                mat = mat_post,
                g = fit_g))
  } else {
    ash_para = modifyList(ash_default,ash_para)
    ATM = do.call(ashr::ash,ash_para)
    # ATM = ash(betahat, sebeta, method="fdr", mixcompdist=mixdist)
    Ef = ATM$result$PosteriorMean
    SDf = ATM$result$PosteriorSD
    Ef2 = SDf^2 + Ef^2
    return(list(Ef = Ef, Ef2 = Ef2))
  }

}

#' title prior and posterior part in objective function
#'
#' description prior and posterior part in objective function
#'
#' @return PrioPost the value for the proir and posterior in objectice function
#'  \itemize{
#'   \item{\code{PrioPost}} {the value for the proir and posterior in objectice function}
#'   \item{\code{penalty}} {penalty term of the g_l or g_f}
#'  }
#' @param mat matrix of flash.data in ash output which is for posterior
#' @param fit_g in fitted.g in ash output which is for prior
#' @keywords internal
#'
# two parts for the likelihood
Fval = function(mat,fit_g){
  # make sure the first row included in the index
  nonzeroindex = unique(c(1,which(fit_g$pi != 0)))
  #prepare for the posterior part
  mat_postmean = mat$comp_postmean[nonzeroindex,]
  mat_postmean2 = mat$comp_postmean2[nonzeroindex,]
  mat_postvar = mat_postmean2 - mat_postmean^2
  mat_postprob = mat$comp_postprob[nonzeroindex,]
  # figure our the dimension
  K = dim(mat_postprob)[1]
  N = dim(mat_postprob)[2]
  if(is.vector(mat_postprob)){
    K = 1
    N = length(mat_postprob)
  }
  # prepare for the prior
  prior_pi = fit_g$pi[nonzeroindex]
  prior_var = (fit_g$sd[nonzeroindex])^2
  mat_priorvar = matrix(rep(prior_var,N),ncol = N)
  mat_priorprob = matrix(rep(prior_pi,N),ncol = N)
  # to get the value
  varodds = mat_priorvar / mat_postvar
  varodds[1,] = 1 # deal with 0/0
  probodds = mat_priorprob / mat_postprob
  ssrodds = mat_postmean2 / mat_priorvar
  ssrodds[1,] = 1  # deal with 0/0
  priorpost = mat_postprob * (log(probodds) - (1/2)*log(varodds) - (1/2)* (ssrodds -1))
  # in case of mat_postprob = 0
  priorpost[which(mat_postprob< 1e-100)] = 0
  PrioPost = sum(priorpost)
  #######
  # now we try to get the penalty term
  # I use the default of setting for lambda_k in https://oup.silverchair-cdn.com/oup/backfile/Content_public/Journal/biostatistics/18/2/10.1093_biostatistics_kxw041/1/kxw041_Supp.pdf?Expires=1498001808&Signature=dtxDDFmYZ1u1jFoXfNg4mqGQnTPkQQMFZgt2-FE9r8GcgRl4DdEaP0oI6kzvUjZB5H3hLZmvAZfuFkt3GxojQJ1eX0y7I9kMJC75VTrxw1Ym~psIe8rdU5Xw4S4KZm79o1cOXBzY1iAORFkVM4z59uP1T1ltFJDu5fhlTYErU3flbiE1ivTm-BYN5P7w8dex-R5MWn98NpMVRUsoHRgbe7FhwavdnQ8QAD7~q1F4DzMyZk09TeNdEc7GNdBKbSOoS9mGGnvrqXvPz1nn~QzJ2U2CMCf04GfsgkclyUnH0aD6Vh1lSAfoEV47ZnT-9~942~f8rMs30H2Kqji7BjlctA__&Key-Pair-Id=APKAIUCZBIA4LVPAVW3Q
  h_value = (10 - 1) * log(fit_g$pi[1])
  return(list(PrioPost = PrioPost, penalty = h_value))
}

#' title conditional likelihood
#'
#' description conditional likelihood in objective function
#'
#' @return c_lik for the onditional likelihood in objectice function
#' @param N  dimension of residual matrix
#' @param P  dimension of residual matrix
#' @param sigmae2_v  residual matrix
#' @param sigmae2  variance structure of error
#' @keywords internal
#'
# this version we need to know the truth of sigmae2, we can use sigmae2_v as the truth
C_likelihood = function(N,P,sigmae2_v,sigmae2){
  if(is.matrix(sigmae2)){
    # print("clik as matrix of sigmae2")
    c_lik = -(1/2) * sum( log(2*pi*sigmae2) + (sigmae2_v)/(sigmae2) )
  }else if(is.vector(sigmae2) & length(sigmae2)==P){
    # print("clik as vector sigmae2")
    # change the format to fit the conditional likelihood
    sigmae2_v = colMeans(sigmae2_v)
    c_lik = -(N/2) * sum( log(2*pi*sigmae2) + (sigmae2_v)/(sigmae2) )
  } else {
    # print("clik as sigmae2 is constant")
    # change the format to fit the conditional likelihood and accelerate the computation.
    sigmae2_v = mean(sigmae2_v)
    c_lik = -(N*P)/2 * ( log(2*pi*sigmae2) + (sigmae2_v)/(sigmae2) )
    # here I want to use fully variantional inference Elogsigmae2 rather logEsigmae2
    # c_lik = -(N*P)/2 * ( log(2*pi*sigmae2_true) + (sigmae2_v)/(sigmae2_true) + log(N*P/2) - digamma(N*P/2) )
  }
  return(list(c_lik = c_lik))
}

#' title objective function in VEM
#'
#' description  objective function
#'
#' @return obj_val value of the objectice function
#' @param N  dimension of residual matrix
#' @param P  dimension of residual matrix
#' @param sigmae2_v  residual matrix
#' @param sigmae2_true  true variance structure (we use the estimated one to replace that if the truth is unknown)
#' @param par_l ash output for l
#' @param par_f ash output for f
#' @param objtype  objective function type,
#' "margin_lik" for conditional likelihood,
#' "lowerbound_lik" for full objective function
#' @keywords internal
#'

# objective function
obj = function(N,P,sigmae2_v,sigmae2,par_f,par_l,objtype = "margin_lik"){
  if(is.list(sigmae2)){
    # print("obj using kronecker product")
    sigmae2 = sigmae2$sig2_l %*% t(sigmae2$sig2_f)
  }
  if(objtype=="lowerbound_lik"){
    priopost_f = Fval(par_f$mat, par_f$g)$PrioPost
    priopost_l = Fval(par_l$mat, par_l$g)$PrioPost
    penalty_f = Fval(par_f$mat, par_f$g)$penalty
    penalty_l = Fval(par_l$mat, par_l$g)$penalty
    c_lik = C_likelihood(N,P,sigmae2_v,sigmae2)$c_lik
    obj_val = c_lik + priopost_l + priopost_f + penalty_l + penalty_f
  } else {
    obj_val = C_likelihood(N,P,sigmae2_v,sigmae2)$c_lik
  }
  return(obj_val)
}

#' title rescale the lambda_l and lambda_f for identifiablity
#'
#' description rescale the lambda_l and lambda_f for identifiablity
#'
#' @return a list of sig2_l and sig2_f for the rescaled variance of the kronecker product
#' @param sig2_l variance of the kronecker product
#' @param sig2_l variance of the kronecker product
#' @keywords internal
#'
rescale_sigmae2 = function(sig2_l,sig2_f){
  norm_l = sqrt(sum(sig2_l^2))
  norm_f = sqrt(sum(sig2_f^2))
  norm_total = norm_l * norm_f
  sig2_l = sig2_l / norm_l
  sig2_f = sig2_f / norm_f
  sig2_l = sig2_l * sqrt(norm_total)
  sig2_f = sig2_f * sqrt(norm_total)
  return(list(sig2_l = sig2_l,sig2_f = sig2_f))
}

#' title initial value for Bayes variance structure estimation for kronecker productor
#'
#' description initial value for Bayes variance structure estimation for kronecker productor
#'
#' @return sig2_out estimated variance structure
#' @param sigmae2_v  residual matrix
#' @keywords internal
#'
inital_Bayes_var = function(sigmae2_v){
  N = dim(sigmae2_v)[1]
  P = dim(sigmae2_v)[2]
  # estimate the initial value of lambda_l and lambda_f
  sig2_l_pre = rowMeans(sigmae2_v)
  sig2_f_pre = colMeans( sigmae2_v / matrix(rep(sig2_l_pre,P), ncol = P) )
  sig2_pre_list = rescale_sigmae2(sig2_l_pre,sig2_f_pre)
  sig2_l_pre = sig2_pre_list$sig2_l
  sig2_f_pre = sig2_pre_list$sig2_f
  # start the iteration
  maxiter = 100
  inital_tol = 1e-3
  tau = 0
  epsilon = 1
  while(epsilon > inital_tol & tau <= maxiter){
    tau = tau + 1
    sig2_l = rowMeans( sigmae2_v / matrix(rep(sig2_f_pre,each = N),ncol = P) )
    sig2_f = colMeans( sigmae2_v / matrix(rep(sig2_l,P), ncol = P) )
    sig2_list = rescale_sigmae2(sig2_l,sig2_f)
    sig2_l = sig2_list$sig2_l
    sig2_f = sig2_list$sig2_f
    epsilon = sqrt(mean((sig2_f - sig2_f_pre)^2 )) + sqrt(mean((sig2_l - sig2_l_pre)^2))
    sig2_l_pre = sig2_l
    sig2_f_pre = sig2_f
  }
  return(list(sig2_l = sig2_l,sig2_f = sig2_f))
}

#' title Bayes variance structure estimation for kronecker productor
#'
#' description prior and posterior part in objective function
#'
#' @return sig2_out estimated variance structure
#' @param sigmae2_v  residual matrix
#' @param sigmae2 variance structure
#' and it is a list of two vectors in this case.
#' @keywords internal
#'
Bayes_var = function(sigmae2_v,sigmae2 = NA){
  N = dim(sigmae2_v)[1]
  P = dim(sigmae2_v)[2]
  if( is.na(sigmae2) || !is.list(sigmae2) ){
    # we don't know the truth this is in the first iteration
    sigmae2 = inital_Bayes_var(sigmae2_v)
  }
  sig2_l_pre = sigmae2$sig2_l
  sig2_f_pre = sigmae2$sig2_f
  # this has already beed rescaled
  # here we use alpha_l = alpha_f = beta_l = beta_f = 0
  sig2_l = rowMeans( sigmae2_v / matrix(rep(sig2_f_pre,each = N),ncol = P) )
  sig2_f = colMeans( sigmae2_v / matrix(rep(sig2_l,P), ncol = P) )
  #rescaled the variance
  sig2_list = rescale_sigmae2(sig2_l,sig2_f)
  sig2_l = sig2_list$sig2_l
  sig2_f = sig2_list$sig2_f
  # sig2_out = matrix(rep(sig2_l,P),ncol = P) * matrix(rep(sig2_f,each = N),ncol = P)
  return(list(sig2_l = sig2_l,sig2_f = sig2_f))
}

#' title noisy variance structure estimation
#'
#' description prior and posterior part in objective function
#'
#' @return sig2_out estimated variance structure
#' @param sigmae2_v  residual matrix
#' @param sigmae2_true variance structure
#' and it is a list of two vectors in this case.
#' @keywords internal
#'

noisy_var = function(sigmae2_v,sigmae2_true){
  # a reaonable upper bound which control optimal algorithm
  upper_range = abs(mean(sigmae2_v - sigmae2_true)) + sd(sigmae2_v - sigmae2_true)
  # value of likelihood
  # f_lik <- function(x,sigmae2_v_l = sigmae2_v,sigmae2_true_l = sigmae2_true){
  #  -(1/2)*sum( log(2*pi*(x+sigmae2_true_l)) + (1/(x+sigmae2_true_l))*sigmae2_v_l )
  # }
  # I need the negative of the f_lik since optim find the minimum rather than the maximum
  f_lik <- function(x,sigmae2_v_l = sigmae2_v,sigmae2_true_l = sigmae2_true){
    (1/2)*mean( log(2*pi*(x+sigmae2_true_l)) + (1/(x+sigmae2_true_l))*sigmae2_v_l )
  }
  AA <- optim(mean(sigmae2_v - sigmae2_true), f_lik,lower = 0, upper = upper_range,  method = "Brent")
  e_sig = AA$par
  return(e_sig + sigmae2_true)
}

#' title noisy variance structure estimation with noisy variance on each column
#'
#' description prior and posterior part in objective function
#'
#' @return sig2_out estimated variance structure
#' @param sigmae2_v  residual matrix
#' @param sigmae2_true variance structure
#' and it is a list of two vectors in this case.
#' @keywords internal
#'

noisy_var_column = function(sigmae2_v,sigmae2_true){
  # in this case, we need to think about the for each column
  # we can use noisy_var = function(sigmae2_v,sigmae2_true) with the input of each column
  P = dim(sigmae2_true)[2]
  sig2_out = sapply(seq(1,P),function(x){noisy_var(sigmae2_v[,x],sigmae2_true[,x])})
  return(sig2_out)
}

#' title module for estiamtion of the variance structure
#'
#' description estiamtion of the variance structure
#'
#' @return sigmae2 estimated variance structure
#' @param partype parameter type for the variance,
#' "constant" for constant variance,
#' "var_col" for nonconstant variance for column,
#' "known" for the kown variance,
#' "Bayes_var" for Bayes version of the nonconstant variance for row and column
#' "noisy" is noisy version s_ij + sigmae2
#' "noisy_col" noisy version for column s_ij + sigmae2_j
#' @param sigmae2_v  residual matrix
#' @param sigmae2_true  true variance structure
#' @param sigmae2_pre the previouse sigmae2 which is only used in Bayes_var
#' @keywords internal
#'
# sigma estimation function
sigma_est = function(sigmae2_v,sigmae2_true,sigmae2_pre = NA, partype = "constant"){
  if(partype == "var_col"){
    # print("use var col method")
    sigmae2 = colMeans(sigmae2_v)
  } else if(partype == "noisy"){
    # print("use the noisy version")
    sigmae2 = noisy_var(sigmae2_v,sigmae2_true)
  } else if(partype == "noisy_col"){
    sigmae2 = noisy_var_column(sigmae2_v,sigmae2_true)
  }else if (partype == "Bayes_var"){
    # here sigmae2_true is a list
    # print("use the kronecker product")
    sigmae2 = Bayes_var(sigmae2_v,sigmae2_pre)
  }else if (partype == "known"){
    # print("use the known sigmae")
    sigmae2 = sigmae2_true
  } else {
    # print("use the mean value for sigmae")
    # this is for constant case
    sigmae2 = mean(sigmae2_v)
  }
  return(sigmae2)
}

#' title empirical sample variance matrix estimation
#'
#' description this function is to calculate the empirical sample variance matrix sigmae2_v
#'
#' @return sigmae2_v  empirical sample variance matrix
#' @param Y which is the residual
#' @param fl_list this is a list for the El Ef El2 and Ef2 from other factors
#' @param El mean estimation for the current loading
#' @param Ef mean estimation for the current factor
#' @param El2 second momnet estimation for the current factor
#' @param Ef2 second moment estimation for the current loading
#' @keywords internal
#'
sigmae2_v_est = function(Y,El,Ef,El2,Ef2,fl_list=list()){
  # in this version of function we haven't included the missing value case in
  if(length(fl_list)==0){
    # this is for the rank one case since there are no other factors
    # here Y is just the original data
    sigmae2_v =  Y^2 - 2*Y*(El %*% t(Ef)) + (El2 %*% t(Ef2))
  }else{
    # this is the for the rank more than one we have other factors
    # now the Y is residual matrix rather the original matrix
    # to get the original data we use Yhat
    Yhat = Y + fl_list$El %*% t(fl_list$Ef)
    # the residual matrix should be
    # this is for E(l_1f_1+l_2f_2+...+l_kf_k)^2
    fl_norm = (El%*%t(Ef) + fl_list$El%*%t(fl_list$Ef))^2 -
      (El^2 %*% t(Ef^2) + (fl_list$El)^2 %*% t((fl_list$Ef)^2)) +
      (El2 %*% t(Ef2) + fl_list$El2 %*% t(fl_list$Ef2))
    sigmae2_v = Yhat^2 - 2* Yhat * (El%*%t(Ef) + fl_list$El%*%t(fl_list$Ef)) + fl_norm
  }
  return(sigmae2_v)
}

#' title one step update in flash iteration using ash
#'
#' description one step update in flash iteration using ash
#' @return list of factor, loading and variance of noise matrix
#'  \itemize{
#'   \item{\code{El}} {is a N vector for mean of loadings}
#'   \item{\code{El2}} {is a N vector for second moment of loadings}
#'   \item{\code{Ef}} {is a N vector for mean of factors}
#'   \item{\code{Ef2}} {is a N vector for second moment of factors}
#'   \item{\code{sigmae2_v}}{is a N by P matrix for residual square}
#'   \item{\code{sigmae2_true}}{is a N by P matrix for estimated value for the variance structure}
#'   \item{\code{obj_val}}{the value of objectice function}
#'  }
#' @param Y the data matrix
#' @param N dimension of Y
#' @param P dimension of Y
#' @param El mean for the loadings
#' @param El2 second moment for the loadings
#' @param Ef mean for the factors
#' @param Ef2 second moment for the factors
#' @param sigmae2_v residual square
#' @param sigmae2_true the (true) known variance structure
#' Here, sigmae2 is the estimated variance structure in each step
#' sigmae2_true is the truth we know, some times sigmae2 is noisy version of sigmae2_true
#' @param sigmae2 the estimation of the variance structure
#' @param nonnegative if the facotor and loading are nonnegative or not.
#' TRUE for nonnegative
#' FALSE for no constraint
#' @param partype parameter type for the variance,
#' "constant" for constant variance,
#' "var_col" for nonconstant variance for column,
#' "known" for the kown variance,
#' "Bayes_var" for Bayes version of the nonconstant variance for row and column
#' "loganova" is anova estiamtion for the log residual square
#' @param objtype  objective function type,
#' "margin_lik" for conditional likelihood,
#' "lowerbound_lik" for full objective function
#' @param fix_factor whether the factor is fixed or not
#' TRUE for fix_factor
#' FALSE for non-constraint
#' @keywords internal
#'
# one step update function
one_step_update = function(Y, El, El2, Ef, Ef2,
                           N, P,
                           sigmae2_v, sigmae2_true,
                           sigmae2,
                           nonnegative = FALSE,
                           partype = "constant",
                           objtype = "margin_lik",
                           fix_factor = FALSE,
                           ash_para = list(),
                           fl_list=list()){
  # if fix_factor is True, please choose objtype = "margin_lik"
  output = ifelse(objtype == "lowerbound_lik", "matrix", "mean")
  # deal with the missing value
  na_index_Y = is.na(Y)
  is_missing = any(na_index_Y)  # keep the missing result
  if(is_missing){
    # print("missing value use the EY as missing Y")
    Y[na_index_Y] = (El %*% t(Ef))[na_index_Y]
  }
  if(fix_factor){
    # actually we need do nothing here
    output = "mean"
    objtype = "margin_lik"
  }else{
    # estimate the variance structure
    # sigma_est = function(sigmae2_v,sigmae2_true,sigmae2_pre = NA, partype = "constant")
    sigmae2 = sigma_est(sigmae2_v,sigmae2_true,sigmae2 ,partype)
    # for the list case in kronecker product
    if(is.list(sigmae2)){
      # print("sigmae is kronecker product in estimating sigmae2 in one step update")
      sigmae2 = sigmae2$sig2_l %*% t(sigmae2$sig2_f)
    }
    # Y = lf^T + E and ATM is for l given f, for f given l we need y^T = fl^T + E^T
    # sigmae2_input = ifelse(is.matrix(sigmae2),t(sigmae2),sigmae2) is wrong here
    if(is.matrix(sigmae2)){
      # print("sigmae is matrix in estimating sigmae2 in one step update")
      sigmae2_input = t(sigmae2)
    }else{
      sigmae2_input = sigmae2
    }
    par_f = ATM_r1(t(Y), El, El2, sigmae2_input,
                   col_var = "column", nonnegative,
                   output,partype,ash_para)
    Ef = par_f$Ef
    Ef2 = par_f$Ef2
    # if the Ef is zeros ,just return zeros
    if(sum(Ef^2) <= 1e-12){
      El = rep(0,length(El))
      Ef = rep(0,length(Ef))
      sigmae2_v =  sigmae2_v_est(Y,El,Ef,El^2,Ef^2,fl_list)
      sigmae2 = sigma_est(sigmae2_v,sigmae2_true,sigmae2 ,partype)
      obj_val = obj(N, P, sigmae2_v, sigmae2, par_f=NA, par_l=NA, objtype="margin_lik")
      return(list(El = El, El2 = El^2,
                  Ef = Ef, Ef2 = Ef^2,
                  sigmae2_v = sigmae2_v,
                  sigmae2 = sigmae2,
                  obj_val = obj_val))
    }
    # update the Y and the Y^2 in this if else block
    if(is_missing){
      # the missing varianece is just sigmae2_v + sigmae2
      if(is.matrix(sigmae2)){
        sigmae2_impute = sigmae2[na_index_Y]
      }else if(is.vector(sigmae2) & length(sigmae2) == length(Ef)){
        sigmae2_impute = (matrix(rep(sigmae2,each = N),ncol = P))[na_index_Y]
      }else{
        sigmae2_impute = sigmae2
      }
      sigmae2_v = sigmae2_v_est(Y,El,Ef,El2,Ef2,fl_list)
      sigmae2_v[na_index_Y] = sigmae2_v[na_index_Y] + sigmae2_impute
    }else{
      sigmae2_v = sigmae2_v_est(Y,El,Ef,El2,Ef2,fl_list)
    }
  }

  sigmae2 = sigma_est(sigmae2_v,sigmae2_true,sigmae2 ,partype)
  if(is.list(sigmae2)){
    # kronecker product
    sigmae2 = sigmae2$sig2_l %*% t(sigmae2$sig2_f)
  }
  par_l = ATM_r1(Y, Ef, Ef2,
                 sigmae2, col_var = "row",
                 nonnegative, output,
                 partype, ash_para)
  El = par_l$Ef
  El2 = par_l$Ef2
  # if El is zeros just return
  if(sum(El^2) <= 1e-12){
    El = rep(0,length(El))
    Ef = rep(0,length(Ef))
    sigmae2_v = sigmae2_v_est(Y,El,Ef,El^2,Ef^2,fl_list)
    sigmae2 = sigma_est(sigmae2_v,sigmae2_true,sigmae2 ,partype)
    obj_val = obj(N, P, sigmae2_v, sigmae2, par_f=NA, par_l=NA, objtype="margin_lik")
    return(list(El = El, El2 = El^2,
                Ef = Ef, Ef2 = Ef^2,
                sigmae2_v = sigmae2_v,
                sigmae2 = sigmae2,
                obj_val = obj_val))
  }
  if(is_missing){
    # the missing varianece is just sigmae2_v + sigmae2
    if(is.matrix(sigmae2)){
      sigmae2_impute = sigmae2[na_index_Y]
    }else if(is.vector(sigmae2) & length(sigmae2) == length(Ef)){
      sigmae2_impute = (matrix(rep(sigmae2,each = N),ncol = P))[na_index_Y]
    }else{
      sigmae2_impute = sigmae2
    }
    sigmae2_v = sigmae2_v_est(Y,El,Ef,El2,Ef2,fl_list)
    sigmae2_v[na_index_Y] = sigmae2_v[na_index_Y] + sigmae2_impute
  }else{
    sigmae2_v =  sigmae2_v_est(Y,El,Ef,El2,Ef2,fl_list)
  }
  #use the estiamtion as the truth
  sigmae2 = sigma_est(sigmae2_v,sigmae2_true,sigmae2 ,partype)
  obj_val = obj(N, P, sigmae2_v, sigmae2, par_f, par_l, objtype)

  return(list(El = El, El2 = El2,
              Ef = Ef, Ef2 = Ef2,
              sigmae2_v = sigmae2_v,
              sigmae2 = sigmae2,
              obj_val = obj_val))
}

#' inital value for flash
#'
#' description inital value for flash
#'
#' @return list of factor, loading and variance of noise matrix
#'  \itemize{
#'   \item{\code{El}} {is a N vector for mean of loadings}
#'   \item{\code{El2}} {is a N vector for second moment of loadings}
#'   \item{\code{Ef}} {is a N vector for mean of factors}
#'   \item{\code{Ef2}} {is a N vector for second moment of factors}
#'   \item{\code{sigmae2_v}}{is a N by P matrix for residual square}
#'   \item{\code{sigmae2_true}}{is a N by P matrix for estimated value for the variance structure}
#'  }
#' @param Y the data matrix
#' @param nonnegative if the facotor and loading are nonnegative or not.
#' TRUE for nonnegative
#' FALSE for no constraint
#' @param fix_factor whether the factor is fixed or not
#' TRUE for fix_factor
#' FALSE for non-constraint
#' @param factor_value is the factor value if the factor is fixed
#' @param fl_list this is a list for the El, El2, Ef and Ef2 for other factor loadings from the model.
#'                in the flash for backfitting, we need those to get the residuals and estimation for variance
#' @param initial_list_r1 this is a list for the inital values of current El Ef El2 and Ef2.
#' @param fix_initial this is a indicator for the initial value is fix or not.
#' @keywords internal
#'
initial_value = function(Y, nonnegative = FALSE,
                         factor_value = NA,fix_factor = FALSE,
                         initial_list_r1 = list(), fix_initial = FALSE,
                         fl_list=list()){
  # use the total mean as the estimated missing value
  na_index_Y = is.na(Y)
  is_missing = any(na_index_Y)
  if(is_missing){
    # this is the initialization for the missing value
    Y[na_index_Y] = mean(Y, na.rm = TRUE)
  }
  # the flash with plug-in missing value
  if(fix_initial){
    El = initial_list_r1$El
    Ef = initial_list_r1$Ef
    El2 = initial_list_r1$El2
    Ef2 = initial_list_r1$Ef2
  }else if(fix_factor){
    Ef = factor_value
    Ef2 = Ef^2
    El = as.vector( (Y %*% Ef) / (sum(Ef2)) )
    El2 = El^2
  }else{
    # El = svd(Y)$u[,1]
    El = as.vector(irlba::irlba(Y,nv = 0,nu = 1)$u)
    # the nonnegative value need positive inital value
    if(nonnegative){
      # El = abs(El)
      El = El * (El > 0)
    }
    El2 = El^2
    Ef = as.vector(t(El)%*%Y)
    Ef2 = Ef^2
  }
  # residual matrix initialization
  # here we don't have any information about the sigmae2_true
  if(is_missing){
    # here we don't have any information about the variance and the Y as well
    # since the previoust the Y is impute by the mean of observed Ys so I would use the mean of Y here as well.
    # this initialization still need to be considered more.
    sigmae2_v =  sigmae2_v_est(Y,El,Ef,El2,Ef2,fl_list)
  }else{
    sigmae2_v =  sigmae2_v_est(Y,El,Ef,El2,Ef2,fl_list)
  }
  return(list(El = El, El2 = El^2,
              Ef = Ef, Ef2 = Ef^2,
              sigmae2_v = sigmae2_v))
}

#' FLASH
#'
#' factor loading adaptive shrinkage rank one version
#' @return list of factor, loading and variance of noise matrix
#'  \itemize{
#'   \item{\code{El}} {is a N vector for mean of loadings}
#'   \item{\code{Ef}} {is a N vector for mean of factors}
#'   \item{\code{sigmae2}}{is a N by P matrix for estimated value for the variance structure}
#'  }
#' @param Y the data matrix
#' @param tol is for the tolerence for convergence in iterations and ash
#' @param maciter_r1 is maximum of the iteration times for rank one case
#' @param sigmae2_true true value for the variance structure
#' @param nonnegative if the facotor and loading are nonnegative or not.
#' TRUE for nonnegative
#' FALSE for no constraint
#' @param partype parameter type for the variance,
#' "constant" for constant variance,
#' "var_col" for nonconstant variance for column,
#' "known" for the kown variance,
#' "Bayes_var" for Bayes version of the nonconstant variance for row and column
#' "loganova" is anova estiamtion for the log residual square
#' @param objtype  objective function type,
#' "margin_lik" for conditional likelihood,
#' "lowerbound_lik" for full objective function
#' @param fix_factor whether the factor is fixed or not
#' TRUE for fix_factor
#' FALSE for non-constraint
#' @param factor_value is the factor value if the factor is fixed
#' @param ash_para is the parameters list for ash
#' @param fl_list is a list containing all the informations from other factors ans loadings
#' l is loadings, f is factors, l2 and f2 are corresponding second moments.
#' priorpost_vec is expectation of log piropost ratio
#' clik is conditional likelihood (marginal likelihood)
#'
#' @details flash privide rank one matrix decomposition with variational EM algorithm.
#'
#' @export flash
#'
#' @importFrom ashr ash
#'
flash = function(Y, tol=1e-5, maxiter_r1 = 500,
                 partype = c("constant","known","Bayes_var","var_col","noisy","noisy_col"),
                 sigmae2_true = NA,
                 factor_value = NA,fix_factor = FALSE,
                 initial_list_r1 = list(), fix_initial = FALSE,
                 nonnegative = FALSE,
                 objtype = c("margin_lik","lowerbound_lik"),
                 ash_para = list(),
                 fl_list=list()){
  # match the parameters
  partype = match.arg(partype, c("constant","known","Bayes_var","var_col","noisy","noisy_col"))
  objtype = match.arg(objtype, c("margin_lik","lowerbound_lik"))

  # check the input
  if( any(!is.na(sigmae2_true)) & !(partype %in% c("known","noisy","noisy_col")) ){
    stop("You should choose partype = 'known' or 'noisy' ,'noisy_col',if you want to input the value of sigmae2_true")
  }

  N = dim(Y)[1]
  P = dim(Y)[2]

  # to get the inital values
  g_initial = initial_value(Y, nonnegative,factor_value,fix_factor,initial_list_r1, fix_initial,fl_list)
  El = g_initial$El
  Ef = g_initial$Ef
  El2 = g_initial$El2
  Ef2 = g_initial$Ef2
  sigmae2_v = g_initial$sigmae2_v

  # start iteration
  # in the first iteration, there is no value foe estimated sigmae2
  g_update = one_step_update(Y, El, El2, Ef, Ef2,
                             N, P,
                             sigmae2_v, sigmae2_true,
                             sigmae2 = NA,
                             nonnegative ,
                             partype ,
                             objtype ,
                             fix_factor,
                             ash_para,
                             fl_list)
  # parameters updates
  El = g_update$El
  El2 = g_update$El2
  Ef = g_update$Ef
  Ef2 = g_update$Ef2
  sigmae2_v = g_update$sigmae2_v
  sigmae2 = g_update$sigmae2
  obj_val = g_update$obj_val

  # track the objective value
  obj_val_track = c(obj_val)

  # we should also return when the first run get all zeros
  if(sum(El^2)==0 || sum(Ef^2)==0){
    sigmae2 = sigma_est(sigmae2_v,sigmae2_true,sigmae2 ,partype)
    if(is.list(sigmae2)){
      # print("here using kronecker product")
      sigmae2 = sigmae2$sig2_l %*% t(sigmae2$sig2_f)
    }
    # add one more output for greedy algorithm which not useful here
    c_lik_val = C_likelihood(N,P,sigmae2_v,sigmae2)$c_lik
    # the above value is not useful, but is helpful to get the postprior value
    # since obj_val = c_lik_value + priorpost_l + priorpost_f
    return(list(l = El, f = Ef, l2 = El2, f2 = Ef2,
                sigmae2 = sigmae2,
                obj_val = obj_val,
                c_lik_val = c_lik_val))
  }

  epsilon = 1
  tau = 1
  while(epsilon >= tol & tau < maxiter_r1){
    tau = tau + 1
    pre_obj = obj_val

    g_update = one_step_update(Y, El, El2, Ef, Ef2,
                               N, P,
                               sigmae2_v, sigmae2_true,
                               sigmae2,
                               nonnegative ,
                               partype ,
                               objtype ,
                               fix_factor,
                               ash_para,
                               fl_list)
    # parameters updates
    El = g_update$El
    El2 = g_update$El2
    Ef = g_update$Ef
    Ef2 = g_update$Ef2
    sigmae2_v = g_update$sigmae2_v
    sigmae2 = g_update$sigmae2
    obj_val = g_update$obj_val

    if(sum(El^2)==0 || sum(Ef^2)==0){
      El = rep(0,length(El))
      Ef = rep(0,length(Ef))
      break
    }
    epsilon = abs(pre_obj - obj_val)
    obj_val_track = c(obj_val_track,obj_val)
  }
  sigmae2 = sigma_est(sigmae2_v,sigmae2_true,sigmae2 ,partype)
  if(is.list(sigmae2)){
    # print("here, using kronecker product")
    sigmae2 = sigmae2$sig2_l %*% t(sigmae2$sig2_f)
  }
  # add one more output for greedy algorithm which not useful here
  c_lik_val = C_likelihood(N,P,sigmae2_v,sigmae2)$c_lik
  # the above value is not useful, but is helpful to get the postprior value
  # since obj_val = c_lik_value + priorpost_l + priorpost_f
  return(list(l = El, f = Ef, l2 = El2, f2 = Ef2,
              sigmae2 = sigmae2,
              obj_val = obj_val,
              c_lik_val = c_lik_val,
              obj_val_track = obj_val_track))
}

```

```{r}
sim_K = function(K, N, P, SF, SL, signal,noise){
  E = matrix(rnorm(N*P,0,noise),nrow=N)
  Y = E
  L_true = array(0, dim = c(N,K))
  F_true = array(0, dim = c(P,K))

  for(k in 1:K){
    lstart = rnorm(N, 0, signal)
    fstart = rnorm(P, 0, signal)

    index = sample(seq(1:N),(N*SL))
    lstart[index] = 0
    index = sample(seq(1:P),(P*SF))
    fstart[index] = 0

    L_true[,k] = lstart
    F_true[,k] = fstart

    Y = Y + lstart %*% t(fstart)
  }
  return(list(Y = Y, L_true = L_true, F_true = F_true, Error = E))
}
```

we check that the new objective function with penalty term does monotonically increase.

```{r}
set.seed(99)
K = 1
N = 20
P = 30
SF = 0.5
SL = 0.6
signal = 1
noise = 1
par(mfrow = c(2, 2))
par(cex = 0.6)
par(mar = c(3, 3, 0.8, 0.8), oma = c(1, 1, 1, 1))
data = sim_K(K,N, P, SF , SL , signal ,noise )
Y = data$Y
E = data$Error
g1 = flash(Y,objtype = "l")
plot(g1$obj_val_track,main = "fdr method with penalized obj",type = "l")
g2 = flashr::flash(Y,objtype = "l")
plot(g2$obj_val_track,main = "fdr method without penalized obj",type = "l")
```

we can see that the old version of code without penalty doesn't have monotonically increasing objective function, but the new version code with penalty term does.

## try the same problem: rank 0 vs rank 1 with  fdr method.

```{r,echo = FALSE}

rank0check = function(K=1,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 0,noise = 1, mtype = "fdr"){
  data = sim_K(K,N, P, SF , SL , signal ,noise )
  Y = data$Y
  E = data$Error
  if(mtype == "fdr"){
    g1 = flash(Y,objtype = "l")
  }else{
    g1 = flash(Y,objtype = "l",ash_para = list(method = "shrink"))
  }

  RMSE = sqrt(mean((Y - g1$l %*% t(g1$f) -E)^2)) 
  obj_val = g1$obj_val
  # to get the rank zero objective value
  sigmae2_v0 = Y^2
  fit_g0 = list(pi=c(1,0,0),sd = c(0,0.1,1))
  mat0 = list(comp_postmean = matrix(0,ncol= N,nrow = 3),
              comp_postmean2 = matrix(0,ncol= N,nrow = 3),
              comp_postprob = matrix(0,ncol= N,nrow = 3))
  mat0$comp_postprob[1,] = 1
  par_l0 = list(g = fit_g0, mat = mat0)
  fit_g0 = list(pi=c(1,0,0),sd = c(0,0.1,1))
  mat0 = list(comp_postmean = matrix(0,ncol= P,nrow = 3),
              comp_postmean2 = matrix(0,ncol= P,nrow = 3),
              comp_postprob = matrix(0,ncol= P,nrow = 3))
  mat0$comp_postprob[1,] = 1
  par_f0 = list(g = fit_g0, mat = mat0)
  obj0 = obj(N,P,sigmae2_v0,sigmae2 = mean(sigmae2_v0),par_f0,par_l0,objtype = "lowerbound_lik")

  return(list(rmse = RMSE,obj_val = obj_val,obj_0 = obj0))
}

sim_rank0 = function(K=1,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 0,noise = 1,mtype = "fdr"){
  T = 200
  RMSE = rep(NA,T)
  sign_obj = rep(NA,T)
  for(t in 1:T){
    g0 = rank0check(K = 1,N, P, SF, SL , signal,noise,mtype = mtype)
    RMSE[t] = g0$rmse
    sign_obj[t] = sign(g0$obj_val - g0$obj_0)
  }
  par(mfrow = c(2, 2))
  par(cex = 0.6)
  par(mar = c(3, 3, 0.8, 0.8), oma = c(1, 1, 1, 1))
  boxplot(RMSE,main = "RMSE")
  hist(sign_obj, breaks = 3,main = "sign of difference of obj value")
  hist(RMSE,breaks = 50,main = "RMSE")
  plot(sign_obj,RMSE, main = "sign of obj diff vs RMSE")
}

#### old code

rank0check_old = function(K=1,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 0,noise = 1, mtype = "shrink"){
  data = sim_K(K,N, P, SF , SL , signal ,noise )
  Y = data$Y
  E = data$Error
  if(mtype == "fdr"){
    g1 = flashr::flash(Y,objtype = "l")
  }else{
    g1 = flashr::flash(Y,objtype = "l",ash_para = list(method = "shrink"))
  }

  RMSE = sqrt(mean((Y - g1$l %*% t(g1$f) -E)^2)) 
  obj_val = g1$obj_val
  # to get the rank zero objective value
  sigmae2_v0 = Y^2
  fit_g0 = list(pi=c(1,0,0),sd = c(0,0.1,1))
  mat0 = list(comp_postmean = matrix(0,ncol= N,nrow = 3),
              comp_postmean2 = matrix(0,ncol= N,nrow = 3),
              comp_postprob = matrix(0,ncol= N,nrow = 3))
  mat0$comp_postprob[1,] = 1
  par_l0 = list(g = fit_g0, mat = mat0)
  fit_g0 = list(pi=c(1,0,0),sd = c(0,0.1,1))
  mat0 = list(comp_postmean = matrix(0,ncol= P,nrow = 3),
              comp_postmean2 = matrix(0,ncol= P,nrow = 3),
              comp_postprob = matrix(0,ncol= P,nrow = 3))
  mat0$comp_postprob[1,] = 1
  par_f0 = list(g = fit_g0, mat = mat0)
  obj0 = obj(N,P,sigmae2_v0,sigmae2 = mean(sigmae2_v0),par_f0,par_l0,objtype = "lowerbound_lik")

  return(list(rmse = RMSE,obj_val = obj_val,obj_0 = obj0))
}

sim_rank0_old = function(K=1,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 0,noise = 1,mtype = "shrink"){
  T = 200
  RMSE = rep(NA,T)
  sign_obj = rep(NA,T)
  for(t in 1:T){
    g0 = rank0check_old(K = 1,N, P, SF, SL , signal,noise,mtype = mtype)
    RMSE[t] = g0$rmse
    sign_obj[t] = sign(g0$obj_val - g0$obj_0)
  }
  par(mfrow = c(2, 2))
  par(cex = 0.6)
  par(mar = c(3, 3, 0.8, 0.8), oma = c(1, 1, 1, 1))
  boxplot(RMSE,main = "RMSE")
  hist(sign_obj, breaks = 3,main = "sign of difference of obj value")
  hist(RMSE,breaks = 50,main = "RMSE")
  plot(sign_obj,RMSE, main = "sign of obj diff vs RMSE")
}
```

#### case 1: rank 0

we compare the old code with our new code, we can see that most of the cases a

*shrink method without penalty*

```{r}
set.seed(99)
sim_rank0_old(K=1,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 0,noise = 1)
```

*fdr method with penalty*

```{r}
set.seed(99)
sim_rank0(K=1,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 0,noise = 1)
```

we can see the improvement that the proportion of P(estimated structure better than rank 0 structure) is much larger.

#### case 2: rank 1, very sparse with large noise

*shrink method without penalty*

```{r}
set.seed(99)
sim_rank0_old(K=1,N = 100, P=200, SF = 0.8, SL = 0.8, signal = 1,noise = 2)
```

*fdr method with penalty*

```{r}
set.seed(99)
sim_rank0(K=1,N = 100, P=200, SF = 0.8, SL = 0.8, signal = 1,noise = 2)
```

we can see that the results are improved since the proportion of P(estimated structure better than rank 0 structure) is much larger.

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```

### 2017-06-20 analysis/Obj_summary.Rmd
Files commited by wflow_commit.

--- 2017-06-20 analysis/Objcheck_2.Rmd
@@
 ---
-title: "objective function with the penalty term"
-author: "wei"
+title: "simple summary of objective function for rank 1 vs rank 0"
+author: "Wei Wang"
 date: YYYY-MM-DD
@@
 
-In this case, we would like to try the objective function with penalty term.
-
-I would like to modify the flash code in this document.
-
-```{r,echo=FALSE}
-# ATM function
-#' title ash type model for l
-#'
-#' description use ash type model to maxmization
-#'
-#' @return list of factor, loading and variance of noise matrix
-#'  \itemize{
-#'   \item{\code{El}} {is a N vector for mean of loadings}
-#'   \item{\code{El2}} {is a N vector for second moment of loadings}
-#'  }
-#' @param Y is data matrix
-#' @param Ef is mean for the factor
-#' @param Ef2 is second moment for the factor
-#' @param sigmae2 is variance structure for noise matrix
-#' @param col_var is just for column variance, column is for column, row is for row
-#' @param nonnegative is flag for whether output nonnegative value
-#' @param output is format of output, "mean" is for mean value, "matrix" is for flash data matrix in ash
-#'
-#' @keywords internal
-#'
-
-ATM_r1 = function(Y, Ef, Ef2,
-                  sigmae2, col_var = "row",
-                  nonnegative = FALSE, output = "mean",
-                  partype = "constant",ash_para = list()){
-  # this part is preparing betahat and sebeta for ash
-  if(is.matrix(sigmae2)){
-    # print("sigmae2 is a matrix")
-    # this is for all variance are different
-    sum_Ef2 = (1/sigmae2) %*% Ef2
-    sum_Ef2 = as.vector(sum_Ef2)
-    sebeta = sqrt(1/(sum_Ef2))
-    betahat = as.vector( (Y/sigmae2) %*% Ef ) / (sum_Ef2)
-    betahat=as.vector(betahat)
-  } else if(is.vector(sigmae2) & length(sigmae2) == length(Ef)){
-    # this is for the non-constant variance in column
-    # print("sigmae2 is vector")
-    if(col_var == "row"){
-      sum_Ef2 = (1/sigmae2) * Ef2
-      sum_Ef2 = sum(sum_Ef2)
-      sebeta = sqrt(1/(sum_Ef2))
-      betahat = as.vector( Y %*% (Ef/sigmae2) ) / (sum_Ef2)
-      betahat=as.vector(betahat)
-    }else {
-      # for column
-      # here I have alread change the dimension by taking transpose to Y
-      sum_Ef2 = sum(Ef2)
-      sebeta = sqrt( sigmae2/(sum_Ef2) )
-      betahat = (t(Ef) %*% t(Y)) / (sum_Ef2)
-      betahat=as.vector(betahat)
-    }
-  }else{
-    # for constant case
-    # print("sigmae2 is a constant")
-    sum_Ef2 = sum(Ef2)
-    sebeta = sqrt(sigmae2/(sum_Ef2))
-    betahat = (t(Ef) %*% t(Y)) / (sum_Ef2)
-    betahat=as.vector(betahat)
-  }
-  # set ash default setting up
-  ash_default = list(betahat = betahat, sebeta = sebeta,
-                     method = "fdr", mixcompdist = "normal")
-  # ATM update
-  # decide the sign for output
-  if(nonnegative){
-    ash_default$mixcompdist = "+uniform"
-  }
-  # decide the output to decide the convergence criterion
-  if(output == "matrix"){
-    ash_para$outputlevel = 4
-    ash_para = modifyList(ash_default,ash_para)
-    ATM = do.call(ashr::ash,ash_para)
-    # ATM = ash(betahat, sebeta, method="fdr", mixcompdist=mixdist,outputlevel=4)
-    Ef = ATM$result$PosteriorMean
-    SDf = ATM$result$PosteriorSD
-    Ef2 = SDf^2 + Ef^2
-    mat_post = ATM$flash_data
-    fit_g = ATM$fitted_g
-    return(list(Ef = Ef,
-                Ef2 = Ef2,
-                mat = mat_post,
-                g = fit_g))
-  } else {
-    ash_para = modifyList(ash_default,ash_para)
-    ATM = do.call(ashr::ash,ash_para)
-    # ATM = ash(betahat, sebeta, method="fdr", mixcompdist=mixdist)
-    Ef = ATM$result$PosteriorMean
-    SDf = ATM$result$PosteriorSD
-    Ef2 = SDf^2 + Ef^2
-    return(list(Ef = Ef, Ef2 = Ef2))
-  }
-
-}
-
-#' title prior and posterior part in objective function
-#'
-#' description prior and posterior part in objective function
-#'
-#' @return PrioPost the value for the proir and posterior in objectice function
-#'  \itemize{
-#'   \item{\code{PrioPost}} {the value for the proir and posterior in objectice function}
-#'   \item{\code{penalty}} {penalty term of the g_l or g_f}
-#'  }
-#' @param mat matrix of flash.data in ash output which is for posterior
-#' @param fit_g in fitted.g in ash output which is for prior
-#' @keywords internal
-#'
-# two parts for the likelihood
-Fval = function(mat,fit_g){
-  # make sure the first row included in the index
-  nonzeroindex = unique(c(1,which(fit_g$pi != 0)))
-  #prepare for the posterior part
-  mat_postmean = mat$comp_postmean[nonzeroindex,]
-  mat_postmean2 = mat$comp_postmean2[nonzeroindex,]
-  mat_postvar = mat_postmean2 - mat_postmean^2
-  mat_postprob = mat$comp_postprob[nonzeroindex,]
-  # figure our the dimension
-  K = dim(mat_postprob)[1]
-  N = dim(mat_postprob)[2]
-  if(is.vector(mat_postprob)){
-    K = 1
-    N = length(mat_postprob)
-  }
-  # prepare for the prior
-  prior_pi = fit_g$pi[nonzeroindex]
-  prior_var = (fit_g$sd[nonzeroindex])^2
-  mat_priorvar = matrix(rep(prior_var,N),ncol = N)
-  mat_priorprob = matrix(rep(prior_pi,N),ncol = N)
-  # to get the value
-  varodds = mat_priorvar / mat_postvar
-  varodds[1,] = 1 # deal with 0/0
-  probodds = mat_priorprob / mat_postprob
-  ssrodds = mat_postmean2 / mat_priorvar
-  ssrodds[1,] = 1  # deal with 0/0
-  priorpost = mat_postprob * (log(probodds) - (1/2)*log(varodds) - (1/2)* (ssrodds -1))
-  # in case of mat_postprob = 0
-  priorpost[which(mat_postprob< 1e-100)] = 0
-  PrioPost = sum(priorpost)
-  #######
-  # now we try to get the penalty term
-  # I use the default of setting for lambda_k in https://oup.silverchair-cdn.com/oup/backfile/Content_public/Journal/biostatistics/18/2/10.1093_biostatistics_kxw041/1/kxw041_Supp.pdf?Expires=1498001808&Signature=dtxDDFmYZ1u1jFoXfNg4mqGQnTPkQQMFZgt2-FE9r8GcgRl4DdEaP0oI6kzvUjZB5H3hLZmvAZfuFkt3GxojQJ1eX0y7I9kMJC75VTrxw1Ym~psIe8rdU5Xw4S4KZm79o1cOXBzY1iAORFkVM4z59uP1T1ltFJDu5fhlTYErU3flbiE1ivTm-BYN5P7w8dex-R5MWn98NpMVRUsoHRgbe7FhwavdnQ8QAD7~q1F4DzMyZk09TeNdEc7GNdBKbSOoS9mGGnvrqXvPz1nn~QzJ2U2CMCf04GfsgkclyUnH0aD6Vh1lSAfoEV47ZnT-9~942~f8rMs30H2Kqji7BjlctA__&Key-Pair-Id=APKAIUCZBIA4LVPAVW3Q
-  h_value = (10 - 1) * log(fit_g$pi[1])
-  return(list(PrioPost = PrioPost, penalty = h_value))
-}
-
-#' title conditional likelihood
-#'
-#' description conditional likelihood in objective function
-#'
-#' @return c_lik for the onditional likelihood in objectice function
-#' @param N  dimension of residual matrix
-#' @param P  dimension of residual matrix
-#' @param sigmae2_v  residual matrix
-#' @param sigmae2  variance structure of error
-#' @keywords internal
-#'
-# this version we need to know the truth of sigmae2, we can use sigmae2_v as the truth
-C_likelihood = function(N,P,sigmae2_v,sigmae2){
-  if(is.matrix(sigmae2)){
-    # print("clik as matrix of sigmae2")
-    c_lik = -(1/2) * sum( log(2*pi*sigmae2) + (sigmae2_v)/(sigmae2) )
-  }else if(is.vector(sigmae2) & length(sigmae2)==P){
-    # print("clik as vector sigmae2")
-    # change the format to fit the conditional likelihood
-    sigmae2_v = colMeans(sigmae2_v)
-    c_lik = -(N/2) * sum( log(2*pi*sigmae2) + (sigmae2_v)/(sigmae2) )
-  } else {
-    # print("clik as sigmae2 is constant")
-    # change the format to fit the conditional likelihood and accelerate the computation.
-    sigmae2_v = mean(sigmae2_v)
-    c_lik = -(N*P)/2 * ( log(2*pi*sigmae2) + (sigmae2_v)/(sigmae2) )
-    # here I want to use fully variantional inference Elogsigmae2 rather logEsigmae2
-    # c_lik = -(N*P)/2 * ( log(2*pi*sigmae2_true) + (sigmae2_v)/(sigmae2_true) + log(N*P/2) - digamma(N*P/2) )
-  }
-  return(list(c_lik = c_lik))
-}
-
-#' title objective function in VEM
-#'
-#' description  objective function
-#'
-#' @return obj_val value of the objectice function
-#' @param N  dimension of residual matrix
-#' @param P  dimension of residual matrix
-#' @param sigmae2_v  residual matrix
-#' @param sigmae2_true  true variance structure (we use the estimated one to replace that if the truth is unknown)
-#' @param par_l ash output for l
-#' @param par_f ash output for f
-#' @param objtype  objective function type,
-#' "margin_lik" for conditional likelihood,
-#' "lowerbound_lik" for full objective function
-#' @keywords internal
-#'
-
-# objective function
-obj = function(N,P,sigmae2_v,sigmae2,par_f,par_l,objtype = "margin_lik"){
-  if(is.list(sigmae2)){
-    # print("obj using kronecker product")
-    sigmae2 = sigmae2$sig2_l %*% t(sigmae2$sig2_f)
-  }
-  if(objtype=="lowerbound_lik"){
-    priopost_f = Fval(par_f$mat, par_f$g)$PrioPost
-    priopost_l = Fval(par_l$mat, par_l$g)$PrioPost
-    penalty_f = Fval(par_f$mat, par_f$g)$penalty
-    penalty_l = Fval(par_l$mat, par_l$g)$penalty
-    c_lik = C_likelihood(N,P,sigmae2_v,sigmae2)$c_lik
-    obj_val = c_lik + priopost_l + priopost_f + penalty_l + penalty_f
-  } else {
-    obj_val = C_likelihood(N,P,sigmae2_v,sigmae2)$c_lik
-  }
-  return(obj_val)
-}
-
-#' title rescale the lambda_l and lambda_f for identifiablity
-#'
-#' description rescale the lambda_l and lambda_f for identifiablity
-#'
-#' @return a list of sig2_l and sig2_f for the rescaled variance of the kronecker product
-#' @param sig2_l variance of the kronecker product
-#' @param sig2_l variance of the kronecker product
-#' @keywords internal
-#'
-rescale_sigmae2 = function(sig2_l,sig2_f){
-  norm_l = sqrt(sum(sig2_l^2))
-  norm_f = sqrt(sum(sig2_f^2))
-  norm_total = norm_l * norm_f
-  sig2_l = sig2_l / norm_l
-  sig2_f = sig2_f / norm_f
-  sig2_l = sig2_l * sqrt(norm_total)
-  sig2_f = sig2_f * sqrt(norm_total)
-  return(list(sig2_l = sig2_l,sig2_f = sig2_f))
-}
-
-#' title initial value for Bayes variance structure estimation for kronecker productor
-#'
-#' description initial value for Bayes variance structure estimation for kronecker productor
-#'
-#' @return sig2_out estimated variance structure
-#' @param sigmae2_v  residual matrix
-#' @keywords internal
-#'
-inital_Bayes_var = function(sigmae2_v){
-  N = dim(sigmae2_v)[1]
-  P = dim(sigmae2_v)[2]
-  # estimate the initial value of lambda_l and lambda_f
-  sig2_l_pre = rowMeans(sigmae2_v)
-  sig2_f_pre = colMeans( sigmae2_v / matrix(rep(sig2_l_pre,P), ncol = P) )
-  sig2_pre_list = rescale_sigmae2(sig2_l_pre,sig2_f_pre)
-  sig2_l_pre = sig2_pre_list$sig2_l
-  sig2_f_pre = sig2_pre_list$sig2_f
-  # start the iteration
-  maxiter = 100
-  inital_tol = 1e-3
-  tau = 0
-  epsilon = 1
-  while(epsilon > inital_tol & tau <= maxiter){
-    tau = tau + 1
-    sig2_l = rowMeans( sigmae2_v / matrix(rep(sig2_f_pre,each = N),ncol = P) )
-    sig2_f = colMeans( sigmae2_v / matrix(rep(sig2_l,P), ncol = P) )
-    sig2_list = rescale_sigmae2(sig2_l,sig2_f)
-    sig2_l = sig2_list$sig2_l
-    sig2_f = sig2_list$sig2_f
-    epsilon = sqrt(mean((sig2_f - sig2_f_pre)^2 )) + sqrt(mean((sig2_l - sig2_l_pre)^2))
-    sig2_l_pre = sig2_l
-    sig2_f_pre = sig2_f
-  }
-  return(list(sig2_l = sig2_l,sig2_f = sig2_f))
-}
-
-#' title Bayes variance structure estimation for kronecker productor
-#'
-#' description prior and posterior part in objective function
-#'
-#' @return sig2_out estimated variance structure
-#' @param sigmae2_v  residual matrix
-#' @param sigmae2 variance structure
-#' and it is a list of two vectors in this case.
-#' @keywords internal
-#'
-Bayes_var = function(sigmae2_v,sigmae2 = NA){
-  N = dim(sigmae2_v)[1]
-  P = dim(sigmae2_v)[2]
-  if( is.na(sigmae2) || !is.list(sigmae2) ){
-    # we don't know the truth this is in the first iteration
-    sigmae2 = inital_Bayes_var(sigmae2_v)
-  }
-  sig2_l_pre = sigmae2$sig2_l
-  sig2_f_pre = sigmae2$sig2_f
-  # this has already beed rescaled
-  # here we use alpha_l = alpha_f = beta_l = beta_f = 0
-  sig2_l = rowMeans( sigmae2_v / matrix(rep(sig2_f_pre,each = N),ncol = P) )
-  sig2_f = colMeans( sigmae2_v / matrix(rep(sig2_l,P), ncol = P) )
-  #rescaled the variance
-  sig2_list = rescale_sigmae2(sig2_l,sig2_f)
-  sig2_l = sig2_list$sig2_l
-  sig2_f = sig2_list$sig2_f
-  # sig2_out = matrix(rep(sig2_l,P),ncol = P) * matrix(rep(sig2_f,each = N),ncol = P)
-  return(list(sig2_l = sig2_l,sig2_f = sig2_f))
-}
-
-#' title noisy variance structure estimation
-#'
-#' description prior and posterior part in objective function
-#'
-#' @return sig2_out estimated variance structure
-#' @param sigmae2_v  residual matrix
-#' @param sigmae2_true variance structure
-#' and it is a list of two vectors in this case.
-#' @keywords internal
-#'
-
-noisy_var = function(sigmae2_v,sigmae2_true){
-  # a reaonable upper bound which control optimal algorithm
-  upper_range = abs(mean(sigmae2_v - sigmae2_true)) + sd(sigmae2_v - sigmae2_true)
-  # value of likelihood
-  # f_lik <- function(x,sigmae2_v_l = sigmae2_v,sigmae2_true_l = sigmae2_true){
-  #  -(1/2)*sum( log(2*pi*(x+sigmae2_true_l)) + (1/(x+sigmae2_true_l))*sigmae2_v_l )
-  # }
-  # I need the negative of the f_lik since optim find the minimum rather than the maximum
-  f_lik <- function(x,sigmae2_v_l = sigmae2_v,sigmae2_true_l = sigmae2_true){
-    (1/2)*mean( log(2*pi*(x+sigmae2_true_l)) + (1/(x+sigmae2_true_l))*sigmae2_v_l )
-  }
-  AA <- optim(mean(sigmae2_v - sigmae2_true), f_lik,lower = 0, upper = upper_range,  method = "Brent")
-  e_sig = AA$par
-  return(e_sig + sigmae2_true)
-}
-
-#' title noisy variance structure estimation with noisy variance on each column
-#'
-#' description prior and posterior part in objective function
-#'
-#' @return sig2_out estimated variance structure
-#' @param sigmae2_v  residual matrix
-#' @param sigmae2_true variance structure
-#' and it is a list of two vectors in this case.
-#' @keywords internal
-#'
-
-noisy_var_column = function(sigmae2_v,sigmae2_true){
-  # in this case, we need to think about the for each column
-  # we can use noisy_var = function(sigmae2_v,sigmae2_true) with the input of each column
-  P = dim(sigmae2_true)[2]
-  sig2_out = sapply(seq(1,P),function(x){noisy_var(sigmae2_v[,x],sigmae2_true[,x])})
-  return(sig2_out)
-}
-
-#' title module for estiamtion of the variance structure
-#'
-#' description estiamtion of the variance structure
-#'
-#' @return sigmae2 estimated variance structure
-#' @param partype parameter type for the variance,
-#' "constant" for constant variance,
-#' "var_col" for nonconstant variance for column,
-#' "known" for the kown variance,
-#' "Bayes_var" for Bayes version of the nonconstant variance for row and column
-#' "noisy" is noisy version s_ij + sigmae2
-#' "noisy_col" noisy version for column s_ij + sigmae2_j
-#' @param sigmae2_v  residual matrix
-#' @param sigmae2_true  true variance structure
-#' @param sigmae2_pre the previouse sigmae2 which is only used in Bayes_var
-#' @keywords internal
-#'
-# sigma estimation function
-sigma_est = function(sigmae2_v,sigmae2_true,sigmae2_pre = NA, partype = "constant"){
-  if(partype == "var_col"){
-    # print("use var col method")
-    sigmae2 = colMeans(sigmae2_v)
-  } else if(partype == "noisy"){
-    # print("use the noisy version")
-    sigmae2 = noisy_var(sigmae2_v,sigmae2_true)
-  } else if(partype == "noisy_col"){
-    sigmae2 = noisy_var_column(sigmae2_v,sigmae2_true)
-  }else if (partype == "Bayes_var"){
-    # here sigmae2_true is a list
-    # print("use the kronecker product")
-    sigmae2 = Bayes_var(sigmae2_v,sigmae2_pre)
-  }else if (partype == "known"){
-    # print("use the known sigmae")
-    sigmae2 = sigmae2_true
-  } else {
-    # print("use the mean value for sigmae")
-    # this is for constant case
-    sigmae2 = mean(sigmae2_v)
-  }
-  return(sigmae2)
-}
-
-#' title empirical sample variance matrix estimation
-#'
-#' description this function is to calculate the empirical sample variance matrix sigmae2_v
-#'
-#' @return sigmae2_v  empirical sample variance matrix
-#' @param Y which is the residual
-#' @param fl_list this is a list for the El Ef El2 and Ef2 from other factors
-#' @param El mean estimation for the current loading
-#' @param Ef mean estimation for the current factor
-#' @param El2 second momnet estimation for the current factor
-#' @param Ef2 second moment estimation for the current loading
-#' @keywords internal
-#'
-sigmae2_v_est = function(Y,El,Ef,El2,Ef2,fl_list=list()){
-  # in this version of function we haven't included the missing value case in
-  if(length(fl_list)==0){
-    # this is for the rank one case since there are no other factors
-    # here Y is just the original data
-    sigmae2_v =  Y^2 - 2*Y*(El %*% t(Ef)) + (El2 %*% t(Ef2))
-  }else{
-    # this is the for the rank more than one we have other factors
-    # now the Y is residual matrix rather the original matrix
-    # to get the original data we use Yhat
-    Yhat = Y + fl_list$El %*% t(fl_list$Ef)
-    # the residual matrix should be
-    # this is for E(l_1f_1+l_2f_2+...+l_kf_k)^2
-    fl_norm = (El%*%t(Ef) + fl_list$El%*%t(fl_list$Ef))^2 -
-      (El^2 %*% t(Ef^2) + (fl_list$El)^2 %*% t((fl_list$Ef)^2)) +
-      (El2 %*% t(Ef2) + fl_list$El2 %*% t(fl_list$Ef2))
-    sigmae2_v = Yhat^2 - 2* Yhat * (El%*%t(Ef) + fl_list$El%*%t(fl_list$Ef)) + fl_norm
-  }
-  return(sigmae2_v)
-}
-
-#' title one step update in flash iteration using ash
-#'
-#' description one step update in flash iteration using ash
-#' @return list of factor, loading and variance of noise matrix
-#'  \itemize{
-#'   \item{\code{El}} {is a N vector for mean of loadings}
-#'   \item{\code{El2}} {is a N vector for second moment of loadings}
-#'   \item{\code{Ef}} {is a N vector for mean of factors}
-#'   \item{\code{Ef2}} {is a N vector for second moment of factors}
-#'   \item{\code{sigmae2_v}}{is a N by P matrix for residual square}
-#'   \item{\code{sigmae2_true}}{is a N by P matrix for estimated value for the variance structure}
-#'   \item{\code{obj_val}}{the value of objectice function}
-#'  }
-#' @param Y the data matrix
-#' @param N dimension of Y
-#' @param P dimension of Y
-#' @param El mean for the loadings
-#' @param El2 second moment for the loadings
-#' @param Ef mean for the factors
-#' @param Ef2 second moment for the factors
-#' @param sigmae2_v residual square
-#' @param sigmae2_true the (true) known variance structure
-#' Here, sigmae2 is the estimated variance structure in each step
-#' sigmae2_true is the truth we know, some times sigmae2 is noisy version of sigmae2_true
-#' @param sigmae2 the estimation of the variance structure
-#' @param nonnegative if the facotor and loading are nonnegative or not.
-#' TRUE for nonnegative
-#' FALSE for no constraint
-#' @param partype parameter type for the variance,
-#' "constant" for constant variance,
-#' "var_col" for nonconstant variance for column,
-#' "known" for the kown variance,
-#' "Bayes_var" for Bayes version of the nonconstant variance for row and column
-#' "loganova" is anova estiamtion for the log residual square
-#' @param objtype  objective function type,
-#' "margin_lik" for conditional likelihood,
-#' "lowerbound_lik" for full objective function
-#' @param fix_factor whether the factor is fixed or not
-#' TRUE for fix_factor
-#' FALSE for non-constraint
-#' @keywords internal
-#'
-# one step update function
-one_step_update = function(Y, El, El2, Ef, Ef2,
-                           N, P,
-                           sigmae2_v, sigmae2_true,
-                           sigmae2,
-                           nonnegative = FALSE,
-                           partype = "constant",
-                           objtype = "margin_lik",
-                           fix_factor = FALSE,
-                           ash_para = list(),
-                           fl_list=list()){
-  # if fix_factor is True, please choose objtype = "margin_lik"
-  output = ifelse(objtype == "lowerbound_lik", "matrix", "mean")
-  # deal with the missing value
-  na_index_Y = is.na(Y)
-  is_missing = any(na_index_Y)  # keep the missing result
-  if(is_missing){
-    # print("missing value use the EY as missing Y")
-    Y[na_index_Y] = (El %*% t(Ef))[na_index_Y]
-  }
-  if(fix_factor){
-    # actually we need do nothing here
-    output = "mean"
-    objtype = "margin_lik"
-  }else{
-    # estimate the variance structure
-    # sigma_est = function(sigmae2_v,sigmae2_true,sigmae2_pre = NA, partype = "constant")
-    sigmae2 = sigma_est(sigmae2_v,sigmae2_true,sigmae2 ,partype)
-    # for the list case in kronecker product
-    if(is.list(sigmae2)){
-      # print("sigmae is kronecker product in estimating sigmae2 in one step update")
-      sigmae2 = sigmae2$sig2_l %*% t(sigmae2$sig2_f)
-    }
-    # Y = lf^T + E and ATM is for l given f, for f given l we need y^T = fl^T + E^T
-    # sigmae2_input = ifelse(is.matrix(sigmae2),t(sigmae2),sigmae2) is wrong here
-    if(is.matrix(sigmae2)){
-      # print("sigmae is matrix in estimating sigmae2 in one step update")
-      sigmae2_input = t(sigmae2)
-    }else{
-      sigmae2_input = sigmae2
-    }
-    par_f = ATM_r1(t(Y), El, El2, sigmae2_input,
-                   col_var = "column", nonnegative,
-                   output,partype,ash_para)
-    Ef = par_f$Ef
-    Ef2 = par_f$Ef2
-    # if the Ef is zeros ,just return zeros
-    if(sum(Ef^2) <= 1e-12){
-      El = rep(0,length(El))
-      Ef = rep(0,length(Ef))
-      sigmae2_v =  sigmae2_v_est(Y,El,Ef,El^2,Ef^2,fl_list)
-      sigmae2 = sigma_est(sigmae2_v,sigmae2_true,sigmae2 ,partype)
-      obj_val = obj(N, P, sigmae2_v, sigmae2, par_f=NA, par_l=NA, objtype="margin_lik")
-      return(list(El = El, El2 = El^2,
-                  Ef = Ef, Ef2 = Ef^2,
-                  sigmae2_v = sigmae2_v,
-                  sigmae2 = sigmae2,
-                  obj_val = obj_val))
-    }
-    # update the Y and the Y^2 in this if else block
-    if(is_missing){
-      # the missing varianece is just sigmae2_v + sigmae2
-      if(is.matrix(sigmae2)){
-        sigmae2_impute = sigmae2[na_index_Y]
-      }else if(is.vector(sigmae2) & length(sigmae2) == length(Ef)){
-        sigmae2_impute = (matrix(rep(sigmae2,each = N),ncol = P))[na_index_Y]
-      }else{
-        sigmae2_impute = sigmae2
-      }
-      sigmae2_v = sigmae2_v_est(Y,El,Ef,El2,Ef2,fl_list)
-      sigmae2_v[na_index_Y] = sigmae2_v[na_index_Y] + sigmae2_impute
-    }else{
-      sigmae2_v = sigmae2_v_est(Y,El,Ef,El2,Ef2,fl_list)
-    }
-  }
-
-  sigmae2 = sigma_est(sigmae2_v,sigmae2_true,sigmae2 ,partype)
-  if(is.list(sigmae2)){
-    # kronecker product
-    sigmae2 = sigmae2$sig2_l %*% t(sigmae2$sig2_f)
-  }
-  par_l = ATM_r1(Y, Ef, Ef2,
-                 sigmae2, col_var = "row",
-                 nonnegative, output,
-                 partype, ash_para)
-  El = par_l$Ef
-  El2 = par_l$Ef2
-  # if El is zeros just return
-  if(sum(El^2) <= 1e-12){
-    El = rep(0,length(El))
-    Ef = rep(0,length(Ef))
-    sigmae2_v = sigmae2_v_est(Y,El,Ef,El^2,Ef^2,fl_list)
-    sigmae2 = sigma_est(sigmae2_v,sigmae2_true,sigmae2 ,partype)
-    obj_val = obj(N, P, sigmae2_v, sigmae2, par_f=NA, par_l=NA, objtype="margin_lik")
-    return(list(El = El, El2 = El^2,
-                Ef = Ef, Ef2 = Ef^2,
-                sigmae2_v = sigmae2_v,
-                sigmae2 = sigmae2,
-                obj_val = obj_val))
-  }
-  if(is_missing){
-    # the missing varianece is just sigmae2_v + sigmae2
-    if(is.matrix(sigmae2)){
-      sigmae2_impute = sigmae2[na_index_Y]
-    }else if(is.vector(sigmae2) & length(sigmae2) == length(Ef)){
-      sigmae2_impute = (matrix(rep(sigmae2,each = N),ncol = P))[na_index_Y]
-    }else{
-      sigmae2_impute = sigmae2
-    }
-    sigmae2_v = sigmae2_v_est(Y,El,Ef,El2,Ef2,fl_list)
-    sigmae2_v[na_index_Y] = sigmae2_v[na_index_Y] + sigmae2_impute
-  }else{
-    sigmae2_v =  sigmae2_v_est(Y,El,Ef,El2,Ef2,fl_list)
-  }
-  #use the estiamtion as the truth
-  sigmae2 = sigma_est(sigmae2_v,sigmae2_true,sigmae2 ,partype)
-  obj_val = obj(N, P, sigmae2_v, sigmae2, par_f, par_l, objtype)
-
-  return(list(El = El, El2 = El2,
-              Ef = Ef, Ef2 = Ef2,
-              sigmae2_v = sigmae2_v,
-              sigmae2 = sigmae2,
-              obj_val = obj_val))
-}
-
-#' inital value for flash
-#'
-#' description inital value for flash
-#'
-#' @return list of factor, loading and variance of noise matrix
-#'  \itemize{
-#'   \item{\code{El}} {is a N vector for mean of loadings}
-#'   \item{\code{El2}} {is a N vector for second moment of loadings}
-#'   \item{\code{Ef}} {is a N vector for mean of factors}
-#'   \item{\code{Ef2}} {is a N vector for second moment of factors}
-#'   \item{\code{sigmae2_v}}{is a N by P matrix for residual square}
-#'   \item{\code{sigmae2_true}}{is a N by P matrix for estimated value for the variance structure}
-#'  }
-#' @param Y the data matrix
-#' @param nonnegative if the facotor and loading are nonnegative or not.
-#' TRUE for nonnegative
-#' FALSE for no constraint
-#' @param fix_factor whether the factor is fixed or not
-#' TRUE for fix_factor
-#' FALSE for non-constraint
-#' @param factor_value is the factor value if the factor is fixed
-#' @param fl_list this is a list for the El, El2, Ef and Ef2 for other factor loadings from the model.
-#'                in the flash for backfitting, we need those to get the residuals and estimation for variance
-#' @param initial_list_r1 this is a list for the inital values of current El Ef El2 and Ef2.
-#' @param fix_initial this is a indicator for the initial value is fix or not.
-#' @keywords internal
-#'
-initial_value = function(Y, nonnegative = FALSE,
-                         factor_value = NA,fix_factor = FALSE,
-                         initial_list_r1 = list(), fix_initial = FALSE,
-                         fl_list=list()){
-  # use the total mean as the estimated missing value
-  na_index_Y = is.na(Y)
-  is_missing = any(na_index_Y)
-  if(is_missing){
-    # this is the initialization for the missing value
-    Y[na_index_Y] = mean(Y, na.rm = TRUE)
-  }
-  # the flash with plug-in missing value
-  if(fix_initial){
-    El = initial_list_r1$El
-    Ef = initial_list_r1$Ef
-    El2 = initial_list_r1$El2
-    Ef2 = initial_list_r1$Ef2
-  }else if(fix_factor){
-    Ef = factor_value
-    Ef2 = Ef^2
-    El = as.vector( (Y %*% Ef) / (sum(Ef2)) )
-    El2 = El^2
-  }else{
-    # El = svd(Y)$u[,1]
-    El = as.vector(irlba::irlba(Y,nv = 0,nu = 1)$u)
-    # the nonnegative value need positive inital value
-    if(nonnegative){
-      # El = abs(El)
-      El = El * (El > 0)
-    }
-    El2 = El^2
-    Ef = as.vector(t(El)%*%Y)
-    Ef2 = Ef^2
-  }
-  # residual matrix initialization
-  # here we don't have any information about the sigmae2_true
-  if(is_missing){
-    # here we don't have any information about the variance and the Y as well
-    # since the previoust the Y is impute by the mean of observed Ys so I would use the mean of Y here as well.
-    # this initialization still need to be considered more.
-    sigmae2_v =  sigmae2_v_est(Y,El,Ef,El2,Ef2,fl_list)
-  }else{
-    sigmae2_v =  sigmae2_v_est(Y,El,Ef,El2,Ef2,fl_list)
-  }
-  return(list(El = El, El2 = El^2,
-              Ef = Ef, Ef2 = Ef^2,
-              sigmae2_v = sigmae2_v))
-}
-
-#' FLASH
-#'
-#' factor loading adaptive shrinkage rank one version
-#' @return list of factor, loading and variance of noise matrix
-#'  \itemize{
-#'   \item{\code{El}} {is a N vector for mean of loadings}
-#'   \item{\code{Ef}} {is a N vector for mean of factors}
-#'   \item{\code{sigmae2}}{is a N by P matrix for estimated value for the variance structure}
-#'  }
-#' @param Y the data matrix
-#' @param tol is for the tolerence for convergence in iterations and ash
-#' @param maciter_r1 is maximum of the iteration times for rank one case
-#' @param sigmae2_true true value for the variance structure
-#' @param nonnegative if the facotor and loading are nonnegative or not.
-#' TRUE for nonnegative
-#' FALSE for no constraint
-#' @param partype parameter type for the variance,
-#' "constant" for constant variance,
-#' "var_col" for nonconstant variance for column,
-#' "known" for the kown variance,
-#' "Bayes_var" for Bayes version of the nonconstant variance for row and column
-#' "loganova" is anova estiamtion for the log residual square
-#' @param objtype  objective function type,
-#' "margin_lik" for conditional likelihood,
-#' "lowerbound_lik" for full objective function
-#' @param fix_factor whether the factor is fixed or not
-#' TRUE for fix_factor
-#' FALSE for non-constraint
-#' @param factor_value is the factor value if the factor is fixed
-#' @param ash_para is the parameters list for ash
-#' @param fl_list is a list containing all the informations from other factors ans loadings
-#' l is loadings, f is factors, l2 and f2 are corresponding second moments.
-#' priorpost_vec is expectation of log piropost ratio
-#' clik is conditional likelihood (marginal likelihood)
-#'
-#' @details flash privide rank one matrix decomposition with variational EM algorithm.
-#'
-#' @export flash
-#'
-#' @importFrom ashr ash
-#'
-flash = function(Y, tol=1e-5, maxiter_r1 = 500,
-                 partype = c("constant","known","Bayes_var","var_col","noisy","noisy_col"),
-                 sigmae2_true = NA,
-                 factor_value = NA,fix_factor = FALSE,
-                 initial_list_r1 = list(), fix_initial = FALSE,
-                 nonnegative = FALSE,
-                 objtype = c("margin_lik","lowerbound_lik"),
-                 ash_para = list(),
-                 fl_list=list()){
-  # match the parameters
-  partype = match.arg(partype, c("constant","known","Bayes_var","var_col","noisy","noisy_col"))
-  objtype = match.arg(objtype, c("margin_lik","lowerbound_lik"))
-
-  # check the input
-  if( any(!is.na(sigmae2_true)) & !(partype %in% c("known","noisy","noisy_col")) ){
-    stop("You should choose partype = 'known' or 'noisy' ,'noisy_col',if you want to input the value of sigmae2_true")
-  }
-
-  N = dim(Y)[1]
-  P = dim(Y)[2]
-
-  # to get the inital values
-  g_initial = initial_value(Y, nonnegative,factor_value,fix_factor,initial_list_r1, fix_initial,fl_list)
-  El = g_initial$El
-  Ef = g_initial$Ef
-  El2 = g_initial$El2
-  Ef2 = g_initial$Ef2
-  sigmae2_v = g_initial$sigmae2_v
-
-  # start iteration
-  # in the first iteration, there is no value foe estimated sigmae2
-  g_update = one_step_update(Y, El, El2, Ef, Ef2,
-                             N, P,
-                             sigmae2_v, sigmae2_true,
-                             sigmae2 = NA,
-                             nonnegative ,
-                             partype ,
-                             objtype ,
-                             fix_factor,
-                             ash_para,
-                             fl_list)
-  # parameters updates
-  El = g_update$El
-  El2 = g_update$El2
-  Ef = g_update$Ef
-  Ef2 = g_update$Ef2
-  sigmae2_v = g_update$sigmae2_v
-  sigmae2 = g_update$sigmae2
-  obj_val = g_update$obj_val
-
-  # track the objective value
-  obj_val_track = c(obj_val)
-
-  # we should also return when the first run get all zeros
-  if(sum(El^2)==0 || sum(Ef^2)==0){
-    sigmae2 = sigma_est(sigmae2_v,sigmae2_true,sigmae2 ,partype)
-    if(is.list(sigmae2)){
-      # print("here using kronecker product")
-      sigmae2 = sigmae2$sig2_l %*% t(sigmae2$sig2_f)
-    }
-    # add one more output for greedy algorithm which not useful here
-    c_lik_val = C_likelihood(N,P,sigmae2_v,sigmae2)$c_lik
-    # the above value is not useful, but is helpful to get the postprior value
-    # since obj_val = c_lik_value + priorpost_l + priorpost_f
-    return(list(l = El, f = Ef, l2 = El2, f2 = Ef2,
-                sigmae2 = sigmae2,
-                obj_val = obj_val,
-                c_lik_val = c_lik_val))
-  }
-
-  epsilon = 1
-  tau = 1
-  while(epsilon >= tol & tau < maxiter_r1){
-    tau = tau + 1
-    pre_obj = obj_val
-
-    g_update = one_step_update(Y, El, El2, Ef, Ef2,
-                               N, P,
-                               sigmae2_v, sigmae2_true,
-                               sigmae2,
-                               nonnegative ,
-                               partype ,
-                               objtype ,
-                               fix_factor,
-                               ash_para,
-                               fl_list)
-    # parameters updates
-    El = g_update$El
-    El2 = g_update$El2
-    Ef = g_update$Ef
-    Ef2 = g_update$Ef2
-    sigmae2_v = g_update$sigmae2_v
-    sigmae2 = g_update$sigmae2
-    obj_val = g_update$obj_val
-
-    if(sum(El^2)==0 || sum(Ef^2)==0){
-      El = rep(0,length(El))
-      Ef = rep(0,length(Ef))
-      break
-    }
-    epsilon = abs(pre_obj - obj_val)
-    obj_val_track = c(obj_val_track,obj_val)
-  }
-  sigmae2 = sigma_est(sigmae2_v,sigmae2_true,sigmae2 ,partype)
-  if(is.list(sigmae2)){
-    # print("here, using kronecker product")
-    sigmae2 = sigmae2$sig2_l %*% t(sigmae2$sig2_f)
-  }
-  # add one more output for greedy algorithm which not useful here
-  c_lik_val = C_likelihood(N,P,sigmae2_v,sigmae2)$c_lik
-  # the above value is not useful, but is helpful to get the postprior value
-  # since obj_val = c_lik_value + priorpost_l + priorpost_f
-  return(list(l = El, f = Ef, l2 = El2, f2 = Ef2,
-              sigmae2 = sigmae2,
-              obj_val = obj_val,
-              c_lik_val = c_lik_val,
-              obj_val_track = obj_val_track))
-}
-
-```
-
-```{r}
-sim_K = function(K, N, P, SF, SL, signal,noise){
-  E = matrix(rnorm(N*P,0,noise),nrow=N)
-  Y = E
-  L_true = array(0, dim = c(N,K))
-  F_true = array(0, dim = c(P,K))
-
-  for(k in 1:K){
-    lstart = rnorm(N, 0, signal)
-    fstart = rnorm(P, 0, signal)
-
-    index = sample(seq(1:N),(N*SL))
-    lstart[index] = 0
-    index = sample(seq(1:P),(P*SF))
-    fstart[index] = 0
-
-    L_true[,k] = lstart
-    F_true[,k] = fstart
-
-    Y = Y + lstart %*% t(fstart)
-  }
-  return(list(Y = Y, L_true = L_true, F_true = F_true, Error = E))
-}
-```
-
-we check that the new objective function with penalty term does monotonically increase.
-
-```{r}
-set.seed(99)
-K = 1
-N = 20
-P = 30
-SF = 0.5
-SL = 0.6
-signal = 1
-noise = 1
-par(mfrow = c(2, 2))
-par(cex = 0.6)
-par(mar = c(3, 3, 0.8, 0.8), oma = c(1, 1, 1, 1))
-data = sim_K(K,N, P, SF , SL , signal ,noise )
-Y = data$Y
-E = data$Error
-g1 = flash(Y,objtype = "l")
-plot(g1$obj_val_track,main = "fdr method with penalized obj",type = "l")
-g2 = flashr::flash(Y,objtype = "l")
-plot(g2$obj_val_track,main = "fdr method without penalized obj",type = "l")
-```
-
-we can see that the old version of code without penalty doesn't have monotonically increasing objective function, but the new version code with penalty term does.
-
-## try the same problem: rank 0 vs rank 1 with  fdr method.
-
-```{r,echo = FALSE}
-
-rank0check = function(K=1,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 0,noise = 1, mtype = "fdr"){
-  data = sim_K(K,N, P, SF , SL , signal ,noise )
-  Y = data$Y
-  E = data$Error
-  if(mtype == "fdr"){
-    g1 = flash(Y,objtype = "l")
-  }else{
-    g1 = flash(Y,objtype = "l",ash_para = list(method = "shrink"))
-  }
-
-  RMSE = sqrt(mean((Y - g1$l %*% t(g1$f) -E)^2)) 
-  obj_val = g1$obj_val
-  # to get the rank zero objective value
-  sigmae2_v0 = Y^2
-  fit_g0 = list(pi=c(1,0,0),sd = c(0,0.1,1))
-  mat0 = list(comp_postmean = matrix(0,ncol= N,nrow = 3),
-              comp_postmean2 = matrix(0,ncol= N,nrow = 3),
-              comp_postprob = matrix(0,ncol= N,nrow = 3))
-  mat0$comp_postprob[1,] = 1
-  par_l0 = list(g = fit_g0, mat = mat0)
-  fit_g0 = list(pi=c(1,0,0),sd = c(0,0.1,1))
-  mat0 = list(comp_postmean = matrix(0,ncol= P,nrow = 3),
-              comp_postmean2 = matrix(0,ncol= P,nrow = 3),
-              comp_postprob = matrix(0,ncol= P,nrow = 3))
-  mat0$comp_postprob[1,] = 1
-  par_f0 = list(g = fit_g0, mat = mat0)
-  obj0 = obj(N,P,sigmae2_v0,sigmae2 = mean(sigmae2_v0),par_f0,par_l0,objtype = "lowerbound_lik")
-
-  return(list(rmse = RMSE,obj_val = obj_val,obj_0 = obj0))
-}
-
-sim_rank0 = function(K=1,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 0,noise = 1,mtype = "fdr"){
-  T = 200
-  RMSE = rep(NA,T)
-  sign_obj = rep(NA,T)
-  for(t in 1:T){
-    g0 = rank0check(K = 1,N, P, SF, SL , signal,noise,mtype = mtype)
-    RMSE[t] = g0$rmse
-    sign_obj[t] = sign(g0$obj_val - g0$obj_0)
-  }
-  par(mfrow = c(2, 2))
-  par(cex = 0.6)
-  par(mar = c(3, 3, 0.8, 0.8), oma = c(1, 1, 1, 1))
-  boxplot(RMSE,main = "RMSE")
-  hist(sign_obj, breaks = 3,main = "sign of difference of obj value")
-  hist(RMSE,breaks = 50,main = "RMSE")
-  plot(sign_obj,RMSE, main = "sign of obj diff vs RMSE")
-}
-
-#### old code
-
-rank0check_old = function(K=1,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 0,noise = 1, mtype = "shrink"){
-  data = sim_K(K,N, P, SF , SL , signal ,noise )
-  Y = data$Y
-  E = data$Error
-  if(mtype == "fdr"){
-    g1 = flashr::flash(Y,objtype = "l")
-  }else{
-    g1 = flashr::flash(Y,objtype = "l",ash_para = list(method = "shrink"))
-  }
-
-  RMSE = sqrt(mean((Y - g1$l %*% t(g1$f) -E)^2)) 
-  obj_val = g1$obj_val
-  # to get the rank zero objective value
-  sigmae2_v0 = Y^2
-  fit_g0 = list(pi=c(1,0,0),sd = c(0,0.1,1))
-  mat0 = list(comp_postmean = matrix(0,ncol= N,nrow = 3),
-              comp_postmean2 = matrix(0,ncol= N,nrow = 3),
-              comp_postprob = matrix(0,ncol= N,nrow = 3))
-  mat0$comp_postprob[1,] = 1
-  par_l0 = list(g = fit_g0, mat = mat0)
-  fit_g0 = list(pi=c(1,0,0),sd = c(0,0.1,1))
-  mat0 = list(comp_postmean = matrix(0,ncol= P,nrow = 3),
-              comp_postmean2 = matrix(0,ncol= P,nrow = 3),
-              comp_postprob = matrix(0,ncol= P,nrow = 3))
-  mat0$comp_postprob[1,] = 1
-  par_f0 = list(g = fit_g0, mat = mat0)
-  obj0 = obj(N,P,sigmae2_v0,sigmae2 = mean(sigmae2_v0),par_f0,par_l0,objtype = "lowerbound_lik")
-
-  return(list(rmse = RMSE,obj_val = obj_val,obj_0 = obj0))
-}
-
-sim_rank0_old = function(K=1,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 0,noise = 1,mtype = "shrink"){
-  T = 200
-  RMSE = rep(NA,T)
-  sign_obj = rep(NA,T)
-  for(t in 1:T){
-    g0 = rank0check_old(K = 1,N, P, SF, SL , signal,noise,mtype = mtype)
-    RMSE[t] = g0$rmse
-    sign_obj[t] = sign(g0$obj_val - g0$obj_0)
-  }
-  par(mfrow = c(2, 2))
-  par(cex = 0.6)
-  par(mar = c(3, 3, 0.8, 0.8), oma = c(1, 1, 1, 1))
-  boxplot(RMSE,main = "RMSE")
-  hist(sign_obj, breaks = 3,main = "sign of difference of obj value")
-  hist(RMSE,breaks = 50,main = "RMSE")
-  plot(sign_obj,RMSE, main = "sign of obj diff vs RMSE")
-}
-```
-
-#### case 1: rank 0
-
-we compare the old code with our new code, we can see that most of the cases a
-
-*shrink method without penalty*
-
-```{r}
-set.seed(99)
-sim_rank0_old(K=1,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 0,noise = 1)
-```
-
-*fdr method with penalty*
-
-```{r}
-set.seed(99)
-sim_rank0(K=1,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 0,noise = 1)
-```
-
-we can see the improvement that the proportion of P(estimated structure better than rank 0 structure) is much larger.
-
-#### case 2: rank 1, very sparse with large noise
-
-*shrink method without penalty*
-
-```{r}
-set.seed(99)
-sim_rank0_old(K=1,N = 100, P=200, SF = 0.8, SL = 0.8, signal = 1,noise = 2)
-```
-
-*fdr method with penalty*
-
-```{r}
-set.seed(99)
-sim_rank0(K=1,N = 100, P=200, SF = 0.8, SL = 0.8, signal = 1,noise = 2)
-```
-
-we can see that the results are improved since the proportion of P(estimated structure better than rank 0 structure) is much larger.
-
 ## Session information

### 2017-06-20 analysis/index.Rmd
Files commited by wflow_commit.

@@
 Click on this [load reslut of GTEx data](GTExresult.html) to see more details.
+
+Click on this [pairwise plot](pairwiseplotofflashgtexzscore.html) to see more details.
+
+### Issues
+
+Click on this [gap issue](ObjGap.html) to see more details. (test)
+
+Click on this [summary of the rank0 vs rank1](Obj_summary.html) to see more details.
+
+- Click on this [rank 0 vs rank 1](Objcheck_1.html) to see more details.
+- Click on this [rank 0 vs rank 1: fdr with penalty](Objcheck_2.html) to see more details.
+
+### meeting summary
+
+[04-26](meetnotes.html)

### 2017-06-20 analysis/Obj_summary.Rmd
Files commited by wflow_commit.

@@
 
+## rank 0 vs rank 1
+
+this document is based on Matthew's suggestion:
+
+> If you simulate rank 0, and run greedy with K=1:
+>
+> - how often does the K=1 converge to a point mass at 0?
+>
+> - how often does the K=1 converge to a lower bound that is worse than the lower bound for K=0?
+>
+> Then you could do the same question for method =fdr (but with this penalized objective function).
+
+### shrink method without penalty term in objective function
+
+we find that we might need fdr method and we should use the objective funciton with penalty.[rank 0 vs rank 1](Objcheck_1.html)
+
+### check the objective function with penalty
+
+In this case, we check the objective funciton with penalty term, and it does monotonically increase.
+
+### fdr method with penalty term in objective function
+
+we can see the improvement that the proportion of P(estimated structure better than rank 0 structure) is much larger. [rank 0 vs rank 1: fdr with penalty](Objcheck_2.html)
+
 ## Session information

### 2017-06-20 analysis/Objcheck_2.Rmd
Files commited by wflow_commit.

@@
   RMSE = rep(NA,T)
-  sign_obj = rep(NA,T)
+  diff_obj = rep(NA,T)
   for(t in 1:T){
@@
     RMSE[t] = g0$rmse
-    sign_obj[t] = sign(g0$obj_val - g0$obj_0)
+    diff_obj[t] = (g0$obj_val - g0$obj_0)
   }
@@
   boxplot(RMSE,main = "RMSE")
-  hist(sign_obj, breaks = 3,main = "sign of difference of obj value")
+  hist(diff_obj, breaks = 3,main = "sign of difference of obj value")
   hist(RMSE,breaks = 50,main = "RMSE")
-  plot(sign_obj,RMSE, main = "sign of obj diff vs RMSE")
+  plot(diff_obj,RMSE, main = "sign of obj diff vs RMSE")
+  return(c(sum(diff_obj == 0),sum(diff_obj>0),diff_obj<0)/T)
 }

### 2017-06-20 analysis/Objcheck_2.Rmd
Files commited by wflow_commit.

--- 2017-06-20 analysis/Objcheck_2.Rmd
@@
   boxplot(RMSE,main = "RMSE")
-  hist(sign_obj, breaks = 3,main = "sign of difference of obj value")
+  hist(diff_obj, breaks = 3,main = "sign of difference of obj value")
   hist(RMSE,breaks = 50,main = "RMSE")
-  plot(sign_obj,RMSE, main = "sign of obj diff vs RMSE")
+  plot(diff_obj,RMSE, main = "sign of obj diff vs RMSE")
+  return(c(sum(diff_obj == 0),sum(diff_obj>0),sum(diff_obj<0))/T)
 }

### 2017-06-20 analysis/Objcheck_2.Rmd
Files commited by wflow_commit.

@@
 ```{r,echo=FALSE}
-# ATM function
-#' title ash type model for l
-#'
-#' description use ash type model to maxmization
-#'
-#' @return list of factor, loading and variance of noise matrix
-#'  \itemize{
-#'   \item{\code{El}} {is a N vector for mean of loadings}
-#'   \item{\code{El2}} {is a N vector for second moment of loadings}
-#'  }
-#' @param Y is data matrix
-#' @param Ef is mean for the factor
-#' @param Ef2 is second moment for the factor
-#' @param sigmae2 is variance structure for noise matrix
-#' @param col_var is just for column variance, column is for column, row is for row
-#' @param nonnegative is flag for whether output nonnegative value
-#' @param output is format of output, "mean" is for mean value, "matrix" is for flash data matrix in ash
-#'
-#' @keywords internal
-#'
-
-ATM_r1 = function(Y, Ef, Ef2,
-                  sigmae2, col_var = "row",
-                  nonnegative = FALSE, output = "mean",
-                  partype = "constant",ash_para = list()){
-  # this part is preparing betahat and sebeta for ash
-  if(is.matrix(sigmae2)){
-    # print("sigmae2 is a matrix")
-    # this is for all variance are different
-    sum_Ef2 = (1/sigmae2) %*% Ef2
-    sum_Ef2 = as.vector(sum_Ef2)
-    sebeta = sqrt(1/(sum_Ef2))
-    betahat = as.vector( (Y/sigmae2) %*% Ef ) / (sum_Ef2)
-    betahat=as.vector(betahat)
-  } else if(is.vector(sigmae2) & length(sigmae2) == length(Ef)){
-    # this is for the non-constant variance in column
-    # print("sigmae2 is vector")
-    if(col_var == "row"){
-      sum_Ef2 = (1/sigmae2) * Ef2
-      sum_Ef2 = sum(sum_Ef2)
-      sebeta = sqrt(1/(sum_Ef2))
-      betahat = as.vector( Y %*% (Ef/sigmae2) ) / (sum_Ef2)
-      betahat=as.vector(betahat)
-    }else {
-      # for column
-      # here I have alread change the dimension by taking transpose to Y
-      sum_Ef2 = sum(Ef2)
-      sebeta = sqrt( sigmae2/(sum_Ef2) )
-      betahat = (t(Ef) %*% t(Y)) / (sum_Ef2)
-      betahat=as.vector(betahat)
-    }
-  }else{
-    # for constant case
-    # print("sigmae2 is a constant")
-    sum_Ef2 = sum(Ef2)
-    sebeta = sqrt(sigmae2/(sum_Ef2))
-    betahat = (t(Ef) %*% t(Y)) / (sum_Ef2)
-    betahat=as.vector(betahat)
-  }
-  # set ash default setting up
-  ash_default = list(betahat = betahat, sebeta = sebeta,
-                     method = "fdr", mixcompdist = "normal")
-  # ATM update
-  # decide the sign for output
-  if(nonnegative){
-    ash_default$mixcompdist = "+uniform"
-  }
-  # decide the output to decide the convergence criterion
-  if(output == "matrix"){
-    ash_para$outputlevel = 4
-    ash_para = modifyList(ash_default,ash_para)
-    ATM = do.call(ashr::ash,ash_para)
-    # ATM = ash(betahat, sebeta, method="fdr", mixcompdist=mixdist,outputlevel=4)
-    Ef = ATM$result$PosteriorMean
-    SDf = ATM$result$PosteriorSD
-    Ef2 = SDf^2 + Ef^2
-    mat_post = ATM$flash_data
-    fit_g = ATM$fitted_g
-    return(list(Ef = Ef,
-                Ef2 = Ef2,
-                mat = mat_post,
-                g = fit_g))
-  } else {
-    ash_para = modifyList(ash_default,ash_para)
-    ATM = do.call(ashr::ash,ash_para)
-    # ATM = ash(betahat, sebeta, method="fdr", mixcompdist=mixdist)
-    Ef = ATM$result$PosteriorMean
-    SDf = ATM$result$PosteriorSD
-    Ef2 = SDf^2 + Ef^2
-    return(list(Ef = Ef, Ef2 = Ef2))
-  }
-
-}
-
-#' title prior and posterior part in objective function
-#'
-#' description prior and posterior part in objective function
-#'
-#' @return PrioPost the value for the proir and posterior in objectice function
-#'  \itemize{
-#'   \item{\code{PrioPost}} {the value for the proir and posterior in objectice function}
-#'   \item{\code{penalty}} {penalty term of the g_l or g_f}
-#'  }
-#' @param mat matrix of flash.data in ash output which is for posterior
-#' @param fit_g in fitted.g in ash output which is for prior
-#' @keywords internal
-#'
-# two parts for the likelihood
-Fval = function(mat,fit_g){
-  # make sure the first row included in the index
-  nonzeroindex = unique(c(1,which(fit_g$pi != 0)))
-  #prepare for the posterior part
-  mat_postmean = mat$comp_postmean[nonzeroindex,]
-  mat_postmean2 = mat$comp_postmean2[nonzeroindex,]
-  mat_postvar = mat_postmean2 - mat_postmean^2
-  mat_postprob = mat$comp_postprob[nonzeroindex,]
-  # figure our the dimension
-  K = dim(mat_postprob)[1]
-  N = dim(mat_postprob)[2]
-  if(is.vector(mat_postprob)){
-    K = 1
-    N = length(mat_postprob)
-  }
-  # prepare for the prior
-  prior_pi = fit_g$pi[nonzeroindex]
-  prior_var = (fit_g$sd[nonzeroindex])^2
-  mat_priorvar = matrix(rep(prior_var,N),ncol = N)
-  mat_priorprob = matrix(rep(prior_pi,N),ncol = N)
-  # to get the value
-  varodds = mat_priorvar / mat_postvar
-  varodds[1,] = 1 # deal with 0/0
-  probodds = mat_priorprob / mat_postprob
-  ssrodds = mat_postmean2 / mat_priorvar
-  ssrodds[1,] = 1  # deal with 0/0
-  priorpost = mat_postprob * (log(probodds) - (1/2)*log(varodds) - (1/2)* (ssrodds -1))
-  # in case of mat_postprob = 0
-  priorpost[which(mat_postprob< 1e-100)] = 0
-  PrioPost = sum(priorpost)
-  #######
-  # now we try to get the penalty term
-  # I use the default of setting for lambda_k in https://oup.silverchair-cdn.com/oup/backfile/Content_public/Journal/biostatistics/18/2/10.1093_biostatistics_kxw041/1/kxw041_Supp.pdf?Expires=1498001808&Signature=dtxDDFmYZ1u1jFoXfNg4mqGQnTPkQQMFZgt2-FE9r8GcgRl4DdEaP0oI6kzvUjZB5H3hLZmvAZfuFkt3GxojQJ1eX0y7I9kMJC75VTrxw1Ym~psIe8rdU5Xw4S4KZm79o1cOXBzY1iAORFkVM4z59uP1T1ltFJDu5fhlTYErU3flbiE1ivTm-BYN5P7w8dex-R5MWn98NpMVRUsoHRgbe7FhwavdnQ8QAD7~q1F4DzMyZk09TeNdEc7GNdBKbSOoS9mGGnvrqXvPz1nn~QzJ2U2CMCf04GfsgkclyUnH0aD6Vh1lSAfoEV47ZnT-9~942~f8rMs30H2Kqji7BjlctA__&Key-Pair-Id=APKAIUCZBIA4LVPAVW3Q
-  h_value = (10 - 1) * log(fit_g$pi[1])
-  return(list(PrioPost = PrioPost, penalty = h_value))
-}
-
-#' title conditional likelihood
-#'
-#' description conditional likelihood in objective function
-#'
-#' @return c_lik for the onditional likelihood in objectice function
-#' @param N  dimension of residual matrix
-#' @param P  dimension of residual matrix
-#' @param sigmae2_v  residual matrix
-#' @param sigmae2  variance structure of error
-#' @keywords internal
-#'
-# this version we need to know the truth of sigmae2, we can use sigmae2_v as the truth
-C_likelihood = function(N,P,sigmae2_v,sigmae2){
-  if(is.matrix(sigmae2)){
-    # print("clik as matrix of sigmae2")
-    c_lik = -(1/2) * sum( log(2*pi*sigmae2) + (sigmae2_v)/(sigmae2) )
-  }else if(is.vector(sigmae2) & length(sigmae2)==P){
-    # print("clik as vector sigmae2")
-    # change the format to fit the conditional likelihood
-    sigmae2_v = colMeans(sigmae2_v)
-    c_lik = -(N/2) * sum( log(2*pi*sigmae2) + (sigmae2_v)/(sigmae2) )
-  } else {
-    # print("clik as sigmae2 is constant")
-    # change the format to fit the conditional likelihood and accelerate the computation.
-    sigmae2_v = mean(sigmae2_v)
-    c_lik = -(N*P)/2 * ( log(2*pi*sigmae2) + (sigmae2_v)/(sigmae2) )
-    # here I want to use fully variantional inference Elogsigmae2 rather logEsigmae2
-    # c_lik = -(N*P)/2 * ( log(2*pi*sigmae2_true) + (sigmae2_v)/(sigmae2_true) + log(N*P/2) - digamma(N*P/2) )
-  }
-  return(list(c_lik = c_lik))
-}
-
-#' title objective function in VEM
-#'
-#' description  objective function
-#'
-#' @return obj_val value of the objectice function
-#' @param N  dimension of residual matrix
-#' @param P  dimension of residual matrix
-#' @param sigmae2_v  residual matrix
-#' @param sigmae2_true  true variance structure (we use the estimated one to replace that if the truth is unknown)
-#' @param par_l ash output for l
-#' @param par_f ash output for f
-#' @param objtype  objective function type,
-#' "margin_lik" for conditional likelihood,
-#' "lowerbound_lik" for full objective function
-#' @keywords internal
-#'
-
-# objective function
-obj = function(N,P,sigmae2_v,sigmae2,par_f,par_l,objtype = "margin_lik"){
-  if(is.list(sigmae2)){
-    # print("obj using kronecker product")
-    sigmae2 = sigmae2$sig2_l %*% t(sigmae2$sig2_f)
-  }
-  if(objtype=="lowerbound_lik"){
-    priopost_f = Fval(par_f$mat, par_f$g)$PrioPost
-    priopost_l = Fval(par_l$mat, par_l$g)$PrioPost
-    penalty_f = Fval(par_f$mat, par_f$g)$penalty
-    penalty_l = Fval(par_l$mat, par_l$g)$penalty
-    c_lik = C_likelihood(N,P,sigmae2_v,sigmae2)$c_lik
-    obj_val = c_lik + priopost_l + priopost_f + penalty_l + penalty_f
-  } else {
-    obj_val = C_likelihood(N,P,sigmae2_v,sigmae2)$c_lik
-  }
-  return(obj_val)
-}
-
-#' title rescale the lambda_l and lambda_f for identifiablity
-#'
-#' description rescale the lambda_l and lambda_f for identifiablity
-#'
-#' @return a list of sig2_l and sig2_f for the rescaled variance of the kronecker product
-#' @param sig2_l variance of the kronecker product
-#' @param sig2_l variance of the kronecker product
-#' @keywords internal
-#'
-rescale_sigmae2 = function(sig2_l,sig2_f){
-  norm_l = sqrt(sum(sig2_l^2))
-  norm_f = sqrt(sum(sig2_f^2))
-  norm_total = norm_l * norm_f
-  sig2_l = sig2_l / norm_l
-  sig2_f = sig2_f / norm_f
-  sig2_l = sig2_l * sqrt(norm_total)
-  sig2_f = sig2_f * sqrt(norm_total)
-  return(list(sig2_l = sig2_l,sig2_f = sig2_f))
-}
-
-#' title initial value for Bayes variance structure estimation for kronecker productor
-#'
-#' description initial value for Bayes variance structure estimation for kronecker productor
-#'
-#' @return sig2_out estimated variance structure
-#' @param sigmae2_v  residual matrix
-#' @keywords internal
-#'
-inital_Bayes_var = function(sigmae2_v){
-  N = dim(sigmae2_v)[1]
-  P = dim(sigmae2_v)[2]
-  # estimate the initial value of lambda_l and lambda_f
-  sig2_l_pre = rowMeans(sigmae2_v)
-  sig2_f_pre = colMeans( sigmae2_v / matrix(rep(sig2_l_pre,P), ncol = P) )
-  sig2_pre_list = rescale_sigmae2(sig2_l_pre,sig2_f_pre)
-  sig2_l_pre = sig2_pre_list$sig2_l
-  sig2_f_pre = sig2_pre_list$sig2_f
-  # start the iteration
-  maxiter = 100
-  inital_tol = 1e-3
-  tau = 0
-  epsilon = 1
-  while(epsilon > inital_tol & tau <= maxiter){
-    tau = tau + 1
-    sig2_l = rowMeans( sigmae2_v / matrix(rep(sig2_f_pre,each = N),ncol = P) )
-    sig2_f = colMeans( sigmae2_v / matrix(rep(sig2_l,P), ncol = P) )
-    sig2_list = rescale_sigmae2(sig2_l,sig2_f)
-    sig2_l = sig2_list$sig2_l
-    sig2_f = sig2_list$sig2_f
-    epsilon = sqrt(mean((sig2_f - sig2_f_pre)^2 )) + sqrt(mean((sig2_l - sig2_l_pre)^2))
-    sig2_l_pre = sig2_l
-    sig2_f_pre = sig2_f
-  }
-  return(list(sig2_l = sig2_l,sig2_f = sig2_f))
-}
-
-#' title Bayes variance structure estimation for kronecker productor
-#'
-#' description prior and posterior part in objective function
-#'
-#' @return sig2_out estimated variance structure
-#' @param sigmae2_v  residual matrix
-#' @param sigmae2 variance structure
-#' and it is a list of two vectors in this case.
-#' @keywords internal
-#'
-Bayes_var = function(sigmae2_v,sigmae2 = NA){
-  N = dim(sigmae2_v)[1]
-  P = dim(sigmae2_v)[2]
-  if( is.na(sigmae2) || !is.list(sigmae2) ){
-    # we don't know the truth this is in the first iteration
-    sigmae2 = inital_Bayes_var(sigmae2_v)
-  }
-  sig2_l_pre = sigmae2$sig2_l
-  sig2_f_pre = sigmae2$sig2_f
-  # this has already beed rescaled
-  # here we use alpha_l = alpha_f = beta_l = beta_f = 0
-  sig2_l = rowMeans( sigmae2_v / matrix(rep(sig2_f_pre,each = N),ncol = P) )
-  sig2_f = colMeans( sigmae2_v / matrix(rep(sig2_l,P), ncol = P) )
-  #rescaled the variance
-  sig2_list = rescale_sigmae2(sig2_l,sig2_f)
-  sig2_l = sig2_list$sig2_l
-  sig2_f = sig2_list$sig2_f
-  # sig2_out = matrix(rep(sig2_l,P),ncol = P) * matrix(rep(sig2_f,each = N),ncol = P)
-  return(list(sig2_l = sig2_l,sig2_f = sig2_f))
-}
-
-#' title noisy variance structure estimation
-#'
-#' description prior and posterior part in objective function
-#'
-#' @return sig2_out estimated variance structure
-#' @param sigmae2_v  residual matrix
-#' @param sigmae2_true variance structure
-#' and it is a list of two vectors in this case.
-#' @keywords internal
-#'
-
-noisy_var = function(sigmae2_v,sigmae2_true){
-  # a reaonable upper bound which control optimal algorithm
-  upper_range = abs(mean(sigmae2_v - sigmae2_true)) + sd(sigmae2_v - sigmae2_true)
-  # value of likelihood
-  # f_lik <- function(x,sigmae2_v_l = sigmae2_v,sigmae2_true_l = sigmae2_true){
-  #  -(1/2)*sum( log(2*pi*(x+sigmae2_true_l)) + (1/(x+sigmae2_true_l))*sigmae2_v_l )
-  # }
-  # I need the negative of the f_lik since optim find the minimum rather than the maximum
-  f_lik <- function(x,sigmae2_v_l = sigmae2_v,sigmae2_true_l = sigmae2_true){
-    (1/2)*mean( log(2*pi*(x+sigmae2_true_l)) + (1/(x+sigmae2_true_l))*sigmae2_v_l )
-  }
-  AA <- optim(mean(sigmae2_v - sigmae2_true), f_lik,lower = 0, upper = upper_range,  method = "Brent")
-  e_sig = AA$par
-  return(e_sig + sigmae2_true)
-}
-
-#' title noisy variance structure estimation with noisy variance on each column
-#'
-#' description prior and posterior part in objective function
-#'
-#' @return sig2_out estimated variance structure
-#' @param sigmae2_v  residual matrix
-#' @param sigmae2_true variance structure
-#' and it is a list of two vectors in this case.
-#' @keywords internal
-#'
-
-noisy_var_column = function(sigmae2_v,sigmae2_true){
-  # in this case, we need to think about the for each column
-  # we can use noisy_var = function(sigmae2_v,sigmae2_true) with the input of each column
-  P = dim(sigmae2_true)[2]
-  sig2_out = sapply(seq(1,P),function(x){noisy_var(sigmae2_v[,x],sigmae2_true[,x])})
-  return(sig2_out)
-}
-
-#' title module for estiamtion of the variance structure
-#'
-#' description estiamtion of the variance structure
-#'
-#' @return sigmae2 estimated variance structure
-#' @param partype parameter type for the variance,
-#' "constant" for constant variance,
-#' "var_col" for nonconstant variance for column,
-#' "known" for the kown variance,
-#' "Bayes_var" for Bayes version of the nonconstant variance for row and column
-#' "noisy" is noisy version s_ij + sigmae2
-#' "noisy_col" noisy version for column s_ij + sigmae2_j
-#' @param sigmae2_v  residual matrix
-#' @param sigmae2_true  true variance structure
-#' @param sigmae2_pre the previouse sigmae2 which is only used in Bayes_var
-#' @keywords internal
-#'
-# sigma estimation function
-sigma_est = function(sigmae2_v,sigmae2_true,sigmae2_pre = NA, partype = "constant"){
-  if(partype == "var_col"){
-    # print("use var col method")
-    sigmae2 = colMeans(sigmae2_v)
-  } else if(partype == "noisy"){
-    # print("use the noisy version")
-    sigmae2 = noisy_var(sigmae2_v,sigmae2_true)
-  } else if(partype == "noisy_col"){
-    sigmae2 = noisy_var_column(sigmae2_v,sigmae2_true)
-  }else if (partype == "Bayes_var"){
-    # here sigmae2_true is a list
-    # print("use the kronecker product")
-    sigmae2 = Bayes_var(sigmae2_v,sigmae2_pre)
-  }else if (partype == "known"){
-    # print("use the known sigmae")
-    sigmae2 = sigmae2_true
-  } else {
-    # print("use the mean value for sigmae")
-    # this is for constant case
-    sigmae2 = mean(sigmae2_v)
-  }
-  return(sigmae2)
-}
-
-#' title empirical sample variance matrix estimation
-#'
-#' description this function is to calculate the empirical sample variance matrix sigmae2_v
-#'
-#' @return sigmae2_v  empirical sample variance matrix
-#' @param Y which is the residual
-#' @param fl_list this is a list for the El Ef El2 and Ef2 from other factors
-#' @param El mean estimation for the current loading
-#' @param Ef mean estimation for the current factor
-#' @param El2 second momnet estimation for the current factor
-#' @param Ef2 second moment estimation for the current loading
-#' @keywords internal
-#'
-sigmae2_v_est = function(Y,El,Ef,El2,Ef2,fl_list=list()){
-  # in this version of function we haven't included the missing value case in
-  if(length(fl_list)==0){
-    # this is for the rank one case since there are no other factors
-    # here Y is just the original data
-    sigmae2_v =  Y^2 - 2*Y*(El %*% t(Ef)) + (El2 %*% t(Ef2))
-  }else{
-    # this is the for the rank more than one we have other factors
-    # now the Y is residual matrix rather the original matrix
-    # to get the original data we use Yhat
-    Yhat = Y + fl_list$El %*% t(fl_list$Ef)
-    # the residual matrix should be
-    # this is for E(l_1f_1+l_2f_2+...+l_kf_k)^2
-    fl_norm = (El%*%t(Ef) + fl_list$El%*%t(fl_list$Ef))^2 -
-      (El^2 %*% t(Ef^2) + (fl_list$El)^2 %*% t((fl_list$Ef)^2)) +
-      (El2 %*% t(Ef2) + fl_list$El2 %*% t(fl_list$Ef2))
-    sigmae2_v = Yhat^2 - 2* Yhat * (El%*%t(Ef) + fl_list$El%*%t(fl_list$Ef)) + fl_norm
-  }
-  return(sigmae2_v)
-}
-
-#' title one step update in flash iteration using ash
-#'
-#' description one step update in flash iteration using ash
-#' @return list of factor, loading and variance of noise matrix
-#'  \itemize{
-#'   \item{\code{El}} {is a N vector for mean of loadings}
-#'   \item{\code{El2}} {is a N vector for second moment of loadings}
-#'   \item{\code{Ef}} {is a N vector for mean of factors}
-#'   \item{\code{Ef2}} {is a N vector for second moment of factors}
-#'   \item{\code{sigmae2_v}}{is a N by P matrix for residual square}
-#'   \item{\code{sigmae2_true}}{is a N by P matrix for estimated value for the variance structure}
-#'   \item{\code{obj_val}}{the value of objectice function}
-#'  }
-#' @param Y the data matrix
-#' @param N dimension of Y
-#' @param P dimension of Y
-#' @param El mean for the loadings
-#' @param El2 second moment for the loadings
-#' @param Ef mean for the factors
-#' @param Ef2 second moment for the factors
-#' @param sigmae2_v residual square
-#' @param sigmae2_true the (true) known variance structure
-#' Here, sigmae2 is the estimated variance structure in each step
-#' sigmae2_true is the truth we know, some times sigmae2 is noisy version of sigmae2_true
-#' @param sigmae2 the estimation of the variance structure
-#' @param nonnegative if the facotor and loading are nonnegative or not.
-#' TRUE for nonnegative
-#' FALSE for no constraint
-#' @param partype parameter type for the variance,
-#' "constant" for constant variance,
-#' "var_col" for nonconstant variance for column,
-#' "known" for the kown variance,
-#' "Bayes_var" for Bayes version of the nonconstant variance for row and column
-#' "loganova" is anova estiamtion for the log residual square
-#' @param objtype  objective function type,
-#' "margin_lik" for conditional likelihood,
-#' "lowerbound_lik" for full objective function
-#' @param fix_factor whether the factor is fixed or not
-#' TRUE for fix_factor
-#' FALSE for non-constraint
-#' @keywords internal
-#'
-# one step update function
-one_step_update = function(Y, El, El2, Ef, Ef2,
-                           N, P,
-                           sigmae2_v, sigmae2_true,
-                           sigmae2,
-                           nonnegative = FALSE,
-                           partype = "constant",
-                           objtype = "margin_lik",
-                           fix_factor = FALSE,
-                           ash_para = list(),
-                           fl_list=list()){
-  # if fix_factor is True, please choose objtype = "margin_lik"
-  output = ifelse(objtype == "lowerbound_lik", "matrix", "mean")
-  # deal with the missing value
-  na_index_Y = is.na(Y)
-  is_missing = any(na_index_Y)  # keep the missing result
-  if(is_missing){
-    # print("missing value use the EY as missing Y")
-    Y[na_index_Y] = (El %*% t(Ef))[na_index_Y]
-  }
-  if(fix_factor){
-    # actually we need do nothing here
-    output = "mean"
-    objtype = "margin_lik"
-  }else{
-    # estimate the variance structure
-    # sigma_est = function(sigmae2_v,sigmae2_true,sigmae2_pre = NA, partype = "constant")
-    sigmae2 = sigma_est(sigmae2_v,sigmae2_true,sigmae2 ,partype)
-    # for the list case in kronecker product
-    if(is.list(sigmae2)){
-      # print("sigmae is kronecker product in estimating sigmae2 in one step update")
-      sigmae2 = sigmae2$sig2_l %*% t(sigmae2$sig2_f)
-    }
-    # Y = lf^T + E and ATM is for l given f, for f given l we need y^T = fl^T + E^T
-    # sigmae2_input = ifelse(is.matrix(sigmae2),t(sigmae2),sigmae2) is wrong here
-    if(is.matrix(sigmae2)){
-      # print("sigmae is matrix in estimating sigmae2 in one step update")
-      sigmae2_input = t(sigmae2)
-    }else{
-      sigmae2_input = sigmae2
-    }
-    par_f = ATM_r1(t(Y), El, El2, sigmae2_input,
-                   col_var = "column", nonnegative,
-                   output,partype,ash_para)
-    Ef = par_f$Ef
-    Ef2 = par_f$Ef2
-    # if the Ef is zeros ,just return zeros
-    if(sum(Ef^2) <= 1e-12){
-      El = rep(0,length(El))
-      Ef = rep(0,length(Ef))
-      sigmae2_v =  sigmae2_v_est(Y,El,Ef,El^2,Ef^2,fl_list)
-      sigmae2 = sigma_est(sigmae2_v,sigmae2_true,sigmae2 ,partype)
-      obj_val = obj(N, P, sigmae2_v, sigmae2, par_f=NA, par_l=NA, objtype="margin_lik")
-      return(list(El = El, El2 = El^2,
-                  Ef = Ef, Ef2 = Ef^2,
-                  sigmae2_v = sigmae2_v,
-                  sigmae2 = sigmae2,
-                  obj_val = obj_val))
-    }
-    # update the Y and the Y^2 in this if else block
-    if(is_missing){
-      # the missing varianece is just sigmae2_v + sigmae2
-      if(is.matrix(sigmae2)){
-        sigmae2_impute = sigmae2[na_index_Y]
-      }else if(is.vector(sigmae2) & length(sigmae2) == length(Ef)){
-        sigmae2_impute = (matrix(rep(sigmae2,each = N),ncol = P))[na_index_Y]
-      }else{
-        sigmae2_impute = sigmae2
-      }
-      sigmae2_v = sigmae2_v_est(Y,El,Ef,El2,Ef2,fl_list)
-      sigmae2_v[na_index_Y] = sigmae2_v[na_index_Y] + sigmae2_impute
-    }else{
-      sigmae2_v = sigmae2_v_est(Y,El,Ef,El2,Ef2,fl_list)
-    }
-  }
-
-  sigmae2 = sigma_est(sigmae2_v,sigmae2_true,sigmae2 ,partype)
-  if(is.list(sigmae2)){
-    # kronecker product
-    sigmae2 = sigmae2$sig2_l %*% t(sigmae2$sig2_f)
-  }
-  par_l = ATM_r1(Y, Ef, Ef2,
-                 sigmae2, col_var = "row",
-                 nonnegative, output,
-                 partype, ash_para)
-  El = par_l$Ef
-  El2 = par_l$Ef2
-  # if El is zeros just return
-  if(sum(El^2) <= 1e-12){
-    El = rep(0,length(El))
-    Ef = rep(0,length(Ef))
-    sigmae2_v = sigmae2_v_est(Y,El,Ef,El^2,Ef^2,fl_list)
-    sigmae2 = sigma_est(sigmae2_v,sigmae2_true,sigmae2 ,partype)
-    obj_val = obj(N, P, sigmae2_v, sigmae2, par_f=NA, par_l=NA, objtype="margin_lik")
-    return(list(El = El, El2 = El^2,
-                Ef = Ef, Ef2 = Ef^2,
-                sigmae2_v = sigmae2_v,
-                sigmae2 = sigmae2,
-                obj_val = obj_val))
-  }
-  if(is_missing){
-    # the missing varianece is just sigmae2_v + sigmae2
-    if(is.matrix(sigmae2)){
-      sigmae2_impute = sigmae2[na_index_Y]
-    }else if(is.vector(sigmae2) & length(sigmae2) == length(Ef)){
-      sigmae2_impute = (matrix(rep(sigmae2,each = N),ncol = P))[na_index_Y]
-    }else{
-      sigmae2_impute = sigmae2
-    }
-    sigmae2_v = sigmae2_v_est(Y,El,Ef,El2,Ef2,fl_list)
-    sigmae2_v[na_index_Y] = sigmae2_v[na_index_Y] + sigmae2_impute
-  }else{
-    sigmae2_v =  sigmae2_v_est(Y,El,Ef,El2,Ef2,fl_list)
-  }
-  #use the estiamtion as the truth
-  sigmae2 = sigma_est(sigmae2_v,sigmae2_true,sigmae2 ,partype)
-  obj_val = obj(N, P, sigmae2_v, sigmae2, par_f, par_l, objtype)
-
-  return(list(El = El, El2 = El2,
-              Ef = Ef, Ef2 = Ef2,
-              sigmae2_v = sigmae2_v,
-              sigmae2 = sigmae2,
-              obj_val = obj_val))
-}
-
-#' inital value for flash
-#'
-#' description inital value for flash
-#'
-#' @return list of factor, loading and variance of noise matrix
-#'  \itemize{
-#'   \item{\code{El}} {is a N vector for mean of loadings}
-#'   \item{\code{El2}} {is a N vector for second moment of loadings}
-#'   \item{\code{Ef}} {is a N vector for mean of factors}
-#'   \item{\code{Ef2}} {is a N vector for second moment of factors}
-#'   \item{\code{sigmae2_v}}{is a N by P matrix for residual square}
-#'   \item{\code{sigmae2_true}}{is a N by P matrix for estimated value for the variance structure}
-#'  }
-#' @param Y the data matrix
-#' @param nonnegative if the facotor and loading are nonnegative or not.
-#' TRUE for nonnegative
-#' FALSE for no constraint
-#' @param fix_factor whether the factor is fixed or not
-#' TRUE for fix_factor
-#' FALSE for non-constraint
-#' @param factor_value is the factor value if the factor is fixed
-#' @param fl_list this is a list for the El, El2, Ef and Ef2 for other factor loadings from the model.
-#'                in the flash for backfitting, we need those to get the residuals and estimation for variance
-#' @param initial_list_r1 this is a list for the inital values of current El Ef El2 and Ef2.
-#' @param fix_initial this is a indicator for the initial value is fix or not.
-#' @keywords internal
-#'
-initial_value = function(Y, nonnegative = FALSE,
-                         factor_value = NA,fix_factor = FALSE,
-                         initial_list_r1 = list(), fix_initial = FALSE,
-                         fl_list=list()){
-  # use the total mean as the estimated missing value
-  na_index_Y = is.na(Y)
-  is_missing = any(na_index_Y)
-  if(is_missing){
-    # this is the initialization for the missing value
-    Y[na_index_Y] = mean(Y, na.rm = TRUE)
-  }
-  # the flash with plug-in missing value
-  if(fix_initial){
-    El = initial_list_r1$El
-    Ef = initial_list_r1$Ef
-    El2 = initial_list_r1$El2
-    Ef2 = initial_list_r1$Ef2
-  }else if(fix_factor){
-    Ef = factor_value
-    Ef2 = Ef^2
-    El = as.vector( (Y %*% Ef) / (sum(Ef2)) )
-    El2 = El^2
-  }else{
-    # El = svd(Y)$u[,1]
-    El = as.vector(irlba::irlba(Y,nv = 0,nu = 1)$u)
-    # the nonnegative value need positive inital value
-    if(nonnegative){
-      # El = abs(El)
-      El = El * (El > 0)
-    }
-    El2 = El^2
-    Ef = as.vector(t(El)%*%Y)
-    Ef2 = Ef^2
-  }
-  # residual matrix initialization
-  # here we don't have any information about the sigmae2_true
-  if(is_missing){
-    # here we don't have any information about the variance and the Y as well
-    # since the previoust the Y is impute by the mean of observed Ys so I would use the mean of Y here as well.
-    # this initialization still need to be considered more.
-    sigmae2_v =  sigmae2_v_est(Y,El,Ef,El2,Ef2,fl_list)
-  }else{
-    sigmae2_v =  sigmae2_v_est(Y,El,Ef,El2,Ef2,fl_list)
-  }
-  return(list(El = El, El2 = El^2,
-              Ef = Ef, Ef2 = Ef^2,
-              sigmae2_v = sigmae2_v))
-}
-
-#' FLASH
-#'
-#' factor loading adaptive shrinkage rank one version
-#' @return list of factor, loading and variance of noise matrix
-#'  \itemize{
-#'   \item{\code{El}} {is a N vector for mean of loadings}
-#'   \item{\code{Ef}} {is a N vector for mean of factors}
-#'   \item{\code{sigmae2}}{is a N by P matrix for estimated value for the variance structure}
-#'  }
-#' @param Y the data matrix
-#' @param tol is for the tolerence for convergence in iterations and ash
-#' @param maciter_r1 is maximum of the iteration times for rank one case
-#' @param sigmae2_true true value for the variance structure
-#' @param nonnegative if the facotor and loading are nonnegative or not.
-#' TRUE for nonnegative
-#' FALSE for no constraint
-#' @param partype parameter type for the variance,
-#' "constant" for constant variance,
-#' "var_col" for nonconstant variance for column,
-#' "known" for the kown variance,
-#' "Bayes_var" for Bayes version of the nonconstant variance for row and column
-#' "loganova" is anova estiamtion for the log residual square
-#' @param objtype  objective function type,
-#' "margin_lik" for conditional likelihood,
-#' "lowerbound_lik" for full objective function
-#' @param fix_factor whether the factor is fixed or not
-#' TRUE for fix_factor
-#' FALSE for non-constraint
-#' @param factor_value is the factor value if the factor is fixed
-#' @param ash_para is the parameters list for ash
-#' @param fl_list is a list containing all the informations from other factors ans loadings
-#' l is loadings, f is factors, l2 and f2 are corresponding second moments.
-#' priorpost_vec is expectation of log piropost ratio
-#' clik is conditional likelihood (marginal likelihood)
-#'
-#' @details flash privide rank one matrix decomposition with variational EM algorithm.
-#'
-#' @export flash
-#'
-#' @importFrom ashr ash
-#'
-flash = function(Y, tol=1e-5, maxiter_r1 = 500,
-                 partype = c("constant","known","Bayes_var","var_col","noisy","noisy_col"),
-                 sigmae2_true = NA,
-                 factor_value = NA,fix_factor = FALSE,
-                 initial_list_r1 = list(), fix_initial = FALSE,
-                 nonnegative = FALSE,
-                 objtype = c("margin_lik","lowerbound_lik"),
-                 ash_para = list(),
-                 fl_list=list()){
-  # match the parameters
-  partype = match.arg(partype, c("constant","known","Bayes_var","var_col","noisy","noisy_col"))
-  objtype = match.arg(objtype, c("margin_lik","lowerbound_lik"))
-
-  # check the input
-  if( any(!is.na(sigmae2_true)) & !(partype %in% c("known","noisy","noisy_col")) ){
-    stop("You should choose partype = 'known' or 'noisy' ,'noisy_col',if you want to input the value of sigmae2_true")
-  }
-
-  N = dim(Y)[1]
-  P = dim(Y)[2]
-
-  # to get the inital values
-  g_initial = initial_value(Y, nonnegative,factor_value,fix_factor,initial_list_r1, fix_initial,fl_list)
-  El = g_initial$El
-  Ef = g_initial$Ef
-  El2 = g_initial$El2
-  Ef2 = g_initial$Ef2
-  sigmae2_v = g_initial$sigmae2_v
-
-  # start iteration
-  # in the first iteration, there is no value foe estimated sigmae2
-  g_update = one_step_update(Y, El, El2, Ef, Ef2,
-                             N, P,
-                             sigmae2_v, sigmae2_true,
-                             sigmae2 = NA,
-                             nonnegative ,
-                             partype ,
-                             objtype ,
-                             fix_factor,
-                             ash_para,
-                             fl_list)
-  # parameters updates
-  El = g_update$El
-  El2 = g_update$El2
-  Ef = g_update$Ef
-  Ef2 = g_update$Ef2
-  sigmae2_v = g_update$sigmae2_v
-  sigmae2 = g_update$sigmae2
-  obj_val = g_update$obj_val
-
-  # track the objective value
-  obj_val_track = c(obj_val)
-
-  # we should also return when the first run get all zeros
-  if(sum(El^2)==0 || sum(Ef^2)==0){
-    sigmae2 = sigma_est(sigmae2_v,sigmae2_true,sigmae2 ,partype)
-    if(is.list(sigmae2)){
-      # print("here using kronecker product")
-      sigmae2 = sigmae2$sig2_l %*% t(sigmae2$sig2_f)
-    }
-    # add one more output for greedy algorithm which not useful here
-    c_lik_val = C_likelihood(N,P,sigmae2_v,sigmae2)$c_lik
-    # the above value is not useful, but is helpful to get the postprior value
-    # since obj_val = c_lik_value + priorpost_l + priorpost_f
-    return(list(l = El, f = Ef, l2 = El2, f2 = Ef2,
-                sigmae2 = sigmae2,
-                obj_val = obj_val,
-                c_lik_val = c_lik_val))
-  }
-
-  epsilon = 1
-  tau = 1
-  while(epsilon >= tol & tau < maxiter_r1){
-    tau = tau + 1
-    pre_obj = obj_val
-
-    g_update = one_step_update(Y, El, El2, Ef, Ef2,
-                               N, P,
-                               sigmae2_v, sigmae2_true,
-                               sigmae2,
-                               nonnegative ,
-                               partype ,
-                               objtype ,
-                               fix_factor,
-                               ash_para,
-                               fl_list)
-    # parameters updates
-    El = g_update$El
-    El2 = g_update$El2
-    Ef = g_update$Ef
-    Ef2 = g_update$Ef2
-    sigmae2_v = g_update$sigmae2_v
-    sigmae2 = g_update$sigmae2
-    obj_val = g_update$obj_val
-
-    if(sum(El^2)==0 || sum(Ef^2)==0){
-      El = rep(0,length(El))
-      Ef = rep(0,length(Ef))
-      break
-    }
-    epsilon = abs(pre_obj - obj_val)
-    obj_val_track = c(obj_val_track,obj_val)
-  }
-  sigmae2 = sigma_est(sigmae2_v,sigmae2_true,sigmae2 ,partype)
-  if(is.list(sigmae2)){
-    # print("here, using kronecker product")
-    sigmae2 = sigmae2$sig2_l %*% t(sigmae2$sig2_f)
-  }
-  # add one more output for greedy algorithm which not useful here
-  c_lik_val = C_likelihood(N,P,sigmae2_v,sigmae2)$c_lik
-  # the above value is not useful, but is helpful to get the postprior value
-  # since obj_val = c_lik_value + priorpost_l + priorpost_f
-  return(list(l = El, f = Ef, l2 = El2, f2 = Ef2,
-              sigmae2 = sigmae2,
-              obj_val = obj_val,
-              c_lik_val = c_lik_val,
-              obj_val_track = obj_val_track))
-}
-
+source("../code/flash_reg_170620.R")
 ```
@@
 
+### check the new obj function
+
 we check that the new objective function with penalty term does monotonically increase.
@@
 set.seed(99)
-K = 1
-N = 20
-P = 30
-SF = 0.5
-SL = 0.6
-signal = 1
-noise = 1
-par(mfrow = c(2, 2))
+objcheck_1 = function(K=1,N = 20, P=30, SF = 0.5, SL = 0.6, signal = 1,noise = 1){
+  par(mfrow = c(3, 2))
   par(cex = 0.6)
@@
   plot(g1$obj_val_track,main = "fdr method with penalized obj",type = "l")
-g2 = flashr::flash(Y,objtype = "l")
-plot(g2$obj_val_track,main = "fdr method without penalized obj",type = "l")
+  g2 =  flash(Y,objtype = "l",ash_para = list(method = "shrink"))
+  plot(g2$obj_val_track,main = "shrink method without penalized obj",type = "l")
+  g3 =  flashr::flash(Y,objtype = "l",ash_para = list(method = "shrink"))
+  plot(g3$obj_val_track,main = "shrink method without penalized obj(old)",type = "l")
+  g4 =  flashr::flash(Y,objtype = "l")
+  plot(g4$obj_val_track,main = "fdr method without penalized obj",type = "l")
+  plot(c(g1$obj_val,g2$obj_val,g3$obj_val,g4$obj_val),main = "obj value")
+  RMSE = c(sqrt(mean((Y - g1$l %*% t(g1$f) -E)^2)) ,
+           sqrt(mean((Y - g2$l %*% t(g2$f) -E)^2)) ,
+           sqrt(mean((Y - g3$l %*% t(g3$f) -E)^2)) ,
+           sqrt(mean((Y - g4$l %*% t(g4$f) -E)^2)) )
+  plot(RMSE,main = "RMSE")
+}
+```
+
+```{r}
+objcheck_1(K=1,N = 20, P=30, SF = 0.5, SL = 0.6, signal = 1,noise = 1)
+```
+
+```{r}
+objcheck_1(K=1,N = 200, P=300, SF = 0.8, SL = 0.6, signal = 1,noise = 1)
+```
+
+```{r}
+objcheck_1(K=1,N = 200, P=300, SF = 0.9, SL = 0.8, signal = 1,noise = 1)
+```
+
+```{r}
+objcheck_1(K=1,N = 200, P=300, SF = 0.5, SL = 0.9, signal = 1,noise = 3)
+```
+
+```{r}
+objcheck_1(K=1,N = 100, P=200, SF = 0.8, SL = 0.9, signal = 1,noise = 3)
 ```
@@
 we can see that the old version of code without penalty doesn't have monotonically increasing objective function, but the new version code with penalty term does.
+
+*now we have two methods (fdr, shrink), and both of them have monotonically increasing corresponding objective function in each case.*
 

### 2017-06-20 analysis/Objcheck_2.Rmd
Files commited by wflow_commit.

@@
   RMSE = sqrt(mean((Y - g1$l %*% t(g1$f) -E)^2)) 
+  RMSE0 = sqrt(mean((Y  -E)^2)) 
   obj_val = g1$obj_val
@@
   par_f0 = list(g = fit_g0, mat = mat0)
-  obj0 = obj(N,P,sigmae2_v0,sigmae2 = mean(sigmae2_v0),par_f0,par_l0,objtype = "lowerbound_lik")
+  obj0 = obj(N,P,sigmae2_v0,sigmae2 = mean(sigmae2_v0),par_f0,par_l0,objtype = "lowerbound_lik",ash_para = list(method = NULL))
 
-  return(list(rmse = RMSE,obj_val = obj_val,obj_0 = obj0))
+  return(list(rmse = RMSE,rmse0 = RMSE0,obj_val = obj_val,obj_0 = obj0))
 }
@@
   boxplot(RMSE,main = "RMSE")
-  hist(diff_obj, breaks = 3,main = "sign of difference of obj value")
+  hist(diff_obj, breaks = 50,main = "sign of difference of obj value")
   hist(RMSE,breaks = 50,main = "RMSE")
@@
 
-#### old code
-
-rank0check_old = function(K=1,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 0,noise = 1, mtype = "shrink"){
-  data = sim_K(K,N, P, SF , SL , signal ,noise )
-  Y = data$Y
-  E = data$Error
-  if(mtype == "fdr"){
-    g1 = flashr::flash(Y,objtype = "l")
-  }else{
-    g1 = flashr::flash(Y,objtype = "l",ash_para = list(method = "shrink"))
-  }
-
-  RMSE = sqrt(mean((Y - g1$l %*% t(g1$f) -E)^2)) 
-  obj_val = g1$obj_val
-  # to get the rank zero objective value
-  sigmae2_v0 = Y^2
-  fit_g0 = list(pi=c(1,0,0),sd = c(0,0.1,1))
-  mat0 = list(comp_postmean = matrix(0,ncol= N,nrow = 3),
-              comp_postmean2 = matrix(0,ncol= N,nrow = 3),
-              comp_postprob = matrix(0,ncol= N,nrow = 3))
-  mat0$comp_postprob[1,] = 1
-  par_l0 = list(g = fit_g0, mat = mat0)
-  fit_g0 = list(pi=c(1,0,0),sd = c(0,0.1,1))
-  mat0 = list(comp_postmean = matrix(0,ncol= P,nrow = 3),
-              comp_postmean2 = matrix(0,ncol= P,nrow = 3),
-              comp_postprob = matrix(0,ncol= P,nrow = 3))
-  mat0$comp_postprob[1,] = 1
-  par_f0 = list(g = fit_g0, mat = mat0)
-  obj0 = obj(N,P,sigmae2_v0,sigmae2 = mean(sigmae2_v0),par_f0,par_l0,objtype = "lowerbound_lik")
-
-  return(list(rmse = RMSE,obj_val = obj_val,obj_0 = obj0))
-}
-
-sim_rank0_old = function(K=1,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 0,noise = 1,mtype = "shrink"){
-  T = 200
-  RMSE = rep(NA,T)
-  diff_obj = rep(NA,T)
-  for(t in 1:T){
-    g0 = rank0check_old(K = 1,N, P, SF, SL , signal,noise,mtype = mtype)
-    RMSE[t] = g0$rmse
-    diff_obj[t] = (g0$obj_val - g0$obj_0)
-  }
-  par(mfrow = c(2, 2))
-  par(cex = 0.6)
-  par(mar = c(3, 3, 0.8, 0.8), oma = c(1, 1, 1, 1))
-  boxplot(RMSE,main = "RMSE")
-  hist(diff_obj, breaks = 3,main = "sign of difference of obj value")
-  hist(RMSE,breaks = 50,main = "RMSE")
-  plot(diff_obj,RMSE, main = "sign of obj diff vs RMSE")
-  return(c(sum(diff_obj == 0),sum(diff_obj>0),sum(diff_obj<0))/T)
-}
 ```
@@
 set.seed(99)
-sim_rank0_old(K=1,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 0,noise = 1)
+sim_rank0(K=1,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 0,noise = 1,mtype = "shrink")
 ```
@@
 set.seed(99)
-sim_rank0_old(K=1,N = 100, P=200, SF = 0.8, SL = 0.8, signal = 1,noise = 2)
+sim_rank0(K=1,N = 100, P=200, SF = 0.8, SL = 0.8, signal = 1,noise = 2,mtype = "shrink")
 ```

### 2017-06-21 analysis/Obj_summary.Rmd
Files commited by wflow_commit.

@@
 
-In this case, we check the objective funciton with penalty term, and it does monotonically increase.
+In this case, we check the objective funciton with penalty term, and it does monotonically increase. [rank 0 vs rank 1: fdr with penalty](Objcheck_2.html)
+
+**now we have two methods (fdr, shrink), and both of them have monotonically increasing corresponding objective function in each case.**
+
+> i) fit flash (with or without penalty)
+>
+> ii) check if flash-obtained objective function exceeds the K=0 objective.
+(note: not the limit of the objective as g_f, g_l -> 0, but the actual objective at g_f=g_l=0).
+If not, set result to g_f=g_l = 0.
+If so, keep g_f, g_l as output by flash.
+>
+> -- Matthew's advice.
+
+- we  have done the (i), and we found that it is reasonable to do the (ii).
+This is because we find that $diff(LB(fitted) - LB(rank=0))$ is negatively correlated with RMSE, which means larger obj value goes with smaller RMSE.
+So it makes sense we cut the obj value off at the rank 0 model.
+- (ii) is easy to implement.
+
+#### questions
+
+1. how to deal with $(\lambda_0 - 1) log \pi_0$ when $\pi_0 = 0$?
+2. now we only have two cases: shrink with $\lambda_0 = 1$ and fdr with $\lambda_0 = 10$. Do we need more choice of $\lambda_0$?
 

### 2017-06-28 code/flash_reg_170620.R
add penalty

@@ -1,0 +1,859 @@
+# ATM function
+#' title ash type model for l
+#'
+#' description use ash type model to maxmization
+#'
+#' @return list of factor, loading and variance of noise matrix
+#'  \itemize{
+#'   \item{\code{El}} {is a N vector for mean of loadings}
+#'   \item{\code{El2}} {is a N vector for second moment of loadings}
+#'  }
+#' @param Y is data matrix
+#' @param Ef is mean for the factor
+#' @param Ef2 is second moment for the factor
+#' @param sigmae2 is variance structure for noise matrix
+#' @param col_var is just for column variance, column is for column, row is for row
+#' @param nonnegative is flag for whether output nonnegative value
+#' @param output is format of output, "mean" is for mean value, "matrix" is for flash data matrix in ash
+#'
+#' @keywords internal
+#'
+
+ATM_r1 = function(Y, Ef, Ef2,
+                  sigmae2, col_var = "row",
+                  nonnegative = FALSE, output = "mean",
+                  partype = "constant",ash_para = list()){
+  # this part is preparing betahat and sebeta for ash
+  if(is.matrix(sigmae2)){
+    # print("sigmae2 is a matrix")
+    # this is for all variance are different
+    sum_Ef2 = (1/sigmae2) %*% Ef2
+    sum_Ef2 = as.vector(sum_Ef2)
+    sebeta = sqrt(1/(sum_Ef2))
+    betahat = as.vector( (Y/sigmae2) %*% Ef ) / (sum_Ef2)
+    betahat=as.vector(betahat)
+  } else if(is.vector(sigmae2) & length(sigmae2) == length(Ef)){
+    # this is for the non-constant variance in column
+    # print("sigmae2 is vector")
+    if(col_var == "row"){
+      sum_Ef2 = (1/sigmae2) * Ef2
+      sum_Ef2 = sum(sum_Ef2)
+      sebeta = sqrt(1/(sum_Ef2))
+      betahat = as.vector( Y %*% (Ef/sigmae2) ) / (sum_Ef2)
+      betahat=as.vector(betahat)
+    }else {
+      # for column
+      # here I have alread change the dimension by taking transpose to Y
+      sum_Ef2 = sum(Ef2)
[…]

### 2017-06-30 analysis/flash_alpha_versino.Rmd
Files commited by wflow_commit.

--- 2017-06-20 analysis/Objcheck_2.Rmd
@@
 ---
-title: "objective function with the penalty term"
-author: "wei"
+title: "flash alpha version"
+author: "Wei"
 date: YYYY-MM-DD
@@
 
-In this case, we would like to try the objective function with penalty term.
-
-I would like to modify the flash code in this document.
-
-```{r,echo=FALSE}
+```{r}
 source("../code/flash_reg_170620.R")
@@
 
-### check the new obj function
-
-we check that the new objective function with penalty term does monotonically increase.
-
 ```{r}
@@
 ```{r}
-objcheck_1(K=1,N = 20, P=30, SF = 0.5, SL = 0.6, signal = 1,noise = 1)
-```
-
-```{r}
-objcheck_1(K=1,N = 200, P=300, SF = 0.8, SL = 0.6, signal = 1,noise = 1)
-```
-
-```{r}
-objcheck_1(K=1,N = 200, P=300, SF = 0.9, SL = 0.8, signal = 1,noise = 1)
-```
-
-```{r}
-objcheck_1(K=1,N = 200, P=300, SF = 0.5, SL = 0.9, signal = 1,noise = 3)
-```
-
-```{r}
-objcheck_1(K=1,N = 100, P=200, SF = 0.8, SL = 0.9, signal = 1,noise = 3)
-```
-
-we can see that the old version of code without penalty doesn't have monotonically increasing objective function, but the new version code with penalty term does.
-
-*now we have two methods (fdr, shrink), and both of them have monotonically increasing corresponding objective function in each case.*
-
-## try the same problem: rank 0 vs rank 1 with  fdr method.
-
-```{r,echo = FALSE}
-
-rank0check = function(K=1,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 0,noise = 1, mtype = "fdr"){
+# test the obj0
+K=1;N = 200; P=300; SF = 0.5; SL = 0.6; signal = 0.1; noise = 1
 data = sim_K(K,N, P, SF , SL , signal ,noise )
@@
 E = data$Error
-  if(mtype == "fdr"){
 g1 = flash(Y,objtype = "l")
-  }else{
-    g1 = flash(Y,objtype = "l",ash_para = list(method = "shrink"))
-  }
+g2 = flash(Y,objtype = "m")
 
-  RMSE = sqrt(mean((Y - g1$l %*% t(g1$f) -E)^2)) 
-  RMSE0 = sqrt(mean((Y  -E)^2)) 
-  obj_val = g1$obj_val
-  # to get the rank zero objective value
 sigmae2_v0 = Y^2
@@
   obj0 = obj(N,P,sigmae2_v0,sigmae2 = mean(sigmae2_v0),par_f0,par_l0,objtype = "lowerbound_lik",ash_para = list(method = NULL))
-
-  return(list(rmse = RMSE,rmse0 = RMSE0,obj_val = obj_val,obj_0 = obj0))
-}
-
-sim_rank0 = function(K=1,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 0,noise = 1,mtype = "fdr"){
-  T = 200
-  RMSE = rep(NA,T)
-  diff_obj = rep(NA,T)
-  for(t in 1:T){
-    g0 = rank0check(K = 1,N, P, SF, SL , signal,noise,mtype = mtype)
-    RMSE[t] = g0$rmse
-    diff_obj[t] = (g0$obj_val - g0$obj_0)
-  }
-  par(mfrow = c(2, 2))
-  par(cex = 0.6)
-  par(mar = c(3, 3, 0.8, 0.8), oma = c(1, 1, 1, 1))
-  boxplot(RMSE,main = "RMSE")
-  hist(diff_obj, breaks = 50,main = "sign of difference of obj value")
-  hist(RMSE,breaks = 50,main = "RMSE")
-  plot(diff_obj,RMSE, main = "sign of obj diff vs RMSE")
-  return(c(sum(diff_obj == 0),sum(diff_obj>0),sum(diff_obj<0))/T)
-}
-
+  Fval(par_f0$mat, par_f0$g,ash_para=list(method = NULL))
 ```
-
-#### case 1: rank 0
-
-we compare the old code with our new code, we can see that most of the cases a
-
-*shrink method without penalty*
-
-```{r}
-set.seed(99)
-sim_rank0(K=1,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 0,noise = 1,mtype = "shrink")
-```
-
-*fdr method with penalty*
-
-```{r}
-set.seed(99)
-sim_rank0(K=1,N = 100, P=200, SF = 0.5, SL = 0.5, signal = 0,noise = 1)
-```
-
-we can see the improvement that the proportion of P(estimated structure better than rank 0 structure) is much larger.
-
-#### case 2: rank 1, very sparse with large noise
-
-*shrink method without penalty*
-
-```{r}
-set.seed(99)
-sim_rank0(K=1,N = 100, P=200, SF = 0.8, SL = 0.8, signal = 1,noise = 2,mtype = "shrink")
-```
-
-*fdr method with penalty*
-
-```{r}
-set.seed(99)
-sim_rank0(K=1,N = 100, P=200, SF = 0.8, SL = 0.8, signal = 1,noise = 2)
-```
-
-we can see that the results are improved since the proportion of P(estimated structure better than rank 0 structure) is much larger.
 

### 2017-06-30 code/flash_reg_170620.R
need to check the missing value

@@ -659,6 +659,7 @@ initial_value = function(Y, nonnegative = FALSE,
   if(is_missing){
     # this is the initialization for the missing value
     Y[na_index_Y] = mean(Y, na.rm = TRUE)
+    # Y[na_index_Y] = 0
   }
   # the flash with plug-in missing value
   if(fix_initial){
@@ -697,6 +698,46 @@ initial_value = function(Y, nonnegative = FALSE,
               Ef = Ef, Ef2 = Ef^2,
               sigmae2_v = sigmae2_v))
 }
+
+
+#' objective function value at zero
+#'
+#' @return list of factor, loading and variance of noise matrix
+#'  \itemize{
+#'   \item{\code{obj_val}} {objective function value at zero}
+#'  }
+#' @param Y the data matrix
+#' @param sigmae2_true true value for the variance structure
+#' @param partype parameter type for the variance,
+#' "constant" for constant variance,
+#' "var_col" for nonconstant variance for column,
+#' "known" for the kown variance,
+#' "Bayes_var" for Bayes version of the nonconstant variance for row and column
+#' "loganova" is anova estiamtion for the log residual square
+#' @param fl_list is a list containing all the informations from other factors ans loadings
+#' l is loadings, f is factors, l2 and f2 are corresponding second moments.
+#' priorpost_vec is expectation of log piropost ratio
+#' clik is conditional likelihood (marginal likelihood)
+#'
+#' @details objective function value at zero, use to check the rank one estimation and rank zero estimation results.
+#'
+
+obj_zero = function(Y,N,P,sigmae2_true,fl_list,partype){
+  na_index_Y = is.na(Y)
+  is_missing = any(na_index_Y)
+  if(is_missing){
+    # this is the initialization for the missing value
+    Y[na_index_Y] = mean(Y, na.rm = TRUE)
+  }
+  El = rep(0,N)
+  Ef = rep(0,P)
+  sigmae2_v = sigmae2_v_est(Y,El,Ef,El^2,Ef^2,fl_list)
+  sigmae2 = sigma_est(sigmae2_v,sigmae2_true, NA ,partype)
[…]

### 2017-09-16 analysis/flash_beta_test.Rmd
Files commited by wflow_commit.

---
title: "test flashr2"
author: "Wei"
date: 09-12
output: 
  html_document:
    code_folding: hide
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

### this is code for simulated data

```{r}
sim_hd = function(N, P, SF, SL, signal, a = rchisq(N,3),b = rchisq(P,1),mu = 0){

  E = matrix(rep(0,N*P),nrow=N)
  sig2_true = matrix(rep(0,N*P),nrow=N)
  for(i in 1:N){
    for(j in 1:P){
      sig2_true[i,j] = mu + a[i] * b[j]
      E[i,j] = rnorm(1,0,sqrt(mu + a[i] * b[j]))
    }
  }

  K=1
  lstart = rnorm(N, 0, signal)

  fstart = rnorm(P, 0, signal)

  index = sample(seq(1:N),(N*SL))
  lstart[index] = 0
  index = sample(seq(1:P),(P*SF))
  fstart[index] = 0

  Y = lstart %*% t(fstart) + E

  return(list(Y = Y, L_true = lstart, F_true = fstart, Error = E,sig2_true = sig2_true))
}
sim_K = function(K, N, P, SF, SL, signal,noise){
  E = matrix(rnorm(N*P,0,noise),nrow=N)
  Y = E
  L_true = array(0, dim = c(N,K))
  F_true = array(0, dim = c(P,K))

  for(k in 1:K){
    lstart = rnorm(N, 0, signal)
    fstart = rnorm(P, 0, signal)

    index = sample(seq(1:N),(N*SL))
    lstart[index] = 0
    index = sample(seq(1:P),(P*SF))
    fstart[index] = 0

    L_true[,k] = lstart
    F_true[,k] = fstart

    Y = Y + lstart %*% t(fstart)
  }
  return(list(Y = Y, L_true = L_true, F_true = F_true, Error = E))
}
```

```{r}
library(flashr2)
```

## check the prediction performance, this is something I don't understand

In this case I run flash for different situation and I found the old version has smaller RMSE.

```{r}
library(PMA)
library(softImpute)
set.seed(999)
N = 100
P = 200
dat = sim_K(K=1,N, P, SF = 0.5, SL = 0.5, signal = 1,noise = 5)
# a = rchisq(N,3)
# b = rchisq(P,1)
# data = sim_hd(N, P, SF = 0.5, SL = 0.5, signal = 1, a ,b,mu = 0)
Y = dat$Y
```

```{r}
data = set_flash_data(Y)
g1= flash_r1(data,verbose=TRUE)
gg1 = flashr::flash(Y)
sum((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(gg1$l %*% t(gg1$f)))^2)
sum((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(g1$EL %*% t(g1$EF)))^2)
sum((as.vector(dat$L_true %*% t(dat$F_true))-0)^2)
```

## missing value

```{r}
library(PMA)
library(softImpute)
set.seed(999)
N = 100
P = 200
dat = sim_K(K=1,N, P, SF = 0.5, SL = 0.5, signal = 3,noise = 5)
# a = rchisq(N,3)
# b = rchisq(P,1)
# data = sim_hd(N, P, SF = 0.5, SL = 0.5, signal = 1, a ,b,mu = 0)
Y = dat$Y
Y_miss = Y
Y_miss[1:50,1:100] = NA
Y_miss[51:100,101:200] = NA
```

```{r}
data.miss = set_flash_data(Y_miss)
g1.miss = flash_r1(data.miss,verbose=TRUE)
gg1.miss = flashr::flash(Y_miss)
sum((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(gg1.miss$l %*% t(gg1.miss$f)))^2)
sum((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(g1.miss$EL %*% t(g1.miss$EF)))^2)
sum((as.vector(dat$L_true %*% t(dat$F_true))-0)^2)
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```

### 2017-09-16 analysis/flash_beta_test.Rmd
Files commited by wflow_commit.

@@
 
+I print the obj value to check if it is increase.
+
 ```{r}
@@
 Y = dat$Y
+Y_miss = Y
+Y_miss[1:50,1:100] = NA
+Y_miss[51:100,101:200] = NA
 ```
@@
 
+### for this case, when data are missing, the obj is not increase
+
+the final result is rank zero, so I am not sure it is a problem
+
+```{r}
+data.miss = set_flash_data(Y_miss)
+g1.miss = flash_r1(data.miss,verbose=TRUE)
+```
+
 ## missing value
+
+In this case, I have two problems:
+
+1. The Obj function seems not increasing.
+
+2. the RMSE seems a bit large for flashr2. I think the $\tau$ estimation is not that precise.
 

### 2017-09-16 analysis/index.Rmd
Files commited by wflow_commit.

@@
 
-Click on this [summary of the rank0 vs rank1](Obj_summary.html) to see more details.
+Click on this [summary of the rank0 vs rank1](Obj_summary.html) to see a simple summary to the following two items.
 
@@
 
-### meeting summary
+## flashr2
 
-[04-26](meetnotes.html)
+This is a simple test [first try](flash_beta_test.html)
+Initialization issue: [softimpute as initialization](missingimputeinitial.html)

### 2017-09-16 analysis/missingimputeinitial.Rmd
Files commited by wflow_commit.

--- 2017-09-16 analysis/flash_beta_test.Rmd
@@
 ---
-title: "test flashr2"
-author: "Wei"
-date: 09-12
+title: "Initialization for missing data"
+author: "Wei Wang"
+date: 09-16
 output: 
@@
 <!-- Add your analysis here -->
-
-### this is code for simulated data
 
@@
 
-## check the prediction performance, this is something I don't understand
+## same data set as in "flash_beta_test"
 
-In this case I run flash for different situation and I found the old version has smaller RMSE.
-
-```{r}
-library(PMA)
-library(softImpute)
-set.seed(999)
-N = 100
-P = 200
-dat = sim_K(K=1,N, P, SF = 0.5, SL = 0.5, signal = 1,noise = 5)
-# a = rchisq(N,3)
-# b = rchisq(P,1)
-# data = sim_hd(N, P, SF = 0.5, SL = 0.5, signal = 1, a ,b,mu = 0)
-Y = dat$Y
-```
-
-```{r}
-data = set_flash_data(Y)
-g1= flash_r1(data,verbose=TRUE)
-gg1 = flashr::flash(Y)
-sum((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(gg1$l %*% t(gg1$f)))^2)
-sum((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(g1$EL %*% t(g1$EF)))^2)
-sum((as.vector(dat$L_true %*% t(dat$F_true))-0)^2)
-```
-
-## missing value
-
-```{r}
-library(PMA)
+```{r,message=F,warning=F}
 library(softImpute)
@@
 
+Here I just run flash on this data sets four times and provide different result due to the unstable initial value.
+
 ```{r}
+par(mfrow = c(2,2))
+for(i in 1:4){
   data.miss = set_flash_data(Y_miss)
-g1.miss = flash_r1(data.miss,verbose=TRUE)
-gg1.miss = flashr::flash(Y_miss)
-sum((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(gg1.miss$l %*% t(gg1.miss$f)))^2)
-sum((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(g1.miss$EL %*% t(g1.miss$EF)))^2)
-sum((as.vector(dat$L_true %*% t(dat$F_true))-0)^2)
+  g1.miss = flash_r1(data.miss,verbose=F)
+  plot(as.vector(dat$L_true %*% t(dat$F_true)), as.vector(g1.miss$EL %*% t(g1.miss$EF)), xlab = "truth",ylab = "estimated", main = "vectorized LF")
+}
+```
+
+## I don't think softimpute is good initialization method (or maybe I don't understand this method very well)
+
+Here is a example of use softimpute. maybe I was not using it in the right way.
+
+Here I use default `lambda`, and we can try `type = "svd"`. I tried different combination and this method is still not stable.
+
+```{r}
+par(mfrow = c(2,2))
+for(i in 1:4){
+  gsoft = softImpute(Y_miss, rank.max = 1,type = "als")
+  sfl = (gsoft$u)
+  sff = gsoft$d * (gsoft$v)
+  plot(as.vector(dat$L_true %*% t(dat$F_true)), as.vector(sfl %*% t(sff)), xlab = "truth",ylab = "estimated", main = "vectorized LF")
+}
 ```

### 2017-09-16 analysis/flash_beta_test.Rmd
Files commited by wflow_commit.

@@
 
-```{r}
+```{r,message=FALSE,results=FALSE}
 library(PMA)
@@
 g1= flash_r1(data,verbose=TRUE)
-gg1 = flashr::flash(Y)
+gg1 = flashr::flash(Y,ash_para_l = list(method = "shrink"),ash_para_f = list(method = "shrink"))
 sum((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(gg1$l %*% t(gg1$f)))^2)
@@
 g1.miss = flash_r1(data.miss,verbose=TRUE)
-gg1.miss = flashr::flash(Y_miss)
+gg1.miss = flashr::flash(Y_miss,ash_para_l = list(method = "shrink"),ash_para_f = list(method = "shrink"))
 sum((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(gg1.miss$l %*% t(gg1.miss$f)))^2)
@@
 
+## compare versions
+
+In this case, we set flash use `shrink` method and compare with the new version.
+
+```{r}
+sparse_level_l = seq(0.1,0.9,0.2)
+sparse_level_f = seq(0.1,0.9,0.2)
+signal_level = seq(1,5)
+RMSE_1 = rep(NA,5*5*5)
+RMSE_2 = rep(NA,5*5*5)
+for(i in 1:5){
+  for(j in 1:5){
+    for(k in 1:5){
+      set.seed(999)
+      N = 100
+      P = 200
+      dat = sim_K(K=1,N, P, SF = sparse_level_f[i],
+                  SL = sparse_level_l[j], 
+                  signal = signal_level[k],
+                  noise = 5)
+      Y = dat$Y
+      data = set_flash_data(Y)
+      g1= flash_r1(data,verbose=TRUE)
+      gg1 = flashr::flash(Y,ash_para_l = list(method = "shrink"),ash_para_f = list(method = "shrink"))
+      RMSE_1[25*(i-1) + 5*(j-1) + k] = 
+        sqrt(mean((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(gg1$l %*% t(gg1$f)))^2))/sqrt(mean((as.vector(dat$L_true %*% t(dat$F_true))-0)^2))
+      RMSE_2[25*(i-1) + 5*(j-1) + k] = 
+        sqrt(mean((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(g1$EL %*% t(g1$EF)))^2))/sqrt(mean((as.vector(dat$L_true %*% t(dat$F_true))-0)^2))
+    }
+  }
+}
+
+plot(RMSE_1 - RMSE_2)
+```
+
+To check the extreme case, and I found it is because flashr doesn't output rand zero estimation where zero estimation is better.
+
+```{r}
+i = 5
+j = 5
+k = 1
+set.seed(999)
+N = 100
+P = 200
+dat = sim_K(K=1,N, P, SF = sparse_level_f[i],
+            SL = sparse_level_l[j], 
+            signal = signal_level[k],
+            noise = 5)
+Y = dat$Y
+data = set_flash_data(Y)
+g1= flash_r1(data,verbose=TRUE)
+gg1 = flashr::flash(Y,ash_para_l = list(method = "shrink"),ash_para_f = list(method = "shrink"))
+
+  sqrt(mean((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(gg1$l %*% t(gg1$f)))^2))/sqrt(mean((as.vector(dat$L_true %*% t(dat$F_true))-0)^2))
+
+  sqrt(mean((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(g1$EL %*% t(g1$EF)))^2))/sqrt(mean((as.vector(dat$L_true %*% t(dat$F_true))-0)^2))
+
+```
 ## Session information

### 2017-09-17 analysis/flash_beta_test.Rmd
Files commited by wflow_commit.

--- 2017-09-16 analysis/flash_beta_test.Rmd
@@
 data = set_flash_data(Y)
-g1= flash_r1(data,verbose=TRUE)
-gg1 = flashr::flash(Y)
+g1= flash_r1(data,verbose=F)
+gg1 = flashr::flash(Y,ash_para_l = list(method = "shrink"),ash_para_f = list(method = "shrink"))
 sum((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(gg1$l %*% t(gg1$f)))^2)
@@
 data.miss = set_flash_data(Y_miss)
-g1.miss = flash_r1(data.miss,verbose=TRUE)
+g1.miss = flash_r1(data.miss,verbose=F)
 ```
@@
 data.miss = set_flash_data(Y_miss)
-g1.miss = flash_r1(data.miss,verbose=TRUE)
-gg1.miss = flashr::flash(Y_miss)
+g1.miss = flash_r1(data.miss,verbose=F)
+gg1.miss = flashr::flash(Y_miss,ash_para_l = list(method = "shrink"),ash_para_f = list(method = "shrink"))
 sum((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(gg1.miss$l %*% t(gg1.miss$f)))^2)
@@
 
+## compare versions
+
+In this case, we set flash use `shrink` method and compare with the new version.
+
+```{r}
+sparse_level_l = seq(0.1,0.9,0.2)
+sparse_level_f = seq(0.1,0.9,0.2)
+signal_level = seq(1,5)
+RMSE_1 = rep(NA,5*5*5)
+RMSE_2 = rep(NA,5*5*5)
+for(i in 1:5){
+  for(j in 1:5){
+    for(k in 1:5){
+      set.seed(999)
+      N = 100
+      P = 200
+      dat = sim_K(K=1,N, P, SF = sparse_level_f[i],
+                  SL = sparse_level_l[j], 
+                  signal = signal_level[k],
+                  noise = 5)
+      Y = dat$Y
+      data = set_flash_data(Y)
+      g1= flash_r1(data,verbose=F)
+      gg1 = flashr::flash(Y,ash_para_l = list(method = "shrink"),ash_para_f = list(method = "shrink"))
+      RMSE_1[25*(i-1) + 5*(j-1) + k] = 
+        sqrt(mean((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(gg1$l %*% t(gg1$f)))^2))/sqrt(mean((as.vector(dat$L_true %*% t(dat$F_true))-0)^2))
+      RMSE_2[25*(i-1) + 5*(j-1) + k] = 
+        sqrt(mean((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(g1$EL %*% t(g1$EF)))^2))/sqrt(mean((as.vector(dat$L_true %*% t(dat$F_true))-0)^2))
+    }
+  }
+}
+
+plot(RMSE_1 - RMSE_2)
+```
+
+To check the extreme case, and I found it is because flashr doesn't output rand zero estimation where zero estimation is better.
+
+```{r}
+i = 5
+j = 5
+k = 1
+set.seed(999)
+N = 100
+P = 200
+dat = sim_K(K=1,N, P, SF = sparse_level_f[i],
+            SL = sparse_level_l[j], 
+            signal = signal_level[k],
+            noise = 5)
+Y = dat$Y
+data = set_flash_data(Y)
+g1= flash_r1(data,verbose=F)
+gg1 = flashr::flash(Y,ash_para_l = list(method = "shrink"),ash_para_f = list(method = "shrink"))
+
+  sqrt(mean((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(gg1$l %*% t(gg1$f)))^2))/sqrt(mean((as.vector(dat$L_true %*% t(dat$F_true))-0)^2))
+
+  sqrt(mean((as.vector(dat$L_true %*% t(dat$F_true))-as.vector(g1$EL %*% t(g1$EF)))^2))/sqrt(mean((as.vector(dat$L_true %*% t(dat$F_true))-0)^2))
+
+```
 ## Session information

### 2017-09-17 analysis/NULLcheck.Rmd
Files commited by wflow_commit.

---
title: "null check for flashr2"
author: "wei"
date: 09-16
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

```{r}
library(flashr2)
library(MASS)
```

```{r}
########  spiked covariance model

set.seed(99)
T = 1000
lambda_3 = rep(NA,T)
flash_rank_3= rep(1,T)
for(t in 1:T){
  P = 300
  N = 100
  Sigma_0 = diag(rep(1,P))
  X = mvrnorm(n = N,mu = rep(0,P), Sigma = Sigma_0)
  lambda_3[t] =  svd(X)$d[1]
  data = set_flash_data(X)
  g1= flash_r1(data,verbose=F)
  if(sum(g1$EL) ==0 )  {flash_rank_3[t] = 0}
}

set.seed(99)
T = 1000
lambda_4 = rep(NA,T)
flash_rank_4= rep(1,T)
for(t in 1:T){
  P = 300
  N = 100
  Sigma_0 = diag(rep(1,P))
  nu = rnorm(P, 0, (6)/(sqrt(P)))
  index_nu = sample(seq(1:P),(P*0.95))
  nu[index_nu] = 0
  Sigma_1 = Sigma_0 + nu %*% t(nu)
  X = mvrnorm(n = N,mu = rep(0,P), Sigma = Sigma_1)
  lambda_4[t] =  svd(X)$d[1]
  data = set_flash_data(X)
  g1= flash_r1(data,verbose=F)
  if(sum(g1$EL) ==0 )  {flash_rank_4[t] = 0}
}

lambda_4_0 = lambda_4[which(flash_rank_4 ==0)]
lambda_4_1 = lambda_4[which(flash_rank_4 ==1)]

library(ggplot2)
dat = data.frame(lambda = c(lambda_3,lambda_4_0,lambda_4_1), name = c(rep("null",length(lambda_3)), rep("alt 0", length(lambda_4_0)),rep("alt 1", length(lambda_4_1)) ))
ggplot(dat,aes(x=lambda)) +
  geom_histogram(data=subset(dat,name == 'null'),fill = "red", alpha = 0.2) +
  # geom_histogram(data=subset(dat,name == 'alt 0'),fill = "blue", alpha = 0.2) +
  geom_histogram(data=subset(dat,name %in% c('alt 1','alt 0')),fill = "green", alpha = 0.2)
```

-red: null
-green: alternative

```{r}
library(ggplot2)
dat = data.frame(lambda = c(lambda_3,lambda_4_0,lambda_4_1), name = c(rep("null",length(lambda_3)), rep("alt 0", length(lambda_4_0)),rep("alt 1", length(lambda_4_1)) ))
ggplot(dat,aes(x=lambda)) +
  geom_histogram(data=subset(dat,name == 'null'),fill = "red", alpha = 0.2) +
  geom_histogram(data=subset(dat,name == 'alt 0'),fill = "blue", alpha = 0.2) +
  geom_histogram(data=subset(dat,name == 'alt 1'),fill = "green", alpha = 0.2)

```

-red: null
-green: alternative estimated as rank 1
-bule: alternative estimated as rank 0

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```

### 2017-09-17 analysis/index.Rmd
Files commited by wflow_commit.

@@
 This is a simple test [first try](flash_beta_test.html)
+
 Initialization issue: [softimpute as initialization](missingimputeinitial.html)
+
+Click here to find [null check for flashr2](NULLcheck.html)

### 2017-09-17 analysis/NULLcheck.Rmd
Files commited by wflow_commit.

@@
 T = 1000
-lambda_3 = rep(NA,T)
-flash_rank_3= rep(1,T)
+lambda_0 = rep(NA,T)
+flash_rank_0= rep(1,T)
+flash2_rank_0= rep(1,T)
 for(t in 1:T){
@@
   X = mvrnorm(n = N,mu = rep(0,P), Sigma = Sigma_0)
-  lambda_3[t] =  svd(X)$d[1]
+  lambda_0[t] =  svd(X)$d[1]
   data = set_flash_data(X)
   g1= flash_r1(data,verbose=F)
-  if(sum(g1$EL) ==0 )  {flash_rank_3[t] = 0}
+  g2 = flashr::flash(X)
+  if(sum(g1$EL) ==0 )  {flash2_rank_0[t] = 0}
+  if(sum(g2$l) ==0 )  {flash_rank_0[t] = 0}
 }
@@
 T = 1000
-lambda_4 = rep(NA,T)
-flash_rank_4= rep(1,T)
+lambda_1 = rep(NA,T)
+flash_rank_1= rep(1,T)
+flash2_rank_1= rep(1,T)
 for(t in 1:T){
@@
   X = mvrnorm(n = N,mu = rep(0,P), Sigma = Sigma_1)
-  lambda_4[t] =  svd(X)$d[1]
+  lambda_1[t] =  svd(X)$d[1]
   data = set_flash_data(X)
   g1= flash_r1(data,verbose=F)
-  if(sum(g1$EL) ==0 )  {flash_rank_4[t] = 0}
+  g2 = flashr::flash(X)
+  if(sum(g1$EL) ==0 )  {flash2_rank_1[t] = 0}
+  if(sum(g2$l) ==0 )  {flash_rank_1[t] = 0}
 }
 
-lambda_4_0 = lambda_4[which(flash_rank_4 ==0)]
-lambda_4_1 = lambda_4[which(flash_rank_4 ==1)]
+lambda_1_0_0 = lambda_0[which(flash_rank_0 ==0)]
+lambda_1_0_1 = lambda_0[which(flash_rank_0 ==1)]
+lambda_2_0_0 = lambda_0[which(flash2_rank_0 ==0)]
+lambda_2_0_1 = lambda_0[which(flash2_rank_0 ==1)]
 
-library(ggplot2)
-dat = data.frame(lambda = c(lambda_3,lambda_4_0,lambda_4_1), name = c(rep("null",length(lambda_3)), rep("alt 0", length(lambda_4_0)),rep("alt 1", length(lambda_4_1)) ))
-ggplot(dat,aes(x=lambda)) +
-  geom_histogram(data=subset(dat,name == 'null'),fill = "red", alpha = 0.2) +
-  # geom_histogram(data=subset(dat,name == 'alt 0'),fill = "blue", alpha = 0.2) +
-  geom_histogram(data=subset(dat,name %in% c('alt 1','alt 0')),fill = "green", alpha = 0.2)
+lambda_1_1_0 = lambda_1[which(flash_rank_1 ==0)]
+lambda_1_1_1 = lambda_1[which(flash_rank_1 ==1)]
+lambda_2_1_0 = lambda_1[which(flash2_rank_1 ==0)]
+lambda_2_1_1 = lambda_1[which(flash2_rank_1 ==1)]
+
 ```
 
--red: null
--green: alternative
+## model
+
+In this case the data sets are:
+$$H_0: X \sim N(0, \Sigma_0)$$
+$$H_1: X \sim N(0,\Sigma_1)$$
+where $\Sigma_0 = I$ and $\Sigma_1 = I + \nu \nu^T$
 
@@
 library(ggplot2)
-dat = data.frame(lambda = c(lambda_3,lambda_4_0,lambda_4_1), name = c(rep("null",length(lambda_3)), rep("alt 0", length(lambda_4_0)),rep("alt 1", length(lambda_4_1)) ))
+dat = data.frame(lambda = c(lambda_0,lambda_1), name = c(rep("null",length(lambda_0)), rep("alt", length(lambda_1)) ))
 ggplot(dat,aes(x=lambda)) +
   geom_histogram(data=subset(dat,name == 'null'),fill = "red", alpha = 0.2) +
+  geom_histogram(data=subset(dat,name == 'alt'),fill = "green", alpha = 0.2) 
+```
+
+* red: null
+* green: alternative
+
+### flashr
+
+```{r}
+library(ggplot2)
+dat = data.frame(lambda = c(lambda_1_0_0,lambda_1_0_1,lambda_1_1_0,lambda_1_1_1), 
+                 name = c(rep("null 0",length(lambda_1_0_0)),
+                          rep("null 1",length(lambda_1_0_1)),
+                          rep("alt 0",length(lambda_1_1_0)),
+                          rep("alt 1",length(lambda_1_1_1))))
+ggplot(dat,aes(x=lambda)) +
+  geom_histogram(data=subset(dat,name == 'null 0'),fill = "red", alpha = 0.2) +
+  geom_histogram(data=subset(dat,name == 'null 1'),fill = "black", alpha = 0.2) +
   geom_histogram(data=subset(dat,name == 'alt 0'),fill = "blue", alpha = 0.2) +
@@
 
--red: null
--green: alternative estimated as rank 1
--bule: alternative estimated as rank 0
+* red: null as rank 0
+* black: null as rank 1
+* blue: alternative as 0
+* green: alternative as 1
+
+### flashr2
+
+```{r}
+library(ggplot2)
+dat2 = data.frame(lambda = c(lambda_2_0_0,lambda_2_0_1,lambda_2_1_0,lambda_2_1_1), 
+                 name = c(rep("null 0",length(lambda_2_0_0)),
+                          rep("null 1",length(lambda_2_0_1)),
+                          rep("alt 0",length(lambda_2_1_0)),
+                          rep("alt 1",length(lambda_2_1_1))))
+ggplot(dat,aes(x=lambda)) +
+  geom_histogram(data=subset(dat2,name == 'null 0'),fill = "red", alpha = 0.2) +
+  geom_histogram(data=subset(dat2,name == 'null 1'),fill = "black", alpha = 0.2) +
+  geom_histogram(data=subset(dat2,name == 'alt 0'),fill = "blue", alpha = 0.2) +
+  geom_histogram(data=subset(dat2,name == 'alt 1'),fill = "green", alpha = 0.2)
+
+```
+
+* red: null as rank 0
+* black: null as rank 1
+* blue: alternative as 0
+* green: alternative as 1
 

### 2017-09-17 analysis/nullcheck2.Rmd
Files commited by wflow_commit.

--- 2017-09-17 analysis/NULLcheck.Rmd
@@
 ---
-title: "null check for flashr2"
-author: "wei"
-date: 09-16
+title: "NUll check for flashr2"
+author: "Wei Wang"
+date: 09-17
 output: html_document
@@
   g1= flash_r1(data,verbose=F)
-  g2 = flashr::flash(X)
+  g2 = flashr::flash(X,objtype = "l",
+                     ash_para_l = list(nullweight = 1 ),
+                     ash_para_f = list(nullweight = 1 ))
   if(sum(g1$EL) ==0 )  {flash2_rank_0[t] = 0}
@@
   Sigma_0 = diag(rep(1,P))
-  nu = rnorm(P, 0, (6)/(sqrt(P)))
-  index_nu = sample(seq(1:P),(P*0.95))
-  nu[index_nu] = 0
+  nu = rnorm(P, 0, (1)/(sqrt(P)))
+  # index_nu = sample(seq(1:P),(P*0.95))
+  # nu[index_nu] = 0
   Sigma_1 = Sigma_0 + nu %*% t(nu)
@@
   g1= flash_r1(data,verbose=F)
-  g2 = flashr::flash(X)
+  g2 = flashr::flash(X,objtype = "l",
+                     ash_para_l = list(nullweight = 1 ),
+                     ash_para_f = list(nullweight = 1 ))
   if(sum(g1$EL) ==0 )  {flash2_rank_1[t] = 0}

### 2017-09-29 analysis/applyflahr2onsimulations.Rmd
Files commited by wflow_commit.

---
title: "apply flashr2 on the paper"
author: "Wei Wang"
date: 09-28
output: 
  html_document:
    code_folding: hide
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

## rank 1 data set

### simulated data

```{r,echo=FALSE}
library("MASS")
betaO=function(P,betapi,betasigma,propt){
  idx=rmultinom(P,size=1,prob=betapi)
  K=length(betapi)
  s=array(0,dim=c(K,K))
  diag(s)=betasigma
  bnorm=mvrnorm(P,rep(0,K),s)
  betaO=apply(bnorm*t(idx),1,sum)
  betaO=betaO*rbinom(P,1,propt)
  return(betaO)
}

datamaker = function(N,P,l_pi,l_se,l_sp,f_pi,f_se,f_sp,sigmae){
  # here I would like to fix the sparsity of L and F which is simple
  # if we need to do futher experiment
  L_true = betaO(N,l_pi,l_se,l_sp)
  F_true = betaO(P,f_pi,f_se,f_sp)
  E = matrix(rnorm(N*P,0,sigmae),ncol = P)
  Y = L_true %*% t(F_true) + E
  return(list(Y = Y,L_true = L_true,F_true = F_true))
}
# I think we can try some situation
# pve is small due to the

# this can be a example of the bi sparse factor and loadings

CVPMD=function(Y,c_u,c_v){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = matrix(sample(P,P),ncol = 5)
  rowindex = matrix(sample(N,N),ncol = 5)

  missing= array(0,dim = c(5,N,P))
  foldindex = array(0,dim = c(5,5,2))
  for(i in 1:5){
    for(j in 1:5){
      foldindex[i,j,1] = i
      foldindex[i,j,2] = (i+j) %% 5
    }
  }
  foldindex[which(foldindex == 0)] = 5
  for(i in 1:5){
    missing[i, , ] = Y
    for(j in 1:5){
      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
    }
  }
  n_u = length(c_u)
  n_v = length(c_v)
  CVRMSE = array(0,dim = c(n_u,n_v))
  minrmse = Inf
  opt_u = 0
  opt_v = 0
  for(t_u in 1:n_u){
    for(t_v in 1:n_v){
      rmse = rep(0,5)
      for(i in 1:5){
        out = PMD(missing[i,,], sumabsu = c_u[t_u], sumabsv = c_v[t_v])
        misshat = out$d * out$u %*% t(out$v)
        for(j in 1:5){
          rmse[i] = rmse[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] - 
                                     misshat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2)
        }
        rmse[i] = sqrt(rmse[i] / (N * P/5))
      }
      CVRMSE[t_u,t_v] = mean(rmse)
      if(CVRMSE[t_u,t_v] < minrmse){
        minrmse = CVRMSE[t_u,t_v]
        opt_u = c_u[t_u]
        opt_v = c_v[t_v]
      }
    }
  }
  return(list(opt_u = opt_u,opt_v = opt_v))
}

PMA.wrapper = function(Y,ngrids = 10){
  library(PMA)
  N = dim(Y)[1]
  P = dim(Y)[2]
  c_u = seq(1,sqrt(N),length.out = ngrids)
  c_v = seq(1,sqrt(P),length.out = ngrids)
  cvout = CVPMD(Y,c_u,c_v)
  out = PMD(Y,sumabsu = cvout$opt_u, sumabsv = cvout$opt_v )
  residual_PMD = Y - out$d * out$u %*% t(out$v)
  return(list(d = out$d, u = out$u, v = out$v, residual = residual_PMD))
}
```

### sparse

```{r}
library(cowplot)
set.seed(99)
L_se = c( 0.25, 0.5, 1, 2, 4)
L_pi = c(1, 1,1,1,1)
L_pi = L_pi / sum(L_pi)
N = 200
P = 300
Data = datamaker(N,P,L_pi,L_se,0.1,c(1),c(1),1,sqrt(1))
#Data = datamaker(N,P,L_pi,L_se,0.36,c(1),c(1),1,sqrt(1))
dim(Data$Y)
# = flashr::flash(Data$Y)
data = flashr2::flash_set_data(Data$Y)
gf = flashr2::flash_r1(data,verbose=FALSE)

PMDlog = capture.output({gPMD = invisible( PMA.wrapper(Data$Y)) })
MSE_PMD = sqrt(mean(( gPMD$d *(gPMD$u %*% t(gPMD$v)) - Data$L_true%*%t(Data$F_true))^2))/sqrt(mean((Data$L_true%*%t(Data$F_true))^2))

library(ggplot2)
library(grid)
# zoom bounding box
xlim <- c(-0.08,0.08); ylim <- c(-0.08,0.08)

mat = cbind(Data$L_true,gf$EL,gPMD$u)
mat = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x] * sign(cor(mat[,x],mat[,1])) })
mat1 = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x]/sqrt(sum((mat[,x])^2)) } )
colnames(mat1) = c("truth","FLASH","PMD")
df1 = data.frame(mat1)

base1 <- ggplot(df1, aes(x = truth, y = FLASH)) +
  geom_point()  +
  theme_bw()+ geom_abline(intercept = 0,slope = 1)
zoomed1 = ggplot(df1, aes(x = truth, y = FLASH)) + 
    geom_point() + coord_cartesian(xlim=xlim, ylim=ylim)+ theme_grey()+ geom_abline(intercept = 0,slope = 1)

g1 <- ggplotGrob(zoomed1)
base1 +annotation_custom(grob = g1, xmin = -0.6, xmax = -0.1, ymin = -0.05, ymax = 0.45)

base2 <- ggplot(df1, aes(x = truth, y = PMD)) +
  geom_point()  +
  theme_bw()+ geom_abline(intercept = 0,slope = 1)
zoomed2 = ggplot(df1, aes(x = truth, y = PMD)) + 
    geom_point() + coord_cartesian(xlim=xlim, ylim=ylim)+ theme_grey() + geom_abline(intercept = 0,slope = 1)

g2 <- ggplotGrob(zoomed2)
base2 +annotation_custom(grob = g2, xmin = -0.6, xmax = -0.1, ymin = -0.05, ymax = 0.45)
plot_grid(base1, base2, zoomed1,zoomed2, labels = "AUTO", ncol = 2)

gsfamix_1 = readRDS("../data/simulation/rankone/highsparse/sfamix_res.rds")
gsvd = svd(Data$Y)
gsfa = readRDS("../data/simulation/rankone/highsparse/sfa_res.rds")

# to get the KSVD result
# try other methods
library("R.matlab")
# saveRDS(Data,file = "../data/simulation/rankone/mediansparse/test_1.rds")
# write the data into matlab
# writeMat("../data/simulation/rankone/highsparse/test_1.mat", data = Data)
# KSVD
res_KSVD = readMat("../data/simulation/rankone/highsparse/KSVDtest_1.mat",sparseMatrixClass="matrix")
Dic = res_KSVD$result[[1]]
loading_Dic = res_KSVD$result[[2]]
MSE_KSVD = sqrt(mean(( Dic %*% loading_Dic - Data$L_true%*%t(Data$F_true))^2))/sqrt(mean((Data$L_true%*%t(Data$F_true))^2))

# add ssvd into comparison 
gssvd = ssvd::ssvd(Data$Y,method = "method")
MSE_SSVD = sqrt(mean(( gssvd$d * (gssvd$u %*% t(gssvd$v)) - Data$L_true%*%t(Data$F_true))^2))/sqrt(mean((Data$L_true%*%t(Data$F_true))^2))

library(ggplot2)
library(grid)

mat = cbind(Data$L_true,gf$EL,gPMD$u,gsfa$L,as.vector(gsfamix_1$F),as.vector(Dic),gssvd$u)
mat = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x] * sign(cor(mat[,x],mat[,1])) })
mat1 = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x]/sqrt(sum((mat[,x])^2)) } )
colnames(mat1) = c("truth","FLASH","PMD","SFA","SFAmix","KSVD","SSVD")
df1 = data.frame(mat1)
xybound = c( -0.75, -0.05, -0.05, 0.55)
z_size = c(-0.08,0.08,-0.08,0.08)
zoom_plot = function(df1,i,zoom_size,bound){
  xlim <- c(zoom_size[1],zoom_size[2])
  ylim <- c(zoom_size[3],zoom_size[4])

  base1 <- ggplot(df1, aes(x = truth, y = df1[,i])) +
    geom_point()  + labs(y = colnames(df1)[i]) +
    theme_bw()+ geom_abline(intercept = 0,slope = 1) + coord_cartesian(xlim=c(-0.65,0.5), ylim=c(-0.65,0.5))
  zoomed1 = ggplot(df1, aes(x = truth, y = df1[,i]))  + labs(y = colnames(df1)[i]) + 
    geom_point() + coord_cartesian(xlim=xlim, ylim=ylim)+ theme_grey()+ geom_abline(intercept = 0,slope = 1)

  g1 <- ggplotGrob(zoomed1)
  pic1 = base1 +annotation_custom(grob = g1,  xmin = bound[1], xmax = bound[2], ymin = bound[3], ymax = bound[4])
  return(pic1)
}

pic_list = list() 
for(i in 2:7){
  pic_list[[(i - 1)]] = zoom_plot(df1,i,z_size,xybound)
}
plot_grid(pic_list[[1]], pic_list[[2]], 
          pic_list[[3]], pic_list[[4]], 
          pic_list[[5]], pic_list[[6]],  labels = "AUTO", ncol = 3)
```

### intermediate sparse

```{r}
library(cowplot)
set.seed(99)
L_se = c( 0.25, 0.5, 1, 2, 4)
L_pi = c(1, 1,1,1,1)
L_pi = L_pi / sum(L_pi)
N = 200
P = 300
Data = datamaker(N,P,L_pi,L_se,0.7,c(1),c(1),1,sqrt(16))
dim(Data$Y)
data = flashr2::flash_set_data(Data$Y)
gf = flashr2::flash_r1(data,verbose=FALSE)
PMDlog = capture.output({gPMD = invisible( PMA.wrapper(Data$Y)) })

gsfamix_1 = readRDS("../data/simulation/rankone/mediansparse/sfamix_2.rds")
gsfamix_2 = readRDS("../data/simulation/rankone/mediansparse/sfamix_3.rds")
gsfa = readRDS("../data/simulation/rankone/mediansparse/sfa_res.rds")

# to get the KSVD result
# try other methods
library("R.matlab")
# saveRDS(Data,file = "../data/simulation/rankone/mediansparse/test_1.rds")
# write the data into matlab
# writeMat("../data/simulation/rankone/mediansparse/test_1.mat", data = Data)
# KSVD
res_KSVD = readMat("../data/simulation/rankone/mediansparse/KSVDtest_1.mat",sparseMatrixClass="matrix")
Dic = res_KSVD$result[[1]]
loading_Dic = res_KSVD$result[[2]]
MSE_KSVD = sqrt(mean(( Dic %*% loading_Dic - Data$L_true%*%t(Data$F_true))^2))/sqrt(mean((Data$L_true%*%t(Data$F_true))^2))

# add ssvd into comparison 
gssvd = ssvd::ssvd(Data$Y,method = "method")

library(ggplot2)
library(grid)
# zoom bounding box
xlim <- c(-0.1,0.1); ylim <- c(-0.1,0.1)

mat = cbind(Data$L_true,gf$EL,gPMD$u,gsfa$L,as.vector(gsfamix_1$F),as.vector(gsfamix_2$F),as.vector(Dic))
mat = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x] * sign(cor(mat[,x],mat[,1])) })
mat1 = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x]/sqrt(sum((mat[,x])^2)) } )
colnames(mat1) = c("truth","FLASH","PMD","SFA","SFAmix","SSVD","KSVD")
df1 = data.frame(mat1)

base1 <- ggplot(df1, aes(x = truth, y = FLASH)) +
  geom_point()  +
  theme_bw()+ geom_abline(intercept = 0,slope = 1)
zoomed1 = ggplot(df1, aes(x = truth, y = FLASH)) + 
    geom_point() + coord_cartesian(xlim=xlim, ylim=ylim)+ theme_grey()+ geom_abline(intercept = 0,slope = 1)

g1 <- ggplotGrob(zoomed1)
pic1 = base1 +annotation_custom(grob = g1,  xmin = -0.35, xmax = 0, ymin = 0.05, ymax = 0.4)

# mat = cbind(Data$L_true,gf$l,gPMD$u,gsfa$L,as.vector(gsfamix_1$F),as.vector(gsfamix_2$F),as.vector(Dic))
mat = cbind(Data$L_true,gf$EL,gPMD$u,gsfa$L,as.vector(gsfamix_1$F),as.vector(gssvd$u),as.vector(Dic))
mat = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x] * sign(cor(mat[,x],mat[,1])) })
mat1 = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x]/sqrt(sum((mat[,x])^2)) } )
colnames(mat1) = c("truth","FLASH","PMD","SFA","SFAmix","SSVD","KSVD")
df1 = data.frame(mat1)
xybound = c(-0.42,0,0.03,0.38)
z_size = c(-0.07,0.07,-0.07,0.07)
zoom_plot = function(df1,i,zoom_size,bound){
  xlim <- c(zoom_size[1],zoom_size[2])
  ylim <- c(zoom_size[3],zoom_size[4])

  base1 <- ggplot(df1, aes(x = truth, y = df1[,i])) +
    geom_point()  + labs(y = colnames(df1)[i]) +
    theme_bw()+ geom_abline(intercept = 0,slope = 1) + coord_cartesian(xlim=c(-0.36,0.36), ylim=c(-0.36,0.36))
  zoomed1 = ggplot(df1, aes(x = truth, y = df1[,i]))  + labs(y = colnames(df1)[i]) + 
    geom_point() + coord_cartesian(xlim=xlim, ylim=ylim)+ theme_grey()+ geom_abline(intercept = 0,slope = 1)

  g1 <- ggplotGrob(zoomed1)
  pic1 = base1 +annotation_custom(grob = g1,  xmin = bound[1], xmax = bound[2], ymin = bound[3], ymax = bound[4])
  return(pic1)
}

pic_list = list() 
for(i in 2:7){
  pic_list[[(i - 1)]] = zoom_plot(df1,i,z_size,xybound)
}
plot_grid(pic_list[[1]], pic_list[[2]], 
          pic_list[[3]], pic_list[[4]], 
          pic_list[[6]], pic_list[[5]],  labels = "AUTO", ncol = 3)
```

## rank 3 case

```{r}

CVPMD=function(Y,c_s,K){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = matrix(sample(P,P),ncol = 5)
  rowindex = matrix(sample(N,N),ncol = 5)

  missing= array(0,dim = c(5,N,P))
  foldindex = array(0,dim = c(5,5,2))
  for(i in 1:5){
    for(j in 1:5){
      foldindex[i,j,1] = i
      foldindex[i,j,2] = (i+j) %% 5
    }
  }
  foldindex[which(foldindex == 0)] = 5
  for(i in 1:5){
    missing[i, , ] = Y
    for(j in 1:5){
      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
    }
  }
  n_s = length(c_s)
  CVRMSE = rep(NA,n_s)
  minrmse = Inf
  opt_s = 0
  # here we only need one layer
  for(t_s in 1:n_s){
    rmse = rep(0,5)
    for(i in 1:5){
      out = PMD(missing[i,,], sumabs = c_s[t_s], sumabsv = NULL, sumabsu = NULL,K = K)
      misshat = out$u %*%  diag(out$d) %*% t(out$v)
      for(j in 1:5){
        rmse[i] = rmse[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] - 
                                   misshat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2)
      }
      rmse[i] = sqrt(rmse[i] / (N * P/5))
    }
    CVRMSE[t_s] = mean(rmse)
    if(CVRMSE[t_s] < minrmse){
      minrmse = CVRMSE[t_s]
      opt_s = c_s[t_s]
    }
  }

  return(list(opt_s = opt_s ))
}

PMA.wrapper = function(Y,ngrids = 10,K){
  library(PMA)
  N = dim(Y)[1]
  P = dim(Y)[2]
  c_s = seq(0.2,0.8,len=ngrids)
  cvout = CVPMD(Y,c_s, K = K)
  out = PMD(Y,sumabsu = NULL, sumabsv = NULL, sumabs = cvout$opt_s ,K = K)
  residual_PMD = Y -  out$u %*%  diag(out$d) %*% t(out$v)
  return(list(d = out$d, u = out$u, v = out$v, residual = residual_PMD))
}

```

```{r}
set.seed(99)
N = 150
P = 240
K = 3
L_true = array(0,dim = c(N,K))
F_true = array(0,dim = c(P,K))
F_true[1:80,1] = rnorm(80,0,0.5)
F_true[81:160,2] = rnorm(80,0,1)
F_true[161:240,3] = rnorm(80,0,2)
L_true[1:10,1] = rnorm(10,0,2)
L_true[11:60,2] = rnorm(50,0,1)
L_true[61:150,3] = rnorm(90,0,0.5)
G = L_true %*% t(F_true)
# generate Y
E = matrix(rnorm(N*P,0,2),nrow=N)
Y = L_true %*% t(F_true) + E

data = flashr2::flash_set_data(Y)
f_greedy = flashr2::flash_add_greedy(data,Kmax=10)
g_b_flash = flashr2::flash_backfit(data,f_greedy)

outcv = PMA.wrapper(Y,10,K = 3)
library(ssvd)
gssvd = ssvd(Y,r = 3)
```

```{r}
library(reshape2)
melted_cormat_1 <- melt(G, na.rm = TRUE)
melted_cormat_2 <- melt(g_b_flash$EL %*% t(g_b_flash$EF), na.rm = TRUE)
melted_cormat_3 <- melt(outcv$u %*% diag(outcv$d) %*% t(outcv$v), na.rm = TRUE)
melted_cormat_4 <- melt(Y, na.rm = TRUE)
melted_cormat_5 <- melt(gssvd$u %*% diag(gssvd$d) %*% t(gssvd$v), na.rm = TRUE)
# Heatmap
library(ggplot2)
p1 = ggplot(data = melted_cormat_1, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-2,2), space = "Lab") + labs(title = "truth", y = "sample", x = "variable") +
  theme_minimal() + theme(legend.position="none",plot.background = element_rect())

p2 = ggplot(data = melted_cormat_2, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-2,2), space = "Lab") + labs(title = "FLASH",y = "sample", x = "variable") +
  theme_minimal() + theme(legend.position="none",plot.background = element_rect())

p3 = ggplot(data = melted_cormat_3, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-2,2), space = "Lab") + labs(title = "PMD",y = "sample", x = "variable") +
  theme_minimal() + theme(legend.position="none",plot.background = element_rect())

p4 = ggplot(data = melted_cormat_4, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-5,5), space = "Lab") + labs(title = "observed data",y = "sample", x = "variable") +
  theme_minimal() + theme(legend.position="none",plot.background = element_rect())

p5 = ggplot(data = melted_cormat_5, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-5,5), space = "Lab") + labs(title = "SSVD",y = "sample", x = "variable") +
  theme_minimal() + theme(legend.position="none",plot.background = element_rect())

library(cowplot)
plot_grid(p1,p5,p3,p2, ncol = 2)

```

we notice the the flashr2 provide 4 factors in this case while flashr provides 3 factors. I think it is because we use FDR method in flashr method.

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```

### 2017-09-29 analysis/applyflahr2onsimulations.Rmd
Files commited by wflow_commit.

@@
 
-```{r}
+```{r,message=FALSE,warning=FALSE}
 library(cowplot)
@@
 base2 +annotation_custom(grob = g2, xmin = -0.6, xmax = -0.1, ymin = -0.05, ymax = 0.45)
-plot_grid(base1, base2, zoomed1,zoomed2, labels = "AUTO", ncol = 2)
+# plot_grid(base1, base2, zoomed1,zoomed2, labels = "AUTO", ncol = 2)
 
@@
 
-```{r}
+```{r,message=FALSE,warning=FALSE}
 set.seed(99)

### 2017-09-29 analysis/index.Rmd
Files commited by wflow_commit.

@@
 
-Click here to find [null check for flashr2](NULLcheck.html)
+Click here to find the [null check for flashr2](NULLcheck.html) and [densen nu case](nullcheck2.html)
+
+### apply flashr2 for the examples in FLASH paper
+
+#### simulations
+
+Click here to check the [simulatoin result (rank one and rank three)](applyflahr2onsimulations.html)

### 2017-09-29 analysis/GTEXzscoreTemplate.Rmd
Files commited by wflow_commit.

---
title: "Template for GTEx example on cluster"
author: "wei"
date: 09-29
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

## Install the package

1 please git clone the `flashr2` on PPS.

`git clone https://github.com/stephenslab/flashr2.git`

2 install `flashr2` locally

`R CMD build flashr2`

`R CMD INSTALL flashr2_0.2-2.tar.gz`

## prepare the data

### original data

Please read your original data as a $N \times P$ matrix in R.

In this case you can create a folder the restore the data as `.rds` file

```
mkdir testflashr2
cd testflashr2
cp gtexEQTL_zscore.rds GTEX/testflashr2/
```

now you have the data matrix in `gtexEQTL_zscore.rds`

to restore the result and track the errors, you can crest folders

```
mkdir output
mkdir ourlog
```

## run FLASH

Here I take the GTEx EQTL zscore as an example

### creat a .R file

```
library(ashr)
library(flashr2)
load("./gtexEQTL_zscore.rds")
Y = t(zscore)
data = flash_set_data(Y)
f_greedy = flash_add_greedy(data,Kmax=60)
f_greedy_bf = flash_backfit(data,f_greedy)
#f_greedy = flash_add_greedy(data,Kmax=60,var_type = "by_column",ash_param=list(method = "fdr"))
#f_greedy_bf = flash_backfit(data,f_greedy,var_type = "by_column",ash_param=list(method = "fdr"))
saveRDS(f_greedy,file = "./output/gflashvarcol.rds")
saveRDS(f_greedy_bf,file = "./output/bflashvarcol.rds")
```

here 60 is much larger than the sample size. our method doesn't restrict $K < \min(P,N)$. But for this case, 60 is enough. We call this file as `flashwrapper.R`

### creat a .sbatch file

```
#!/bin/bash

#SBATCH --job-name=arrayJob
#SBATCH --output=/home/weidong/HG/flash/data/GTEX/testflashr2/outlog/arrayJob_%A_%a.out
#SBATCH --error=/home/weidong/HG/flash/data/GTEX/testflashr2/outlog/arrayJob_%A_%a.err
#SBATCH --array=1
#SBATCH --time=32:00:00
#SBATCH --partition=mstephens
#SBATCH --nodes=1
#SBATCH --ntasks-per-node=1
#SBATCH --mem-per-cpu=20G

######################
# Begin work section #
######################

# Print this sub-job's task ID
cd /home/weidong/HG/flash/data/GTEX/testflashr2
Rscript --verbose flashwrapper.R
```

you can substitute the folder path `/home/weidong/HG/flash/data/GTEX/testflashr2`. and we call this `.sbatch` file as `runflash.sbatch`

## run

```
sbatch runflash.sbatch 
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```

### 2017-10-01 analysis/GTEXzscoreTemplate.Rmd
Files commited by wflow_commit.

@@
 #SBATCH --array=1
-#SBATCH --time=32:00:00
+#SBATCH --time=30:00:00
 #SBATCH --partition=mstephens
@@
 #SBATCH --ntasks-per-node=1
-#SBATCH --mem-per-cpu=20G
+#SBATCH --mem-per-cpu=8G
 

### 2017-10-04 analysis/templateforOCV.Rmd
Files commited by wflow_commit.

---
title: "apply flashr2 with OCV"
author: "wei wang"
date: 09-30
output: 
  html_document:
    code_folding: hide
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

## preparation

1. creat a folder

```
mkdir Breastcancer
```

2. put your data into this folder

here we use the file 'example.mat'. the reason we use `.mat` file is because there is a matlab package in our comparision.

3. creat the wrapper function in R and matlab

### wrapper function for OCV

```{r}
OCV_index=function(Y,k_fold = 5){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = matrix(sample(P,P),ncol = k_fold)
  rowindex = matrix(sample(N,N),ncol = k_fold)

  foldindex = array(0,dim = c(k_fold,k_fold,2))
  for(i in 1:k_fold){
    for(j in 1:k_fold){
      foldindex[i,j,1] = i
      foldindex[i,j,2] = (i+j) %% k_fold
    }
  }
  foldindex[which(foldindex == 0)] = k_fold

  return(list(foldindex = foldindex, rowindex = rowindex, colindex = colindex))
}
# OCVindex = OCV_index(Y,k_fold = 5)

OCV_data = function(Y,OCVindex,k_fold = 5){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = OCVindex$colindex
  rowindex = OCVindex$rowindex
  foldindex = OCVindex$foldindex
  missing= array(0,dim = c(k_fold,N,P))
  for(i in 1:k_fold){
    missing[i, , ] = Y
    for(j in 1:k_fold){
      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
    }
  }
  return(missing)
}
# OCVdata = OCV_data(Y,OCVindex,k_fold = 5)

OCV_SSE = function(Y,OCVindex,OCVdata,k_fold = 5,method = "flash",Kmax = 50){
  colindex = OCVindex$colindex
  rowindex = OCVindex$rowindex
  foldindex = OCVindex$foldindex
  missing = OCVdata
  SSE = rep(0,k_fold)
  for(i in 1:k_fold){
    miss_hat = call_method(missing[i,,], method = method, Kmax = Kmax)
    for(j in 1:k_fold){
      SSE[i] = SSE[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] -
                               miss_hat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2)
    }
  }
  RMSE = sqrt(sum(SSE)/(dim(Y)[1] * dim(Y)[2]))
  return(RMSE)
}

call_method = function(Y_data,method,Kmax = 50){
  if(method == "flash"){
    data = flashr2::flash_set_data(Y_data)
    f_greedy = flashr2::flash_add_greedy(data,Kmax=Kmax,ash_param=list(method = "fdr"),var_type = "constant")
    Y_hat = f_greedy$EL %*% t(f_greedy$EF)
  }else if(method == "PMD"){
    out = PMA::PMD(Y_data, K = Kmax)
    Y_hat =  out$u %*% diag(out$d) %*% t(out$v)
  }else if(method == "softImpute"){
    gsoft = softImpute(Y_data, rank.max = Kmax)
    Y_hat =  gsoft$u %*% diag(gsoft$d) %*% t(gsoft$v)
  }else{
    stop("the method is not implemented yet, please check it out")
  }
  return(Y_hat)
}

NSF_OCV = function(OCVindex){
  if(file.exists("NBSFAout")){
    unlink("NBSFAout", recursive= T)
  }
  system("mkdir NBSFAout")
  # write the missing index

  writeMat("./NBSFAout/OCVindex.mat", OCVindex = OCVindex)

  system("matlab -nosplash -nodesktop -r \"addpath(\'~/HG/flash/data/OCVmissflashr2/Breastcancer\'); run(\'missvalue.m\');exit;\" ")
  g_nsfa = readMat("./NBSFAout/NSFAresult.mat")
  nsfa_rmse = as.vector(g_nsfa$RMSE)
  return(nsfa_rmse)
}

```

Here you need change the `\"addpath(\'~/HG/flash/data/OCVmissflashr2/Breastcancer\');` to your folder which contains `missvalue.m` file. this file name is `wrapper.R`

### matlab code

```{}
addpath('~/HG/flash/data/missingvalue/methods/NBSF/nsfa-master/');
addpath('~/HG/flash/data/missingvalue/methods/NBSF/nsfa-master/utils/');

% you need to change the path and the data file (not centered)
load ~/HG/flash/data/OCVmissflashr2/Breastcancer/example.mat;
Ycentered=Y-repmat(mean(Y,2),1,size(Y,2));
settings=defaultsettings();
[settings.D,settings.N]=size(Ycentered);
settings.iterations=100;
% now the missing is decide by the this is the missing index.(or you can upload the missing matrix)

load(fullfile(pwd,'NBSFAout/OCVindex.mat'))

missindex = ones(settings.D,settings.N);
k_fold = 5;
foldindex = zeros(k_fold,k_fold,2);
for k = 1:2
    for j = 1:k_fold
        for i = 1:k_fold
            foldindex(i,j,k) = OCVindex.foldindex(i + k_fold*(j - 1) + k_fold * k_fold * (k - 1));
        end
    end
end

SSE = zeros(1,k_fold);
for i = 1:k_fold
    mvmask = missindex;
    for j = 1:k_fold
        mvmask(OCVindex.rowindex(:,foldindex(j,i,1)),OCVindex.colindex(:,foldindex(j,i,2))) = 0;
    end
    initialsample=init_nsfa(settings);
    [finalsample,resultstable]=nsfa(Ycentered,mvmask,initialsample,settings);
    Y_hat = (finalsample.G *  finalsample.X);
    for j = 1:k_fold
        SSE(i) = SSE(i) + ...
            sum(sum((Ycentered(OCVindex.rowindex(:,foldindex(j,i,1)),OCVindex.colindex(:,foldindex(j,i,2)))- ...
            Y_hat(OCVindex.rowindex(:,foldindex(j,i,1)),OCVindex.colindex(:,foldindex(j,i,2)))).^2));
    end

end

RMSE = sqrt(sum(SSE)/(settings.D*settings.N));
RMSE;

save_path = fullfile(pwd,'NBSFAout/NSFAresult');
save(save_path,'RMSE'); 
```

you need change the path as well. this file name is 'missvalue.m'

### creat the run.R file to run the code

In this case, we have problem in running `flashr2` on RCC. so we just try other methods first.

```{}
library(R.matlab)
## run the code
Y_centered = readMat("~/HG/flash/data/OCVmissflashr2/Breastcancer/example.mat")
Y = Y_centered$Y
# in the matlab package of NSF, the use the centered data by rows
N = dim(Y)[1]
P = dim(Y)[2]
Y = Y - rowMeans(Y) %*% t(rep(1,P))
OCVindex = OCV_index(Y,k_fold = 5)
OCVdata = OCV_data(Y,OCVindex,k_fold = 5)
soft_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "softImpute",Kmax = 50)
# flashG_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "flash",Kmax = 50)
pmd_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "PMD",Kmax = 50)
nsfa_rmse = NSF_OCV(OCVindex)
flashG_rmse = mean(c( nsfa_rmse,pmd_rmse,soft_rmse))

result = c(flashG_rmse, nsfa_rmse,pmd_rmse,soft_rmse)
saveRDS(result, "./NBSFAout/output.rds")

```

this file name is `runOCV.R`

### creat folder to track the result

```
mkdir outlog
```

### creat .sbatch file

```
#!/bin/bash

#SBATCH --job-name=arrayJob
#SBATCH --output=/home/weidong/HG/flash/data/OCVmissflashr2/Breastcancer/outlog/arrayJob_%A_%a.out
#SBATCH --error=/home/weidong/HG/flash/data/OCVmissflashr2/Breastcancer/outlog/arrayJob_%A_%a.err
#SBATCH --array=1-100
#SBATCH --time=06:00:00
#SBATCH --partition=broadwl
#SBATCH --ntasks=1
#SBATCH --mem-per-cpu=2000

######################
# Begin work section #
######################

# Print this sub-job's task ID
cd /home/weidong/HG/flash/data/OCVmissflashr2/Breastcancer
mkdir test${SLURM_ARRAY_TASK_ID}
cd test${SLURM_ARRAY_TASK_ID}
Rscript --verbose /home/weidong/HG/flash/data/OCVmissflashr2/Breastcancer/runOCV.R
```

this file is name as `OCV.sbatch`

## run OCV

now in your folder, you should have:

```
[weidong@midway2-login1 Breastcancer]$ ls
OCV.sbatch  example.mat  missvalue.m  outlog  runOCV.R  wrapper.R
```

```
sbatch OCV.sbatch
```

## to get the result

```{}
T = 100
results = matrix(NA,ncol = 4, nrow = T)
for(i in 1:T){
  test_folder = paste("test", i, sep = "")
  out_folder = "NBSFAout"
  out_file = "output.rds"
  file_name = file.path(test_folder,out_folder,out_file)
  results[i,] = try(readRDS(file_name))
}

saveRDS(results,"./boxplot.rds")
```

## plot the result

```{r}
plot_res = function(output,title = "data",legend_position = "none"){
  rmse = as.vector(output)
  N = dim(output)[1]
  methods = rep(c("flash","NBSFA","PMD","softImpute"), each = N)
  df = data.frame(RMSE = rmse, Method = methods )
  p<-ggplot(df, aes(x=Method, y=RMSE, color=Method)) +
  #geom_boxplot()
  geom_violin()+
  ggtitle(title)+theme_gray() +
    theme(legend.position= legend_position, legend.text=element_text(size=15), axis.text.y = element_text(size =6))
  p
}

#gtex_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/boxplot_gtex.rds")
#pg = plot_res(gtex_res,"GTEx data")
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```

### 2017-10-04 analysis/templateforOCV.Rmd
Files commited by wflow_commit.

@@
   p<-ggplot(df, aes(x=Method, y=RMSE, color=Method)) +
-  #geom_boxplot()
-  geom_violin()+
+  geom_boxplot()+
+  # geom_violin()+
   ggtitle(title)+theme_gray() +
@@
 
-#gtex_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/boxplot_gtex.rds")
-#pg = plot_res(gtex_res,"GTEx data")
+GZ_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/gtexzscore_box.rds")
+pg = plot_res(GZ_res,"GTEx data")
+BC_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/BreastCancer_box.rds")
+pb = plot_res(BC_res,"Breast Cancer data")
+PT_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/president_box.rds")
+pp = plot_res(PT_res,"Text data")
+DT_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/denoiseRtumor_box.rds")
+pd = plot_res(DT_res,"Tumor data")
+```
+
+## plots
+
+```{r}
+pg
+pb
+pp
+pd
 ```

### 2017-10-04 analysis/GTEXzscoreTemplate.Rmd
Files commited by wflow_commit.

@@
 
+## plot
+
+```{}
+b_flash = readRDS("../data/GTExdata/res_flashr2/bflashvarcol.rds")
+load("../data/GTExdata/gtexEQTL_zscore.rds")
+ssY = sum(zscore^2)
+K = dim(b_flash$EL)[2] -1
+pve = (sapply(seq(1,K),function(x){ sum(b_flash$EL[,x]^2 %*% t(b_flash$EF[,x]^2)) }))/ssY
+pve = pmax(round(pve,3),0.001)
+
+dat = read.table('../data/GTExColors.txt', sep = '\t', comment.char = '')
+colordata = dat[c(1:6,9:18,21:23,26:30,32,33,35,36,38:53),1:2]
+
+L = b_flash$EL[,1:14]
+library(reshape2)
+data_L = melt(L)
+colnames(data_L) = c("tissue","loading","value")
+library(ggplot2)
+tissue_color = as.character(colordata[,2])
+data_L$tissue = factor(data_L$tissue,levels = 1:44 ,labels = as.character(colordata[,1]) )
+data_L$loading = factor(data_L$loading,levels = 1:14 ,labels = paste("loading",1:14,"; pve:", pve[1:14]))
+
+ggplot(data_L,aes(x = tissue,y = value,fill = factor(tissue) )) +
+  geom_bar(stat = "identity",width = 0.6) +
+  scale_fill_manual(values=tissue_color) +
+  scale_x_discrete(labels = NULL) +
+  theme_grey()+
+  theme(legend.position="right", legend.text=element_text(size=5), axis.text.y = element_text(size = 5)) + 
+  labs(title = "GTEx data", y = "loading values" ,x = "tissues", fill="tissue") +
+  facet_wrap(~loading, ncol = 2, scales = "free_y") +
+  guides(fill = guide_legend(ncol = 1, keyheight = 0.5, keywidth = 0.5))
+  ggsave("flashr2GTEx1.pdf", width = 8, height = 11)
+
+# the 27th factor is zero
+L = b_flash$EL[,15:26]
+library(reshape2)
+data_L = melt(L)
+colnames(data_L) = c("tissue","loading","value")
+library(ggplot2)
+tissue_color = as.character(colordata[,2])
+data_L$tissue = factor(data_L$tissue,levels = 1:44 ,labels = as.character(colordata[,1]) )
+data_L$loading = factor(data_L$loading,levels = 1:12 ,labels = paste("loading",15:26,"; pve:", pve[15:26]))
+
+ggplot(data_L,aes(x = tissue,y = value,fill = factor(tissue) )) +
+  geom_bar(stat = "identity",width = 0.6) +
+  scale_fill_manual(values=tissue_color) +
+  scale_x_discrete(labels = NULL) +
+  theme_grey()+
+  theme(legend.position="right", legend.text=element_text(size=5), axis.text.y = element_text(size = 5)) + 
+  labs(title = "GTEx data", y = "loading values" ,x = "tissues", fill="tissue") +
+  facet_wrap(~loading, ncol = 2, scales = "free_y") +
+  guides(fill = guide_legend(ncol = 1, keyheight = 0.5, keywidth = 0.5))
+  ggsave("flashr2GTEx2.pdf", width = 8, height = 10)
+```
+
+the plots are
+
+[plot1](flashr2GTEx1.pdf)
+
+[plot2](flashr2GTEx2.pdf)
+
 ## Session information

### 2017-10-05 analysis/GTEXzscoreTemplate.Rmd
Files commited by wflow_commit.

@@
 
-[plot1](flashr2GTEx1.pdf)
+<img src="figure/flashr2GTEx1.pdf" alt="the plot is in analysis/flashr2GTEx1.pdf"  width="4200" height="4200">
 
-[plot2](flashr2GTEx2.pdf)
+<img src="figure/flashr2GTEx2.pdf" alt="the plot is in analysis/flashr2GTEx1.pdf"  width="4200" height="4200">
+
+## an example to run SFA on RCC or PPS
+
+1. go to the folder where the data is.
+
+### sfa wrapper
+
+this example is on PPS cluster
+
+```{r}
+SFA.wrapper = function(X, K){
+  N = dim(X)[1]
+  P = dim(X)[2]
+
+  if(file.exists("SFAout")){
+    unlink("SFAout", recursive= T)
+  }
+  system("mkdir SFAout")
+
+  write.table(X,file="dscsfa.txt",row.names=F,col.names=F)
+  # sfa command line on cluster
+  # ~/flash/simulation/methods/sfa/src/sfa -gen ./dscsfamix.txt -g 600 -k 1 -n 200 -iter 100 -rand 999 -o ./SFAout/sfa
+
+  # sqrt(mean(( loadings %*% factor - test$L_true%*%t(test$F_true))^2))/sqrt(mean((test$L_true%*%t(test$F_true))^2))
+  system(paste("~/flash/simulation/methods/sfa/src/sfa", "-gen", "./dscsfa.txt", "-g", N, "-k", K,
+               "-n", P, "-iter", 100, "-rand", 999, "-o", "./SFAout/sfa", sep=" "))
+
+  if(file.info("./SFAout/sfa_F.out")$size == 1){
+    return(list(L = 0, F = 0))
+  } else{
+    Fhat=read.table("./SFAout/sfa_F.out")
+    lambda=read.table("./SFAout/sfa_lambda.out")
+    Fhat=as.matrix(Fhat)
+    lambda=as.matrix(lambda)
+    return(list(L = lambda, F = Fhat))
+  }
+
+}
+```
+
+this is `sfawrapper.R`
+
+### creat run_sfa.R file
+
+```
+load("~/flash/dataanalysis/GTeX/EQTLzscore/gtexEQTL_zscore.rds")
+setwd("~/flash/dataanalysis/GTeX/EQTLzscore/")
+source("./sfawrapper.R")
+source("./SFAmixwrapper.R")
+Y = t(zscore)
+N = dim(Y)[1]
+P = dim(Y)[2]
+
+gsfamix = SFAmix.wrapper(Y,30)
+gsfa = SFA.wrapper(Y, 30)
+saveRDS(gsfa, file = "~/flash/dataanalysis/GTeX/EQTLzscore/sfaGTExzscore.rds")
+saveRDS(gsfamix, file = "~/flash/dataanalysis/GTeX/EQTLzscore/sfamixGTExzscore.rds")
+
+```
+
+and run it
+
+```
+R CMD BATCH run_sfa.R
+```
+
+SFA provide NaN result when set `K = 30` and `K = 26`. (all the $L_i$ and $F_j$ are NaN).
 

### 2017-10-05 analysis/GTEXzscoreTemplate.Rmd
Files commited by wflow_commit.

@@
 }
+
+SFAmix.wrapper = function(X,K){
+  N = dim(X)[1]
+  P = dim(X)[2]
+
+  write.table(X,file="dscsfamix.txt",row.names=F,col.names=F)
+  print(class(X))
+  print(dim(X))
+
+  if(file.exists("SFAmixout")){
+    unlink("SFAmixout", recursive= T)
+  }
+  system("mkdir SFAmixout")
+
+  # ~/mvash/mvsim/SFAmix/SFAmix --nf K --y dscrsfamix.txt --out SFAmixout --sep space
+  # this is run on the PPS cluster
+  system(paste("~/mvash/mvsim/SFAmix/SFAmix","--nf",K,"--y","dscsfamix.txt","--out",
+               "SFAmixout","--sep","space",sep=" "))
+
+  alpha=read.table("./SFAmixout/PSI")
+
+  if(file.info("./SFAmixout/EX")$size == 1){
+    return(list(L = 0, F = 0))
+  } else{
+    Fhat=read.table("./SFAmixout/EX")
+    lambda=read.table("./SFAmixout/LAM")
+    Psi=as.vector(alpha)
+    Fhat=as.matrix(Fhat)
+    lambda=as.matrix(lambda)
+    P=dim(lambda)[2]
+    n=dim(Fhat)[1]
+    return(list(L = Fhat, F = lambda))
+  }
+
+}
+
 ```
@@
 source("./SFAmixwrapper.R")
-Y = t(zscore)
+Y = zscore
 N = dim(Y)[1]
 P = dim(Y)[2]
+gsfa = SFA.wrapper(Y, 26)
+saveRDS(gsfa, file = "~/flash/dataanalysis/GTeX/EQTLzscore/sfaGTExzscore.rds")
 
 gsfamix = SFAmix.wrapper(Y,30)
-gsfa = SFA.wrapper(Y, 30)
-saveRDS(gsfa, file = "~/flash/dataanalysis/GTeX/EQTLzscore/sfaGTExzscore.rds")
 saveRDS(gsfamix, file = "~/flash/dataanalysis/GTeX/EQTLzscore/sfamixGTExzscore.rds")

### 2017-10-05 analysis/GTEXzscoreTemplate.Rmd
Files commited by wflow_commit.

@@
 date: 09-29
-output: html_document
+output: 
+  html_document:
+    code_folding: hide
 ---
@@
 P = dim(Y)[2]
-gsfa = SFA.wrapper(Y, 26)
+gsfa = SFA.wrapper(Y, 30)
 saveRDS(gsfa, file = "~/flash/dataanalysis/GTeX/EQTLzscore/sfaGTExzscore.rds")
 
+Y = t(zscore)
 gsfamix = SFAmix.wrapper(Y,30)
@@
 
-SFA provide NaN result when set `K = 30` and `K = 26`. (all the $L_i$ and $F_j$ are NaN).
+#### remark
+
+For `Y = t(zscore)`, SFA provide NaN result when set `K = 30` and `K = 26`. (all the $L_i$ and $F_j$ are NaN). So I use `Y = zscore` which works in `K = 30` and `K = 26`. In this case, I set `K = 30`
+
+in this case we can get
+
+```
+> gsfa = SFA.wrapper(Y, 30)
+ Setting G to be 16069
+ Setting N to be 44
+Trying to read in matrix with G=16069 and N =44
+Opening matrix ./dscsfa.txt with 16069 rows and 44 columns
+Opening matrix ./dscsfa.txt with 16069 rows and 44 columns
+read in matrix 16069 by 44
+initializing sfa...
+G = 16069
+N = 44
+K = 30
+
+```
+
+But for `SFAmix` we should use `Y = t(zscore)`.
+
+The approaches make sure we can get $(\hat{\sigma}_1^2,\cdots,\hat{\sigma}_j^2,\cdots)$ is a `16069` vector.
+
+### plot SFA result
+
+```{r}
+sfares = readRDS("../data/GTExdata/SFAres/sfaGTExzscore.rds")
+par(mfrow = c(2,3),mar=c(5.1,4.1,4.1,2.1)-1.9)
+for(k in 1:30){
+  barplot(sfares$F[,k],main = paste("factor",k))
+}
+
+```
 

### 2017-10-05 analysis/GTEXzscoreTemplate.Rmd
Files commited by wflow_commit.

@@
 for(k in 1:30){
-  barplot(sfares$F[,k],main = paste("factor",k))
+  barplot(sfares$F[k,],main = paste("factor",k))
 }
@@
 ```
+
+### SFA K=3
+
+```{r}
+g3 = readRDS('../data/GTExdata/SFAres/sfaK_3.rds')
+par(mfrow = c(1,3),mar=c(5.1,4.1,4.1,2.1)-1.9)
+for(k in 1:3){
+  barplot(g3$F[k,],main = paste("factor",k))
+}
+```
+
+### plot SFAmix result
+
+```{r}
+sfamixres = readRDS("../data/GTExdata/SFAres/sfamixGTExzscore.rds")
+dim(sfamixres$L)
+dim(sfamixres$F)
+```
+
+this is only a rank one matrix. I tried two time. two runs give rank one matrix.
 

### 2017-10-05 analysis/GTEXzscoreTemplate.Rmd
Files commited by wflow_commit.

@@
 sfares = readRDS("../data/GTExdata/SFAres/sfaGTExzscore.rds")
+# the F is K by N matrix! not N by K matrix
 par(mfrow = c(2,3),mar=c(5.1,4.1,4.1,2.1)-1.9)

### 2017-10-06 analysis/GTEXzscoreTemplate.Rmd
Files commited by wflow_commit.

@@
 
-### plot SFA result
+### SFA K=3
+
+```{r}
+g3 = readRDS('../data/GTExdata/SFAres/sfaK_3.rds')
+par(mfrow = c(1,3),mar=c(5.1,4.1,4.1,2.1)-1.9)
+for(k in 1:3){
+  barplot(g3$F[k,],main = paste("factor",k))
+}
+```
+
+### plot SFA result K = 30
 
@@
 
-### SFA K=3
+### plot of SFA result with t(Y) K = 10
 
 ```{r}
-g3 = readRDS('../data/GTExdata/SFAres/sfaK_3.rds')
-par(mfrow = c(1,3),mar=c(5.1,4.1,4.1,2.1)-1.9)
-for(k in 1:3){
-  barplot(g3$F[k,],main = paste("factor",k))
+gt10 = readRDS('../data/GTExdata/SFAres/sfaTYK_10.rds')
+
+par(mfrow = c(2,3),mar=c(5.1,4.1,4.1,2.1)-1.9)
+for(k in 1:10){
+  barplot(gt10$L[,k],main = paste("factor",k))
 }
+```
+
+#### problem in K = 20 and K = 30 in this case.
+
+```
+gsl: lu.c:147: ERROR: matrix is singular
+Default GSL error handler invoked.
 ```

### 2017-10-06 analysis/GTEXzscoreTemplate.Rmd
Files commited by wflow_commit.

@@
 
+If you would like to use `t(Y)` you should use `-vn` in the `sfa` command line in order to get $(\hat{\sigma}_1^2,\cdots,\hat{\sigma}_j^2,\cdots)$ is a `16069` vector
+
+you need add `-mg` to get the mean vector with length of 44
+
 ```{r}
-gt10 = readRDS('../data/GTExdata/SFAres/sfaTYK_10.rds')
-
+gt10 = readRDS('../data/GTExdata/SFAres/sfaTK_10m.rds')
+mu = sapply(seq(1:44),function(x){gt10$mu[[x]]})
 par(mfrow = c(2,3),mar=c(5.1,4.1,4.1,2.1)-1.9)
+barplot(mu,main = "mean vector")
 for(k in 1:10){
-  barplot(gt10$L[,k],main = paste("factor",k))
+  barplot(gt10$LF$L[,k],main = paste("factor",k))
 }

### 2017-10-06 analysis/GTEXzscoreTemplate.Rmd
Files commited by wflow_commit.

@@
 ```{r}
+load("../data/GTExdata/gtexEQTL_zscore.rds")
 gt10 = readRDS('../data/GTExdata/SFAres/sfaTK_10m.rds')
+ssy = sum(zscore^2)
+pve = sapply(seq(1:10),function(x){sum((gt10$LF$L[,x] %*% t(gt10$LF$F[x,]))^2)})
+pve = pve/ssy
+
 mu = sapply(seq(1:44),function(x){gt10$mu[[x]]})
@@
 for(k in 1:10){
-  barplot(gt10$LF$L[,k],main = paste("factor",k))
+  barplot(gt10$LF$L[,k],main = paste("factor",k,"pve",round(pve[k],3)))
 }

### 2017-10-06 analysis/templateforOCV.Rmd
Files commited by wflow_commit.

@@
 
+PT_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/president_box.rds")
+pp = plot_res(PT_res,"Text data")
+DT_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/denoiseRtumor_box.rds")
+pd = plot_res(DT_res,"Tumor data")
 GZ_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/gtexzscore_box.rds")
@@
 pb = plot_res(BC_res,"Breast Cancer data")
-PT_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/president_box.rds")
-pp = plot_res(PT_res,"Text data")
-DT_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/denoiseRtumor_box.rds")
-pd = plot_res(DT_res,"Tumor data")
+ML_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/ML100K_box.rds")
+ML_res[c(2,13,17,21,29,37,62,76,77,93,95,100),] = NA
+ML_res = matrix(as.numeric(ML_res),ncol = 4)
+pM = plot_res(ML_res,"Movie Lens data")
 ```
@@
 ```{r}
+pp
+pd
 pg
 pb
-pp
-pd
+pM
 ```

### 2017-10-06 analysis/templateforOCV.Rmd
Files commited by wflow_commit.

@@
 ```{r}
+library(ggplot2)
 plot_res = function(output,title = "data",legend_position = "none"){

### 2017-10-07 analysis/templateforOCV.Rmd
Files commited by wflow_commit.

--- 2017-10-06 analysis/templateforOCV.Rmd
@@
   # geom_violin()+
-  ggtitle(title)+theme_gray() +
+  ggtitle(title) +  theme_bw()+
     theme(legend.position= legend_position, legend.text=element_text(size=15), axis.text.y = element_text(size =6))
@@
 ```{r}
-pp
-pd
-pg
-pb
-pM
+gridExtra::grid.arrange(pp,pd,pg,pb,pM, layout_matrix = rbind(c(1,NA,2),c(NA,5,NA),c(4,NA,3)))
 ```

### 2017-10-11 analysis/nullcheck2.Rmd
Files commited by wflow_commit.

--- 2017-09-17 analysis/NULLcheck.Rmd
@@
 
+## model
+
+In this case the data sets are:
+$$H_0: X \sim N(0, \Sigma_0)$$
+$$H_1: X \sim N(0,\Sigma_1)$$
+where $\Sigma_0 = I$ and $\Sigma_1 = I + \nu \nu^T$
+
+### dense case
+
 ```{r}
@@
 set.seed(99)
-T = 1000
-lambda_3 = rep(NA,T)
-flash_rank_3= rep(1,T)
+T = 300
+lambda_0 = rep(NA,T)
+flash2_rank_0= rep(1,T)
 for(t in 1:T){
@@
   X = mvrnorm(n = N,mu = rep(0,P), Sigma = Sigma_0)
-  lambda_3[t] =  svd(X)$d[1]
-  data = set_flash_data(X)
-  g1= flash_r1(data,verbose=F)
-  if(sum(g1$EL) ==0 )  {flash_rank_3[t] = 0}
+  X = X/sqrt(mean((X)^2))
+  lambda_0[t] =  svd(X)$d[1]
+  data =  flashr2::flash_set_data(X)
+  g1= flashr2::flash_r1(data,verbose=F)
+  if(sum(g1$EL) ==0 )  {flash2_rank_0[t] = 0}
 }
@@
 set.seed(99)
-T = 1000
-lambda_4 = rep(NA,T)
-flash_rank_4= rep(1,T)
+lambda_1 = rep(NA,T)
+flash2_rank_1= rep(1,T)
+for(t in 1:T){
+  P = 300
+  N = 100
+  Sigma_0 = diag(rep(1,P))
+  nu = rnorm(P, 0, (2)/(sqrt(P)))
+  # index_nu = sample(seq(1:P),(P*0.95))
+  # nu[index_nu] = 0
+  Sigma_1 = Sigma_0 + nu %*% t(nu)
+  X = mvrnorm(n = N,mu = rep(0,P), Sigma = Sigma_1)
+  X = X/sqrt(mean((X)^2))
+  lambda_1[t] =  svd(X)$d[1]
+  data = flashr2::flash_set_data(X)
+  g1= flashr2::flash_r1(data,verbose=F)
+  if(sum(g1$EL) ==0 )  {flash2_rank_1[t] = 0}
+}
+
+lambda_0_0 = lambda_0[which(flash2_rank_0 ==0)]
+lambda_0_1 = lambda_0[which(flash2_rank_0 ==1)]
+
+lambda_1_0 = lambda_1[which(flash2_rank_1 ==0)]
+lambda_1_1 = lambda_1[which(flash2_rank_1 ==1)]
+
+simple_roc <- function(labels, scores){
+  labels <- labels[order(scores, decreasing=TRUE)]
+  data.frame(TPR=cumsum(labels)/sum(labels), FPR=cumsum(!labels)/sum(!labels), labels)
+}
+
+lambda.roc = simple_roc(labels = c(rep(0,T),rep(1,T)),scores = c(lambda_0,lambda_1))
+plot(lambda.roc$FPR,lambda.roc$TPR,type = "l",col = "black")
+flash.roc = simple_roc(labels = c(rep(0,T),rep(1,T)),scores = c(flash2_rank_0,flash2_rank_1))
+lines(flash.roc$FPR,flash.roc$TPR,type = "l",col = "red")
+```
+
+### sparse case
+
+```{r}
+########  spiked covariance model
+
+set.seed(99)
+T = 300
+lambda_0 = rep(NA,T)
+flash2_rank_0= rep(1,T)
+for(t in 1:T){
+  P = 300
+  N = 100
+  Sigma_0 = diag(rep(1,P))
+  X = mvrnorm(n = N,mu = rep(0,P), Sigma = Sigma_0)
+  X = X/sqrt(mean((X)^2))
+  lambda_0[t] =  svd(X)$d[1]
+  data =  flashr2::flash_set_data(X)
+  g1= flashr2::flash_r1(data,verbose=F)
+  if(sum(g1$EL) ==0 )  {flash2_rank_0[t] = 0}
+}
+
+set.seed(99)
+lambda_1 = rep(NA,T)
+flash2_rank_1= rep(1,T)
 for(t in 1:T){
@@
   X = mvrnorm(n = N,mu = rep(0,P), Sigma = Sigma_1)
-  lambda_4[t] =  svd(X)$d[1]
-  data = set_flash_data(X)
-  g1= flash_r1(data,verbose=F)
-  if(sum(g1$EL) ==0 )  {flash_rank_4[t] = 0}
+  X = X/sqrt(mean((X)^2))
+  lambda_1[t] =  svd(X)$d[1]
+  data = flashr2::flash_set_data(X)
+  g1= flashr2::flash_r1(data,verbose=F)
+  if(sum(g1$EL) ==0 )  {flash2_rank_1[t] = 0}
 }
 
-lambda_4_0 = lambda_4[which(flash_rank_4 ==0)]
-lambda_4_1 = lambda_4[which(flash_rank_4 ==1)]
+lambda_0_0 = lambda_0[which(flash2_rank_0 ==0)]
+lambda_0_1 = lambda_0[which(flash2_rank_0 ==1)]
 
-library(ggplot2)
-dat = data.frame(lambda = c(lambda_3,lambda_4_0,lambda_4_1), name = c(rep("null",length(lambda_3)), rep("alt 0", length(lambda_4_0)),rep("alt 1", length(lambda_4_1)) ))
-ggplot(dat,aes(x=lambda)) +
-  geom_histogram(data=subset(dat,name == 'null'),fill = "red", alpha = 0.2) +
-  # geom_histogram(data=subset(dat,name == 'alt 0'),fill = "blue", alpha = 0.2) +
-  geom_histogram(data=subset(dat,name %in% c('alt 1','alt 0')),fill = "green", alpha = 0.2)
+lambda_1_0 = lambda_1[which(flash2_rank_1 ==0)]
+lambda_1_1 = lambda_1[which(flash2_rank_1 ==1)]
+
+simple_roc <- function(labels, scores){
+  labels <- labels[order(scores, decreasing=TRUE)]
+  data.frame(TPR=cumsum(labels)/sum(labels), FPR=cumsum(!labels)/sum(!labels), labels)
+}
+
+lambda.roc = simple_roc(labels = c(rep(0,T),rep(1,T)),scores = c(lambda_0,lambda_1))
+plot(lambda.roc$FPR,lambda.roc$TPR,type = "l",col = "black")
+flash.roc = simple_roc(labels = c(rep(0,T),rep(1,T)),scores = c(flash2_rank_0,flash2_rank_1))
+lines(flash.roc$FPR,flash.roc$TPR,type = "l",col = "red")
 ```
 
--red: null
--green: alternative
-
-```{r}
-library(ggplot2)
-dat = data.frame(lambda = c(lambda_3,lambda_4_0,lambda_4_1), name = c(rep("null",length(lambda_3)), rep("alt 0", length(lambda_4_0)),rep("alt 1", length(lambda_4_1)) ))
-ggplot(dat,aes(x=lambda)) +
-  geom_histogram(data=subset(dat,name == 'null'),fill = "red", alpha = 0.2) +
-  geom_histogram(data=subset(dat,name == 'alt 0'),fill = "blue", alpha = 0.2) +
-  geom_histogram(data=subset(dat,name == 'alt 1'),fill = "green", alpha = 0.2)
-
-```
-
--red: null
--green: alternative estimated as rank 1
--bule: alternative estimated as rank 0
+## likelihood ratio
 

### 2017-12-04 analysis/compareOptionsOCV.Rmd
Files commited by wflow_commit.

---
title: "compare the options for missing value MSE"
author: "Wei"
date: 2017-12-04
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

## Movie Lens data

### FDR without null check vs default setting

```{r}
library(ggplot2)
plot_res = function(output,title = "data",legend_position = "none"){
  rmse = as.vector(output)
  N = dim(output)[1]
  methods = rep(c("flash_fdr","flash","NBSFA_fdr","NBSFA","PMD_fdr","PMD","softImpute_fdr","softImpute"), each = N)
  df = data.frame(RMSE = rmse, Method = methods )
  p<-ggplot(df, aes(x=Method, y=RMSE, color=Method)) +
  geom_boxplot()+
  # geom_violin()+
  ggtitle(title) +  theme_bw()+
    theme(legend.position= legend_position, legend.text=element_text(size=15), axis.text.y = element_text(size =6))
  p
}
data1 = readRDS("../data/cmpOptions/ML100K/FDRnocheck_box.rds")
data2 = readRDS("../data/cmpOptions/ML100K/box_defaultseeting.rds")
data1 = matrix(as.numeric(data1),ncol = 4)
data2 = matrix(as.numeric(data2),ncol = 4)
pM = plot_res(cbind(data1[,1],data2[,1],
                    data1[,2],data2[,2],
                    data1[,3],data2[,3],
                    data1[,4],data2[,4]),"Movie Lens data different settings")
pM
```

We run the experiments twice. `flash_fdr` `NSFA_fdr` `PMD_fdr` and `softImpute_fdr` stand for the results from the simulation data set with FDR method and without Nullcheck in FLASH method. And `flash` `NSFA` `PMD` and `softImpute` represent the results from simulation data set with default setting (shrink, normal with Nullcheck) in FLASH.

## GTEx EQTL Z-score

### FDR without null check vs default setting

```{r}
data1 = readRDS("../data/cmpOptions/GETxZscore/FDRnocheck_box.rds")
data2 = readRDS("../data/cmpOptions/GETxZscore/DefaultSetting_box.rds")
data1 = matrix(as.numeric(data1),ncol = 4)
data2 = matrix(as.numeric(data2),ncol = 4)
pG = plot_res(cbind(data1[,1],data2[,1],
                    data1[,2],data2[,2],
                    data1[,3],data2[,3],
                    data1[,4],data2[,4]),"GTEx z-score different settings")
pG
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```

### 2017-12-04 analysis/GSEA_flashr.Rmd

--- 2017-12-04 analysis/compareOptionsOCV.Rmd
@@
 ---
-title: "compare the options for missing value MSE"
+title: "GSEA for flash factors"
 author: "Wei"
-date: 2017-12-04
+date: 10-2017
 output: html_document
@@
 
-## Movie Lens data
+## try the 3rd factor
 
-### FDR without null check vs default setting
+Here, we should preare the list of the gene names.
 
 ```{r}
-library(ggplot2)
-plot_res = function(output,title = "data",legend_position = "none"){
-  rmse = as.vector(output)
-  N = dim(output)[1]
-  methods = rep(c("flash_fdr","flash","NBSFA_fdr","NBSFA","PMD_fdr","PMD","softImpute_fdr","softImpute"), each = N)
-  df = data.frame(RMSE = rmse, Method = methods )
-  p<-ggplot(df, aes(x=Method, y=RMSE, color=Method)) +
-  geom_boxplot()+
-  # geom_violin()+
-  ggtitle(title) +  theme_bw()+
-    theme(legend.position= legend_position, legend.text=element_text(size=15), axis.text.y = element_text(size =6))
-  p
-}
-data1 = readRDS("../data/cmpOptions/ML100K/FDRnocheck_box.rds")
-data2 = readRDS("../data/cmpOptions/ML100K/box_defaultseeting.rds")
-data1 = matrix(as.numeric(data1),ncol = 4)
-data2 = matrix(as.numeric(data2),ncol = 4)
-pM = plot_res(cbind(data1[,1],data2[,1],
-                    data1[,2],data2[,2],
-                    data1[,3],data2[,3],
-                    data1[,4],data2[,4]),"Movie Lens data different settings")
-pM
-```
-
-We run the experiments twice.
-`flash_fdr` `NSFA_fdr` `PMD_fdr` and `softImpute_fdr` stand for the results from the simulation data set with FDR method and without Nullcheck in FLASH method.
-And `flash` `NSFA` `PMD` and `softImpute` represent the results from simulation data set with default setting (shrink, normal with Nullcheck) in FLASH.
-
-## GTEx EQTL Z-score
-
-### FDR without null check vs default setting
-
-```{r}
-data1 = readRDS("../data/cmpOptions/GETxZscore/FDRnocheck_box.rds")
-data2 = readRDS("../data/cmpOptions/GETxZscore/DefaultSetting_box.rds")
-data1 = matrix(as.numeric(data1),ncol = 4)
-data2 = matrix(as.numeric(data2),ncol = 4)
-pG = plot_res(cbind(data1[,1],data2[,1],
-                    data1[,2],data2[,2],
-                    data1[,3],data2[,3],
-                    data1[,4],data2[,4]),"GTEx z-score different settings")
-pG
+res = readRDS("../data/GTEX/bflashvarcol.rds")
+load("../data/GTEX/gtexEQTL_zscore.rds")
+sorted_f3 = sort(res$EF[,3],index.return = T)
+negative_index = sorted_f3$ix[which(sorted_f3$x<0)]
+negative_name = row.names(zscore)[negative_index]
+new_names = sapply(negative_name, function(x){(strsplit(x,"\\.")[[1]])[1]})
+# write.table(new_names,file = "../data/GTEX/f_3_neg.txt",row.names = F,col.names = F)
 ```

### 2017-12-09 analysis/MSErank1sim.Rmd
Files commited by wflow_commit.

---
title: "MSE for rank_1 simulation"
author: "wei wang"
date: 17-12-09
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

```{}
flash.wrapper = function(Y_data){
  # missindex is a matirx with 3 column here: i j x
  # Y has miss value already
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  data =  flashr2::flash_set_data(Y)
  g_flash = flashr2::flash_r1(data,verbose=F)
  Y_hat = g_flash$EL %*% t(g_flash$EF)
  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
  return(RMSE)
}

#here I use SVD instead of KSVD.
SVD.wrapper = function(Y_data){
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  gsvd = svd(Y)
  Y_hat = gsvd$d[1] * (gsvd$u[,1] %*% t(gsvd$v[,1]))
  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
  return(RMSE)
}

SSVD.wrapper = function(Y_data){
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  gssvd = ssvd::ssvd(Y,method = "method")
  Y_hat = gssvd$d * (gssvd$u %*% t(gssvd$v))
  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
  return(RMSE)
}
```

```{}
library(PMA)
library(flashr2)
library(ssvd)
source("~/HG/flash/data/simulations/rankone/MSE/Rfunction.R")
L_se = c( 0.25, 0.5, 1, 2, 4)
L_pi = c(1, 1,1,1,1)
L_pi = L_pi / sum(L_pi)
N = 200
P = 300
# Data = datamaker(N,P,L_pi,L_se,0.7,c(1),c(1),1,sqrt(16))
Data = datamaker(N,P,L_pi,L_se,0.1,c(1),c(1),1,sqrt(1))
RMSE = rep(NA,6)
RMSE[1] = PMA.wrapper(Y_data)
RMSE[2] = flash.wrapper(Y_data)
RMSE[3] = SFA.wrapper(Y_data, K = 1)
RMSE[4] = SFAmix.wrapper(Y_data,K = 1)
RMSE[5] = SVD.wrapper(Y_data)
RMSE[6] = SSVD.wrapper(Y_data)
saveRDS(result, "./output.rds")
```

$$RMSE =\frac{ \sqrt{(Y-\hat{Y} - E )^2}}{\sqrt{(Y - E )^2}}$$

## batch file

```
#!/bin/bash

#SBATCH --job-name=arrayJob
#SBATCH --output=/home/weidong/HG/flash/data/simulations/rankone/MSE/median/outlog/arrayJob_%A_%a.out
#SBATCH --error=/home/weidong/HG/flash/data/simulations/rankone/MSE/median/outlog/arrayJob_%A_%a.err
#SBATCH --array=1-100
#SBATCH --time=02:00:00
#SBATCH --partition=mstephens
#SBATCH --ntasks=1
#SBATCH --mem-per-cpu=2000

######################
# Begin work section #
######################

# Print this sub-job's task ID
cd /home/weidong/HG/flash/data/simulations/rankone/MSE/median
mkdir test${SLURM_ARRAY_TASK_ID}
cd test${SLURM_ARRAY_TASK_ID}
Rscript --verbose /home/weidong/HG/flash/data/simulations/rankone/MSE/median/run.R
```

## run

```
sbatch Jobs.sbatch
```

## plot the result

```
T = 100
results = matrix(NA,ncol = 6, nrow = T)
for(i in 1:T){
  test_folder = paste("test", i, sep = "")
  # out_folder = "NBSFAout"
  out_file = "output.rds"
  file_name = file.path(test_folder,out_file)
  results[i,] = try(readRDS(file_name))
}
```

## RMSE sparse case 100 simulations

```{r}
sparse_res = readRDS("../data/simulation/rankone/RMSE/result_sparse.rds")
colnames(sparse_res) = c("PMD","flash","SFA","SFAmix","SVD","SSVD")
boxplot(sparse_res,main = "sparse case")
```

## RMSE intermediate sparse case 100 simulations

```{r}
sparse_res = readRDS("../data/simulation/rankone/RMSE/result_itermediate.rds")
colnames(sparse_res) = c("PMD","flash","SFA","SFAmix","SVD","SSVD")
boxplot(sparse_res,main = "intermediate sparse case")
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```

### 2017-12-09 analysis/index.Rmd
Files commited by wflow_commit.

@@
 Click here to check the [simulatoin result (rank one and rank three)](applyflahr2onsimulations.html)
+
+Click here to check the [RMSE for the rank one example in the paper](MSErank1sim.html)
+
+#### GTEx Z-score
+
+Here is a example to run `flashr2` on RCC [RCC template](GTEXzscoreTemplate.html)
+
+#### missing data imputation
+
+click here to check the template for [missing data imputation with OCV](templateforOCV.html)
+
+click here to check the comparison of different options for FLASH [different shrink options](compareOptionsOCV.html)

### 2018-01-03 analysis/compare_ash_ebmn.Rmd
Files commited by wflow_commit.

---
title: "compare the different versions of ebnm methods"
author: "Wei"
date: 2018-01-02
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

## for the simulated data

```{r,echo=FALSE}
library("ebnm")
library("ashr")
library("MASS")
betaO=function(P,betapi,betasigma,propt){
  idx=rmultinom(P,size=1,prob=betapi)
  K=length(betapi)
  s=array(0,dim=c(K,K))
  diag(s)=betasigma
  bnorm=mvrnorm(P,rep(0,K),s)
  betaO=apply(bnorm*t(idx),1,sum)
  betaO=betaO*rbinom(P,1,propt)
  return(betaO)
}

datamaker = function(N,P,l_pi,l_se,l_sp,f_pi,f_se,f_sp,sigmae){
  # here I would like to fix the sparsity of L and F which is simple
  # if we need to do futher experiment
  L_true = betaO(N,l_pi,l_se,l_sp)
  F_true = betaO(P,f_pi,f_se,f_sp)
  E = matrix(rnorm(N*P,0,sigmae),ncol = P)
  Y = L_true %*% t(F_true) + E
  return(list(Y = Y,L_true = L_true,F_true = F_true,E = E))
}

```

### sparse loding (simulated data)

we can see there is tiny difference between `flash_r1` and `flash_add_greedy`.
The difference between the `ebnm_pn` and `ebnm_ash` are tiny based on the RMSE ratio.

```{r}
set.seed(99)
L_se = c( 0.25, 0.5, 1, 2, 4)
L_pi = c(1, 1,1,1,1)
L_pi = L_pi / sum(L_pi)
N = 200
P = 300
Data = datamaker(N,P,L_pi,L_se,0.1,c(1),c(1),1,sqrt(1))
library(flashr2)
data = flash_set_data(Data$Y)
f_greedy_pn = flash_add_greedy(data,Kmax=10,ebnm_fn = ebnm_pn)
f_greedy_ash = flash_add_greedy(data,Kmax=10,ebnm_fn = ebnm_ash)
# here flashr2 provide 2 factors but 2nd factor is all zero.
f_pn = flash_r1(data, ebnm_fn = ebnm_pn)
f_ash = flash_r1(data, ebnm_fn = ebnm_ash)
par(mfrow = c(2, 2)) 
plot(f_pn$EL,f_greedy_pn$EL[,1],main = "r1 vs greedy (pn)")
plot(f_pn$EL-f_greedy_pn$EL[,1],main = "difference r1 vs greedy")
# there is tiny difference between greedy algorithm and rank one fucntion
plot(f_ash$EL,f_greedy_ash$EL[,1],main = "r1 vs greedy (ash)")
plot(f_ash$EL+f_greedy_ash$EL[,1],main = "difference r1 vs greedy")
# compare the ash and pn methods
par(mfrow = c(1, 1)) 
plot(f_ash$EL,f_pn$EL,main = "pn vs ash (EL)")
Y_hat_pn = f_pn$EL %*% t(f_pn$EF)
RMSE_pn = sqrt(mean(( Data$Y - Y_hat_pn - Data$E )^2 ))/sqrt(mean(( Data$Y - Data$E )^2 ))
Y_hat_ash = f_ash$EL %*% t(f_ash$EF)
RMSE_ash = sqrt(mean(( Data$Y - Y_hat_ash - Data$E )^2 ))/sqrt(mean(( Data$Y - Data$E )^2 ))
c(RMSE_pn,RMSE_ash)
```

Here we can see that the `ebnm_pn` is slightly better than `ebnm_ash`.

We try repeat this experiment 100 times

```{r}
Sim_RMSE = function(N,P,l_pi,l_se,l_sp,f_pi,f_se,f_sp,sigmae){
  Data = datamaker(N,P,l_pi,l_se,l_sp,f_pi,f_se,f_sp,sigmae)
  data = flash_set_data(Data$Y)
  f_pn = flash_r1(data, ebnm_fn = ebnm_pn)
  f_ash = flash_r1(data, ebnm_fn = ebnm_ash)
  Y_hat_pn = f_pn$EL %*% t(f_pn$EF)
  RMSE_pn = sqrt(mean(( Data$Y - Y_hat_pn - Data$E )^2 ))/sqrt(mean(( Data$Y - Data$E )^2 ))
  Y_hat_ash = f_ash$EL %*% t(f_ash$EF)
  RMSE_ash = sqrt(mean(( Data$Y - Y_hat_ash - Data$E )^2 ))/sqrt(mean(( Data$Y - Data$E )^2 ))
  return(c(RMSE_pn,RMSE_ash))
}
```

```{r}
T = 100
RMSE = array(NA,dim = c(T,2))
L_se = c( 0.25, 0.5, 1, 2, 4)
L_pi = c(1, 1,1,1,1)
L_pi = L_pi / sum(L_pi)
N = 200
P = 300
for(i in 1:T){
  RMSE[i,] = Sim_RMSE(N,P,L_pi,L_se,0.1,c(1),c(1),1,sqrt(1))
}
colnames(RMSE) = c("pn","ash")
par(mfrow = c(1, 2)) 
boxplot(RMSE,main = "pn vs ash")
boxplot(RMSE[,1] - RMSE[,2],main = "difference")
```

### intermediate sparse loding (simulated data)

The difference between the `ebnm_pn` and `ebnm_ash` are also tiny based on the RMSE ratio.

```{r}
library(cowplot)
set.seed(99)
L_se = c( 0.25, 0.5, 1, 2, 4)
L_pi = c(1, 1,1,1,1)
L_pi = L_pi / sum(L_pi)
N = 200
P = 300
Data = datamaker(N,P,L_pi,L_se,0.7,c(1),c(1),1,sqrt(16))
data = flash_set_data(Data$Y)
f_pn = flash_r1(data, ebnm_fn = ebnm_pn)
f_ash = flash_r1(data, ebnm_fn = ebnm_ash)
Y_hat_pn = f_pn$EL %*% t(f_pn$EF)
RMSE_pn = sqrt(mean(( Data$Y - Y_hat_pn - Data$E )^2 ))/sqrt(mean(( Data$Y - Data$E )^2 ))
Y_hat_ash = f_ash$EL %*% t(f_ash$EF)
RMSE_ash = sqrt(mean(( Data$Y - Y_hat_ash - Data$E )^2 ))/sqrt(mean(( Data$Y - Data$E )^2 ))
c(RMSE_pn,RMSE_ash)
```

Here we can see that the `ebnm_pn` is slightly worse than `ebnm_ash`.

```{r}
T = 100
RMSE = array(NA,dim = c(T,2))
L_se = c( 0.25, 0.5, 1, 2, 4)
L_pi = c(1, 1,1,1,1)
L_pi = L_pi / sum(L_pi)
N = 200
P = 300
for(i in 1:T){
  RMSE[i,] = Sim_RMSE(N,P,L_pi,L_se,0.7,c(1),c(1),1,sqrt(16))
}
colnames(RMSE) = c("pn","ash")
par(mfrow = c(1, 2)) 
boxplot(RMSE,main = "pn vs ash")
boxplot(RMSE[,1] - RMSE[,2],main = "difference")
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```

### 2018-01-04 analysis/compare_ash_pn_miss.Rmd
Files commited by wflow_commit.

--- 2017-10-06 analysis/templateforOCV.Rmd
@@
 ---
-title: "apply flashr2 with OCV"
-author: "wei wang"
-date: 09-30
+title: "compare the different versions of ebnm methods on missing data"
+author: "Wei"
+date: 2018-01-02
 output: 
@@
 <!-- Add your analysis here -->
-
-## preparation
-
-1. creat a folder
-
-```
-mkdir Breastcancer
-```
-
-2. put your data into this folder
-
-here we use the file 'example.mat'. the reason we use `.mat` file is because there is a matlab package in our comparision.
-
-3. creat the wrapper function in R and matlab
-
-### wrapper function for OCV
 
@@
 call_method = function(Y_data,method,Kmax = 50){
-  if(method == "flash"){
+  if(method == "ebnm_pn"){
     data = flashr2::flash_set_data(Y_data)
-    f_greedy = flashr2::flash_add_greedy(data,Kmax=Kmax,ash_param=list(method = "fdr"),var_type = "constant")
+    f_greedy = flashr2::flash_add_greedy(data,Kmax=Kmax,ebnm_fn = ebnm_pn,verbose = F)
     Y_hat = f_greedy$EL %*% t(f_greedy$EF)
-  }else if(method == "PMD"){
-    out = PMA::PMD(Y_data, K = Kmax)
-    Y_hat =  out$u %*% diag(out$d) %*% t(out$v)
-  }else if(method == "softImpute"){
-    gsoft = softImpute(Y_data, rank.max = Kmax)
-    Y_hat =  gsoft$u %*% diag(gsoft$d) %*% t(gsoft$v)
+  }else if(method == "ebnm_ash"){
+    data = flashr2::flash_set_data(Y_data)
+    f_greedy = flashr2::flash_add_greedy(data,Kmax=Kmax,ebnm_fn = ebnm_ash,verbose = F)
+    Y_hat = f_greedy$EL %*% t(f_greedy$EF)
   }else{
@@
 }
+```
 
-NSF_OCV = function(OCVindex){
-  if(file.exists("NBSFAout")){
-    unlink("NBSFAout", recursive= T)
-  }
-  system("mkdir NBSFAout")
-  # write the missing index
+## denoise tumor data example in our paper
 
-  writeMat("./NBSFAout/OCVindex.mat", OCVindex = OCVindex)
+```{r}
+set.seed(99)
+library(denoiseR)
+data(tumors)
 
-  system("matlab -nosplash -nodesktop -r \"addpath(\'~/HG/flash/data/OCVmissflashr2/Breastcancer\'); run(\'missvalue.m\');exit;\" ")
-  g_nsfa = readMat("./NBSFAout/NSFAresult.mat")
-  nsfa_rmse = as.vector(g_nsfa$RMSE)
-  return(nsfa_rmse)
-}
+Y = tumors[, -ncol(tumors)]
+Y = as.matrix(Y)
+N = dim(Y)[1]
+P = dim(Y)[2]
+
+OCVindex = OCV_index(Y,k_fold = 5)
+OCVdata = OCV_data(Y,OCVindex,k_fold = 5)
+f_pn = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "ebnm_pn",Kmax = 20)
+# flashG_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "flash",Kmax = 50)
+f_ash = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "ebnm_ash",Kmax = 20)
 
@@
 
-Here you need change the `\"addpath(\'~/HG/flash/data/OCVmissflashr2/Breastcancer\');` to your folder which contains `missvalue.m` file. this file name is `wrapper.R`
+Check the RMSE for the missing value
 
-### matlab code
-
-```{}
-addpath('~/HG/flash/data/missingvalue/methods/NBSF/nsfa-master/');
-addpath('~/HG/flash/data/missingvalue/methods/NBSF/nsfa-master/utils/');
-
-% you need to change the path and the data file (not centered)
-load ~/HG/flash/data/OCVmissflashr2/Breastcancer/example.mat;
-Ycentered=Y-repmat(mean(Y,2),1,size(Y,2));
-settings=defaultsettings();
-[settings.D,settings.N]=size(Ycentered);
-settings.iterations=100;
-% now the missing is decide by the this is the missing index.(or you can upload the missing matrix)
-
-load(fullfile(pwd,'NBSFAout/OCVindex.mat'))
-
-missindex = ones(settings.D,settings.N);
-k_fold = 5;
-foldindex = zeros(k_fold,k_fold,2);
-for k = 1:2
-    for j = 1:k_fold
-        for i = 1:k_fold
-            foldindex(i,j,k) = OCVindex.foldindex(i + k_fold*(j - 1) + k_fold * k_fold * (k - 1));
-        end
-    end
-end
-
-SSE = zeros(1,k_fold);
-for i = 1:k_fold
-    mvmask = missindex;
-    for j = 1:k_fold
-        mvmask(OCVindex.rowindex(:,foldindex(j,i,1)),OCVindex.colindex(:,foldindex(j,i,2))) = 0;
-    end
-    initialsample=init_nsfa(settings);
-    [finalsample,resultstable]=nsfa(Ycentered,mvmask,initialsample,settings);
-    Y_hat = (finalsample.G *  finalsample.X);
-    for j = 1:k_fold
-        SSE(i) = SSE(i) + ...
-            sum(sum((Ycentered(OCVindex.rowindex(:,foldindex(j,i,1)),OCVindex.colindex(:,foldindex(j,i,2)))- ...
-            Y_hat(OCVindex.rowindex(:,foldindex(j,i,1)),OCVindex.colindex(:,foldindex(j,i,2)))).^2));
-    end
-
-end
-
-RMSE = sqrt(sum(SSE)/(settings.D*settings.N));
-RMSE;
-
-save_path = fullfile(pwd,'NBSFAout/NSFAresult');
-save(save_path,'RMSE'); 
+```{r}
+c(f_ash,f_pn)
 ```
 
-you need change the path as well. this file name is 'missvalue.m'
+## Breast cancer data example in our paper
 
-### creat the run.R file to run the code
-
-In this case, we have problem in running `flashr2` on RCC. so we just try other methods first.
-
-```{}
+```{r}
+set.seed(99)
 library(R.matlab)
 ## run the code
-Y_centered = readMat("~/HG/flash/data/OCVmissflashr2/Breastcancer/example.mat")
+Y_centered = readMat("../../ash-sfa/Rcode/postmean/flash_simulation/missingvalue/Breastcancer/example.mat")
 Y = Y_centered$Y
@@
 OCVdata = OCV_data(Y,OCVindex,k_fold = 5)
-soft_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "softImpute",Kmax = 50)
+f_pn = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "ebnm_pn",Kmax = 50)
 # flashG_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "flash",Kmax = 50)
-pmd_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "PMD",Kmax = 50)
-nsfa_rmse = NSF_OCV(OCVindex)
-flashG_rmse = mean(c( nsfa_rmse,pmd_rmse,soft_rmse))
-
-result = c(flashG_rmse, nsfa_rmse,pmd_rmse,soft_rmse)
-saveRDS(result, "./NBSFAout/output.rds")
-
+f_ash = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "ebnm_ash",Kmax = 50)
 ```
 
-this file name is `runOCV.R`
+Check the RMSE for the missing value
 
-### creat folder to track the result
-
-```
-mkdir outlog
+```{r}
+c(f_ash,f_pn)
 ```
 
-### creat .sbatch file
-
-```
-#!/bin/bash
-
-#SBATCH --job-name=arrayJob
-#SBATCH --output=/home/weidong/HG/flash/data/OCVmissflashr2/Breastcancer/outlog/arrayJob_%A_%a.out
-#SBATCH --error=/home/weidong/HG/flash/data/OCVmissflashr2/Breastcancer/outlog/arrayJob_%A_%a.err
-#SBATCH --array=1-100
-#SBATCH --time=06:00:00
-#SBATCH --partition=broadwl
-#SBATCH --ntasks=1
-#SBATCH --mem-per-cpu=2000
-
-######################
-# Begin work section #
-######################
-
-# Print this sub-job's task ID
-cd /home/weidong/HG/flash/data/OCVmissflashr2/Breastcancer
-mkdir test${SLURM_ARRAY_TASK_ID}
-cd test${SLURM_ARRAY_TASK_ID}
-Rscript --verbose /home/weidong/HG/flash/data/OCVmissflashr2/Breastcancer/runOCV.R
-```
-
-this file is name as `OCV.sbatch`
-
-## run OCV
-
-now in your folder, you should have:
-
-```
-[weidong@midway2-login1 Breastcancer]$ ls
-OCV.sbatch  example.mat  missvalue.m  outlog  runOCV.R  wrapper.R
-```
-
-```
-sbatch OCV.sbatch
-```
-
-## to get the result
+### MovieLens data
 
 ```{}
-T = 100
-results = matrix(NA,ncol = 4, nrow = T)
-for(i in 1:T){
-  test_folder = paste("test", i, sep = "")
-  out_folder = "NBSFAout"
-  out_file = "output.rds"
-  file_name = file.path(test_folder,out_folder,out_file)
-  results[i,] = try(readRDS(file_name))
-}
-
-saveRDS(results,"./boxplot.rds")
-```
-
-## plot the result
-
-```{r}
-plot_res = function(output,title = "data",legend_position = "none"){
-  rmse = as.vector(output)
-  N = dim(output)[1]
-  methods = rep(c("flash","NBSFA","PMD","softImpute"), each = N)
-  df = data.frame(RMSE = rmse, Method = methods )
-  p<-ggplot(df, aes(x=Method, y=RMSE, color=Method)) +
-  geom_boxplot()+
-  # geom_violin()+
-  ggtitle(title)+theme_gray() +
-    theme(legend.position= legend_position, legend.text=element_text(size=15), axis.text.y = element_text(size =6))
-  p
-}
-
-PT_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/president_box.rds")
-pp = plot_res(PT_res,"Text data")
-DT_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/denoiseRtumor_box.rds")
-pd = plot_res(DT_res,"Tumor data")
-GZ_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/gtexzscore_box.rds")
-pg = plot_res(GZ_res,"GTEx data")
-BC_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/BreastCancer_box.rds")
-pb = plot_res(BC_res,"Breast Cancer data")
-ML_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/ML100K_box.rds")
-ML_res[c(2,13,17,21,29,37,62,76,77,93,95,100),] = NA
-ML_res = matrix(as.numeric(ML_res),ncol = 4)
-pM = plot_res(ML_res,"Movie Lens data")
-```
-
-## plots
-
-```{r}
-pp
-pd
-pg
-pb
-pM
+set.seed(11)
+library(Matrix)
+ml100K_data = readRDS("../../ash-sfa/Rcode/postmean/flash_simulation/missingvalue/ML100K/biscale_data.rds")
+MLMatrix <- sparseMatrix(i = ml100K_data[,1],
+                         j = ml100K_data[,2],
+                         x = ml100K_data[,3],dims = c(943,1682))
+# turn this sparse matrix into matrix in r
+Y = as.matrix(MLMatrix)
+Y[which(Y == 0)] = NA
+OCVindex = OCV_index(Y,k_fold = 10)
+OCVdata = OCV_data(Y,OCVindex,k_fold = 10)
+f_pn = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "ebnm_pn",Kmax = 10)
+# flashG_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "flash",Kmax = 50)
+f_ash = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "ebnm_ash",Kmax = 10)
 ```

### 2018-01-06 analysis/CVtemplate.Rmd
Files commited by wflow_commit.

---
title: "template for CV in RCC"
author: "Wei"
date: 10-03
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

## preparation

1. creat a file by `mkdir Breastcancer` and put the data file into this folder. in this case, we use matlab function as well, so we use `.mat` file for the data.

2. creat a folder to track the code `mkdir outlog`

## wrapper file

### R functions

```{r}
miss_gen = function(Y, percent = 0.2){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = sample(P,(P*percent))
  rowindex = sample(N,(N*percent))
  missindex = list(colindex = colindex, rowindex = rowindex)
  return(missindex)
}

# methods
flashG_wrapper = function(Y, missindex, rank_max,type = "constant"){
  # delete the value in the missing position
  Y_miss = Y
  Y_miss[missindex$rowindex, missindex$colindex] = NA
  data = flashr2::flash_set_data(Y_miss)
  g_flash = flashr2::flash_add_greedy(data,Kmax=rank_max,ash_param=list(method = "fdr"),var_type = type)
  misshat_flash = g_flash$EL %*% t(g_flash$EF)
  rmse = sqrt(mean((Y[missindex$rowindex, missindex$colindex] - misshat_flash[missindex$rowindex, missindex$colindex])^2))
  return(rmse)
}

flashB_wrapper = function(Y, missindex, rank_max,type = "constant"){
  # delete the value in the missing position
  Y_miss = Y
  Y_miss[missindex$rowindex, missindex$colindex] = NA
  data = flashr2::flash_set_data(Y_miss)
  g_flash = flashr2::flash_add_greedy(data,Kmax=rank_max,ash_param=list(method = "fdr"),var_type = type)
  b_flash = flash_backfit(data,g_flash)
  misshat_flash = b_flash$EL %*% t(b_flash$EF)
  rmse = sqrt(mean((Y[missindex$rowindex, missindex$colindex] - misshat_flash[missindex$rowindex, missindex$colindex])^2))
  return(rmse)
}

softImpute_wrapper = function(Y, missindex, rank_max){
  # delete the value in the missing position
  Y_miss = Y
  Y_miss[missindex$rowindex, missindex$colindex] = NA
  gsoft = softImpute::softImpute(Y_miss, rank.max = rank_max, lambda = 30)
  misshat_soft =  gsoft$u %*% diag(gsoft$d) %*% t(gsoft$v)
  rmse = sqrt(mean((Y[missindex$rowindex, missindex$colindex] - misshat_soft[missindex$rowindex, missindex$colindex])^2))
  return(rmse)
}

PMD_wrapper = function(Y, missindex, rank_max){
  # delete the value in the missing position
  Y_miss = Y
  Y_miss[missindex$rowindex, missindex$colindex] = NA
  out = PMA::PMD(Y_miss, K = rank_max)
  misshat_PMD =  out$u %*% diag(out$d) %*% t(out$v)
  rmse = sqrt(mean((Y[missindex$rowindex, missindex$colindex] - misshat_PMD[missindex$rowindex, missindex$colindex])^2))
  return(rmse)
}

```

```{r}
library(R.matlab)
# call the R functions
source("~/HG/flash/data/CVmissflashr2/template/Rfunction.R")
# read the data
# you need to change the path for different data
Y_data = R.matlab::readMat("~/HG/flash/data/CVmissflashr2/Breastcancer/example.mat")
Y = Y_data$Y
N = dim(Y)[1]
P = dim(Y)[2]
Y = Y - rowMeans(Y) %*% t(rep(1,P))
result = c()
missindex = miss_gen(Y, percent = 0.1)
# creat a folder for the output and matlab in the folder we make
if(file.exists("NBSFAout")){
  unlink("NBSFAout", recursive= T)
}
system("mkdir NBSFAout")
# write the missing index
writeMat("./NBSFAout/missindex.mat", missindex = missindex)

flashG_rmse = flashG_wrapper(Y, missindex, rank_max = 50)
pmd_rmse  = PMD_wrapper(Y, missindex, rank_max= 50 )
soft_rmse = softImpute_wrapper(Y, missindex, rank_max = 50)
# then call the matlab estiamtion 
# this for the folder of the data
system("matlab -nosplash -nodesktop -r \"addpath(\'~/HG/flash/data/CVmissflashr2/template\'); run(\'missvalue.m\');exit;\" ")
g_nsfa = readMat("./NBSFAout/NSFAresult.mat")
nsfa_rmse = as.vector(g_nsfa$rmse.NSFA)

result = c(flashG_rmse, nsfa_rmse,pmd_rmse,soft_rmse)

saveRDS(result, "./NBSFAout/output.rds")

```

### matlab code

```
addpath('~/HG/flash/data/missingvalue/methods/NBSF/nsfa-master/');
addpath('~/HG/flash/data/missingvalue/methods/NBSF/nsfa-master/utils/');

% you need to change the path and the data file (not centered)
load ../example.mat;
Ycentered=Y-repmat(mean(Y,2),1,size(Y,2));
settings=defaultsettings();
[settings.D,settings.N]=size(Ycentered);
settings.iterations=100;
% now the missing is decide by the this is the missing index.(or you can upload the missing matrix)

file_path = fullfile(pwd,'NBSFAout/missindex.mat');
load(file_path);
mvmask = ones(settings.D,settings.N);
mvmask(missindex.rowindex,missindex.colindex) = 0;

initialsample=init_nsfa(settings);
[finalsample,resultstable]=nsfa(Ycentered,mvmask,initialsample,settings);
Y_hat = (finalsample.G *  finalsample.X);

sse_NSFA = sum(sum((Ycentered(missindex.rowindex,missindex.colindex) - Y_hat(missindex.rowindex,missindex.colindex)).^2));
mse_NSFA = sse_NSFA/(length(missindex.rowindex)*length(missindex.colindex))
rmse_NSFA = sqrt(mse_NSFA)

save_path = fullfile(pwd,'NBSFAout/NSFAresult');
save(save_path,'rmse_NSFA'); 
```

## batch file

```
#!/bin/bash

#SBATCH --job-name=arrayJob
#SBATCH --output=/home/weidong/HG/flash/data/CVmissflashr2/Breastcancer/outlog/arrayJob_%A_%a.out
#SBATCH --error=/home/weidong/HG/flash/data/CVmissflashr2/Breastcancer/outlog/arrayJob_%A_%a.err
#SBATCH --array=1-100
#SBATCH --time=06:00:00
#SBATCH --partition=broadwl
#SBATCH --ntasks=1
#SBATCH --mem-per-cpu=6000

######################
# Begin work section #
######################

# Print this sub-job's task ID
cd /home/weidong/HG/flash/data/CVmissflashr2/Breastcancer
mkdir test${SLURM_ARRAY_TASK_ID}
cd test${SLURM_ARRAY_TASK_ID}
Rscript --verbose /home/weidong/HG/flash/data/CVmissflashr2/Breastcancer/runR.R
```

## run

```
sbatch Jobs.sbatch
```

## plot the result

```
T = 100
results = matrix(NA,ncol = 4, nrow = T)
for(i in 1:T){
  test_folder = paste("test", i, sep = "")
  out_folder = "NBSFAout"
  out_file = "output.rds"
  file_name = file.path(test_folder,out_folder,out_file)
  results[i,] = try(readRDS(file_name))
}
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```

### 2018-01-06 analysis/index.Rmd
Files commited by wflow_commit.

@@
 
-click here to check the template for [missing data imputation with OCV](templateforOCV.html)
+click here to check the template for [missing data imputation with OCV](templateforOCV.html) and [CV template](CVtemplate.html)
 
 click here to check the comparison of different options for FLASH [different shrink options](compareOptionsOCV.html)
+
+### different options for EBNM
+
+#### simulation examples in our paper
+
+Click here to check the [simulated examples](compare_ash_ebmn.html)
+
+Click here to check the [missing data examples (first simple comparison)](compare_ash_pn_miss.html)

### 2018-01-06 analysis/CVtemplate.Rmd
Files commited by wflow_commit.

@@
 date: 10-03
-output: html_document
+output: 
+  html_document:
+    code_folding: hide
 ---
@@
 
-```{r}
+```
 library(R.matlab)

### 2018-01-10 analysis/MSErank1sim.Rmd
Files commited by wflow_commit.

@@
 colnames(sparse_res) = c("PMD","flash","SFA","SFAmix","SVD","SSVD")
+sparse_diff = sparse_res - sparse_res[,2]
+
 boxplot(sparse_res,main = "sparse case")
+boxplot(sparse_diff,main = "sparse case (difference)")
+abline(a = 0,b = 0,col= "red")
 ```
@@
 colnames(sparse_res) = c("PMD","flash","SFA","SFAmix","SVD","SSVD")
+sparse_diff = sparse_res - sparse_res[,2]
 boxplot(sparse_res,main = "intermediate sparse case")
+boxplot(sparse_diff,main = "sparse case (difference)")
+abline(a = 0,b = 0,col= "red")
 ```

### 2018-01-10 analysis/PCAresult.Rmd
Files commited by wflow_commit.

---
title: "PCA(SVD) result"
author: "Wei"
date: 2018-01-10
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

```{r}
plot_GTEXres = function(L,pve){
  library(reshape2)
  data_L = melt(L)
  colnames(data_L) = c("tissue","loading","value")
  library(ggplot2)
  tissue_color = as.character(colordata[,2])
  data_L$tissue = factor(data_L$tissue,levels = 1:44 ,labels = as.character(colordata[,1]) )
  data_L$loading = factor(data_L$loading,levels = 1:14 ,labels = paste("loading",1:14,"; pve:", pve[1:14]))

  ggplot(data_L,aes(x = tissue,y = value,fill = factor(tissue) )) +
    geom_bar(stat = "identity",width = 0.6) +
    scale_fill_manual(values=tissue_color) +
    scale_x_discrete(labels = NULL) +
    theme_grey()+
    theme(legend.position="right", legend.text=element_text(size=5), axis.text.y = element_text(size = 5)) + 
    labs(title = "GTEx data", y = "loading values" ,x = "tissues", fill="tissue") +
    facet_wrap(~loading, ncol = 2, scales = "free_y") +
    guides(fill = guide_legend(ncol = 1, keyheight = 0.5, keywidth = 0.5))
}
b_flash = readRDS("../data/GTExdata/res_flashr2/bflashvarcol.rds")
load("../data/GTExdata/gtexEQTL_zscore.rds")
ssY = sum(zscore^2)
K = dim(b_flash$EL)[2] -1
pve = (sapply(seq(1,K),function(x){ sum(b_flash$EL[,x]^2 %*% t(b_flash$EF[,x]^2)) }))/ssY
pve = pmax(round(pve,3),0.001)

dat = read.table('../data/GTExColors.txt', sep = '\t', comment.char = '')
colordata = dat[c(1:6,9:18,21:23,26:30,32,33,35,36,38:53),1:2]
```

```{r}
L = b_flash$EL[,1:8]
plot_GTEXres(L,pve)
L = b_flash$EL[,9:16]
plot_GTEXres(L,pve)

gsvd = svd(zscore)
pve = (sapply(seq(1,K),function(x){ sum((gsvd$d[x] * gsvd$u[,x] %*% t(gsvd$v[,x]))^2) }))/ssY
pve = pmax(round(pve,3),0.001)
L = gsvd$v[,1:8]
plot_GTEXres(L,pve)
L = gsvd$v[,9:16]
plot_GTEXres(L,pve)

```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```

### 2018-01-10 analysis/templateforOCV.Rmd
Files commited by wflow_commit.

@@
 T = 100
-results = matrix(NA,ncol = 4, nrow = T)
+results = matrix(NA,ncol = 22, nrow = T)
 for(i in 1:T){
@@
 
+## code for the tuning parameters from OCV only change one parameter
+
+try this one later.....
+
+```{r}
+CVPMD_softImpute=function(Y,c_s,K,fold = 10, method = "PMD",N_sim = 10){
+  N = dim(Y)[1]
+  P = dim(Y)[2]
+  colindex = matrix(sample(P,P),ncol = fold)
+  rowindex = matrix(sample(N,N),ncol = fold)
+
+  missing= array(0,dim = c(fold,N,P))
+  foldindex = array(0,dim = c(fold,fold,2))
+  for(i in 1:fold){
+    for(j in 1:fold){
+      foldindex[i,j,1] = i
+      foldindex[i,j,2] = (i+j) %% fold
+    }
+  }
+  foldindex[which(foldindex == 0)] = fold
+  for(i in 1:fold){
+    missing[i, , ] = Y
+    for(j in 1:fold){
+      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
+    }
+  }
+  # c_s is the candicate of shrinkage parameter
+  n_s = length(c_s)
+  # rmse for each grids
+  CVRMSE = rep(0,n_s)
+  minrmse = Inf
+  opt_s = 0
+  # for each candidate, we run it N_sim times
+  for(t_s in 1:n_s){
+    # for each grid
+    for(iter in 1:N_sim){
+      # each time we set the rmse to zeros
+      rmse = rep(0,fold)
+      for(i in 1:fold){
+        if(method == "PMD"){
+          out = PMD(missing[i,,], sumabs = c_s[t_s], sumabsv = NULL, sumabsu = NULL,K = K)
+          # out = PMD(missing[i,,], K = K)
+        }else{
+          out = softImpute(Y, rank.max = K,lambda = c_s[t_s]) 
+        }
+        misshat = out$u %*%  diag(out$d) %*% t(out$v)
+        for(j in 1:fold){
+          rmse[i] = rmse[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] - 
+                                     misshat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2)
+        }
+      } #get the result for one run
+      print(sqrt(sum(rmse)/(N*P)))
+      CVRMSE[t_s] = CVRMSE[t_s] + sqrt(sum(rmse)/(N*P))
+    }
+    if(CVRMSE[t_s] < minrmse){
+      minrmse = CVRMSE[t_s]
+      opt_s = c_s[t_s]
+    }
+  }
+  CVRMSE = CVRMSE/N_sim
+  return(list(opt_s = opt_s, output = CVRMSE))
+}
+
+PMA.wrapper = function(Y,ngrids = 10,K,fold = 10){
+  library(PMA)
+  N = dim(Y)[1]
+  P = dim(Y)[2]
+  c_s = seq(0.2,0.99,len=ngrids)
+  cvout = CVPMD_softImpute(Y,c_s,K ,fold , method = "PMD")
+  out = PMD(Y,sumabsu = NULL, sumabsv = NULL, sumabs = cvout$opt_s ,K = K)
+  residual_PMD = Y -  out$u %*%  diag(out$d) %*% t(out$v)
+  return(list(d = out$d, u = out$u, v = out$v, residual = residual_PMD))
+}
+
+softImpute.wrapper = function(Y,ngrids = 10,K,fold = 10){
+  library(softImpute)
+  N = dim(Y)[1]
+  P = dim(Y)[2]
+  c_s = seq(0,10,len=ngrids)
+  cvout = CVPMD_softImpute(Y,c_s,K ,fold , method = "softImpute")
+  out = softImpute(Y, rank.max = K,lambda = cvout$opt_s) 
+  residual_softImpute = Y -  out$u %*%  diag(out$d) %*% t(out$v)
+  return(list(d = out$d, u = out$u, v = out$v, residual = residual_softImpute))
+}
+```
+
+#### new reuslt
+
+```{r}
+library(ggplot2)
+plot_res = function(output,title = "data",legend_position = "none"){
+  rmse = as.vector(output)
+  N = dim(output)[1]
+  pmd_c = sapply(seq(1,10),function(x){paste("PMD",x)})
+  softImpute_c = sapply(seq(1,10),function(x){paste("SF",x)})
+  methods = rep(c("flash","NBSFA",pmd_c,softImpute_c), each = N)
+  df = data.frame(RMSE = rmse, Method = methods )
+  p<-ggplot(df, aes(x=Method, y=RMSE, color=Method)) +
+  geom_boxplot()+
+  # geom_violin()+
+  ggtitle(title) +  theme_bw()+
+    theme(legend.position= legend_position, legend.text=element_text(size=15), axis.text.y = element_text(size =6))
+  p
+}
+```
+
+```{r}
+PT_res = readRDS("../data/missingvalue/box_res_grids_sf_pmd/denoiseTumor_box.rds")
+pt = plot_res(PT_res,"Tumor data")
+pt
+PT_res = readRDS("../data/missingvalue/box_res_grids_sf_pmd/TEXT_prsdt_box.rds")
+pp = plot_res(PT_res,"Text data")
+pp
+```
+
 ## Session information

### 2018-01-10 analysis/templateforOCV.Rmd
Files commited by wflow_commit.

@@
 pp
+PT_res = readRDS("../data/missingvalue/box_res_grids_sf_pmd/BreastCancer_box.rds")
+pb = plot_res(PT_res,"Text data")
+pb
 ```

### 2018-01-17 analysis/templateforOCV.Rmd
Files commited by wflow_commit.

--- 2017-10-07 analysis/templateforOCV.Rmd
@@
 ```{r}
+CVPMD_softImpute=function(Y,c_s,K,fold = 10, method = "PMD"){
+  N = dim(Y)[1]
+  P = dim(Y)[2]
+  colindex = matrix(sample(P,P),ncol = fold)
+  rowindex = matrix(sample(N,N),ncol = fold)
+
+  missing= array(0,dim = c(fold,N,P))
+  foldindex = array(0,dim = c(fold,fold,2))
+  for(i in 1:fold){
+    for(j in 1:fold){
+      foldindex[i,j,1] = i
+      foldindex[i,j,2] = (i+j) %% fold
+    }
+  }
+  foldindex[which(foldindex == 0)] = fold
+  for(i in 1:fold){
+    missing[i, , ] = Y
+    for(j in 1:fold){
+      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
+    }
+    missing[i,,which(colSums(missing[i,,],na.rm = T) ==0)] = Y[,which(colSums(missing[i,,],na.rm = T) ==0)]
+  }
+  # c_s is the candicate of shrinkage parameter
+  n_s = length(c_s)
+  # rmse for each grids
+  CVRMSE = rep(0,n_s)
+  minrmse = Inf
+  opt_s = 0
+  # for each candidate, we run it N_sim times
+  for(t_s in 1:n_s){
+    # for each grid
+    # each time we set the rmse to zeros
+    rmse = rep(0,fold)
+    for(i in 1:fold){
+      if(method == "PMD"){
+        res_log = capture.output({out = PMD(missing[i,,], sumabs = c_s[t_s], sumabsv = NULL, sumabsu = NULL,K = K)})
+      }else{
+        out = softImpute(missing[i,,], rank.max = K,lambda = c_s[t_s])
+      }
+      if(length(out$d)==1){
+        misshat = (out$d) * out$u %*% t(out$v)
+      }else{
+        misshat = out$u %*%  diag(out$d) %*% t(out$v)
+      }
+      for(j in 1:fold){
+        # for each fold j
+        rmse[i] = rmse[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] -
+                                   misshat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2,na.rm = TRUE)
+      }
+    } #get the result for one run
+    CVRMSE[t_s] = CVRMSE[t_s] + sqrt(sum(rmse)/(N*P))
+    if(CVRMSE[t_s] < minrmse){
+      minrmse = CVRMSE[t_s]
+      opt_s = c_s[t_s]
+    }
+  }
+  return(list(opt_s = opt_s, output = CVRMSE))
+}
+
+PMA.wrapper = function(Y,ngrids = 10,K,fold = 10){
+  library(PMA)
+  N = dim(Y)[1]
+  P = dim(Y)[2]
+  c_s = seq(0.3,0.8,len=ngrids)
+  cvout = CVPMD_softImpute(Y,c_s,K ,fold , method = "PMD")
+  res_log = capture.output({out = PMD(Y,sumabsu = NULL, sumabsv = NULL, sumabs = cvout$opt_s ,K = K)})
+  return(list(d = out$d, u = out$u, v = out$v))
+}
+
+softImpute.wrapper = function(Y,ngrids = 10,K,fold = 10){
+  library(softImpute)
+  N = dim(Y)[1]
+  P = dim(Y)[2]
+  c_s = seq(0,100,len=ngrids)
+  cvout = CVPMD_softImpute(Y,c_s,K ,fold , method = "softImpute")
+  out = softImpute(Y, rank.max = K,lambda = cvout$opt_s)
+  return(list(d = out$d, u = out$u, v = out$v))
+}
+
 OCV_index=function(Y,k_fold = 5){
@@
     }
+    missing[i,,which(colSums(missing[i,,],na.rm = T) ==0)] = Y[,which(colSums(missing[i,,],na.rm = T) ==0)]
   }
@@
       SSE[i] = SSE[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] -
-                               miss_hat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2)
+                               miss_hat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2,na.rm = TRUE)
     }
@@
     data = flashr2::flash_set_data(Y_data)
-    f_greedy = flashr2::flash_add_greedy(data,Kmax=Kmax,ash_param=list(method = "fdr"),var_type = "constant")
+    # f_greedy = flashr2::flash_add_greedy(data,Kmax=Kmax,var_type = "constant")
+    f_greedy = flashr2::flash_add_greedy(data,Kmax=Kmax)
     Y_hat = f_greedy$EL %*% t(f_greedy$EF)
+  }else if(method == "flash_wn"){
+    data = flashr2::flash_set_data(Y_data)
+    # f_greedy = flashr2::flash_add_greedy(data,Kmax=Kmax,var_type = "constant",nullcheck=FALSE)
+    f_greedy = flashr2::flash_add_greedy(data,Kmax=Kmax,nullcheck=FALSE)
+    Y_hat = f_greedy$EL %*% t(f_greedy$EF)
+  }else if(method == "flash_bf"){
+    data = flashr2::flash_set_data(Y_data)
+    # f_greedy = flashr2::flash_add_greedy(data,Kmax=Kmax,var_type = "constant")
+    # f = flashr2::flash_backfit(data,f_greedy,var_type = "constant")
+    f_greedy = flashr2::flash_add_greedy(data,Kmax=Kmax)
+    f = flashr2::flash_backfit(data,f_greedy)
+    Y_hat = f$EL %*% t(f$EF)
   }else if(method == "PMD"){
-    out = PMA::PMD(Y_data, K = Kmax)
+    res_log = capture.output({out = PMA.wrapper(Y_data,ngrids = 10,K = Kmax,fold = 10)})
+    if(length(out$d)==1){
+      Y_hat =  (out$d) * out$u %*% t(out$v)
+    }else{
       Y_hat =  out$u %*% diag(out$d) %*% t(out$v)
+    }
+  }else if(method == 'PN'){
+    library(ebnm)
+    library(flashr2)
+    data = flashr2::flash_set_data(Y_data)
+    # f_greedy = flashr2::flash_add_greedy(data,Kmax=Kmax,ebnm_fn = ebnm_pn,var_type = "constant")
+    f_greedy = flashr2::flash_add_greedy(data,Kmax=Kmax,ebnm_fn = ebnm_pn)
+    Y_hat = f_greedy$EL %*% t(f_greedy$EF)
   }else if(method == "softImpute"){
-    gsoft = softImpute(Y_data, rank.max = Kmax)
-    Y_hat =  gsoft$u %*% diag(gsoft$d) %*% t(gsoft$v)
+    gsoft = try(softImpute.wrapper(Y_data,ngrids = 10,K = Kmax,fold = 10))
+    if(length(gsoft$d)==1){
+      Y_hat =  try((gsoft$d) * gsoft$u %*% t(gsoft$v))
     }else{
-    stop("the method is not implemented yet, please check it out")
+      Y_hat =  try(gsoft$u %*% diag(gsoft$d) %*% t(gsoft$v))
+    }
+  }else{
+    # stop("the method is not implemented yet, please check it out")
+    Y_hat = matrix(0,ncol = dim(Y_data)[2],nrow = dim(Y_data)[1])
   }
@@
 
-  system("matlab -nosplash -nodesktop -r \"addpath(\'~/HG/flash/data/OCVmissflashr2/Breastcancer\'); run(\'missvalue.m\');exit;\" ")
+  system("matlab -nosplash -nodesktop -r \"addpath(\'../\'); run(\'missvalue.m\');exit;\" ")
   g_nsfa = readMat("./NBSFAout/NSFAresult.mat")
@@
 ```{}
-addpath('~/HG/flash/data/missingvalue/methods/NBSF/nsfa-master/');
+ddpath('~/HG/flash/data/missingvalue/methods/NBSF/nsfa-master/');
 addpath('~/HG/flash/data/missingvalue/methods/NBSF/nsfa-master/utils/');
@@
 % you need to change the path and the data file (not centered)
-load ~/HG/flash/data/OCVmissflashr2/Breastcancer/example.mat;
-Ycentered=Y-repmat(mean(Y,2),1,size(Y,2));
+load ../Ydata.mat;
+Ycentered=Y;
 settings=defaultsettings();
@@
 
-load(fullfile(pwd,'NBSFAout/OCVindex.mat'))
+file_path = fullfile(pwd,'NBSFAout/OCVindex.mat');
+load(file_path);
 
-missindex = ones(settings.D,settings.N);
-k_fold = 5;
+% missindex = ones(settings.D,settings.N); this is not missing index this
+% is observed index.
+missindex = (~isnan(Y));
+k_fold = 10;
 foldindex = zeros(k_fold,k_fold,2);
@@
     end
+    mvmask(:,sum(mvmask) == 0) = (missindex(:,sum(mvmask) == 0));
     initialsample=init_nsfa(settings);
@@
         SSE(i) = SSE(i) + ...
-            sum(sum((Ycentered(OCVindex.rowindex(:,foldindex(j,i,1)),OCVindex.colindex(:,foldindex(j,i,2)))- ...
+            sum(nansum((Ycentered(OCVindex.rowindex(:,foldindex(j,i,1)),OCVindex.colindex(:,foldindex(j,i,2)))- ...
             Y_hat(OCVindex.rowindex(:,foldindex(j,i,1)),OCVindex.colindex(:,foldindex(j,i,2)))).^2));
@@
 
-RMSE = sqrt(sum(SSE)/(settings.D*settings.N));
+RMSE = sqrt(sum(SSE)/(100000));
 RMSE;
@@
 ```{}
+source("../wrapper.R")
 library(R.matlab)
 ## run the code
-Y_centered = readMat("~/HG/flash/data/OCVmissflashr2/Breastcancer/example.mat")
+Y_centered = readMat("../example.mat")
 Y = Y_centered$Y
@@
 Y = Y - rowMeans(Y) %*% t(rep(1,P))
-OCVindex = OCV_index(Y,k_fold = 5)
-OCVdata = OCV_data(Y,OCVindex,k_fold = 5)
-soft_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "softImpute",Kmax = 50)
-# flashG_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "flash",Kmax = 50)
-pmd_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "PMD",Kmax = 50)
+OCVindex = OCV_index(Y,k_fold = 10)
+OCVdata = OCV_data(Y,OCVindex,k_fold = 10)
+
+PN_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "PN",Kmax = 40)
+flashG_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "flash",Kmax = 40)
+flashGwn_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "flash_wn",Kmax = 40)
+flashB_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "flash_bf",Kmax = 40)
+pmd_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "PMD",Kmax = 40)
+soft_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "softImpute",Kmax = 40)
 nsfa_rmse = NSF_OCV(OCVindex)
-flashG_rmse = mean(c( nsfa_rmse,pmd_rmse,soft_rmse))
 
-result = c(flashG_rmse, nsfa_rmse,pmd_rmse,soft_rmse)
+result = c(PN_rmse,flashG_rmse,flashGwn_rmse,flashB_rmse,nsfa_rmse,pmd_rmse,soft_rmse)
 saveRDS(result, "./NBSFAout/output.rds")
-
 ```
@@
 #SBATCH --job-name=arrayJob
-#SBATCH --output=/home/weidong/HG/flash/data/OCVmissflashr2/Breastcancer/outlog/arrayJob_%A_%a.out
-#SBATCH --error=/home/weidong/HG/flash/data/OCVmissflashr2/Breastcancer/outlog/arrayJob_%A_%a.err
-#SBATCH --array=1-100
-#SBATCH --time=06:00:00
-#SBATCH --partition=broadwl
+#SBATCH --output=./outlog/arrayJob_%A_%a.out
+#SBATCH --error=./outlog/arrayJob_%A_%a.err
+#SBATCH --array=1-20
+#SBATCH --time=26:00:00
+#SBATCH --partition=mstephens
 #SBATCH --ntasks=1
-#SBATCH --mem-per-cpu=2000
+#SBATCH --mem-per-cpu=8000
 
@@
 # Print this sub-job's task ID
-cd /home/weidong/HG/flash/data/OCVmissflashr2/Breastcancer
+module load R
+module load matlab
+
 mkdir test${SLURM_ARRAY_TASK_ID}
 cd test${SLURM_ARRAY_TASK_ID}
-Rscript --verbose /home/weidong/HG/flash/data/OCVmissflashr2/Breastcancer/runOCV.R
+Rscript --verbose ../runOCV.R
 ```
@@
 ```{}
-T = 100
-results = matrix(NA,ncol = 4, nrow = T)
+T = 20
+results = matrix(NA,ncol = 7, nrow = T)
 for(i in 1:T){
@@
 library(ggplot2)
-plot_res = function(output,title = "data",legend_position = "none"){
+plot_res = function(output,title = "data",legend_position = "none", x_label){
   rmse = as.vector(output)
   N = dim(output)[1]
-  methods = rep(c("flash","NBSFA","PMD","softImpute"), each = N)
+  # methods = rep(c("flash","NBSFA","PMD","softImpute"), each = N)
+  methods = rep(x_label, each = N)
   df = data.frame(RMSE = rmse, Method = methods )
@@
 }
-
+x_label = c("flash","NBSFA","PMD","softImpute")
 PT_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/president_box.rds")
-pp = plot_res(PT_res,"Text data")
+pp = plot_res(PT_res,"Text data",x_label = x_label)
 DT_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/denoiseRtumor_box.rds")
-pd = plot_res(DT_res,"Tumor data")
+pd = plot_res(DT_res,"Tumor data",x_label = x_label)
 GZ_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/gtexzscore_box.rds")
-pg = plot_res(GZ_res,"GTEx data")
+pg = plot_res(GZ_res,"GTEx data",x_label = x_label)
 BC_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/BreastCancer_box.rds")
-pb = plot_res(BC_res,"Breast Cancer data")
+pb = plot_res(BC_res,"Breast Cancer data",x_label = x_label)
 ML_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/ML100K_box.rds")
@@
 ML_res = matrix(as.numeric(ML_res),ncol = 4)
-pM = plot_res(ML_res,"Movie Lens data")
+pM = plot_res(ML_res,"Movie Lens data",x_label = x_label)
 ```
@@
 
+## results for different tuning parameters
+
+```{r}
+pmd_c = sapply(seq(1,10),function(x){paste("PMD",x)})
+softImpute_c = sapply(seq(1,10),function(x){paste("SF",x)})
+x_label= c("flash","NBSFA",pmd_c,softImpute_c)
+PT_res = readRDS("../data/missingvalue/box_res_grids_sf_pmd/denoiseTumor_box.rds")
+pt = plot_res(PT_res,"Tumor data",x_label = x_label)
+pt
+PT_res = readRDS("../data/missingvalue/box_res_grids_sf_pmd/TEXT_prsdt_box.rds")
+pp = plot_res(PT_res,"Text data",x_label = x_label)
+pp
+```
+
+## use OCV choose the tuning parameters
+
+### Breast Cancer data
+
+we use 10 grids for `softImpute` and `PMD`
+
+```{r}
+PB_res = readRDS("../data/missingvalue/testingcode/box_Breast.rds")
+x_label= c("PN","flashG","flashGwn","flashB","nsfa","pmd","soft")
+pb = plot_res(PB_res,"Breast Cancer data",x_label = x_label)
+pb
+```
+
+we take the 'flashB' away to compare the rest.
+
+```{r}
+x_label= c("PN","flashG","flashGwn","nsfa","pmd","soft")
+pb = plot_res(PB_res[,-4],"Breast Cancer data",x_label = x_label)
+pb
+```
+
+### denoiseR Tumor data
+
+we use 10 grids for `softImpute` and `PMD`
+
+```{r}
+PT_res = readRDS("../data/missingvalue/testingcode/box_denoiseTumor.rds")
+x_label= c("PN","flashG","flashGwn","flashB","nsfa","pmd","soft")
+pt = plot_res(PT_res,"Tumor data",x_label = x_label)
+pt
+```
+
+### text data
+
+in this data, we add `zero` as use zero values as imputation to compare with other methods.
+
+```{r}
+PT_res = readRDS("../data/missingvalue/testingcode/box_president.rds")
+x_label= c("PN","flashG","flashGwn","flashB","nsfa","pmd","soft","zero")
+pt = plot_res(PT_res,"Text data",x_label = x_label)
+pt
+```
+
 ## Session information

### 2018-01-22 analysis/templateforOCV.Rmd
Files commited by wflow_commit.

@@
     data = flashr2::flash_set_data(Y_data)
-    # f_greedy = flashr2::flash_add_greedy(data,Kmax=Kmax,var_type = "constant")
-    f_greedy = flashr2::flash_add_greedy(data,Kmax=Kmax)
+    f_greedy = flashr2::flash_add_greedy(data,K=Kmax)
     Y_hat = f_greedy$EL %*% t(f_greedy$EF)
@@
     data = flashr2::flash_set_data(Y_data)
-    # f_greedy = flashr2::flash_add_greedy(data,Kmax=Kmax,var_type = "constant",nullcheck=FALSE)
-    f_greedy = flashr2::flash_add_greedy(data,Kmax=Kmax,nullcheck=FALSE)
+    f_greedy = flashr2::flash_add_greedy(data,K=Kmax,nullcheck=FALSE)
     Y_hat = f_greedy$EL %*% t(f_greedy$EF)
+  }else if(method == "flash_gb"){
+    data = flashr2::flash_set_data(Y_data)
+    f_greedy = flashr2::flash_add_greedy(data,K=Kmax)
+    f = flashr2::flash_backfit(data,f_greedy)
+    Y_hat = f$EL %*% t(f$EF)
   }else if(method == "flash_bf"){
     data = flashr2::flash_set_data(Y_data)
-    # f_greedy = flashr2::flash_add_greedy(data,Kmax=Kmax,var_type = "constant")
-    # f = flashr2::flash_backfit(data,f_greedy,var_type = "constant")
-    f_greedy = flashr2::flash_add_greedy(data,Kmax=Kmax)
-    f = flashr2::flash_backfit(data,f_greedy)
+    f_data = flashr2::flash_add_factors_from_data(data,K = Kmax)
+    f = flashr2::flash_backfit(data,f_data)
     Y_hat = f$EL %*% t(f$EF)
@@
     data = flashr2::flash_set_data(Y_data)
-    # f_greedy = flashr2::flash_add_greedy(data,Kmax=Kmax,ebnm_fn = ebnm_pn,var_type = "constant")
-    f_greedy = flashr2::flash_add_greedy(data,Kmax=Kmax,ebnm_fn = ebnm_pn)
+    f_greedy = flashr2::flash_add_greedy(data,K=Kmax,ebnm_fn = ebnm_pn)
     Y_hat = f_greedy$EL %*% t(f_greedy$EF)
+  }else if(method == 'PMD_d'){
+    out = PMA::PMD(Y_data, K = Kmax)
+    Y_hat =  out$u %*% diag(out$d) %*% t(out$v)
+  }else if(method == 'softImpute_d'){
+    gsoft = softImpute::softImpute(Y_data, rank.max = Kmax)
+    Y_hat =  gsoft$u %*% diag(gsoft$d) %*% t(gsoft$v)
   }else if(method == "softImpute"){
@@
     }
+  }else if(method == "SF_flash"){
+    gsoft = try(softImpute.wrapper(Y_data,ngrids = 10,K = Kmax,fold = 10))
+    LL = gsoft$u %*% diag(gsoft$d)
+    FF = gsoft$v
+    data = flashr2::flash_set_data(Y_data)
+    f_data = flashr2::flash_add_lf(data, LL=LL, FF=FF)
+    f = flashr2::flash_backfit(data,f_data)
+    Y_hat = f$EL %*% t(f$EF)
   }else{
@@
 
+labels | null check | backfitting | greedy | ebnm_ash | ebnm_pn 
+------ | ---------- | ----------- | ------ | -------- | ------
+flashG | yes| |yes | yes |
+flashGwn | | | yes|yes |
+flashB | yes| yes| | yes|
+PN | yes | | yes| |yes
+
 ### Breast Cancer data

### 2018-01-25 analysis/setup.Rmd
wflow_publish("setup.Rmd")

--- 2018-01-10 analysis/PCAresult.Rmd
@@
 ---
-title: "PCA(SVD) result"
-author: "Wei"
-date: 2018-01-10
-output: html_document
+title: "Setup for flash project"
+output:
+  html_document:
+    toc: false
 ---
@@
 
-<!-- Add your analysis here -->
+To reproduce our results on your own computer, please follow these
+setup instructions.
 
-```{r}
-plot_GTEXres = function(L,pve){
-  library(reshape2)
-  data_L = melt(L)
-  colnames(data_L) = c("tissue","loading","value")
-  library(ggplot2)
-  tissue_color = as.character(colordata[,2])
-  data_L$tissue = factor(data_L$tissue,levels = 1:44 ,labels = as.character(colordata[,1]) )
-  data_L$loading = factor(data_L$loading,levels = 1:14 ,labels = paste("loading",1:14,"; pve:", pve[1:14]))
+## Setup instructions
 
-  ggplot(data_L,aes(x = tissue,y = value,fill = factor(tissue) )) +
-    geom_bar(stat = "identity",width = 0.6) +
-    scale_fill_manual(values=tissue_color) +
-    scale_x_discrete(labels = NULL) +
-    theme_grey()+
-    theme(legend.position="right", legend.text=element_text(size=5), axis.text.y = element_text(size = 5)) + 
-    labs(title = "GTEx data", y = "loading values" ,x = "tissues", fill="tissue") +
-    facet_wrap(~loading, ncol = 2, scales = "free_y") +
-    guides(fill = guide_legend(ncol = 1, keyheight = 0.5, keywidth = 0.5))
-}
-b_flash = readRDS("../data/GTExdata/res_flashr2/bflashvarcol.rds")
-load("../data/GTExdata/gtexEQTL_zscore.rds")
-ssY = sum(zscore^2)
-K = dim(b_flash$EL)[2] -1
-pve = (sapply(seq(1,K),function(x){ sum(b_flash$EL[,x]^2 %*% t(b_flash$EF[,x]^2)) }))/ssY
-pve = pmax(round(pve,3),0.001)
-
-dat = read.table('../data/GTExColors.txt', sep = '\t', comment.char = '')
-colordata = dat[c(1:6,9:18,21:23,26:30,32,33,35,36,38:53),1:2]
-```
-
-```{r}
-L = b_flash$EL[,1:8]
-plot_GTEXres(L,pve)
-L = b_flash$EL[,9:16]
-plot_GTEXres(L,pve)
-
-gsvd = svd(zscore)
-pve = (sapply(seq(1,K),function(x){ sum((gsvd$d[x] * gsvd$u[,x] %*% t(gsvd$v[,x]))^2) }))/ssY
-pve = pmax(round(pve,3),0.001)
-L = gsvd$v[,1:8]
-plot_GTEXres(L,pve)
-L = gsvd$v[,9:16]
-plot_GTEXres(L,pve)
-
-```
+*Fill out setup instructions in this section. See
+[here](https://stephenslab.github.io/wflow-divvy/setup.html) and
+[here](https://github.com/pcarbo/mouthwash_sims/blob/master/README.md)
+for some ideas.*
 
@@
 
-<!-- Insert the session information into the document -->
+Here are details about the computing environment, including the
+version of R and the R packages, that were used to generate the
+results from the R Markdown documents.
+
 ```{r session-info}

### 2018-01-25 analysis/MSErank1sim.Rmd
Files commited by wflow_commit.

--- 2017-12-09 analysis/MSErank1sim.Rmd
@@
 date: 17-12-09
-output: html_document
+output: 
+  html_document:
+    code_folding: hide
 ---
@@
 
-```{}
+## run the methods
+
+### 1. creat a folder
+
+```
+mkdir rankone
+```
+
+### 2. creat .R file
+
+```{r}
+# this file is for all the R function we need here for sparse factor model
+library("MASS")
+betaO=function(P,betapi,betasigma,propt){
+  idx=rmultinom(P,size=1,prob=betapi)
+  K=length(betapi)
+  s=array(0,dim=c(K,K))
+  diag(s)=betasigma
+  bnorm=mvrnorm(P,rep(0,K),s)
+  betaO=apply(bnorm*t(idx),1,sum)
+  betaO=betaO*rbinom(P,1,propt)
+  return(betaO)
+}
+
+datamaker = function(N,P,l_pi,l_se,l_sp,f_pi,f_se,f_sp,sigmae){
+  # here I would like to fix the sparsity of L and F which is simple
+  # if we need to do futher experiment
+  L_true = betaO(N,l_pi,l_se,l_sp)
+  F_true = betaO(P,f_pi,f_se,f_sp)
+  E = matrix(rnorm(N*P,0,sigmae),ncol = P)
+  Y = L_true %*% t(F_true) + E
+  return(list(Y = Y, L_true = L_true, F_true = F_true, E = E))
+}
+
+CVPMD=function(Y,c_u,c_v){
+  N = dim(Y)[1]
+  P = dim(Y)[2]
+  colindex = matrix(sample(P,P),ncol = 5)
+  rowindex = matrix(sample(N,N),ncol = 5)
+
+  missing= array(0,dim = c(5,N,P))
+  foldindex = array(0,dim = c(5,5,2))
+  for(i in 1:5){
+    for(j in 1:5){
+      foldindex[i,j,1] = i
+      foldindex[i,j,2] = (i+j) %% 5
+    }
+  }
+  foldindex[which(foldindex == 0)] = 5
+  for(i in 1:5){
+    missing[i, , ] = Y
+    for(j in 1:5){
+      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
+    }
+  }
+  n_u = length(c_u)
+  n_v = length(c_v)
+  CVRMSE = array(0,dim = c(n_u,n_v))
+  minrmse = Inf
+  opt_u = 0
+  opt_v = 0
+  for(t_u in 1:n_u){
+    for(t_v in 1:n_v){
+      rmse = rep(0,5)
+      for(i in 1:5){
+        out = PMD(missing[i,,], sumabsu = c_u[t_u], sumabsv = c_v[t_v],center=FALSE)
+        misshat = out$d * out$u %*% t(out$v)
+        for(j in 1:5){
+          rmse[i] = rmse[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] - 
+                                     misshat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2)
+        }
+        rmse[i] = sqrt(rmse[i] / (N * P/5))
+      }
+      CVRMSE[t_u,t_v] = mean(rmse)
+      if(CVRMSE[t_u,t_v] < minrmse){
+        minrmse = CVRMSE[t_u,t_v]
+        opt_u = c_u[t_u]
+        opt_v = c_v[t_v]
+      }
+    }
+  }
+  return(list(opt_u = opt_u,opt_v = opt_v))
+}
+
+# input is a list of Y L_true and F_true which is called Y_data
+PMA.wrapper = function(Y_data,ngrids = 10){
+  library(PMA)
+  Y = Y_data$Y
+  L_true = Y_data$L_true
+  F_true = Y_data$F_true
+  E = Y_data$E
+  N = dim(Y)[1]
+  P = dim(Y)[2]
+  c_u = seq(1,sqrt(N),length.out = ngrids)
+  c_v = seq(1,sqrt(P),length.out = ngrids)
+  cvout = CVPMD(Y,c_u,c_v)
+  out = PMD(Y,sumabsu = cvout$opt_u, sumabsv = cvout$opt_v,center=FALSE )
+  Y_hat = out$d * out$u %*% t(out$v)
+  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
+  return(RMSE)
+}
+
+PMA_d.wrapper = function(Y_data){
+  library(PMA)
+  Y = Y_data$Y
+  L_true = Y_data$L_true
+  F_true = Y_data$F_true
+  E = Y_data$E
+  N = dim(Y)[1]
+  P = dim(Y)[2]
+  cv.out <- PMA::PMD.cv(Y, type="standard", sumabss=seq(0.1, 1, len=20),center=FALSE)
+  out <-  PMA::PMD(Y, type="standard", sumabs=cv.out$bestsumabs, K=1, v=cv.out$v.init,center=FALSE)
+  Y_hat = out$d * out$u %*% t(out$v)
+  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
+  return(RMSE)
+}
+
+PN.wrapper = function(Y_data){
+  Y = Y_data$Y
+  L_true = Y_data$L_true
+  F_true = Y_data$F_true
+  E = Y_data$E
+  N = dim(Y)[1]
+  P = dim(Y)[2]
+  library(ebnm)
+  library(flashr2)
+  data =  flashr2::flash_set_data(Y)
+  g_flash = flashr2::flash_r1(data,verbose=F,var_type = "constant",ebnm_fn = ebnm_pn)
+  Y_hat = g_flash$EL %*% t(g_flash$EF)
+  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
+  return(RMSE)
+}
+
 flash.wrapper = function(Y_data){
@@
   data =  flashr2::flash_set_data(Y)
-  g_flash = flashr2::flash_r1(data,verbose=F)
+  g_flash = flashr2::flash_r1(data,verbose=F,var_type = "constant")
   Y_hat = g_flash$EL %*% t(g_flash$EF)
@@
 
-#here I use SVD instead of KSVD.
+SFA.wrapper = function(Y_data, K){
+  #sfa put sparse on loadings
+  Y = Y_data$Y
+  # N vector
+  L_true = Y_data$L_true
+  # P vector
+  F_true = Y_data$F_true
+  E = Y_data$E
+  N = dim(Y)[1]
+  P = dim(Y)[2]
+
+  if(file.exists("SFAout")){
+    unlink("SFAout", recursive= T)
+  }
+  system("mkdir SFAout")
+
+  write.table(Y, file="dscsfa.txt",row.names=F,col.names=F)
+  # sfa command line on cluster
+  # ~/flash/simulation/methods/sfa/src/sfa -gen ./dscsfamix.txt -g 600 -k 1 -n 200 -iter 100 -rand 999 -o ./SFAout/sfa
+
+  # sqrt(mean(( loadings %*% factor - test$L_true%*%t(test$F_true))^2))/sqrt(mean((test$L_true%*%t(test$F_true))^2))
+  system(paste("~/HG/flash/data/missingvalue/methods/sfa/src/sfa", "-gen", "./dscsfa.txt", "-g", N, "-k", K,
+               "-n", P, "-iter", 100, "-rand", 999, "-o", "./SFAout/sfa", sep=" "))
+
+  if(file.info("./SFAout/sfa_F.out")$size == 1){
+    L_hat = 0
+    F_hat = 0
+    RMSE = 1
+    return(RMSE)
+  } else{
+    Fhat=read.table("./SFAout/sfa_F.out")
+    lambda=read.table("./SFAout/sfa_lambda.out")
+    Fhat=as.matrix(Fhat)
+    lambda=as.matrix(lambda)
+    L_hat = lambda
+    F_hat = Fhat
+
+    Y_hat = L_hat %*% F_hat
+    RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
+    return(RMSE)
+  }
+}
+
+SFAmix.wrapper = function(Y_data,K){
+  Y = t(Y_data$Y)
+  L_true = Y_data$L_true
+  F_true = Y_data$F_true
+  E = t(Y_data$E)
+  N = dim(Y)[1]
+  P = dim(Y)[2]
+
+  write.table(Y,file="dscsfamix.txt",row.names=F,col.names=F)
+  print(class(Y))
+  print(dim(Y))
+
+  if(file.exists("SFAmixout")){
+    unlink("SFAmixout", recursive= T)
+  }
+  system("mkdir SFAmixout")
+
+  # ~/mvash/mvsim/SFAmix/SFAmix --nf K --y dscrsfamix.txt --out SFAmixout --sep space
+  # this is run on the PPS cluster
+  system(paste("~/HG/flash/data/missingvalue/methods/SFAmix/SFAmix",
+               "--nf",K,"--y","dscsfamix.txt","--out",
+               "SFAmixout","--sep","space",sep=" "))
+
+  alpha=read.table("./SFAmixout/PSI")
+
+  if(file.info("./SFAmixout/EX")$size == 1){
+    L_hat = 0
+    F_hat = 0
+    RMSE = 1
+    return(RMSE)
+  } else{
+    Fhat=read.table("./SFAmixout/EX")
+    lambda=read.table("./SFAmixout/LAM")
+    Psi=as.vector(alpha)
+    Fhat=as.matrix(Fhat)
+    lambda=as.matrix(lambda)
+    # return(list(L = Fhat, F = lambda))
+    L_hat = Fhat
+    F_hat = lambda
+
+    Y_hat = L_hat %*% F_hat
+    RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
+    return(RMSE)
+  }
+}
+
 SVD.wrapper = function(Y_data){
@@
 }
+
+CV_softImpute=function(Y,c_s,K,fold = 5, method = "PMD"){
+  N = dim(Y)[1]
+  P = dim(Y)[2]
+  colindex = matrix(sample(P,P),ncol = fold)
+  rowindex = matrix(sample(N,N),ncol = fold)
+
+  missing= array(0,dim = c(fold,N,P))
+  foldindex = array(0,dim = c(fold,fold,2))
+  for(i in 1:fold){
+    for(j in 1:fold){
+      foldindex[i,j,1] = i
+      foldindex[i,j,2] = (i+j) %% fold
+    }
+  }
+  foldindex[which(foldindex == 0)] = fold
+  for(i in 1:fold){
+    missing[i, , ] = Y
+    for(j in 1:fold){
+      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
+    }
+    missing[i,,which(colSums(missing[i,,],na.rm = T) ==0)] = Y[,which(colSums(missing[i,,],na.rm = T) ==0)]
+  }
+  # c_s is the candicate of shrinkage parameter
+  n_s = length(c_s)
+  # rmse for each grids
+  CVRMSE = rep(0,n_s)
+  minrmse = Inf
+  opt_s = 0
+  # for each candidate, we run it N_sim times
+  for(t_s in 1:n_s){
+    # for each grid
+    # each time we set the rmse to zeros
+    rmse = rep(0,fold)
+    for(i in 1:fold){
+      if(method == "PMD"){
+        res_log = capture.output({out = PMD(missing[i,,], sumabs = c_s[t_s], sumabsv = NULL, sumabsu = NULL,K = K)})
+      }else{
+        out = softImpute(missing[i,,], rank.max = K,lambda = c_s[t_s])
+      }
+      if(length(out$d)==1){
+        misshat = (out$d) * out$u %*% t(out$v)
+      }else{
+        misshat = out$u %*%  diag(out$d) %*% t(out$v)
+      }
+      for(j in 1:fold){
+        # for each fold j
+        rmse[i] = rmse[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] -
+                                   misshat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2,na.rm = TRUE)
+      }
+    } #get the result for one run
+    CVRMSE[t_s] = CVRMSE[t_s] + sqrt(sum(rmse)/(N*P))
+    if(CVRMSE[t_s] < minrmse){
+      minrmse = CVRMSE[t_s]
+      opt_s = c_s[t_s]
+    }
+  }
+  return(list(opt_s = opt_s, output = CVRMSE))
+}
+
+softImpute.wrapper = function(Y_data, ngrids = 10, K = 1, fold = 5){
+  library(softImpute)
+  Y = Y_data$Y
+  L_true = Y_data$L_true
+  F_true = Y_data$F_true
+  E = Y_data$E
+  N = dim(Y)[1]
+  P = dim(Y)[2]
+  c_s = seq(0,100,len=ngrids)
+  cvout = CV_softImpute(Y,c_s,K ,fold , method = "softImpute")
+  out = softImpute(Y, rank.max = K,lambda = cvout$opt_s)
+  Y_hat = out$d * out$u %*% t(out$v)
+  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
+  return(RMSE)
+}
+
 ```
+
+name this file as `Rfunction.R`
+
+### creat the folder for different case
+
+```{}
+mkdir intermediate
+```
+
+### creat the run.R file
+
+```{}
+cd intermediate
+```
+
+creat `run.R` file as follows:
 
@@
 library(ssvd)
-source("~/HG/flash/data/simulations/rankone/MSE/Rfunction.R")
+source("../Rfunction.R")
 L_se = c( 0.25, 0.5, 1, 2, 4)
@@
 P = 300
-# Data = datamaker(N,P,L_pi,L_se,0.7,c(1),c(1),1,sqrt(16))
-Data = datamaker(N,P,L_pi,L_se,0.1,c(1),c(1),1,sqrt(1))
+Data = datamaker(N,P,L_pi,L_se,0.7,c(1),c(1),1,sqrt(16))
+# Data = datamaker(N,P,L_pi,L_se,0.1,c(1),c(1),1,sqrt(1))  sparse case
+# Data = datamaker(N,P,L_pi,L_se,1.0,c(1),c(1),1,sqrt(1))  dense case
 RMSE = rep(NA,6)
@@
 
+The score we restroe is as follows:
+
 $$RMSE =\frac{ \sqrt{(Y-\hat{Y} - E )^2}}{\sqrt{(Y - E )^2}}$$
 
-## batch file
+### batch file
 
@@
 #SBATCH --job-name=arrayJob
-#SBATCH --output=/home/weidong/HG/flash/data/simulations/rankone/MSE/median/outlog/arrayJob_%A_%a.out
-#SBATCH --error=/home/weidong/HG/flash/data/simulations/rankone/MSE/median/outlog/arrayJob_%A_%a.err
+#SBATCH --output=./outlog/arrayJob_%A_%a.out
+#SBATCH --error=./outlog/arrayJob_%A_%a.err
 #SBATCH --array=1-100
@@
 # Print this sub-job's task ID
-cd /home/weidong/HG/flash/data/simulations/rankone/MSE/median
 mkdir test${SLURM_ARRAY_TASK_ID}
 cd test${SLURM_ARRAY_TASK_ID}
-Rscript --verbose /home/weidong/HG/flash/data/simulations/rankone/MSE/median/run.R
+Rscript --verbose ../run.R
 ```
 
-## run
+name the above file as `Jobs.sbatch`
 
@@
   test_folder = paste("test", i, sep = "")
-  # out_folder = "NBSFAout"
   out_file = "output.rds"
@@
 }
+
+saveRDS(results,"./RES_rrmse.rds")
 ```
+
+## plot the result
 
@@
 ```{r}
+plot_res = function(output,title = "data",legend_position = "none", methods_name){
+  rmse = as.vector(output)
+  N = dim(output)[1]
+  methods = rep(methods_name, each = N)
+  df = data.frame(RRMSE_diff = rmse, Method = methods )
+  p<-ggplot(df, aes(x=Method, y=RRMSE_diff, color=Method)) +
+  geom_boxplot()+
+  # geom_violin()+
+  ggtitle(title) +  theme_bw()+
+    theme(legend.position= legend_position, legend.text=element_text(size=10),
+          axis.text.y = element_text(size =12),
+          axis.text.x = element_text(size =12,angle = 45, hjust = 1))
+  p
+}
+library(ggplot2)
+```
+
+```{r}
 sparse_res = readRDS("../data/simulation/rankone/RMSE/result_sparse.rds")
-colnames(sparse_res) = c("PMD","flash","SFA","SFAmix","SVD","SSVD")
-boxplot(sparse_res,main = "sparse case")
+colnames(sparse_res) = c("PMD.cv2","flash","SFA","flash.pn","SVD","SSVD","PMD.cv1","SFAmix","SI.cv")
+sparse_diff = sparse_res - sparse_res[,2]
+methods_name = colnames(sparse_diff)
+p1 = plot_res(sparse_diff,title = "Difference from FLASH result (90% zeros)",methods_name = methods_name)
+methods_name = colnames(sparse_diff)[c(1,2,3,4,5,7,9)]
+p1_cut = plot_res(sparse_diff[-22,c(1,2,3,4,5,7,9)],title = "Difference from FLASH result (90% zeros)",methods_name = methods_name)
 ```
@@
 sparse_res = readRDS("../data/simulation/rankone/RMSE/result_itermediate.rds")
-colnames(sparse_res) = c("PMD","flash","SFA","SFAmix","SVD","SSVD")
-boxplot(sparse_res,main = "intermediate sparse case")
+colnames(sparse_res) = c("PMD.cv2","flash","SFA","flash.pn","SVD","SSVD","PMD.cv1","SFAmix","SI.cv")
+sparse_diff = sparse_res - sparse_res[,2]
+# p2 = plot_res(sparse_diff,"Difference from FLASH result (30% zeros)")
+
+methods_name = colnames(sparse_diff)
+p2 = plot_res(sparse_diff,title = "Difference from FLASH result (30% zeros)",methods_name = methods_name)
+methods_name = colnames(sparse_diff)[c(1,2,3,4,5,7,9)]
+p2_cut = plot_res(sparse_diff[,c(1,2,3,4,5,7,9)],title = "Difference from FLASH result (30% zeros)",methods_name = methods_name)
+```
+
+## RMSE dense case 100 simulations
+
+```{r}
+sparse_res = readRDS("../data/simulation/rankone/RMSE/result_dense.rds")
+colnames(sparse_res) = c("PMD.cv2","flash","SFA","flash.pn","SVD","SSVD","PMD.cv1","SFAmix","SI.cv")
+sparse_diff = sparse_res - sparse_res[,2]
+# p3 = plot_res(sparse_diff,"Difference from FLASH result (0% zeros)")
+methods_name = colnames(sparse_diff)
+p3 = plot_res(sparse_diff,title = "Difference from FLASH result (0% zeros)",methods_name = methods_name)
+methods_name = colnames(sparse_diff)[c(1,2,3,4,5,7,9)]
+p3_cut = plot_res(sparse_diff[,c(1,2,3,4,5,7,9)],title = "Difference from FLASH result (0% zeros)",methods_name = methods_name)
+```
+
+```{r}
+# gridExtra::grid.arrange(p1,p2,p3,p1_cut,p2_cut,p3_cut,ncol = 3)
+gridExtra::grid.arrange(p1,p1_cut,p2,p2_cut,p3,p3_cut,ncol = 2)
 ```

### 2018-01-25 analysis/gtex_si.rmd

---
title: "Untitled"
author: "Matthew Stephens"
date: "January 23, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
set.seed(1)
Zscores = readRDS("~/Downloads/MatrixEQTLSumStats.Portable.Z.rds")
Ztop = Zscores$test.z
z.s = softImpute::softImpute(Ztop,rank.max=30,lambda=89)
```

```{r}
#this plotting due to Wei Wang
library(ggplot2)
library(reshape2)
#' @param L a ntissue by K matrix of loadings
#' @param pve a vector of K pve values
#' @param offset a number to add to the loading numbers in titles  (eg set to 9 if plotting loadings 9-x)
plot_GTEXres = function(L,pve,offset=0){
  dat = read.table("GTExColors.txt", sep = '\t', comment.char = '')
  nL = dim(L)[2]
  colordata = dat[c(1:6,9:18,21:23,26:30,32,33,35,36,38:53),1:2]
  data_L = melt(L)
  colnames(data_L) = c("tissue","loading","value")
  tissue_color = as.character(colordata[,2])
  data_L$tissue = factor(data_L$tissue,levels = 1:44 ,labels = as.character(colordata[,1]) )
  data_L$loading = factor(data_L$loading,levels = 1:nL ,labels = paste("loading",offset+(1:nL),"; pve:", round(pve,3)))

  ggplot(data_L,aes(x = tissue,y = value,fill = factor(tissue) )) +
    geom_bar(stat = "identity",width = 0.6) +
    scale_fill_manual(values=tissue_color) +
    scale_x_discrete(labels = NULL) +
    theme_grey()+
    theme(legend.position="right", legend.text=element_text(size=5), axis.text.y = element_text(size = 5)) + 
    labs(title = "GTEx data", y = "loading values" ,x = "tissues", fill="tissue") +
    facet_wrap(~loading, ncol = 2, scales = "free_y") +
    guides(fill = guide_legend(ncol = 1, keyheight = 0.5, keywidth = 0.5))
}

pve.si = z.s$d^2/sum(Ztop^2)
plot_GTEXres(z.s$v[,1:8],pve.si[1:8])
plot_GTEXres(z.s$v[,9:16],pve.si[9:16],9)

```

### 2018-01-25 analysis/simulation3.Rmd
Files commited by wflow_commit.

--- 2017-04-05 analysis/simulation3.Rmd
@@
 date: 2017-04-04
-output: html_document
+output: 
+  html_document:
+    code_folding: hide
 ---
@@
 
-```{r}
-
-CVPMD=function(Y,c_s,K){
-  N = dim(Y)[1]
-  P = dim(Y)[2]
-  colindex = matrix(sample(P,P),ncol = 5)
-  rowindex = matrix(sample(N,N),ncol = 5)
-
-  missing= array(0,dim = c(5,N,P))
-  foldindex = array(0,dim = c(5,5,2))
-  for(i in 1:5){
-    for(j in 1:5){
-      foldindex[i,j,1] = i
-      foldindex[i,j,2] = (i+j) %% 5
-    }
-  }
-  foldindex[which(foldindex == 0)] = 5
-  for(i in 1:5){
-    missing[i, , ] = Y
-    for(j in 1:5){
-      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
-    }
-  }
-  n_s = length(c_s)
-  CVRMSE = rep(NA,n_s)
-  minrmse = Inf
-  opt_s = 0
-  # here we only need one layer
-  for(t_s in 1:n_s){
-    rmse = rep(0,5)
-    for(i in 1:5){
-      out = PMD(missing[i,,], sumabs = c_s[t_s], sumabsv = NULL, sumabsu = NULL,K = K)
-      misshat = out$u %*%  diag(out$d) %*% t(out$v)
-      for(j in 1:5){
-        rmse[i] = rmse[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] - 
-                                   misshat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2)
-      }
-      rmse[i] = sqrt(rmse[i] / (N * P/5))
-    }
-    CVRMSE[t_s] = mean(rmse)
-    if(CVRMSE[t_s] < minrmse){
-      minrmse = CVRMSE[t_s]
-      opt_s = c_s[t_s]
-    }
-  }
-
-  return(list(opt_s = opt_s ))
-}
-
-PMA.wrapper = function(Y,ngrids = 10,K){
-  library(PMA)
-  N = dim(Y)[1]
-  P = dim(Y)[2]
-  c_s = seq(0.2,0.8,len=ngrids)
-  cvout = CVPMD(Y,c_s, K = K)
-  out = PMD(Y,sumabsu = NULL, sumabsv = NULL, sumabs = cvout$opt_s ,K = K)
-  residual_PMD = Y -  out$u %*%  diag(out$d) %*% t(out$v)
-  return(list(d = out$d, u = out$u, v = out$v, residual = residual_PMD))
-}
-
-```
+This file is just generate the structure heatmap of rank 3 data.
 
 ```{r}
-set.seed(99)
+DataMaker = function(sigma = 2,seed_num = 99){
+  set.seed(seed_num)
   N = 150
@@
   # generate Y
-E = matrix(rnorm(N*P,0,2),nrow=N)
+  E = matrix(rnorm(N*P,0,sigma),nrow=N)
   Y = L_true %*% t(F_true) + E
-g_flash = flashr::greedy(Y,K = 5)
-g_b_flash = flashr::backfitting(Y,initial_list = g_flash)
+  return(list(Y = Y, L_true = L_true, F_true = F_true, E = E))
+}
 
-outcv = PMA.wrapper(Y,10,K = 3)
+CVPMD_softImpute=function(Y,c_s,K,fold = 10, method = "PMD"){
+  N = dim(Y)[1]
+  P = dim(Y)[2]
+  colindex = matrix(sample(P,P),ncol = fold)
+  rowindex = matrix(sample(N,N),ncol = fold)
+
+  missing= array(0,dim = c(fold,N,P))
+  foldindex = array(0,dim = c(fold,fold,2))
+  for(i in 1:fold){
+    for(j in 1:fold){
+      foldindex[i,j,1] = i
+      foldindex[i,j,2] = (i+j) %% fold
+    }
+  }
+  foldindex[which(foldindex == 0)] = fold
+  for(i in 1:fold){
+    missing[i, , ] = Y
+    for(j in 1:fold){
+      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
+    }
+    missing[i,,which(colSums(missing[i,,],na.rm = T) ==0)] = Y[,which(colSums(missing[i,,],na.rm = T) ==0)]
+  }
+  # c_s is the candicate of shrinkage parameter
+  n_s = length(c_s)
+  # rmse for each grids
+  CVRMSE = rep(0,n_s)
+  minrmse = Inf
+  opt_s = 0
+  # for each candidate, we run it N_sim times
+  for(t_s in 1:n_s){
+    # for each grid
+    # each time we set the rmse to zeros
+    rmse = rep(0,fold)
+    for(i in 1:fold){
+      if(method == "PMD"){
+        res_log = capture.output({out = PMD(missing[i,,], sumabs = c_s[t_s], sumabsv = NULL, sumabsu = NULL,K = K)})
+      }else{
+        out = softImpute(missing[i,,], rank.max = K,lambda = c_s[t_s])
+      }
+      if(length(out$d)==1){
+        misshat = (out$d) * out$u %*% t(out$v)
+      }else{
+        misshat = out$u %*%  diag(out$d) %*% t(out$v)
+      }
+      for(j in 1:fold){
+        # for each fold j
+        rmse[i] = rmse[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] -
+                                   misshat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2,na.rm = TRUE)
+      }
+    } #get the result for one run
+    CVRMSE[t_s] = CVRMSE[t_s] + sqrt(sum(rmse)/(N*P))
+    if(CVRMSE[t_s] < minrmse){
+      minrmse = CVRMSE[t_s]
+      opt_s = c_s[t_s]
+    }
+  }
+  return(list(opt_s = opt_s, output = CVRMSE))
+}
+
+PMA.wrapper = function(Y_data, ngrids = 10, K=3, fold = 10){
+  library(PMA)
+  Y = Y_data$Y
+  L_true = Y_data$L_true
+  F_true = Y_data$F_true
+  E = Y_data$E
+  N = dim(Y)[1]
+  P = dim(Y)[2]
+  c_s = seq(0.1,0.9,len=ngrids)
+  cvout = CVPMD_softImpute(Y,c_s,K ,fold , method = "PMD")
+  res_log = capture.output({out = PMD(Y,sumabsu = NULL, sumabsv = NULL, sumabs = cvout$opt_s ,K = K)})
+  Y_hat = out$u %*% diag(out$d) %*%  t(out$v)
+  return(Y_hat)
+}
+
+softImpute.wrapper = function(Y_data, ngrids = 10, K = 3, fold = 10){
+  library(softImpute)
+  Y = Y_data$Y
+  L_true = Y_data$L_true
+  F_true = Y_data$F_true
+  E = Y_data$E
+  N = dim(Y)[1]
+  P = dim(Y)[2]
+  c_s = seq(0,100,len=ngrids)
+  cvout = CVPMD_softImpute(Y,c_s,K ,fold , method = "softImpute")
+  out = softImpute(Y, rank.max = K,lambda = cvout$opt_s)
+  if(length(out$d)==1){
+    Y_hat = (out$d) * out$u %*% t(out$v)
+  }else{
+    Y_hat = out$u %*%  diag(out$d) %*% t(out$v)
+  }
+  return(Y_hat)
+}
+
+SVD.wrapper = function(Y_data,K = 3){
+  Y = Y_data$Y
+  L_true = Y_data$L_true
+  F_true = Y_data$F_true
+  E = Y_data$E
+  N = dim(Y)[1]
+  P = dim(Y)[2]
+  gsvd = svd(Y,nu = 3,nv = 3)
+  Y_hat = (gsvd$u[,1:K] %*% diag(gsvd$d[1:K]) %*% t(gsvd$v[,1:K]))
+  return(Y_hat)
+}
+
+SSVD.wrapper = function(Y_data,K = 3){
+  library(ssvd)
+  Y = Y_data$Y
+  L_true = Y_data$L_true
+  F_true = Y_data$F_true
+  E = Y_data$E
+  N = dim(Y)[1]
+  P = dim(Y)[2]
+  gssvd = ssvd::ssvd(Y,method = "method",r = K)
+  Y_hat = (gssvd$u %*% diag(gssvd$d) %*% t(gssvd$v))
+  return(Y_hat)
+}
+
+flashBF.wrapper = function(Y_data,K = 3){
+  Y = Y_data$Y
+  L_true = Y_data$L_true
+  F_true = Y_data$F_true
+  E = Y_data$E
+  N = dim(Y)[1]
+  P = dim(Y)[2]
+  library(ebnm)
+  library(flashr2)
+  data =  flashr2::flash_set_data(Y)
+  f_greedy = flashr2::flash_add_greedy(data,verbose=F,var_type = "constant", K = K)
+  f = flashr2::flash_backfit(data,f_greedy,var_type = "constant")
+  Y_hat = f$EL %*% t(f$EF)
+  return(Y_hat)
+}
+
+flash.wrapper = function(Y_data, K = 3){
+  # missindex is a matirx with 3 column here: i j x
+  # Y has miss value already
+  Y = Y_data$Y
+  L_true = Y_data$L_true
+  F_true = Y_data$F_true
+  E = Y_data$E
+  N = dim(Y)[1]
+  P = dim(Y)[2]
+  data =  flashr2::flash_set_data(Y)
+  g_flash = flashr2::flash_add_greedy(data,verbose=F,var_type = "constant", K = K)
+  Y_hat = g_flash$EL %*% t(g_flash$EF)
+  return(Y_hat)
+}
 ```
@@
 ```{r}
-par(mfrow = c(2,2))
-image( t(apply(t(g_b_flash$l %*% t(g_b_flash$f)),2,rev)),col = grey(seq(0,1,length.out = 5)),main = "FLASH")
-image(t(apply(t(outcv$u %*% diag(outcv$d) %*% t(outcv$v)),2,rev)),col = grey(seq(0,1,length.out = 5)),main = "PMD")
-image(t(apply(t(Y),2,rev)),col = grey(seq(0,1,length.out = 5)),main = "observed data")
-image(t(apply(t(G),2,rev)),col = grey(seq(0,1,length.out = 5)),main = "underlying structure")
-par(mfrow = c(2,2))
-image( t(apply(t(g_b_flash$l %*% t(g_b_flash$f)),2,rev)),col =blues9,main = "FLASH")
-image(t(apply(t(outcv$u %*% diag(outcv$d) %*% t(outcv$v)),2,rev)),col = blues9,main = "PMD")
-image(t(apply(t(Y),2,rev)),col = blues9,main = "observed data")
-image(t(apply(t(G),2,rev)),col = blues9,main = "underlying structure")
+library(PMA)
+library(flashr2)
+library(ssvd)
+library(softImpute)
+Y_data = DataMaker(seed_num = 0)
+
+LF_1 = PMA.wrapper(Y_data)
+LF_2 = flash.wrapper(Y_data)
+LF_3 = flashBF.wrapper(Y_data)
+LF_4 = SVD.wrapper(Y_data)
+LF_5 = SSVD.wrapper(Y_data)
+LF_6 = softImpute.wrapper(Y_data)
+LF_true = Y_data$L_true %*% t(Y_data$F_true)
+
+library(reshape2)
+melted_cormat_1 <- melt(LF_1, na.rm = TRUE)
+melted_cormat_2 <- melt(LF_2, na.rm = TRUE)
+melted_cormat_3 <- melt(LF_3, na.rm = TRUE)
+melted_cormat_4 <- melt(LF_4, na.rm = TRUE)
+melted_cormat_5 <- melt(LF_5, na.rm = TRUE)
+melted_cormat_6 <- melt(LF_6, na.rm = TRUE)
+melted_cormat_true <- melt(LF_true, na.rm = TRUE)
+# Heatmap
+library(ggplot2)
+plot_HP = function(title_name = "truth", melted_cormat){
+  p1 = ggplot(data = melted_cormat, aes(Var2, Var1, fill = value))+
+   geom_tile(color = "white")+
+    scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
+     midpoint = 0, limit = c(-2,2), space = "Lab") + labs(title = title_name, y = "samples", x = "variables") +
+    theme_minimal() + theme(legend.position="none") 
+  p1
+}
+
+p0 = plot_HP(title_name = "Truth", melted_cormat = melted_cormat_true)
+p1 = plot_HP(title_name = "PMD.cv1", melted_cormat = melted_cormat_1)
+p2 = plot_HP(title_name = "flash", melted_cormat = melted_cormat_2)
+p3 = plot_HP(title_name = "flashb", melted_cormat = melted_cormat_3)
+p4 = plot_HP(title_name = "SVD", melted_cormat = melted_cormat_4)
+p5 = plot_HP(title_name = "SSVD", melted_cormat = melted_cormat_5)
+p6 = plot_HP(title_name = "SI.cv", melted_cormat = melted_cormat_6)
 ```
@@
 ```{r}
-library(ggplot2)
-library(cowplot)
-mat = cbind(L_true,g_b_flash$l,outcv$u)
-mat[,8] = -mat[,8]
-mat[,9] = -mat[,9]
-mat[,6] = -mat[,6]
-mat1 = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x]/sqrt(sum((mat[,x])^2)) } )
-colnames(mat1) = c("truth.l1","truth.l2","truth.l3","FLASH.l3","FLASH.l2","FLASH.l1","PMD.u3","PMD.u2","PMD.u1")
-df1 = data.frame(mat1)
-p1_1_l <- ggplot(df1, aes(truth.l1, FLASH.l1)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("loading 1 (FLASH)")
-p1_2_l <- ggplot(df1, aes(truth.l2, FLASH.l2)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("loading 2 (FLASH)")
-p1_3_l <- ggplot(df1, aes(truth.l3, FLASH.l3)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("loading 3 (FLASH)")
-p2_1_l <- ggplot(df1, aes(truth.l1, PMD.u1)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("loading 1 (PMD)")
-p2_2_l <- ggplot(df1, aes(truth.l2, PMD.u2)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("loading 2 (PMD)")
-p2_3_l <- ggplot(df1, aes(truth.l3, PMD.u3)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("loading 3 (PMD)")
-
-library(ggplot2)
-library(cowplot)
-mat = cbind(F_true,g_b_flash$f,outcv$v)
-mat[,8] = -mat[,8]
-mat[,9] = -mat[,9]
-mat[,6] = -mat[,6]
-mat1 = sapply(seq(1,dim(mat)[2]), function(x){ mat[,x]/sqrt(sum((mat[,x])^2)) } )
-colnames(mat1) = c("truth.f1","truth.f2","truth.f3","FLASH.f3","FLASH.f2","FLASH.f1","PMD.v3","PMD.v2","PMD.v1")
-df1 = data.frame(mat1)
-p1_1 <- ggplot(df1, aes(truth.f1, FLASH.f1)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("factor 1 (FLASH)")
-p1_2 <- ggplot(df1, aes(truth.f2, FLASH.f2)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("factor 2 (FLASH)")
-p1_3 <- ggplot(df1, aes(truth.f3, FLASH.f3)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("factor 3 (FLASH)")
-p2_1 <- ggplot(df1, aes(truth.f1, PMD.v1)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("factor 1 (PMD)")
-p2_2 <- ggplot(df1, aes(truth.f2, PMD.v2)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("factor 2 (PMD)")
-p2_3 <- ggplot(df1, aes(truth.f3, PMD.v3)) + geom_point() +  geom_abline(intercept = 0, slope = 1)+theme_gray() + ggtitle("factor 3 (PMD)")
-plot_grid(p1_1_l, p2_1_l, p1_1, p2_1,
-          p1_2_l, p2_2_l, p1_2, p2_2,
-          p1_3_l, p2_3_l, p1_3, p2_3,labels = "" , ncol = 4)
+gridExtra::grid.arrange(p0,p5,p6,p4,p1,p2,ncol = 2)
 ```

### 2018-01-25 analysis/MSErank3sim.Rmd
Files commited by wflow_commit.

---
title: "MSE for rank_1 simulation"
author: "wei wang"
date: 18-01-23
output: 
  html_document:
    code_folding: hide
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

This file is just generate the RRMSE plot of rank 3 data.

```{r}
# the example in our paper using sigma = 2
DataMaker = function(sigma = 2){
  N = 150
  P = 240
  K = 3
  L_true = array(0,dim = c(N,K))
  F_true = array(0,dim = c(P,K))
  F_true[1:80,1] = rnorm(80,0,0.5)
  F_true[81:160,2] = rnorm(80,0,1)
  F_true[161:240,3] = rnorm(80,0,2)
  L_true[1:10,1] = rnorm(10,0,2)
  L_true[11:60,2] = rnorm(50,0,1)
  L_true[61:150,3] = rnorm(90,0,0.5)
  G = L_true %*% t(F_true)
  # generate Y
  E = matrix(rnorm(N*P,0,sigma),nrow=N)
  Y = L_true %*% t(F_true) + E
  return(list(Y = Y, L_true = L_true, F_true = F_true, E = E))
}

CVPMD_softImpute=function(Y,c_s,K,fold = 10, method = "PMD"){
  N = dim(Y)[1]
  P = dim(Y)[2]
  colindex = matrix(sample(P,P),ncol = fold)
  rowindex = matrix(sample(N,N),ncol = fold)

  missing= array(0,dim = c(fold,N,P))
  foldindex = array(0,dim = c(fold,fold,2))
  for(i in 1:fold){
    for(j in 1:fold){
      foldindex[i,j,1] = i
      foldindex[i,j,2] = (i+j) %% fold
    }
  }
  foldindex[which(foldindex == 0)] = fold
  for(i in 1:fold){
    missing[i, , ] = Y
    for(j in 1:fold){
      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
    }
    missing[i,,which(colSums(missing[i,,],na.rm = T) ==0)] = Y[,which(colSums(missing[i,,],na.rm = T) ==0)]
  }
  # c_s is the candicate of shrinkage parameter
  n_s = length(c_s)
  # rmse for each grids
  CVRMSE = rep(0,n_s)
  minrmse = Inf
  opt_s = 0
  # for each candidate, we run it N_sim times
  for(t_s in 1:n_s){
    # for each grid
    # each time we set the rmse to zeros
    rmse = rep(0,fold)
    for(i in 1:fold){
      if(method == "PMD"){
        res_log = capture.output({out = PMD(missing[i,,], sumabs = c_s[t_s], sumabsv = NULL, sumabsu = NULL,K = K)})
      }else{
        out = softImpute(missing[i,,], rank.max = K,lambda = c_s[t_s])
      }
      if(length(out$d)==1){
        misshat = (out$d) * out$u %*% t(out$v)
      }else{
        misshat = out$u %*%  diag(out$d) %*% t(out$v)
      }
      for(j in 1:fold){
        # for each fold j
        rmse[i] = rmse[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] -
                                   misshat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2,na.rm = TRUE)
      }
    } #get the result for one run
    CVRMSE[t_s] = CVRMSE[t_s] + sqrt(sum(rmse)/(N*P))
    if(CVRMSE[t_s] < minrmse){
      minrmse = CVRMSE[t_s]
      opt_s = c_s[t_s]
    }
  }
  return(list(opt_s = opt_s, output = CVRMSE))
}

PMA.wrapper = function(Y_data, ngrids = 10, K=3, fold = 10){
  library(PMA)
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  c_s = seq(0.1,0.9,len=ngrids)
  cvout = CVPMD_softImpute(Y,c_s,K ,fold , method = "PMD")
  res_log = capture.output({out = PMD(Y,sumabsu = NULL, sumabsv = NULL, sumabs = cvout$opt_s ,K = K)})
  Y_hat = out$u %*% diag(out$d) %*%  t(out$v)
  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
  return(RMSE)
}

softImpute.wrapper = function(Y_data, ngrids = 10, K = 3, fold = 10){
  library(softImpute)
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  c_s = seq(0,100,len=ngrids)
  cvout = CVPMD_softImpute(Y,c_s,K ,fold , method = "softImpute")
  out = softImpute(Y, rank.max = K,lambda = cvout$opt_s)
  if(length(out$d)==1){
    Y_hat = (out$d) * out$u %*% t(out$v)
  }else{
    Y_hat = out$u %*%  diag(out$d) %*% t(out$v)
  }
  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
  return(RMSE)
}

SVD.wrapper = function(Y_data,K = 3){
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  gsvd = svd(Y,nu = 3,nv = 3)
  Y_hat = (gsvd$u[,1:K] %*% diag(gsvd$d[1:K]) %*% t(gsvd$v[,1:K]))
  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
  return(RMSE)
}

SSVD.wrapper = function(Y_data,K = 3){
  library(ssvd)
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  gssvd = ssvd::ssvd(Y,method = "method",r = K)
  Y_hat = (gssvd$u %*% diag(gssvd$d) %*% t(gssvd$v))
  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
  return(RMSE)
}

PN.wrapper = function(Y_data,K = 3){
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  library(ebnm)
  library(flashr2)
  data =  flashr2::flash_set_data(Y)
  g_flash = flashr2::flash_add_greedy(data,verbose=F,var_type = "constant",ebnm_fn = ebnm_pn, K = K)
  Y_hat = g_flash$EL %*% t(g_flash$EF)
  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
  return(RMSE)
}

flash.wrapper = function(Y_data, K = 3){
  # missindex is a matirx with 3 column here: i j x
  # Y has miss value already
  Y = Y_data$Y
  L_true = Y_data$L_true
  F_true = Y_data$F_true
  E = Y_data$E
  N = dim(Y)[1]
  P = dim(Y)[2]
  data =  flashr2::flash_set_data(Y)
  g_flash = flashr2::flash_add_greedy(data,verbose=F,var_type = "constant", K = K)
  Y_hat = g_flash$EL %*% t(g_flash$EF)
  RMSE = sqrt(mean(( Y - Y_hat - E )^2 ))/sqrt(mean(( Y - E )^2 ))
  return(RMSE)
}
```

creat `run.R` file as follows:

```{}
library(PMA)
library(flashr2)
library(ssvd)
library(softImpute)
Y_data = DataMaker()
RMSE = rep(NA,6)
RMSE[1] = PMA.wrapper(Y_data)
RMSE[2] = flash.wrapper(Y_data)
RMSE[3] = PN.wrapper(Y_data)
RMSE[4] = SVD.wrapper(Y_data)
RMSE[5] = SSVD.wrapper(Y_data)
RMSE[6] = softImpute.wrapper(Y_data)
saveRDS(RMSE, "./output.rds")
```

```{}
#!/bin/bash

#SBATCH --job-name=arrayJob
#SBATCH --output=./outlog/arrayJob_%A_%a.out
#SBATCH --error=./outlog/arrayJob_%A_%a.err
#SBATCH --array=1-100
#SBATCH --time=02:00:00
#SBATCH --partition=broadwl
#SBATCH --ntasks=1
#SBATCH --mem-per-cpu=2000

######################
# Begin work section #
######################

module load R

# Print this sub-job's task ID
mkdir test${SLURM_ARRAY_TASK_ID}
cd test${SLURM_ARRAY_TASK_ID}
Rscript --verbose ../run.R
```

name the above file as `Jobs.sbatch`

```
sbatch Jobs.sbatch
```

## get the result

```
T = 100
results = matrix(NA,ncol = 6, nrow = T)
for(i in 1:T){
  test_folder = paste("test", i, sep = "")
  out_file = "output.rds"
  file_name = file.path(test_folder,out_file)
  results[i,] = try(readRDS(file_name))
}

saveRDS(results,"./RES_rrmse.rds")
```

## plot the result

```{r}
plot_res = function(output,title = "data",legend_position = "none", methods_name){
  rmse = as.vector(output)
  N = dim(output)[1]
  methods = rep(methods_name, each = N)
  df = data.frame(RRMSE_diff = rmse, Method = methods )
  p<-ggplot(df, aes(x=Method, y=RRMSE_diff, color=Method)) +
  geom_boxplot()+
  # geom_violin()+
  ggtitle(title) +  theme_bw()+
    theme(legend.position= legend_position, legend.text=element_text(size=10),
          plot.title = element_text(size = 18, face = "bold"),
          axis.text.y = element_text(size =15), 
          axis.text.x = element_text(size =15,angle = 45, hjust = 1))
  p
}
library(ggplot2)
```

```{r}
sparse_res = readRDS("../data/simulation/rankthree/boxplot.rds")
colnames(sparse_res) = c("PMD.cv1","flash","flash.pn","SVD","SSVD","SI.cv")
sparse_diff = sparse_res - sparse_res[,2]
# p3 = plot_res(sparse_diff,"Difference from FLASH result (0% zeros)")
methods_name = colnames(sparse_diff)
p3 = plot_res(sparse_diff,title = "Rank 3 example",methods_name = methods_name)
```

```{r}
p3
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```

### 2018-01-25 analysis/templateforOCV.Rmd
Files commited by wflow_commit.

@@
 
-Here you need change the `\"addpath(\'~/HG/flash/data/OCVmissflashr2/Breastcancer\');` to your folder which contains `missvalue.m` file. this file name is `wrapper.R`
+This file name is `wrapper.R`
 
 ### matlab code
+
+In this folder
 
@@
 
-RMSE = sqrt(sum(SSE)/(100000));
+RMSE = sqrt(sum(SSE)/(settings.D*settings.N));
 RMSE;
@@
 
-you need change the path as well. this file name is 'missvalue.m'
+you need change the path `~/HG/flash/data/missingvalue/methods/NBSF/nsfa-master/` and `~/HG/flash/data/missingvalue/methods/NBSF/nsfa-master/utils/` which have the NBSFA matlab code.
 
@@
 ```{r}
+# this is the independent result
+
 library(ggplot2)
@@
   N = dim(output)[1]
-  # methods = rep(c("flash","NBSFA","PMD","softImpute"), each = N)
   methods = rep(x_label, each = N)
@@
     ggtitle(title) +  theme_bw()+
-    theme(legend.position= legend_position, legend.text=element_text(size=15), axis.text.y = element_text(size =6))
+    theme(legend.position= legend_position, legend.text=element_text(size=15),
+          plot.title = element_text(size = 15, face = "bold"),
+          axis.text.y = element_text(size =12.6),
+          axis.text.x = element_text(size =12.6,angle = 45, hjust = 1))
   p
 }
-x_label = c("flash","NBSFA","PMD","softImpute")
-PT_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/president_box.rds")
-pp = plot_res(PT_res,"Text data",x_label = x_label)
-DT_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/denoiseRtumor_box.rds")
-pd = plot_res(DT_res,"Tumor data",x_label = x_label)
-GZ_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/gtexzscore_box.rds")
-pg = plot_res(GZ_res,"GTEx data",x_label = x_label)
-BC_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/BreastCancer_box.rds")
-pb = plot_res(BC_res,"Breast Cancer data",x_label = x_label)
+
+setwd("~/HG/flash_workflow/analysis")
+
+##############  MovieLens data
+
 ML_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/ML100K_box.rds")
@@
 ML_res = matrix(as.numeric(ML_res),ncol = 4)
-pM = plot_res(ML_res,"Movie Lens data",x_label = x_label)
-```
+# c("flash","NBSFA","PMD","softImpute")
+ml_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/OCVNew20rep/movies100K/boxplot_cv.rds")
+# flashG_rmse,flashGwn_rmse,pmd_rmse,soft_rmse
+ml_res[1:20,] = sqrt((ml_res[1:20,])^2 * (943*1682)/(100000))
+ml_res = ml_res[1:20,]
+ML_res = ML_res[-c(2,13,17,21,29,37,62,76,77,93,95,100),]
+ML_res = ML_res[1:20,]
+# add PN result
+ML_pn_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/OCVNew20rep/movies100K/box_cv_pn.rds")
+# the result order would be flashG PMD PMDcv NSFA SF SFcv
+output = cbind(ML_pn_res[,1],ML_pn_res[,2],ML_res[,3],ml_res[,3],ML_res[,2],ML_res[,4],ml_res[,4])
 
-## plots
+x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
+plotM = plot_res(output,"MovieLens data",x_label = x_label)
 
-```{r}
-gridExtra::grid.arrange(pp,pd,pg,pb,pM, layout_matrix = rbind(c(1,NA,2),c(NA,5,NA),c(4,NA,3)))
-```
+############ Breast Cancer data
 
-## results for different tuning parameters
+BC_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/BreastCancer_box.rds")
+BC_res = BC_res[1:20,]
+# c("flash","NBSFA","PMD","softImpute")
+bc_res = readRDS("../data/missingvalue/testingcode/box_Breast.rds")
+# c("PN","flashG","flashGwn","flashB","nsfa","pmd","soft")
 
-```{r}
-pmd_c = sapply(seq(1,10),function(x){paste("PMD",x)})
-softImpute_c = sapply(seq(1,10),function(x){paste("SF",x)})
-x_label= c("flash","NBSFA",pmd_c,softImpute_c)
-PT_res = readRDS("../data/missingvalue/box_res_grids_sf_pmd/denoiseTumor_box.rds")
-pt = plot_res(PT_res,"Tumor data",x_label = x_label)
-pt
-PT_res = readRDS("../data/missingvalue/box_res_grids_sf_pmd/TEXT_prsdt_box.rds")
-pp = plot_res(PT_res,"Text data",x_label = x_label)
-pp
-```
+# the result order would be flashG PMD PMDcv NSFA SF SFcv
+output = cbind(bc_res[,2],bc_res[,1],BC_res[,3],bc_res[,6],BC_res[,2],BC_res[,4],bc_res[,7])
+x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
+plotB = plot_res(output,"Breast Cancer data",x_label = x_label)
 
-## use OCV choose the tuning parameters
+############ GTEx data
 
-labels | null check | backfitting | greedy | ebnm_ash | ebnm_pn 
------- | ---------- | ----------- | ------ | -------- | ------
-flashG | yes| |yes | yes |
-flashGwn | | | yes|yes |
-flashB | yes| yes| | yes|
-PN | yes | | yes| |yes
+GZ_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/gtexzscore_box.rds")
+GZ_res = GZ_res[1:20,]
+# c("flash","NBSFA","PMD","softImpute")
+gz_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/OCVNew20rep/gtexzscore/box_cv.rds")
+# flashG_rmse,nsfa_rmse,pmd_rmse,soft_rmse
+gz_pn_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/OCVNew20rep/gtexzscore/box_pn.rds")
+gz_pn_res[c(23,25,29,34),] = NA
+gz_pn_res = matrix(as.numeric(gz_pn_res[21:40,]),ncol = 2)
 
-### Breast Cancer data
+# the result order would be flashG PMD PMDcv NSFA SF SFcv
+output = cbind(gz_res[,1],gz_pn_res[,2],GZ_res[,3],gz_res[,3],GZ_res[,2],GZ_res[,4],gz_res[,4])
+x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
+plotG = plot_res(output,"GTEx data",x_label = x_label)
 
-we use 10 grids for `softImpute` and `PMD`
+############ Text data
 
-```{r}
-PB_res = readRDS("../data/missingvalue/testingcode/box_Breast.rds")
-x_label= c("PN","flashG","flashGwn","flashB","nsfa","pmd","soft")
-pb = plot_res(PB_res,"Breast Cancer data",x_label = x_label)
-pb
-```
+PT_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/president_box.rds")
+PT_res = PT_res[1:20,]
+# c("flash","NBSFA","PMD","softImpute")
+pt_res = readRDS("../data/missingvalue/testingcode/box_president.rds")
+# c("PN","flashG","flashGwn","flashB","nsfa","pmd","soft","zero")
 
-we take the 'flashB' away to compare the rest.
+# the result order would be flashG PMD PMDcv NSFA SF SFcv
+output = cbind(pt_res[,3],pt_res[,1],PT_res[,3],pt_res[,6],PT_res[,2],PT_res[,4],pt_res[,7])
+x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
+plotP = plot_res(output,"Text data",x_label = x_label)
 
-```{r}
-x_label= c("PN","flashG","flashGwn","nsfa","pmd","soft")
-pb = plot_res(PB_res[,-4],"Breast Cancer data",x_label = x_label)
-pb
-```
+############ Tumor data
 
-### denoiseR Tumor data
+DT_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/denoiseRtumor_box.rds")
+DT_res = DT_res[1:20,]
+# c("flash","NBSFA","PMD","softImpute")
+dt_res = readRDS("../data/missingvalue/testingcode/box_denoiseTumor.rds")
+# c("PN","flashG","flashGwn","flashB","nsfa","pmd","soft")
 
-we use 10 grids for `softImpute` and `PMD`
+# the result order would be flashG PMD PMDcv NSFA SF SFcv
+output = cbind(dt_res[,2],dt_res[,1],DT_res[,3],dt_res[,6],DT_res[,2],DT_res[,4],dt_res[,7])
+x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
+plotT = plot_res(output,"Tumor data",x_label = x_label)
 
-```{r}
-PT_res = readRDS("../data/missingvalue/testingcode/box_denoiseTumor.rds")
-x_label= c("PN","flashG","flashGwn","flashB","nsfa","pmd","soft")
-pt = plot_res(PT_res,"Tumor data",x_label = x_label)
-pt
-```
+# gridExtra::grid.arrange(plotP,plotT,plotG,plotB,plotM, layout_matrix = rbind(c(1,NA,2),c(NA,5,NA),c(4,NA,3)))
+gridExtra::grid.arrange(plotM,plotT,plotG,plotB,plotP, layout_matrix = rbind(c(1,2,3),c(4,5,NA) ))
 
-### text data
-
-in this data, we add `zero` as use zero values as imputation to compare with other methods.
-
-```{r}
-PT_res = readRDS("../data/missingvalue/testingcode/box_president.rds")
-x_label= c("PN","flashG","flashGwn","flashB","nsfa","pmd","soft","zero")
-pt = plot_res(PT_res,"Text data",x_label = x_label)
-pt
 ```

### 2018-01-25 analysis/OCVmissdataSetting.Rmd
Files commited by wflow_commit.

---
title: "Some useful information of OCV for missing data"
author: "wei wang"
date: 2018-01-25
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

## settings for the 5 data sets in our paper.

## other experiments during the test

```{r}
library(ggplot2)
plot_res = function(output,title = "data",legend_position = "none", x_label){
  rmse = as.vector(output)
  N = dim(output)[1]
  # methods = rep(c("flash","NBSFA","PMD","softImpute"), each = N)
  methods = rep(x_label, each = N)
  df = data.frame(RMSE = rmse, Method = methods )
  p<-ggplot(df, aes(x=Method, y=RMSE, color=Method)) +
  geom_boxplot()+
  # geom_violin()+
  ggtitle(title) +  theme_bw()+
    theme(legend.position= legend_position, legend.text=element_text(size=15), axis.text.y = element_text(size =6))
  p
}
x_label = c("flash","NBSFA","PMD","softImpute")
PT_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/president_box.rds")
pp = plot_res(PT_res,"Text data",x_label = x_label)
DT_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/denoiseRtumor_box.rds")
pd = plot_res(DT_res,"Tumor data",x_label = x_label)
GZ_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/gtexzscore_box.rds")
pg = plot_res(GZ_res,"GTEx data",x_label = x_label)
BC_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/BreastCancer_box.rds")
pb = plot_res(BC_res,"Breast Cancer data",x_label = x_label)
ML_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/ML100K_box.rds")
ML_res[c(2,13,17,21,29,37,62,76,77,93,95,100),] = NA
ML_res = matrix(as.numeric(ML_res),ncol = 4)
pM = plot_res(ML_res,"Movie Lens data",x_label = x_label)
```

## plots

```{r}
gridExtra::grid.arrange(pp,pd,pg,pb,pM, layout_matrix = rbind(c(1,NA,2),c(NA,5,NA),c(4,NA,3)))

```

## results for different tuning parameters

```{r}
pmd_c = sapply(seq(1,10),function(x){paste("PMD",x)})
softImpute_c = sapply(seq(1,10),function(x){paste("SF",x)})
x_label= c("flash","NBSFA",pmd_c,softImpute_c)
PT_res = readRDS("../data/missingvalue/box_res_grids_sf_pmd/denoiseTumor_box.rds")
pt = plot_res(PT_res,"Tumor data",x_label = x_label)
pt
PT_res = readRDS("../data/missingvalue/box_res_grids_sf_pmd/TEXT_prsdt_box.rds")
pp = plot_res(PT_res,"Text data",x_label = x_label)
pp
```

## use OCV choose the tuning parameters

labels | null check | backfitting | greedy | ebnm_ash | ebnm_pn 
------ | ---------- | ----------- | ------ | -------- | ------
flashG | yes| |yes | yes |
flashGwn | | | yes|yes |
flashB | yes| yes| | yes|
PN | yes | | yes| |yes

### Breast Cancer data

we use 10 grids for `softImpute` and `PMD`

```{r}
PB_res = readRDS("../data/missingvalue/testingcode/box_Breast.rds")
x_label= c("PN","flashG","flashGwn","flashB","nsfa","pmd","soft")
pb = plot_res(PB_res,"Breast Cancer data",x_label = x_label)
pb
```

we take the 'flashB' away to compare the rest.

```{r}
x_label= c("PN","flashG","flashGwn","nsfa","pmd","soft")
pb = plot_res(PB_res[,-4],"Breast Cancer data",x_label = x_label)
pb
```

### denoiseR Tumor data

we use 10 grids for `softImpute` and `PMD`

```{r}
PT_res = readRDS("../data/missingvalue/testingcode/box_denoiseTumor.rds")
x_label= c("PN","flashG","flashGwn","flashB","nsfa","pmd","soft")
pt = plot_res(PT_res,"Tumor data",x_label = x_label)
pt
```

### text data

in this data, we add `zero` as use zero values as imputation to compare with other methods.

```{r}
PT_res = readRDS("../data/missingvalue/testingcode/box_president.rds")
x_label= c("PN","flashG","flashGwn","flashB","nsfa","pmd","soft","zero")
pt = plot_res(PT_res,"Text data",x_label = x_label)
pt
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```

### 2018-01-25 analysis/OCVmissdataSetting.Rmd
Files commited by wflow_commit.

@@
 ## settings for the 5 data sets in our paper.
+
+### GTEx zsocre
+
+```{r}
+library(R.matlab)
+Y_centered = readMat("../data/output/missingdata/GTExZsocre/example.mat")
+Y = Y_centered$Ycentered
+```
+
+### MovieLens
+
+```{r}
+library(methods)
+library(R.matlab)
+library(Matrix)
+## run the code
+ml100K_data = readRDS("../data/output/missingdata/MovieLens/scale_data.rds")
+MLMatrix <- sparseMatrix(i = ml100K_data[,1],
+                         j = ml100K_data[,2],
+                         x = ml100K_data[,3],dims = c(943,1682))
+# turn this sparse matrix into matrix in r
+Y = as.matrix(MLMatrix)
+Y[which(Y == 0)] = NA
+# writeMat("~/HG/flash/data/OCVmissflashr2/ML100K_scaled/Ydata.mat", Y = Y)
+```
+
+### denoiseR tumor data
+
+```{r}
+library(R.matlab)
+Y_centered = readMat("../data/output/missingdata/DenoiseRtumor/example.mat")
+Y = Y_centered$Ycentered
+```
+
+### denoiseR text data
+
+```{r}
+library(R.matlab)
+## run the code
+Y_centered = readMat("../data/output/missingdata/DenoiseRtext/example.mat")
+Y = Y_centered$Yscaled
+```
+
+### Breast cancer data
+
+```{r}
+library(R.matlab)
+## run the code
+Y_centered = readMat("../data/output/missingdata/BreastCancer/example.mat")
+Y = Y_centered$Y
+# in the matlab package of NSF, the use the centered data by rows
+N = dim(Y)[1]
+P = dim(Y)[2]
+Y = Y - rowMeans(Y) %*% t(rep(1,P))
+```
 

### 2018-01-25 analysis/templateforOCV.Rmd
Files commited by wflow_commit.

@@
 
-ML_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/ML100K_box.rds")
+ML_res = readRDS("../data/output/missingdata/MovieLens/ML100K_box.rds")
 ML_res[c(2,13,17,21,29,37,62,76,77,93,95,100),] = NA
@@
 # c("flash","NBSFA","PMD","softImpute")
-ml_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/OCVNew20rep/movies100K/boxplot_cv.rds")
+ml_res = readRDS("../data/output/missingdata/MovieLens/boxplot_cv.rds")
 # flashG_rmse,flashGwn_rmse,pmd_rmse,soft_rmse
@@
 # add PN result
-ML_pn_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/OCVNew20rep/movies100K/box_cv_pn.rds")
+ML_pn_res = readRDS("../data/output/missingdata/MovieLens/box_cv_pn.rds")
 # the result order would be flashG PMD PMDcv NSFA SF SFcv
@@
 
-BC_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/BreastCancer_box.rds")
+BC_res = readRDS("../data/output/missingdata/BreastCancer/BreastCancer_box.rds")
 BC_res = BC_res[1:20,]
 # c("flash","NBSFA","PMD","softImpute")
-bc_res = readRDS("../data/missingvalue/testingcode/box_Breast.rds")
+bc_res = readRDS("../data/output/missingdata/BreastCancer/box_Breast.rds")
 # c("PN","flashG","flashGwn","flashB","nsfa","pmd","soft")
@@
 
-GZ_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/gtexzscore_box.rds")
+GZ_res = readRDS("../data/output/missingdata/GTExZsocre/gtexzscore_box.rds")
 GZ_res = GZ_res[1:20,]
 # c("flash","NBSFA","PMD","softImpute")
-gz_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/OCVNew20rep/gtexzscore/box_cv.rds")
+gz_res = readRDS("../data/output/missingdata/GTExZsocre/box_cv.rds")
 # flashG_rmse,nsfa_rmse,pmd_rmse,soft_rmse
-gz_pn_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/OCVNew20rep/gtexzscore/box_pn.rds")
+gz_pn_res = readRDS("../data/output/missingdata/GTExZsocre/box_pn.rds")
 gz_pn_res[c(23,25,29,34),] = NA
@@
 
-PT_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/president_box.rds")
+PT_res = readRDS("../data/output/missingdata/DenoiseRtext/president_box.rds")
 PT_res = PT_res[1:20,]
 # c("flash","NBSFA","PMD","softImpute")
-pt_res = readRDS("../data/missingvalue/testingcode/box_president.rds")
+pt_res = readRDS("../data/output/missingdata/DenoiseRtext/box_president.rds")
 # c("PN","flashG","flashGwn","flashB","nsfa","pmd","soft","zero")
@@
 
-DT_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/denoiseRtumor_box.rds")
+DT_res = readRDS("../data/output/missingdata/DenoiseRtumor/denoiseRtumor_box.rds")
 DT_res = DT_res[1:20,]
 # c("flash","NBSFA","PMD","softImpute")
-dt_res = readRDS("../data/missingvalue/testingcode/box_denoiseTumor.rds")
+dt_res = readRDS("../data/output/missingdata/DenoiseRtumor/box_denoiseTumor.rds")
 # c("PN","flashG","flashGwn","flashB","nsfa","pmd","soft")
@@
 
+## other information
+
+please click here to check other information [Missing data OCV settings](OCVmissdataSetting.html)
+
 ## Session information

### 2018-01-27 analysis/index.Rmd
Files commited by wflow_commit.
Updated NOTES.txt.

--- 2017-12-09 analysis/index.Rmd
@@
 
-### simulation studies
+## Simulation studies
 
@@
 
-Click on this [Questions on Algorithm](first-analysis.html) to see more details.
-
-### flash on the chipseq data from Kevin
-
-Click on this [three data sets on the 4500 rows chr1 var_col](sd_chr1_asinh.html) to see more details.
-
-[ns](draft.html)
-
-Click on this [load reslut of GTEx data](GTExresult.html) to see more details.
-
-Click on this [pairwise plot](pairwiseplotofflashgtexzscore.html) to see more details.
-
-### Issues
-
-Click on this [gap issue](ObjGap.html) to see more details. (test)
-
-Click on this [summary of the rank0 vs rank1](Obj_summary.html) to see a simple summary to the following two items.
-
-- Click on this [rank 0 vs rank 1](Objcheck_1.html) to see more details.
-- Click on this [rank 0 vs rank 1: fdr with penalty](Objcheck_2.html) to see more details.
-
-## flashr2
-
-This is a simple test [first try](flash_beta_test.html)
-
-Initialization issue: [softimpute as initialization](missingimputeinitial.html)
-
-Click here to find the [null check for flashr2](NULLcheck.html) and [densen nu case](nullcheck2.html)
-
-### apply flashr2 for the examples in FLASH paper
+### Apply flashr2 for the examples in FLASH paper
 
@@
 
+Click here to check the [RMSE for the rank three example in the paper](simulation3.html) and [RRMSE](MSErank3sim.html)
+
 #### GTEx Z-score
@@
 
+compare with PCA result[compare with SVD](PCAresult.html)
+
 #### missing data imputation
 
-click here to check the template for [missing data imputation with OCV](templateforOCV.html)
+click here to check the template for [missing data imputation with OCV](templateforOCV.html) and [CV template](CVtemplate.html)
 

### 2018-01-27 analysis/oldanalysis.Rmd
Files commited by wflow_commit.

---
title: "old experiments"
author: "wei wang"
date: 2018-01-27
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

Click on this [Questions on Algorithm](first-analysis.html) to see more details.

### flash on the chipseq data from Kevin

Click on this [three data sets on the 4500 rows chr1 var_col](sd_chr1_asinh.html) to see more details.

[ns](draft.html)

Click on this [load reslut of GTEx data](GTExresult.html) to see more details.

Click on this [pairwise plot](pairwiseplotofflashgtexzscore.html) to see more details.

### Issues

Click on this [gap issue](ObjGap.html) to see more details. (test)

Click on this [summary of the rank0 vs rank1](Obj_summary.html) to see a simple summary to the following two items.

- Click on this [rank 0 vs rank 1](Objcheck_1.html) to see more details.
- Click on this [rank 0 vs rank 1: fdr with penalty](Objcheck_2.html) to see more details.

## flashr2

This is a simple test [first try](flash_beta_test.html)

Initialization issue: [softimpute as initialization](missingimputeinitial.html)

Click here to find the [null check for flashr2](NULLcheck.html) and [densen nu case](nullcheck2.html)

### different options for EBNM

#### simulation examples in our paper

Click here to check the [simulated examples](compare_ash_ebmn.html)

Click here to check the [missing data examples (first simple comparison)](compare_ash_pn_miss.html)

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```

### 2018-01-27 analysis/MSErank1sim.Rmd
Files commited by wflow_commit.

@@
 # Data = datamaker(N,P,L_pi,L_se,0.1,c(1),c(1),1,sqrt(1))  sparse case
-# Data = datamaker(N,P,L_pi,L_se,1.0,c(1),c(1),1,sqrt(1))  dense case
+# Data = datamaker(N,P,L_pi,L_se,1.0,c(1),c(1),1,sqrt(25))  dense case
 RMSE = rep(NA,6)
@@
 
-## plot the result
+## get the result
 
@@
 ```{r}
-plot_res = function(output,title = "data",legend_position = "none", methods_name){
+plot_res = function(output,title = "data",legend_position = "none", methods_name,myColors){
   rmse = as.vector(output)
@@
   # geom_violin()+
-  ggtitle(title) +  theme_bw()+
+  ggtitle(title) +  theme_bw()+ scale_color_manual(values=myColors)+
     theme(legend.position= legend_position, legend.text=element_text(size=10),
@@
 methods_name = colnames(sparse_diff)
-p1 = plot_res(sparse_diff,title = "Difference from FLASH result (90% zeros)",methods_name = methods_name)
+fill_color = c("red","yellow3", "springgreen", "springgreen3", "cyan","cyan3", "deepskyblue","violet","purple")
+p1 = plot_res(sparse_diff,title = "Difference from FLASH result (90% zeros)",
+              methods_name = methods_name, myColors = fill_color)
 methods_name = colnames(sparse_diff)[c(1,2,3,4,5,7,9)]
-p1_cut = plot_res(sparse_diff[-22,c(1,2,3,4,5,7,9)],title = "Difference from FLASH result (90% zeros)",methods_name = methods_name)
+p1_cut = plot_res(sparse_diff[-22,c(1,2,3,4,5,7,9)],title = "Difference from FLASH result (90% zeros)",methods_name = methods_name,myColors = fill_color[c(1,2,3,4,5,7,9)])
 ```
@@
 methods_name = colnames(sparse_diff)
-p2 = plot_res(sparse_diff,title = "Difference from FLASH result (30% zeros)",methods_name = methods_name)
+p2 = plot_res(sparse_diff,title = "Difference from FLASH result (30% zeros)",methods_name = methods_name,myColors = fill_color)
 methods_name = colnames(sparse_diff)[c(1,2,3,4,5,7,9)]
-p2_cut = plot_res(sparse_diff[,c(1,2,3,4,5,7,9)],title = "Difference from FLASH result (30% zeros)",methods_name = methods_name)
+p2_cut = plot_res(sparse_diff[,c(1,2,3,4,5,7,9)],title = "Difference from FLASH result (30% zeros)",methods_name = methods_name,myColors = fill_color[c(1,2,3,4,5,7,9)])
 ```
@@
 methods_name = colnames(sparse_diff)
-p3 = plot_res(sparse_diff,title = "Difference from FLASH result (0% zeros)",methods_name = methods_name)
+p3 = plot_res(sparse_diff,title = "Difference from FLASH result (0% zeros)",methods_name = methods_name,myColors = fill_color )
 methods_name = colnames(sparse_diff)[c(1,2,3,4,5,7,9)]
-p3_cut = plot_res(sparse_diff[,c(1,2,3,4,5,7,9)],title = "Difference from FLASH result (0% zeros)",methods_name = methods_name)
+p3_cut = plot_res(sparse_diff[,c(1,2,3,4,5,7,9)],title = "Difference from FLASH result (0% zeros)",methods_name = methods_name,myColors = fill_color[c(1,2,3,4,5,7,9)])
 ```

### 2018-01-27 analysis/MSErank3sim.Rmd
Files commited by wflow_commit.

@@
 ```{r}
-sparse_res = readRDS("../data/simulation/rankthree/boxplot.rds")
+sparse_res = readRDS("../data/output/simulation/rankthree/boxplot.rds")
 colnames(sparse_res) = c("PMD.cv1","flash","flash.pn","SVD","SSVD","SI.cv")
@@
 methods_name = colnames(sparse_diff)
-p3 = plot_res(sparse_diff,title = "Rank 3 example",methods_name = methods_name)
+fill_color =  c("red","yellow3", "springgreen", "springgreen3", "cyan","cyan3", "deepskyblue","violet","purple")
+p3 = plot_res(sparse_diff,title = "Rank 3 example",methods_name = methods_name,
+              myColors = fill_color[c(1,2,3,7,8,9)])
 ```

### 2018-01-27 analysis/softImputelambda.Rmd
Files commited by wflow_commit.

--- 2018-01-22 analysis/templateforOCV.Rmd
@@
 ---
-title: "apply flashr2 with OCV"
-author: "wei wang"
-date: 09-30
+title: "the choice of softImpte lambda"
+author: "Wei Wang"
+date: 2018-01-27
 output: 
@@
 
-## preparation
+## Document
 
-1. creat a folder
+Based on the document of the authors ![https://cran.r-project.org/web/packages/softImpute/softImpute.pdf], we should choose lambda ($\lambda$) as "Ideally lambda should be chosen so that the solution reached has rank slightly
+less than rank.max."
 
-```
-mkdir Breastcancer
-```
-
-2. put your data into this folder
-
-here we use the file 'example.mat'. the reason we use `.mat` file is because there is a matlab package in our comparision.
-
-3. creat the wrapper function in R and matlab
-
-### wrapper function for OCV
+## Brease Cancer data in flash paper
 
@@
   return(list(opt_s = opt_s, output = CVRMSE))
-}
-
-PMA.wrapper = function(Y,ngrids = 10,K,fold = 10){
-  library(PMA)
-  N = dim(Y)[1]
-  P = dim(Y)[2]
-  c_s = seq(0.3,0.8,len=ngrids)
-  cvout = CVPMD_softImpute(Y,c_s,K ,fold , method = "PMD")
-  res_log = capture.output({out = PMD(Y,sumabsu = NULL, sumabsv = NULL, sumabs = cvout$opt_s ,K = K)})
-  return(list(d = out$d, u = out$u, v = out$v))
 }
@@
 
-NSF_OCV = function(OCVindex){
-  if(file.exists("NBSFAout")){
-    unlink("NBSFAout", recursive= T)
-  }
-  system("mkdir NBSFAout")
-  # write the missing index
-
-  writeMat("./NBSFAout/OCVindex.mat", OCVindex = OCVindex)
-
-  system("matlab -nosplash -nodesktop -r \"addpath(\'../\'); run(\'missvalue.m\');exit;\" ")
-  g_nsfa = readMat("./NBSFAout/NSFAresult.mat")
-  nsfa_rmse = as.vector(g_nsfa$RMSE)
-  return(nsfa_rmse)
-}
-
 ```
 
-Here you need change the `\"addpath(\'~/HG/flash/data/OCVmissflashr2/Breastcancer\');` to your folder which contains `missvalue.m` file. this file name is `wrapper.R`
+read the data
 
-### matlab code
-
-```{}
-ddpath('~/HG/flash/data/missingvalue/methods/NBSF/nsfa-master/');
-addpath('~/HG/flash/data/missingvalue/methods/NBSF/nsfa-master/utils/');
-
-% you need to change the path and the data file (not centered)
-load ../Ydata.mat;
-Ycentered=Y;
-settings=defaultsettings();
-[settings.D,settings.N]=size(Ycentered);
-settings.iterations=100;
-% now the missing is decide by the this is the missing index.(or you can upload the missing matrix)
-
-file_path = fullfile(pwd,'NBSFAout/OCVindex.mat');
-load(file_path);
-
-% missindex = ones(settings.D,settings.N); this is not missing index this
-% is observed index.
-missindex = (~isnan(Y));
-k_fold = 10;
-foldindex = zeros(k_fold,k_fold,2);
-for k = 1:2
-    for j = 1:k_fold
-        for i = 1:k_fold
-            foldindex(i,j,k) = OCVindex.foldindex(i + k_fold*(j - 1) + k_fold * k_fold * (k - 1));
-        end
-    end
-end
-
-SSE = zeros(1,k_fold);
-for i = 1:k_fold
-    mvmask = missindex;
-    for j = 1:k_fold
-        mvmask(OCVindex.rowindex(:,foldindex(j,i,1)),OCVindex.colindex(:,foldindex(j,i,2))) = 0;
-    end
-    mvmask(:,sum(mvmask) == 0) = (missindex(:,sum(mvmask) == 0));
-    initialsample=init_nsfa(settings);
-    [finalsample,resultstable]=nsfa(Ycentered,mvmask,initialsample,settings);
-    Y_hat = (finalsample.G *  finalsample.X);
-    for j = 1:k_fold
-        SSE(i) = SSE(i) + ...
-            sum(nansum((Ycentered(OCVindex.rowindex(:,foldindex(j,i,1)),OCVindex.colindex(:,foldindex(j,i,2)))- ...
-            Y_hat(OCVindex.rowindex(:,foldindex(j,i,1)),OCVindex.colindex(:,foldindex(j,i,2)))).^2));
-    end
-
-end
-
-RMSE = sqrt(sum(SSE)/(100000));
-RMSE;
-
-save_path = fullfile(pwd,'NBSFAout/NSFAresult');
-save(save_path,'RMSE');
-```
-
-you need change the path as well. this file name is 'missvalue.m'
-
-### creat the run.R file to run the code
-
-In this case, we have problem in running `flashr2` on RCC. so we just try other methods first.
-
-```{}
-source("../wrapper.R")
+```{r}
 library(R.matlab)
 ## run the code
-Y_centered = readMat("../example.mat")
+Y_centered = readMat("../data/output/missingdata/BreastCancer/example.mat")
 Y = Y_centered$Y
@@
 Y = Y - rowMeans(Y) %*% t(rep(1,P))
+```
+
+first we try to get the lambda based on the authors suggestion.
+
+```{r}
 OCVindex = OCV_index(Y,k_fold = 10)
@@
 
-PN_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "PN",Kmax = 40)
-flashG_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "flash",Kmax = 40)
-flashGwn_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "flash_wn",Kmax = 40)
-flashB_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "flash_bf",Kmax = 40)
-pmd_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "PMD",Kmax = 40)
-soft_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "softImpute",Kmax = 40)
-nsfa_rmse = NSF_OCV(OCVindex)
+colindex = OCVindex$colindex
+rowindex = OCVindex$rowindex
+foldindex = OCVindex$foldindex
+missing = OCVdata
+Y_data = missing[1,,]
 
-result = c(PN_rmse,flashG_rmse,flashGwn_rmse,flashB_rmse,nsfa_rmse,pmd_rmse,soft_rmse)
-saveRDS(result, "./NBSFAout/output.rds")
+out = softImpute(Y_data, rank.max = 40,lambda = 80)
+length(out$d)
+out = softImpute(Y_data, rank.max = 40,lambda = 30)
+length(out$d)
+out = softImpute(Y_data, rank.max = 40,lambda = 20)
+length(out$d)
+out = softImpute(Y_data, rank.max = 40,lambda = 10)
+length(out$d)
+out = softImpute(Y_data, rank.max = 40,lambda = 9)
+length(out$d)
 ```
 
-this file name is `runOCV.R`
+so it seems 10 or 11 should be a good choice which "Ideally lambda should be chosen so that the solution reached has rank slightly less than rank.max"
 
-### creat folder to track the result
+## OCV choice
 
-```
-mkdir outlog
+for the grids we choose in our paper, we use 0-100 with ten grids
+
+```{r}
+library(softImpute)
+N = dim(Y_data)[1]
+P = dim(Y_data)[2]
+ngrids = 10
+fold = 10
+K = 40
+c_s = seq(0,100,len=ngrids)
+cvout = CVPMD_softImpute(Y_data,c_s,K ,fold , method = "softImpute")
+cvout
 ```
 
-### creat .sbatch file
+the result fit the authors' suggestion well (we choose $\lambda = 11$).
 
-```
-#!/bin/bash
+But can we try smaller grids with OCV
 
-#SBATCH --job-name=arrayJob
-#SBATCH --output=./outlog/arrayJob_%A_%a.out
-#SBATCH --error=./outlog/arrayJob_%A_%a.err
-#SBATCH --array=1-20
-#SBATCH --time=26:00:00
-#SBATCH --partition=mstephens
-#SBATCH --ntasks=1
-#SBATCH --mem-per-cpu=8000
-
-######################
-# Begin work section #
-######################
-
-# Print this sub-job's task ID
-module load R
-module load matlab
-
-mkdir test${SLURM_ARRAY_TASK_ID}
-cd test${SLURM_ARRAY_TASK_ID}
-Rscript --verbose ../runOCV.R
+```{r}
+library(softImpute)
+N = dim(Y_data)[1]
+P = dim(Y_data)[2]
+ngrids = 10
+fold = 10
+K = 40
+c_s = seq(0,15,len=ngrids)
+cvout = CVPMD_softImpute(Y_data,c_s,K ,fold , method = "softImpute")
+cvout
 ```
 
-this file is name as `OCV.sbatch`
+At this time, we can see that we choose $\lambda = 5$ this time.
 
-## run OCV
+## check the result
 
-now in your folder, you should have:
-
-```
-[weidong@midway2-login1 Breastcancer]$ ls
-OCV.sbatch  example.mat  missvalue.m  outlog  runOCV.R  wrapper.R
+```{r}
+out = softImpute(Y_data, rank.max = 40,lambda = 10)
+miss_hat = out$u %*% diag(out$d) %*% t(out$v)
+SSE = 0
+for(j in 1:fold){
+      SSE = SSE + sum((Y[rowindex[,foldindex[j,1,1]],colindex[,foldindex[j,1,2]]] -
+                               miss_hat[rowindex[,foldindex[j,1,1]],colindex[,foldindex[j,1,2]]])^2,na.rm = TRUE)
+}
+sqrt(SSE/sum(is.na(Y_data)))
 ```
 
-```
-sbatch OCV.sbatch
+```{r}
+out = softImpute(Y_data, rank.max = 40,lambda = 5)
+miss_hat = out$u %*% diag(out$d) %*% t(out$v)
+SSE = 0
+for(j in 1:fold){
+      SSE = SSE + sum((Y[rowindex[,foldindex[j,1,1]],colindex[,foldindex[j,1,2]]] -
+                               miss_hat[rowindex[,foldindex[j,1,1]],colindex[,foldindex[j,1,2]]])^2,na.rm = TRUE)
+}
+sqrt(SSE/sum(is.na(Y_data)))
 ```
 
-## to get the result
-
-```{}
-T = 20
-results = matrix(NA,ncol = 7, nrow = T)
-for(i in 1:T){
-  test_folder = paste("test", i, sep = "")
-  out_folder = "NBSFAout"
-  out_file = "output.rds"
-  file_name = file.path(test_folder,out_folder,out_file)
-  results[i,] = try(readRDS(file_name))
-}
-
-saveRDS(results,"./boxplot.rds")
-```
-
-## plot the result
-
-```{r}
-library(ggplot2)
-plot_res = function(output,title = "data",legend_position = "none", x_label){
-  rmse = as.vector(output)
-  N = dim(output)[1]
-  # methods = rep(c("flash","NBSFA","PMD","softImpute"), each = N)
-  methods = rep(x_label, each = N)
-  df = data.frame(RMSE = rmse, Method = methods )
-  p<-ggplot(df, aes(x=Method, y=RMSE, color=Method)) +
-  geom_boxplot()+
-  # geom_violin()+
-  ggtitle(title) +  theme_bw()+
-    theme(legend.position= legend_position, legend.text=element_text(size=15), axis.text.y = element_text(size =6))
-  p
-}
-x_label = c("flash","NBSFA","PMD","softImpute")
-PT_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/president_box.rds")
-pp = plot_res(PT_res,"Text data",x_label = x_label)
-DT_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/denoiseRtumor_box.rds")
-pd = plot_res(DT_res,"Tumor data",x_label = x_label)
-GZ_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/gtexzscore_box.rds")
-pg = plot_res(GZ_res,"GTEx data",x_label = x_label)
-BC_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/BreastCancer_box.rds")
-pb = plot_res(BC_res,"Breast Cancer data",x_label = x_label)
-ML_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/ML100K_box.rds")
-ML_res[c(2,13,17,21,29,37,62,76,77,93,95,100),] = NA
-ML_res = matrix(as.numeric(ML_res),ncol = 4)
-pM = plot_res(ML_res,"Movie Lens data",x_label = x_label)
-```
-
-## plots
-
-```{r}
-gridExtra::grid.arrange(pp,pd,pg,pb,pM, layout_matrix = rbind(c(1,NA,2),c(NA,5,NA),c(4,NA,3)))
-```
-
-## results for different tuning parameters
-
-```{r}
-pmd_c = sapply(seq(1,10),function(x){paste("PMD",x)})
-softImpute_c = sapply(seq(1,10),function(x){paste("SF",x)})
-x_label= c("flash","NBSFA",pmd_c,softImpute_c)
-PT_res = readRDS("../data/missingvalue/box_res_grids_sf_pmd/denoiseTumor_box.rds")
-pt = plot_res(PT_res,"Tumor data",x_label = x_label)
-pt
-PT_res = readRDS("../data/missingvalue/box_res_grids_sf_pmd/TEXT_prsdt_box.rds")
-pp = plot_res(PT_res,"Text data",x_label = x_label)
-pp
-```
-
-## use OCV choose the tuning parameters
-
-labels | null check | backfitting | greedy | ebnm_ash | ebnm_pn 
------- | ---------- | ----------- | ------ | -------- | ------
-flashG | yes| |yes | yes |
-flashGwn | | | yes|yes |
-flashB | yes| yes| | yes|
-PN | yes | | yes| |yes
-
-### Breast Cancer data
-
-we use 10 grids for `softImpute` and `PMD`
-
-```{r}
-PB_res = readRDS("../data/missingvalue/testingcode/box_Breast.rds")
-x_label= c("PN","flashG","flashGwn","flashB","nsfa","pmd","soft")
-pb = plot_res(PB_res,"Breast Cancer data",x_label = x_label)
-pb
-```
-
-we take the 'flashB' away to compare the rest.
-
-```{r}
-x_label= c("PN","flashG","flashGwn","nsfa","pmd","soft")
-pb = plot_res(PB_res[,-4],"Breast Cancer data",x_label = x_label)
-pb
-```
-
-### denoiseR Tumor data
-
-we use 10 grids for `softImpute` and `PMD`
-
-```{r}
-PT_res = readRDS("../data/missingvalue/testingcode/box_denoiseTumor.rds")
-x_label= c("PN","flashG","flashGwn","flashB","nsfa","pmd","soft")
-pt = plot_res(PT_res,"Tumor data",x_label = x_label)
-pt
-```
-
-### text data
-
-in this data, we add `zero` as use zero values as imputation to compare with other methods.
-
-```{r}
-PT_res = readRDS("../data/missingvalue/testingcode/box_president.rds")
-x_label= c("PN","flashG","flashGwn","flashB","nsfa","pmd","soft","zero")
-pt = plot_res(PT_res,"Text data",x_label = x_label)
-pt
-```
+we can see that the result is slightly improved.
 

### 2018-01-28 analysis/GTEXzscoreTemplate.Rmd
Files commited by wflow_commit.

@@
 
-```{}
+```{r}
 b_flash = readRDS("../data/GTExdata/res_flashr2/bflashvarcol.rds")
@@
 data_L$tissue = factor(data_L$tissue,levels = 1:44 ,labels = as.character(colordata[,1]) )
-data_L$loading = factor(data_L$loading,levels = 1:14 ,labels = paste("loading",1:14,"; pve:", pve[1:14]))
+data_L$loading = factor(data_L$loading,levels = 1:14 ,labels = paste("Factor",1:14,"; pve:", pve[1:14]))
 
@@
   theme_grey()+
-  theme(legend.position="right", legend.text=element_text(size=5), axis.text.y = element_text(size = 5)) + 
-  labs(title = "GTEx data", y = "loading values" ,x = "tissues", fill="tissue") +
+  theme(legend.position="right", legend.text=element_text(size=9), axis.text.y = element_text(size = 5)) + 
+  labs(title = "GTEx data", y = "factor values" ,x = "tissues", fill="tissue") +
   facet_wrap(~loading, ncol = 2, scales = "free_y") +
-  guides(fill = guide_legend(ncol = 1, keyheight = 0.5, keywidth = 0.5))
-  ggsave("flashr2GTEx1.pdf", width = 8, height = 11)
+  guides(fill = guide_legend(ncol = 1, keyheight = 0.8, keywidth = 0.3))
+  #ggsave("flashr2GTEx1.pdf", width = 8, height = 11)
 
@@
 data_L$tissue = factor(data_L$tissue,levels = 1:44 ,labels = as.character(colordata[,1]) )
-data_L$loading = factor(data_L$loading,levels = 1:12 ,labels = paste("loading",15:26,"; pve:", pve[15:26]))
+data_L$loading = factor(data_L$loading,levels = 1:12 ,labels = paste("Factor",15:26,"; pve:", pve[15:26]))
 
@@
   theme_grey()+
-  theme(legend.position="right", legend.text=element_text(size=5), axis.text.y = element_text(size = 5)) + 
-  labs(title = "GTEx data", y = "loading values" ,x = "tissues", fill="tissue") +
+  theme(legend.position="right", legend.text=element_text(size=9), axis.text.y = element_text(size = 5)) + 
+  labs(title = "GTEx data", y = "factor values" ,x = "tissues", fill="tissue") +
   facet_wrap(~loading, ncol = 2, scales = "free_y") +
-  guides(fill = guide_legend(ncol = 1, keyheight = 0.5, keywidth = 0.5))
-  ggsave("flashr2GTEx2.pdf", width = 8, height = 10)
+  guides(fill = guide_legend(ncol = 1, keyheight = 0.8, keywidth = 0.3))
+  #ggsave("flashr2GTEx2.pdf", width = 8, height = 10)
 ```

### 2018-01-28 analysis/simulation3.Rmd
Files commited by wflow_commit.

@@
 ```{r}
-gridExtra::grid.arrange(p0,p5,p6,p4,p1,p2,ncol = 2)
+gridExtra::grid.arrange(p0,p2,p1,p6,p5,p4,ncol = 2)
+```
+
+## average of 100 simulations
+
+```{r}
+HMatrix = readRDS("../data/simulation/rankthree/HeatMap.rds")
+T = 100
+AVG_HM = matrix(0,nrow = 36000,ncol = 7)
+for(i in 1:T){
+  AVG_HM = AVG_HM + HMatrix[i,,]
+}
+AVG_HM = AVG_HM/100
+```
+
+```{r}
+LF_1 = matrix(AVG_HM[,1],ncol = 240, nrow = 150)
+LF_2 = matrix(AVG_HM[,2],ncol = 240, nrow = 150)
+LF_3 = matrix(AVG_HM[,3],ncol = 240, nrow = 150)
+LF_4 = matrix(AVG_HM[,4],ncol = 240, nrow = 150)
+LF_5 = matrix(AVG_HM[,5],ncol = 240, nrow = 150)
+LF_6 = matrix(AVG_HM[,6],ncol = 240, nrow = 150)
+LF_true = matrix(AVG_HM[,7],ncol = 240, nrow = 150)
+library(reshape2)
+melted_cormat_1 <- melt(LF_1, na.rm = TRUE)
+melted_cormat_2 <- melt(LF_2, na.rm = TRUE)
+melted_cormat_3 <- melt(LF_3, na.rm = TRUE)
+melted_cormat_4 <- melt(LF_4, na.rm = TRUE)
+melted_cormat_5 <- melt(LF_5, na.rm = TRUE)
+melted_cormat_6 <- melt(LF_6, na.rm = TRUE)
+melted_cormat_true <- melt(LF_true, na.rm = TRUE)
+# Heatmap
+library(ggplot2)
+plot_HP = function(title_name = "truth", melted_cormat){
+  p1 = ggplot(data = melted_cormat, aes(Var2, Var1, fill = value))+
+   geom_tile(color = "white")+
+    scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
+     midpoint = 0, limit = c(-0.25,0.25), space = "Lab") + labs(title = title_name, y = "samples", x = "variables") +
+    theme_minimal() + theme(legend.position="none") 
+  p1
+}
+
+p0 = plot_HP(title_name = "Truth", melted_cormat = melted_cormat_true)
+p1 = plot_HP(title_name = "PMD.cv1", melted_cormat = melted_cormat_1)
+p2 = plot_HP(title_name = "flash", melted_cormat = melted_cormat_2)
+p3 = plot_HP(title_name = "flas_pn", melted_cormat = melted_cormat_3)
+p4 = plot_HP(title_name = "SVD", melted_cormat = melted_cormat_4)
+p5 = plot_HP(title_name = "SSVD", melted_cormat = melted_cormat_5)
+p6 = plot_HP(title_name = "SI.cv", melted_cormat = melted_cormat_6)
+
+```
+
+```{r}
+gridExtra::grid.arrange(p0,p2,p3,p1,p6,p5,p4,layout_matrix = rbind(c(NA,1,NA),c(2,3,4),c(5,6,7)))
 ```

### 2018-01-30 analysis/gtex_si.rmd
make the legend text size large

@@
     theme_grey()+
-    theme(legend.position="right", legend.text=element_text(size=5), axis.text.y = element_text(size = 5)) + 
+    theme(legend.position="right", legend.text=element_text(size=9), axis.text.y = element_text(size = 5)) + 
     labs(title = "GTEx data", y = "loading values" ,x = "tissues", fill="tissue") +
     facet_wrap(~loading, ncol = 2, scales = "free_y") +
-    guides(fill = guide_legend(ncol = 1, keyheight = 0.5, keywidth = 0.5))
+    guides(fill = guide_legend(ncol = 1, keyheight = 0.8, keywidth = 0.3))
 }

### 2018-02-03 analysis/gtex_si.rmd
wflow_publish("gtex_si.rmd")

@@
 ---
-title: "Untitled"
+title: "Run SoftImpute on GTEX data"
 author: "Matthew Stephens"
@@
 
+Load the data and run:
+
 ```{r}
 set.seed(1)
-Zscores = readRDS("~/Downloads/MatrixEQTLSumStats.Portable.Z.rds")
+Zscores = readRDS("../data/MatrixEQTLSumStats.Portable.Z.rds")
 Ztop = Zscores$test.z
@@
 
+Now plotting:
 ```{r}
@@
 plot_GTEXres(z.s$v[,1:8],pve.si[1:8])
-plot_GTEXres(z.s$v[,9:16],pve.si[9:16],9)
+plot_GTEXres(z.s$v[,9:16],pve.si[9:16],8)
 

### 2018-02-11 analysis/RsessionInfo.Rmd
Files commited by wflow_commit.

--- 2018-01-27 analysis/softImputelambda.Rmd
@@
 ---
-title: "the choice of softImpte lambda"
+title: "R Session Info"
 author: "Wei Wang"
-date: 2018-01-27
-output: 
-  html_document:
-    code_folding: hide
+date: 
+output: html_document
 ---
@@
 
-## Document
-
-Based on the document of the authors ![https://cran.r-project.org/web/packages/softImpute/softImpute.pdf], we should choose lambda ($\lambda$) as "Ideally lambda should be chosen so that the solution reached has rank slightly
-less than rank.max."
-
-## Brease Cancer data in flash paper
-
 ```{r}
-CVPMD_softImpute=function(Y,c_s,K,fold = 10, method = "PMD"){
-  N = dim(Y)[1]
-  P = dim(Y)[2]
-  colindex = matrix(sample(P,P),ncol = fold)
-  rowindex = matrix(sample(N,N),ncol = fold)
-
-  missing= array(0,dim = c(fold,N,P))
-  foldindex = array(0,dim = c(fold,fold,2))
-  for(i in 1:fold){
-    for(j in 1:fold){
-      foldindex[i,j,1] = i
-      foldindex[i,j,2] = (i+j) %% fold
-    }
-  }
-  foldindex[which(foldindex == 0)] = fold
-  for(i in 1:fold){
-    missing[i, , ] = Y
-    for(j in 1:fold){
-      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
-    }
-    missing[i,,which(colSums(missing[i,,],na.rm = T) ==0)] = Y[,which(colSums(missing[i,,],na.rm = T) ==0)]
-  }
-  # c_s is the candicate of shrinkage parameter
-  n_s = length(c_s)
-  # rmse for each grids
-  CVRMSE = rep(0,n_s)
-  minrmse = Inf
-  opt_s = 0
-  # for each candidate, we run it N_sim times
-  for(t_s in 1:n_s){
-    # for each grid
-    # each time we set the rmse to zeros
-    rmse = rep(0,fold)
-    for(i in 1:fold){
-      if(method == "PMD"){
-        res_log = capture.output({out = PMD(missing[i,,], sumabs = c_s[t_s], sumabsv = NULL, sumabsu = NULL,K = K)})
-      }else{
-        out = softImpute(missing[i,,], rank.max = K,lambda = c_s[t_s])
-      }
-      if(length(out$d)==1){
-        misshat = (out$d) * out$u %*% t(out$v)
-      }else{
-        misshat = out$u %*%  diag(out$d) %*% t(out$v)
-      }
-      for(j in 1:fold){
-        # for each fold j
-        rmse[i] = rmse[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] -
-                                   misshat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2,na.rm = TRUE)
-      }
-    } #get the result for one run
-    CVRMSE[t_s] = CVRMSE[t_s] + sqrt(sum(rmse)/(N*P))
-    if(CVRMSE[t_s] < minrmse){
-      minrmse = CVRMSE[t_s]
-      opt_s = c_s[t_s]
-    }
-  }
-  return(list(opt_s = opt_s, output = CVRMSE))
-}
-
-softImpute.wrapper = function(Y,ngrids = 10,K,fold = 10){
+library(flashr)
+library(ssvd)
+library(methods)
+library(PMA)
+library(R.matlab)
+library(Matrix)
 library(softImpute)
-  N = dim(Y)[1]
-  P = dim(Y)[2]
-  c_s = seq(0,100,len=ngrids)
-  cvout = CVPMD_softImpute(Y,c_s,K ,fold , method = "softImpute")
-  out = softImpute(Y, rank.max = K,lambda = cvout$opt_s)
-  return(list(d = out$d, u = out$u, v = out$v))
-}
-
-OCV_index=function(Y,k_fold = 5){
-  N = dim(Y)[1]
-  P = dim(Y)[2]
-  colindex = matrix(sample(P,P),ncol = k_fold)
-  rowindex = matrix(sample(N,N),ncol = k_fold)
-
-  foldindex = array(0,dim = c(k_fold,k_fold,2))
-  for(i in 1:k_fold){
-    for(j in 1:k_fold){
-      foldindex[i,j,1] = i
-      foldindex[i,j,2] = (i+j) %% k_fold
-    }
-  }
-  foldindex[which(foldindex == 0)] = k_fold
-
-  return(list(foldindex = foldindex, rowindex = rowindex, colindex = colindex))
-}
-# OCVindex = OCV_index(Y,k_fold = 5)
-
-OCV_data = function(Y,OCVindex,k_fold = 5){
-  N = dim(Y)[1]
-  P = dim(Y)[2]
-  colindex = OCVindex$colindex
-  rowindex = OCVindex$rowindex
-  foldindex = OCVindex$foldindex
-  missing= array(0,dim = c(k_fold,N,P))
-  for(i in 1:k_fold){
-    missing[i, , ] = Y
-    for(j in 1:k_fold){
-      missing[i,rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] = NA
-    }
-    missing[i,,which(colSums(missing[i,,],na.rm = T) ==0)] = Y[,which(colSums(missing[i,,],na.rm = T) ==0)]
-  }
-  return(missing)
-}
-# OCVdata = OCV_data(Y,OCVindex,k_fold = 5)
-
-OCV_SSE = function(Y,OCVindex,OCVdata,k_fold = 5,method = "flash",Kmax = 50){
-  colindex = OCVindex$colindex
-  rowindex = OCVindex$rowindex
-  foldindex = OCVindex$foldindex
-  missing = OCVdata
-  SSE = rep(0,k_fold)
-  for(i in 1:k_fold){
-    miss_hat = call_method(missing[i,,], method = method, Kmax = Kmax)
-    for(j in 1:k_fold){
-      SSE[i] = SSE[i] + sum((Y[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]] -
-                               miss_hat[rowindex[,foldindex[j,i,1]],colindex[,foldindex[j,i,2]]])^2,na.rm = TRUE)
-    }
-  }
-  RMSE = sqrt(sum(SSE)/(dim(Y)[1] * dim(Y)[2]))
-  return(RMSE)
-}
-
-call_method = function(Y_data,method,Kmax = 50){
-  if(method == "flash"){
-    data = flashr2::flash_set_data(Y_data)
-    f_greedy = flashr2::flash_add_greedy(data,K=Kmax)
-    Y_hat = f_greedy$EL %*% t(f_greedy$EF)
-  }else if(method == "flash_wn"){
-    data = flashr2::flash_set_data(Y_data)
-    f_greedy = flashr2::flash_add_greedy(data,K=Kmax,nullcheck=FALSE)
-    Y_hat = f_greedy$EL %*% t(f_greedy$EF)
-  }else if(method == "flash_gb"){
-    data = flashr2::flash_set_data(Y_data)
-    f_greedy = flashr2::flash_add_greedy(data,K=Kmax)
-    f = flashr2::flash_backfit(data,f_greedy)
-    Y_hat = f$EL %*% t(f$EF)
-  }else if(method == "flash_bf"){
-    data = flashr2::flash_set_data(Y_data)
-    f_data = flashr2::flash_add_factors_from_data(data,K = Kmax)
-    f = flashr2::flash_backfit(data,f_data)
-    Y_hat = f$EL %*% t(f$EF)
-  }else if(method == "PMD"){
-    res_log = capture.output({out = PMA.wrapper(Y_data,ngrids = 10,K = Kmax,fold = 10)})
-    if(length(out$d)==1){
-      Y_hat =  (out$d) * out$u %*% t(out$v)
-    }else{
-      Y_hat =  out$u %*% diag(out$d) %*% t(out$v)
-    }
-  }else if(method == 'PN'){
-    library(ebnm)
-    library(flashr2)
-    data = flashr2::flash_set_data(Y_data)
-    f_greedy = flashr2::flash_add_greedy(data,K=Kmax,ebnm_fn = ebnm_pn)
-    Y_hat = f_greedy$EL %*% t(f_greedy$EF)
-  }else if(method == 'PMD_d'){
-    out = PMA::PMD(Y_data, K = Kmax)
-    Y_hat =  out$u %*% diag(out$d) %*% t(out$v)
-  }else if(method == 'softImpute_d'){
-    gsoft = softImpute::softImpute(Y_data, rank.max = Kmax)
-    Y_hat =  gsoft$u %*% diag(gsoft$d) %*% t(gsoft$v)
-  }else if(method == "softImpute"){
-    gsoft = try(softImpute.wrapper(Y_data,ngrids = 10,K = Kmax,fold = 10))
-    if(length(gsoft$d)==1){
-      Y_hat =  try((gsoft$d) * gsoft$u %*% t(gsoft$v))
-    }else{
-      Y_hat =  try(gsoft$u %*% diag(gsoft$d) %*% t(gsoft$v))
-    }
-  }else if(method == "SF_flash"){
-    gsoft = try(softImpute.wrapper(Y_data,ngrids = 10,K = Kmax,fold = 10))
-    LL = gsoft$u %*% diag(gsoft$d)
-    FF = gsoft$v
-    data = flashr2::flash_set_data(Y_data)
-    f_data = flashr2::flash_add_lf(data, LL=LL, FF=FF)
-    f = flashr2::flash_backfit(data,f_data)
-    Y_hat = f$EL %*% t(f$EF)
-  }else{
-    # stop("the method is not implemented yet, please check it out")
-    Y_hat = matrix(0,ncol = dim(Y_data)[2],nrow = dim(Y_data)[1])
-  }
-  return(Y_hat)
-}
-
+library(ggplot2)
 ```
-
-read the data
-
-```{r}
-library(R.matlab)
-## run the code
-Y_centered = readMat("../data/output/missingdata/BreastCancer/example.mat")
-Y = Y_centered$Y
-# in the matlab package of NSF, the use the centered data by rows
-N = dim(Y)[1]
-P = dim(Y)[2]
-Y = Y - rowMeans(Y) %*% t(rep(1,P))
-```
-
-first we try to get the lambda based on the authors suggestion.
-
-```{r}
-OCVindex = OCV_index(Y,k_fold = 10)
-OCVdata = OCV_data(Y,OCVindex,k_fold = 10)
-
-colindex = OCVindex$colindex
-rowindex = OCVindex$rowindex
-foldindex = OCVindex$foldindex
-missing = OCVdata
-Y_data = missing[1,,]
-
-out = softImpute(Y_data, rank.max = 40,lambda = 80)
-length(out$d)
-out = softImpute(Y_data, rank.max = 40,lambda = 30)
-length(out$d)
-out = softImpute(Y_data, rank.max = 40,lambda = 20)
-length(out$d)
-out = softImpute(Y_data, rank.max = 40,lambda = 10)
-length(out$d)
-out = softImpute(Y_data, rank.max = 40,lambda = 9)
-length(out$d)
-```
-
-so it seems 10 or 11 should be a good choice which "Ideally lambda should be chosen so that the solution reached has rank slightly less than rank.max"
-
-## OCV choice
-
-for the grids we choose in our paper, we use 0-100 with ten grids
-
-```{r}
-library(softImpute)
-N = dim(Y_data)[1]
-P = dim(Y_data)[2]
-ngrids = 10
-fold = 10
-K = 40
-c_s = seq(0,100,len=ngrids)
-cvout = CVPMD_softImpute(Y_data,c_s,K ,fold , method = "softImpute")
-cvout
-```
-
-the result fit the authors' suggestion well (we choose $\lambda = 11$).
-
-But can we try smaller grids with OCV
-
-```{r}
-library(softImpute)
-N = dim(Y_data)[1]
-P = dim(Y_data)[2]
-ngrids = 10
-fold = 10
-K = 40
-c_s = seq(0,15,len=ngrids)
-cvout = CVPMD_softImpute(Y_data,c_s,K ,fold , method = "softImpute")
-cvout
-```
-
-At this time, we can see that we choose $\lambda = 5$ this time.
-
-## check the result
-
-```{r}
-out = softImpute(Y_data, rank.max = 40,lambda = 10)
-miss_hat = out$u %*% diag(out$d) %*% t(out$v)
-SSE = 0
-for(j in 1:fold){
-      SSE = SSE + sum((Y[rowindex[,foldindex[j,1,1]],colindex[,foldindex[j,1,2]]] -
-                               miss_hat[rowindex[,foldindex[j,1,1]],colindex[,foldindex[j,1,2]]])^2,na.rm = TRUE)
-}
-sqrt(SSE/sum(is.na(Y_data)))
-```
-
-```{r}
-out = softImpute(Y_data, rank.max = 40,lambda = 5)
-miss_hat = out$u %*% diag(out$d) %*% t(out$v)
-SSE = 0
-for(j in 1:fold){
-      SSE = SSE + sum((Y[rowindex[,foldindex[j,1,1]],colindex[,foldindex[j,1,2]]] -
-                               miss_hat[rowindex[,foldindex[j,1,1]],colindex[,foldindex[j,1,2]]])^2,na.rm = TRUE)
-}
-sqrt(SSE/sum(is.na(Y_data)))
-```
-
-we can see that the result is slightly improved.
 

### 2018-02-11 analysis/gtex_si.rmd
Files commited by wflow_commit.

--- 2018-01-30 analysis/gtex_si.rmd
@@
   data_L$tissue = factor(data_L$tissue,levels = 1:44 ,labels = as.character(colordata[,1]) )
-  data_L$loading = factor(data_L$loading,levels = 1:nL ,labels = paste("loading",offset+(1:nL),"; pve:", round(pve,3)))
+  data_L$loading = factor(data_L$loading,levels = 1:nL ,labels = paste("factor",offset+(1:nL),"; pve:", round(pve,3)))
 
@@
     theme(legend.position="right", legend.text=element_text(size=9), axis.text.y = element_text(size = 5)) + 
-    labs(title = "GTEx data", y = "loading values" ,x = "tissues", fill="tissue") +
+    labs(title = "GTEx data", y = "factor values" ,x = "tissues", fill="tissue") +
     facet_wrap(~loading, ncol = 2, scales = "free_y") +

### 2018-02-11 analysis/CVtemplate.Rmd
Files commited by wflow_commit.

--- 2018-01-06 analysis/CVtemplate.Rmd
@@
 author: "Wei"
-date: 10-03
-output: html_document
+output: 
+  html_document:
+    code_folding: hide
 ---
@@
   Y_miss[missindex$rowindex, missindex$colindex] = NA
-  data = flashr2::flash_set_data(Y_miss)
-  g_flash = flashr2::flash_add_greedy(data,Kmax=rank_max,ash_param=list(method = "fdr"),var_type = type)
+  data = flashr::flash_set_data(Y_miss)
+  g_flash = flashr::flash_add_greedy(data,Kmax=rank_max,ash_param=list(method = "fdr"),var_type = type)
   misshat_flash = g_flash$EL %*% t(g_flash$EF)
@@
   Y_miss[missindex$rowindex, missindex$colindex] = NA
-  data = flashr2::flash_set_data(Y_miss)
-  g_flash = flashr2::flash_add_greedy(data,Kmax=rank_max,ash_param=list(method = "fdr"),var_type = type)
+  data = flashr::flash_set_data(Y_miss)
+  g_flash = flashr::flash_add_greedy(data,Kmax=rank_max,ash_param=list(method = "fdr"),var_type = type)
   b_flash = flash_backfit(data,g_flash)
@@
 
-```{r}
+```
 library(R.matlab)
 # call the R functions
-source("~/HG/flash/data/CVmissflashr2/template/Rfunction.R")
+source("~/HG/flash/data/CVmissflashr/template/Rfunction.R")
 # read the data
 # you need to change the path for different data
-Y_data = R.matlab::readMat("~/HG/flash/data/CVmissflashr2/Breastcancer/example.mat")
+Y_data = R.matlab::readMat("~/HG/flash/data/CVmissflashr/Breastcancer/example.mat")
 Y = Y_data$Y
@@
 # this for the folder of the data
-system("matlab -nosplash -nodesktop -r \"addpath(\'~/HG/flash/data/CVmissflashr2/template\'); run(\'missvalue.m\');exit;\" ")
+system("matlab -nosplash -nodesktop -r \"addpath(\'~/HG/flash/data/CVmissflashr/template\'); run(\'missvalue.m\');exit;\" ")
 g_nsfa = readMat("./NBSFAout/NSFAresult.mat")
@@
 #SBATCH --job-name=arrayJob
-#SBATCH --output=/home/weidong/HG/flash/data/CVmissflashr2/Breastcancer/outlog/arrayJob_%A_%a.out
-#SBATCH --error=/home/weidong/HG/flash/data/CVmissflashr2/Breastcancer/outlog/arrayJob_%A_%a.err
+#SBATCH --output=/home/weidong/HG/flash/data/CVmissflashr/Breastcancer/outlog/arrayJob_%A_%a.out
+#SBATCH --error=/home/weidong/HG/flash/data/CVmissflashr/Breastcancer/outlog/arrayJob_%A_%a.err
 #SBATCH --array=1-100
@@
 # Print this sub-job's task ID
-cd /home/weidong/HG/flash/data/CVmissflashr2/Breastcancer
+cd /home/weidong/HG/flash/data/CVmissflashr/Breastcancer
 mkdir test${SLURM_ARRAY_TASK_ID}
 cd test${SLURM_ARRAY_TASK_ID}
-Rscript --verbose /home/weidong/HG/flash/data/CVmissflashr2/Breastcancer/runR.R
+Rscript --verbose /home/weidong/HG/flash/data/CVmissflashr/Breastcancer/runR.R
 ```

### 2018-02-11 analysis/GTEXzscoreTemplate.Rmd
Files commited by wflow_commit.

@@
 author: "wei"
-date: 09-29
 output: 
@@
 
-1 please git clone the `flashr2` on PPS.
+1 please git clone the `flashr` on PPS.
 
-`git clone https://github.com/stephenslab/flashr2.git`
+`git clone https://github.com/stephenslab/flashr.git`
 
-2 install `flashr2` locally
+2 install `flashr` locally
 
-`R CMD build flashr2`
+`R CMD build flashr`
 
-`R CMD INSTALL flashr2_0.2-2.tar.gz`
+`R CMD INSTALL flashr_0.2-2.tar.gz`
 
@@
 ```
-mkdir testflashr2
-cd testflashr2
-cp gtexEQTL_zscore.rds GTEX/testflashr2/
+mkdir testflashr
+cd testflashr
+cp gtexEQTL_zscore.rds GTEX/testflashr/
 ```
@@
 library(ashr)
-library(flashr2)
+library(flashr)
 load("./gtexEQTL_zscore.rds")
@@
 #SBATCH --job-name=arrayJob
-#SBATCH --output=/home/weidong/HG/flash/data/GTEX/testflashr2/outlog/arrayJob_%A_%a.out
-#SBATCH --error=/home/weidong/HG/flash/data/GTEX/testflashr2/outlog/arrayJob_%A_%a.err
+#SBATCH --output=/home/weidong/HG/flash/data/GTEX/testflashr/outlog/arrayJob_%A_%a.out
+#SBATCH --error=/home/weidong/HG/flash/data/GTEX/testflashr/outlog/arrayJob_%A_%a.err
 #SBATCH --array=1
@@
 # Print this sub-job's task ID
-cd /home/weidong/HG/flash/data/GTEX/testflashr2
+cd /home/weidong/HG/flash/data/GTEX/testflashr
 Rscript --verbose flashwrapper.R
@@
 
-you can substitute the folder path `/home/weidong/HG/flash/data/GTEX/testflashr2`. and we call this `.sbatch` file as `runflash.sbatch`
+you can substitute the folder path `/home/weidong/HG/flash/data/GTEX/testflashr`. and we call this `.sbatch` file as `runflash.sbatch`
 
@@
 ## plot
+
+Get the result from `flashr` as `bflashvarcol.rds` (backfitting result).
 
@@
   guides(fill = guide_legend(ncol = 1, keyheight = 0.8, keywidth = 0.3))
-  #ggsave("flashr2GTEx1.pdf", width = 8, height = 11)
+  #ggsave("flashrGTEx1.pdf", width = 8, height = 11)
 
@@
   guides(fill = guide_legend(ncol = 1, keyheight = 0.8, keywidth = 0.3))
-  #ggsave("flashr2GTEx2.pdf", width = 8, height = 10)
+  #ggsave("flashrGTEx2.pdf", width = 8, height = 10)
 ```
 
-the plots are
-
-<img src="figure/flashr2GTEx1.pdf" alt="the plot is in analysis/flashr2GTEx1.pdf"  width="4200" height="4200">
-
-<img src="figure/flashr2GTEx2.pdf" alt="the plot is in analysis/flashr2GTEx1.pdf"  width="4200" height="4200">
-
-## an example to run SFA on RCC or PPS
+## An example to run SFA on RCC or PPS
 

### 2018-02-11 analysis/MSErank1sim.Rmd
Files commited by wflow_commit.

--- 2018-01-25 analysis/MSErank1sim.Rmd
@@
 author: "wei wang"
-date: 17-12-09
 output: 
@@
   library(ebnm)
-  library(flashr2)
-  data =  flashr2::flash_set_data(Y)
-  g_flash = flashr2::flash_r1(data,verbose=F,var_type = "constant",ebnm_fn = ebnm_pn)
+  library(flashr)
+  data =  flashr::flash_set_data(Y)
+  g_flash = flashr::flash_r1(data,verbose=F,var_type = "constant",ebnm_fn = ebnm_pn)
   Y_hat = g_flash$EL %*% t(g_flash$EF)
@@
   P = dim(Y)[2]
-  data =  flashr2::flash_set_data(Y)
-  g_flash = flashr2::flash_r1(data,verbose=F,var_type = "constant")
+  data =  flashr::flash_set_data(Y)
+  g_flash = flashr::flash_r1(data,verbose=F,var_type = "constant")
   Y_hat = g_flash$EL %*% t(g_flash$EF)
@@
 library(PMA)
-library(flashr2)
+library(flashr)
 library(ssvd)
@@
 ```{r}
-plot_res = function(output,title = "data",legend_position = "none", methods_name){
+plot_res = function(output,title = "data",legend_position = "none", methods_name,myColors){
+  library(scales)
+  modulus_trans <- function(lambda){
+   trans_new("modulus",
+             transform = function(y){
+                if(lambda != 0){
+                   yt <- sign(y) * abs(y)^(lambda)
+                } else {
+                   yt = sign(y) * (log(abs(y) + 1))
+                }
+                return(yt)
+             },
+             inverse = function(yt){
+                if(lambda != 0){
+                   y <- abs(yt)^(1/lambda) * sign(yt)
+                } else {
+                   y <- (exp(abs(yt)) - 1) * sign(yt)
+
+                }
+                return(y)
+             }
+             )
+  }
   rmse = as.vector(output)
@@
   # geom_violin()+
-  ggtitle(title) +  theme_bw()+
+  ggtitle(title) +  theme_bw()+ scale_color_manual(values=myColors)+
     theme(legend.position= legend_position, legend.text=element_text(size=10),
-          axis.text.y = element_text(size =12),
-          axis.text.x = element_text(size =12,angle = 45, hjust = 1))
+          axis.text.y = element_text(size =12.9),
+          axis.text.x = element_text(size =15,angle = 45, hjust = 1),
+          plot.title = element_text(size = 12.9, face = "bold")) + 
+           scale_y_continuous(limits = c(-0.02,0.8), trans = modulus_trans(lambda = 0.5))
   p
@@
 methods_name = colnames(sparse_diff)
-p1 = plot_res(sparse_diff,title = "Difference from FLASH result (90% zeros)",methods_name = methods_name)
+fill_color = c("red","yellow3", "springgreen", "springgreen3", "cyan","cyan3", "deepskyblue","violet","purple")
+p1 = plot_res(sparse_diff,title = "Difference from flash result (90% zeros)",
+              methods_name = methods_name, myColors = fill_color)
 methods_name = colnames(sparse_diff)[c(1,2,3,4,5,7,9)]
-p1_cut = plot_res(sparse_diff[-22,c(1,2,3,4,5,7,9)],title = "Difference from FLASH result (90% zeros)",methods_name = methods_name)
+p1_cut = plot_res(sparse_diff[-22,c(1,2,3,4,5,7,9)],title = "Difference from FLASH result (90% zeros)",methods_name = methods_name,myColors = fill_color[c(1,2,3,4,5,7,9)])
 ```
@@
 methods_name = colnames(sparse_diff)
-p2 = plot_res(sparse_diff,title = "Difference from FLASH result (30% zeros)",methods_name = methods_name)
+p2 = plot_res(sparse_diff,title = "Difference from flash result (30% zeros)",methods_name = methods_name,myColors = fill_color)
 methods_name = colnames(sparse_diff)[c(1,2,3,4,5,7,9)]
-p2_cut = plot_res(sparse_diff[,c(1,2,3,4,5,7,9)],title = "Difference from FLASH result (30% zeros)",methods_name = methods_name)
+p2_cut = plot_res(sparse_diff[,c(1,2,3,4,5,7,9)],title = "Difference from FLASH result (30% zeros)",methods_name = methods_name,myColors = fill_color[c(1,2,3,4,5,7,9)])
 ```
@@
 methods_name = colnames(sparse_diff)
-p3 = plot_res(sparse_diff,title = "Difference from FLASH result (0% zeros)",methods_name = methods_name)
+p3 = plot_res(sparse_diff,title = "Difference from flash result (0% zeros)",methods_name = methods_name,myColors = fill_color )
 methods_name = colnames(sparse_diff)[c(1,2,3,4,5,7,9)]
-p3_cut = plot_res(sparse_diff[,c(1,2,3,4,5,7,9)],title = "Difference from FLASH result (0% zeros)",methods_name = methods_name)
+p3_cut = plot_res(sparse_diff[,c(1,2,3,4,5,7,9)],title = "Difference from FLASH result (0% zeros)",methods_name = methods_name,myColors = fill_color[c(1,2,3,4,5,7,9)])
 ```
@@
 # gridExtra::grid.arrange(p1,p2,p3,p1_cut,p2_cut,p3_cut,ncol = 3)
-gridExtra::grid.arrange(p1,p1_cut,p2,p2_cut,p3,p3_cut,ncol = 2)
+# gridExtra::grid.arrange(p1,p1_cut,p2,p2_cut,p3,p3_cut,ncol = 2)
+gridExtra::grid.arrange(p1,p2,p3,ncol = 3)
 ```

### 2018-02-11 analysis/MSErank3sim.Rmd
Files commited by wflow_commit.

--- 2018-01-25 analysis/MSErank3sim.Rmd
@@
 author: "wei wang"
-date: 18-01-23
 output: 
@@
   library(ebnm)
-  library(flashr2)
-  data =  flashr2::flash_set_data(Y)
-  g_flash = flashr2::flash_add_greedy(data,verbose=F,var_type = "constant",ebnm_fn = ebnm_pn, K = K)
+  library(flashr)
+  data =  flashr::flash_set_data(Y)
+  g_flash = flashr::flash_add_greedy(data,verbose=F,var_type = "constant",ebnm_fn = ebnm_pn, K = K)
   Y_hat = g_flash$EL %*% t(g_flash$EF)
@@
   P = dim(Y)[2]
-  data =  flashr2::flash_set_data(Y)
-  g_flash = flashr2::flash_add_greedy(data,verbose=F,var_type = "constant", K = K)
+  data =  flashr::flash_set_data(Y)
+  g_flash = flashr::flash_add_greedy(data,verbose=F,var_type = "constant", K = K)
   Y_hat = g_flash$EL %*% t(g_flash$EF)
@@
   # geom_violin()+
-  ggtitle(title) +  theme_bw()+
+  ggtitle(title) +  theme_bw()+ scale_color_manual(values=myColors)+
     theme(legend.position= legend_position, legend.text=element_text(size=10),
-          plot.title = element_text(size = 18, face = "bold"),
-          axis.text.y = element_text(size =15), 
-          axis.text.x = element_text(size =15,angle = 45, hjust = 1))
+          plot.title = element_text(size = 12.9, face = "bold"),
+          axis.text.y = element_text(size =12), 
+          axis.text.x = element_text(size =12,angle = 45, hjust = 1))
   p
@@
 methods_name = colnames(sparse_diff)
-p3 = plot_res(sparse_diff,title = "Rank 3 example",methods_name = methods_name)
+fill_color =  c("red","yellow3", "springgreen", "springgreen3", "cyan","cyan3", "deepskyblue","violet","purple")
+p3mse = plot_res(sparse_diff,title = "Box plot of RRMSE compared with flash",methods_name = methods_name,
+              myColors = fill_color[c(1,2,3,7,8,9)])
 ```
@@
 ```{r}
-p3
+p3mse
 ```

### 2018-02-11 analysis/index.Rmd
Files commited by wflow_commit.

@@
 
+[old experiments](oldanalysis.html)
+
 ## Simulation studies
@@
 
-### Apply flashr2 for the examples in FLASH paper
+### Apply flashr for the examples in FLASH paper
 
 #### simulations
-
-Click here to check the [simulatoin result (rank one and rank three)](applyflahr2onsimulations.html)
 
@@
 
-Here is a example to run `flashr2` on RCC [RCC template](GTEXzscoreTemplate.html)
+Here is a example to run `flashr` on RCC [RCC template](GTEXzscoreTemplate.html)
 
@@
 click here to check the template for [missing data imputation with OCV](templateforOCV.html) and [CV template](CVtemplate.html)
-
-click here to check the comparison of different options for FLASH [different shrink options](compareOptionsOCV.html)

### 2018-02-11 analysis/setup.Rmd
Files commited by wflow_commit.

@@
 
-*Fill out setup instructions in this section. See
-[here](https://stephenslab.github.io/wflow-divvy/setup.html) and
-[here](https://github.com/pcarbo/mouthwash_sims/blob/master/README.md)
-for some ideas.*
+### Install software and R packages
+
+1. Install [R](https://cran.r-project.org).
+
+2. Install [MATLAB](https://www.mathworks.com/products/matlab.html)
+
+3. Install the required R packages (run the following commands in the R).
+
+```R
+install.packages(c("devtools", "ashr", "PMA","ssvd","softImpute","Matrix","R.matlab","ggplot2","denoiseR"))
+devtools::install_github("stephenslab/flashr")
+```
+
+4. Install MATLAB package [NBSFA](https://github.com/davidaknowles/nsfa)
+
+5. Install packages [sfa](http://stephenslab.uchicago.edu/software.html#sfa) and [SFAmix](http://beehive.cs.princeton.edu/software/)
+
+### Data sets
+
+All the data sets used in the paper are in [data files](https://github.com/NKweiwang/flash_workflow/tree/master/data/output/missingdata)
 
@@
 
+```{r,echo=FALSE}
+library(flashr)
+library(ssvd)
+library(methods)
+library(PMA)
+library(R.matlab)
+library(Matrix)
+library(softImpute)
+library(ggplot2)
+library(denoiseR)
+```
+
 ```{r session-info}

### 2018-02-11 analysis/simulation3.Rmd
Files commited by wflow_commit.

--- 2018-01-25 analysis/simulation3.Rmd
@@
 author: "Wei Wang"
-date: 2017-04-04
 output: 
@@
   library(ebnm)
-  library(flashr2)
-  data =  flashr2::flash_set_data(Y)
-  f_greedy = flashr2::flash_add_greedy(data,verbose=F,var_type = "constant", K = K)
-  f = flashr2::flash_backfit(data,f_greedy,var_type = "constant")
+  library(flashr)
+  data =  flashr::flash_set_data(Y)
+  f_greedy = flashr::flash_add_greedy(data,verbose=F,var_type = "constant", K = K)
+  f = flashr::flash_backfit(data,f_greedy,var_type = "constant")
   Y_hat = f$EL %*% t(f$EF)
@@
 
+## average of 100 simulations
+
 ```{r}
-library(PMA)
-library(flashr2)
-library(ssvd)
-library(softImpute)
-Y_data = DataMaker(seed_num = 0)
+HMatrix = readRDS("../data/simulation/rankthree/HeatMap.rds")
+T = 100
+AVG_HM = matrix(0,nrow = 36000,ncol = 7)
+for(i in 1:T){
+  AVG_HM = AVG_HM + abs(HMatrix[i,,])
+}
+AVG_HM = AVG_HM/100
+```
 
-LF_1 = PMA.wrapper(Y_data)
-LF_2 = flash.wrapper(Y_data)
-LF_3 = flashBF.wrapper(Y_data)
-LF_4 = SVD.wrapper(Y_data)
-LF_5 = SSVD.wrapper(Y_data)
-LF_6 = softImpute.wrapper(Y_data)
-LF_true = Y_data$L_true %*% t(Y_data$F_true)
-
+```{r}
+LF_1 = matrix(AVG_HM[,1],ncol = 240, nrow = 150)
+LF_2 = matrix(AVG_HM[,2],ncol = 240, nrow = 150)
+LF_3 = matrix(AVG_HM[,3],ncol = 240, nrow = 150)
+LF_4 = matrix(AVG_HM[,4],ncol = 240, nrow = 150)
+LF_5 = matrix(AVG_HM[,5],ncol = 240, nrow = 150)
+LF_6 = matrix(AVG_HM[,6],ncol = 240, nrow = 150)
+LF_true = matrix(AVG_HM[,7],ncol = 240, nrow = 150)
 library(reshape2)
@@
    geom_tile(color = "white")+
-    scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
-     midpoint = 0, limit = c(-2,2), space = "Lab") + labs(title = title_name, y = "samples", x = "variables") +
-    theme_minimal() + theme(legend.position="none") 
+    #scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
+     #midpoint = 0, limit = c(-0.81,0.81), space = "Lab") + labs(title = title_name, y = "samples", x = "variables") +
+    scale_fill_gradient2(low = "white", high = "grey6", mid = "grey9", 
+     midpoint = 0.41, limit = c(0,0.82), space = "Lab") + labs(title = title_name, y = "samples", x = "variables") +
+    theme_minimal() + theme(legend.position="none",plot.title = element_text(size = 12.9, face = "bold")) 
   p1
@@
 
-p0 = plot_HP(title_name = "Truth", melted_cormat = melted_cormat_true)
+p0 = plot_HP(title_name = "Latent structure (truth)", melted_cormat = melted_cormat_true)
 p1 = plot_HP(title_name = "PMD.cv1", melted_cormat = melted_cormat_1)
 p2 = plot_HP(title_name = "flash", melted_cormat = melted_cormat_2)
-p3 = plot_HP(title_name = "flashb", melted_cormat = melted_cormat_3)
+p3 = plot_HP(title_name = "flash_pn", melted_cormat = melted_cormat_3)
 p4 = plot_HP(title_name = "SVD", melted_cormat = melted_cormat_4)
@@
 p6 = plot_HP(title_name = "SI.cv", melted_cormat = melted_cormat_6)
+
 ```
@@
 ```{r}
-gridExtra::grid.arrange(p0,p5,p6,p4,p1,p2,ncol = 2)
+gridExtra::grid.arrange(p2,p3,p1,p6,p5,p4,ncol = 3)
 ```

### 2018-02-11 analysis/templateforOCV.Rmd
Files commited by wflow_commit.

@@
 ---
-title: "apply flashr2 with OCV"
+title: "apply flashr with OCV"
 author: "wei wang"
-date: 09-30
 output: 
@@
   if(method == "flash"){
-    data = flashr2::flash_set_data(Y_data)
-    f_greedy = flashr2::flash_add_greedy(data,K=Kmax)
+    data = flashr::flash_set_data(Y_data)
+    f_greedy = flashr::flash_add_greedy(data,K=Kmax)
     Y_hat = f_greedy$EL %*% t(f_greedy$EF)
   }else if(method == "flash_wn"){
-    data = flashr2::flash_set_data(Y_data)
-    f_greedy = flashr2::flash_add_greedy(data,K=Kmax,nullcheck=FALSE)
+    data = flashr::flash_set_data(Y_data)
+    f_greedy = flashr::flash_add_greedy(data,K=Kmax,nullcheck=FALSE)
     Y_hat = f_greedy$EL %*% t(f_greedy$EF)
   }else if(method == "flash_gb"){
-    data = flashr2::flash_set_data(Y_data)
-    f_greedy = flashr2::flash_add_greedy(data,K=Kmax)
-    f = flashr2::flash_backfit(data,f_greedy)
+    data = flashr::flash_set_data(Y_data)
+    f_greedy = flashr::flash_add_greedy(data,K=Kmax)
+    f = flashr::flash_backfit(data,f_greedy)
     Y_hat = f$EL %*% t(f$EF)
   }else if(method == "flash_bf"){
-    data = flashr2::flash_set_data(Y_data)
-    f_data = flashr2::flash_add_factors_from_data(data,K = Kmax)
-    f = flashr2::flash_backfit(data,f_data)
+    data = flashr::flash_set_data(Y_data)
+    f_data = flashr::flash_add_factors_from_data(data,K = Kmax)
+    f = flashr::flash_backfit(data,f_data)
     Y_hat = f$EL %*% t(f$EF)
@@
     library(ebnm)
-    library(flashr2)
-    data = flashr2::flash_set_data(Y_data)
-    f_greedy = flashr2::flash_add_greedy(data,K=Kmax,ebnm_fn = ebnm_pn)
+    library(flashr)
+    data = flashr::flash_set_data(Y_data)
+    f_greedy = flashr::flash_add_greedy(data,K=Kmax,ebnm_fn = ebnm_pn)
     Y_hat = f_greedy$EL %*% t(f_greedy$EF)
@@
     FF = gsoft$v
-    data = flashr2::flash_set_data(Y_data)
-    f_data = flashr2::flash_add_lf(data, LL=LL, FF=FF)
-    f = flashr2::flash_backfit(data,f_data)
+    data = flashr::flash_set_data(Y_data)
+    f_data = flashr::flash_add_lf(data, LL=LL, FF=FF)
+    f = flashr::flash_backfit(data,f_data)
     Y_hat = f$EL %*% t(f$EF)
@@
 
-In this case, we have problem in running `flashr2` on RCC. so we just try other methods first.
+In this case, we have problem in running `flashr` on RCC. so we just try other methods first.
 
@@
 saveRDS(result, "./NBSFAout/output.rds")
+print(sessionInfo())
 ```
@@
 library(ggplot2)
-plot_res = function(output,title = "data",legend_position = "none", x_label){
+plot_res = function(output,title = "data",legend_position = "none", x_label, myColors){
   rmse = as.vector(output)
@@
     # geom_violin()+
-    ggtitle(title) +  theme_bw()+
+    ggtitle(title) +  theme_bw()+  scale_color_manual(values=myColors)+
     theme(legend.position= legend_position, legend.text=element_text(size=15),
@@
 setwd("~/HG/flash_workflow/analysis")
+
+fill_color = c("red","yellow3", "tan1","springgreen3", "springgreen","deepskyblue3","deepskyblue")
 
@@
 x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
-plotM = plot_res(output,"MovieLens data",x_label = x_label)
+plotM = plot_res(output,"MovieLens data",x_label = x_label,myColors = fill_color)
 
@@
 x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
-plotB = plot_res(output,"Breast Cancer data",x_label = x_label)
+plotB = plot_res(output,"Breast cancer data",x_label = x_label,myColors = fill_color)
 
@@
 x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
-plotG = plot_res(output,"GTEx data",x_label = x_label)
+plotG = plot_res(output,"GTEx data",x_label = x_label,myColors = fill_color)
 
@@
 x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
-plotP = plot_res(output,"Text data",x_label = x_label)
+plotP = plot_res(output,"Presidential address data",x_label = x_label,myColors = fill_color)
 
@@
 x_label= c("flash","flash.pn","PMD","PMD.cv1","NBSFA","SI","SI.cv")
-plotT = plot_res(output,"Tumor data",x_label = x_label)
+plotT = plot_res(output,"Tumor data",x_label = x_label,myColors = fill_color)
+```
 
+```{r}
 # gridExtra::grid.arrange(plotP,plotT,plotG,plotB,plotM, layout_matrix = rbind(c(1,NA,2),c(NA,5,NA),c(4,NA,3)))
 gridExtra::grid.arrange(plotM,plotT,plotG,plotB,plotP, layout_matrix = rbind(c(1,2,3),c(4,5,NA) ))
-
 ```

### 2018-02-11 analysis/OCVmissdataSetting.Rmd
Files commited by wflow_commit.

@@
 author: "wei wang"
-date: 2018-01-25
 output: html_document
@@
 <!-- Add your analysis here -->
+
+## software
+
+### SFAmix
+
+http://beehive.cs.princeton.edu/software/
+
+### sfa
+
+http://stephenslab.uchicago.edu/software.html#sfa
+
+### NBSFA
+
+https://github.com/davidaknowles/nsfa
+
+### R packages
+
+```{r}
+library(ssvd)
+library(PMA)
+library(softImpute)
+```
 
